### 2020-02-05 paper/paper.tex
Report outline

\documentclass[a4paper]{article}

\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{float}
\usepackage{hyperref}
\setlength\parindent{0pt}

\title{Footpath Planning for Example-driven Procedural Road Networks}

\author{Alexander Hjelm}

\date{\today}

\begin{document}
\maketitle

\section{Abstract}

\begin{figure}[H]
\centering
\includegraphics[height=0.5\textheight]{test_image}
\caption{En typisk rymdhiss enligt Wikipedia. Bildkälla (\cite{sampleitem1}).}
\label{fig:space}
\end{figure}

This report covers... \cite{sampleitem2}

\begin{thebibliography}{9}
\bibitem{sampleitem1}
Title 1, \\ 
\href{https://en.wikipedia.org/wiki/Geostationary\_ orbit}{https://en.wikipedia.org/wiki/Geostationary\_ orbit}.

\bibitem{sampleitem2}
Title 2, \\ 
\href{https://sv.wikipedia.org/wiki/SpaceX}{https://sv.wikipedia.org/wiki/SpaceX}.

\end{thebibliography}

\end{document}

### 2020-02-05 paper/paper.tex
Paper: Introduciton content

@@
 
+\section(Introduction)
+
+\subsection(Research Question)
+
+This research question deals with the procedural generation of road networks from example data sets, and the edge cases that arise when dealing with footpath planning.
+A common problem is that footpaths are in their nature placed densely.
+How can the program identify critical areas where the footpath width is so large that it collides with existing roads?
+Can this be done so efficiently that the example-driven program can still run in real time?
+An additional question that arises is how much road data will be necessary to create believable procedural road networks for new zones in the Stockholm metropolitan area.
+How can this data be sampled and processed efficiently?
+
+The previous work by Nishida et.
+al.
+has shown that it is possible to create believable procedural road maps from examples.
+Their program has created a network of 35 km road in less than a second, so for the application of zoning new city areas my hypothesis is that such a program can be made to run in real time with such minimal waiting time that it can be used in a real application by a city planner.
+
+The challenge lies in how to extract the footpath data and tie it together with the existing road map.
+the work by Nishida et.
+al.
+only deals with arterial and secondary roads, and I will be adding a tertiary layer of roads, wich is highly possible, but may increase the data consumption and program running time.
+The same applies for the statistical model that identifies colliding roads, which I will be adding on top of that.
+
+\subsection(Problem Constraints)
+
+\begin{enumerate}
+\item The road network is assumed to be a 2D simple graph.
+This means that any overlapping roads, such as tunnels and bridges that cross over street-level roads, will be eliminated from the training set.
+Any self-connected nodes will also be eliminated from the training set.
+\item It will be assumed that the city map consists of only arterial (primary) roads, secondary roads and footpaths.
+\item It will be assumed that the primary, secondary and tertiatry roads all have a fixed width. Primary roads have a set width, and secondary roads and footpaths as well.
+\item It will be assumed that the use case of the application is for generating areas with the size of a neighbourhood or a few city blocks.
+The area that the user will sketch will typically be around 0.1 km2.
+\item The program will not handle terrain features and altitude. It will be assume that the road mesh is projected on a flat 2D surface.
+\end{enumerate}
+
+\subsection(Evaluation)
+
+The question will be examined primarily by creating a piece of software that can generate urban maps with primary and secondary roads, as well as footpaths, from example.
+I will experiment with different methods of collecting road data as well as different sample amounts and densities.
+
+The evaluation plan will consist of programmatic evaluations and quantifyable metrics.
+My main idea is to sample metrics about intersections only (number of connecting roads, the angles between them, distance to connecting intersections, curviness of roads, etc).
+If I can sample all that and create a distribution for a few different cities, I have the two following hypothesises:
+  1. There should be selection of features that produce an observable difference in the distributions between cities that look vastly different.
+  2. If I let my program generate a large number of city parcels from a dataset, the resulting feature distribution should look like to the source dataset, and unlike datasets of other cities.
+These distributions could be created for primary, secondary roads and footpaths separately.
+
+\subsection(Background)
+
+Procedural modeling of road networks is of interest to the fields of urban planning and urban architecture.
+In recent years there has been an advent of interactive tools that aid urban planners in placing or generating features of an urban plan, particularly roads and road networks.
+Additionally, the topic is of interest for 3D content creators in for example the fields of video game design or 3D animation.
+A trend in these fields is using methods of procedural modeling to create large amounts of 3D content quickly and efficiently, while requiring as little hand modeling as possible.
+Studios who create large worlds are often interested in solutions that generate large road networks without any hand modeling.
+
+\subsection(ESAL, KTH)
+The thesis work will be carried out at the Embodied Social Agents Lab (ESAL) at the Department of Computational Science and Technology (CST).
+The lab has before been working on systems for generating procedural urban environments, and are interested in the possibility of generating more detailed maps than have been done before, using example-based methods to preserve the aesthetic look of a greater metropolitan area.
+The lab is particularly interested in the generation of footpaths, since most commercial software that import road maps only import arterial and secondary car roads, and do not include pedestrian walkways, cycle paths and such smaller roads.
+% TODO: any official description of ESAL that can be used here?
+
 \begin{thebibliography}{9}

### 2020-02-05 paper/paper.tex
Paper section oops

@@
 
-\section(Introduction)
+\section{Introduction}
 
-\subsection(Research Question)
+\subsection{Research Question}
 
@@
 
-\subsection(Problem Constraints)
+\subsection{Problem Constraints}
 
@@
 
-\subsection(Evaluation)
+\subsection{Evaluation}
 
@@
 
-\subsection(Background)
+\subsection{Background}
 
@@
 
-\subsection(ESAL, KTH)
+\subsection{ESAL, KTH}
 The thesis work will be carried out at the Embodied Social Agents Lab (ESAL) at the Department of Computational Science and Technology (CST).

### 2020-02-05 unity-project/Assets/Scripts/GameManager.cs
Frame code for generating road meshes from points

@@ -15,6 +15,7 @@ public class GameManager : MonoBehaviour
 
     private RoadNodeCollection roadNodeCollection;
     private List<RoadNode> roadNodesList;
+    private Dictionary<RoadNode, RoadMesh> roadMeshes = new Dictionary<RoadNode, RoadMesh>();
 
     void Start()
     {
@@ -35,6 +36,17 @@ public class GameManager : MonoBehaviour
             }
         }
         roadNodesList = roadNodeCollection.BuildToList();
+
+        // Generate meshes and store them
+        foreach(RoadNode node in roadNodesList)
+        {
+            GameObject roadMeshObj = new GameObject();
+            RoadMesh roadMesh = roadMeshObj.AddComponent(typeof(RoadMesh)) as RoadMesh;
+            // TODO: Add more points here, go through all neighbours as long as there are only 2 for every node
+            // That way we can generate all simple roads
+            roadMesh.GenerateMeshFromPoints(new List<Vector2>() {node.GetPosAsVector2()});
+            roadMeshes[node] = roadMesh;
+        }
     }
 
     void Update()

### 2020-02-05 unity-project/Assets/Scripts/RoadMesh.cs

@@ -1,0 +1,18 @@
+﻿using System.Collections;
+using System.Collections.Generic;
+using UnityEngine;
+
+public class RoadMesh : MonoBehaviour
+{
+    private MeshRenderer _meshRenderer;
+
+    public void Awake()
+    {
+        _meshRenderer = gameObject.AddComponent(typeof(MeshRenderer)) as MeshRenderer;
+    }
+
+    public void GenerateMeshFromPoints(List<Vector2> points)
+    {
+
+    }
+}

### 2020-02-05 unity-project/Assets/Scripts/RoadNode.cs

@@ -1,4 +1,5 @@
 ﻿using System.Collections.Generic;
+using UnityEngine;
 
 public class RoadNode
 {
@@ -26,6 +27,11 @@ public class RoadNode
     public float GetY()
     {
         return _yCoord;
+    }
+
+    public Vector2 GetPosAsVector2()
+    {
+        return new Vector2(_xCoord, _yCoord);
     }
 
     public List<RoadNode> GetNeighbours()

### 2020-02-05 unity-project/Assets/Scripts/RoadMesh.cs
Add a mesh filter component to the newly created road meshes

@@ -4,15 +4,20 @@ using UnityEngine;
 
 public class RoadMesh : MonoBehaviour
 {
-    private MeshRenderer _meshRenderer;
+    private MeshFilter _meshFilter;
 
     public void Awake()
     {
-        _meshRenderer = gameObject.AddComponent(typeof(MeshRenderer)) as MeshRenderer;
+        _meshFilter = gameObject.AddComponent(typeof(MeshFilter)) as MeshFilter;
+        gameObject.AddComponent(typeof(MeshRenderer));
     }
 
     public void GenerateMeshFromPoints(List<Vector2> points)
     {
-
+        Mesh mesh = new Mesh();
+        GetComponent<MeshFilter>().mesh = mesh;
+        //mesh.vertices = newVertices;
+        //mesh.uv = newUV;
+        //mesh.triangles = newTriangles;
     }
 }

### 2020-03-06 paper/paper.tex
Paper: two columns
Add motivation and a few implementation steps to the report

@@
 \usepackage{hyperref}
+\usepackage{multicol}
 \setlength\parindent{0pt}
@@
 This report covers... \cite{sampleitem2}
+
+\begin{multicols}{2}
 
@@
 
+With that said, Nishida et. al. present an excellent solution for generating new road patches that comform to an underlying terrain function.
+
 \subsection{Evaluation}
@@
 
+Motivation
+
+Navigation agents, they might fail navigating when faced with a too narrow path or intersecting obstacles
+Using a road network to creates an underlying navmesh
+
 \subsection{ESAL, KTH}
@@
 % TODO: any official description of ESAL that can be used here?
+
+\section(OSM Data)
+
+To separate sidewalks from detached footpaths, I used the fact that OSM generally provides sidewalks with the same name as the adjacent motorway.
+Thus they can be paired together and measured against each other.
+
+\section{Implementation}
+
+In this implementation, roads of different types were allowed to merge into one another to form common intersections.
+To facilitate this, it was neccessary to work with a path-by-path representation of the point data.
+When two path share a node with common coordinates, they form an intersection which can be of multiple different types.
+
+\end{multicols}
 

### 2020-03-13 paper/paper.tex
Update paper with info on SLU data and coordinate system

@@
 This report covers... \cite{sampleitem2}
-
-\begin{multicols}{2}
 
@@
 
+\subsection(SLU Data)
+
+The SLU map is maintained by The Swedish National Land Survey (Swedish: Lantmäteriet).
+The SLU data is aqqcuired by land surveying methods such as GPS or DGNSS positioning, or by reproduction of features from ortophoto or stereo mapping from 3D aerial images.
+The map is updated continously by Lantmäteriet, in conjoncture with the forming or reforming of property.
+Property limits and building features have a position accuracy requirement of 5 meters.
+(Lantmäteriet, 2019)
+
+\subsection(A word on coordinate systems)
+
+The SLU dataset comes in the SWEREF 99 TM (Lantmäteriet, 2019).
+SWEREF 99 TM is a projected coordinate system and there is no linear transformation to WGS 48, which OpenStreetMap uses (OSM 19).
+The coordinate conversions for this paper were obtained using proj: a Linux commandline application for geospatial coordinate conversion.
+
 \section{Implementation}
@@
 
-\end{multicols}
+\begin{thebibliography}{9}
 
-\begin{thebibliography}{9}
-\bibitem{sampleitem1}
-Title 1, \\ 
-\href{https://en.wikipedia.org/wiki/Geostationary\_ orbit}{https://en.wikipedia.org/wiki/Geostationary\_ orbit}.
+\bibitem{lantmateriet-19}
+(Lantmäteriet, 2019), Produktbeskrivning: GSD-Fastighetskartan vektor \\ 
+\href{https://www.lantmateriet.se/globalassets/kartor-och-geografisk-information/kartor/fastshmi.pdf}.
 
-\bibitem{sampleitem2}
-Title 2, \\ 
-\href{https://sv.wikipedia.org/wiki/SpaceX}{https://sv.wikipedia.org/wiki/SpaceX}.
+\bibitem{osm-19}
+(OSM 19) Converting to WGS84, accessed 13-03-2020\\ 
+\href{https://wiki.openstreetmap.org/wiki/Converting_to_WGS84}.
 

### 2020-03-16 paper/paper.tex
Add sections about accuracy classes and datasets

@@
 
-\section(OSM Data)
+\section(The geodata used in this project)
+
+For the purpose of this project, the OpenStreetMap road dataset has been used in the presentation of the road network around Stockholm, and in the generative process of new road networks.
+To assess and correct path widths and path node locations, the OpenStreetMap building footprint dataset has also been used.
+To make an assessment on the quality and precision of the OpenStreetMap data, the SLU map from the Swedish National Land Survey was used in comparison with the OpenStreetMap map.
+The SLU map conatins only building footprints and property limits.
+Both map segments were obtained around the Stockholm greater metropolitan area.
+The exact map segments that were extracted were in both cases rectangles with the following coordinates:.
+TODO: Get coordinates
+
+\subsection(OSM Data)
+
+OpenStreetMap road data and building footprints are commonly obtained by manually tracing features in commercially available satellite images.
+Such images have a limit on their resolution which puts a theoretical limit on the accuracy of map features compared to their real-world equivalents (Haklay 2010).
+Particularily the high resolution imagery from Bing in 2010 led to an increase in building information in OpenStreetMap (Fan et al, 2014).
 
@@
 
+The SLU map contains 
 The SLU map is maintained by The Swedish National Land Survey (Swedish: Lantmäteriet).
@@
 
-The SLU dataset comes in the SWEREF 99 TM (Lantmäteriet, 2019).
+The SLU dataset is delivered in the SWEREF 99 TM (Lantmäteriet, 2019).
 SWEREF 99 TM is a projected coordinate system and there is no linear transformation to WGS 48, which OpenStreetMap uses (OSM 19).
 The coordinate conversions for this paper were obtained using proj: a Linux commandline application for geospatial coordinate conversion.
+
+\section{Evaluation of the quality of geodata}
+
+(Haklay 2010) specifices a comprehesive list of 8 accuracy classes when it comes to evaluating the quality of geographic information.
+
+\begin{itemize}
+  \item Lineage. This is the historical aspect of the dataset, which concerns the collection process and evolution.
+  \item Positional accuracy. This relates the coordinate value of an object in the database to the actual location of the ground in the real world.
+  \item Attibute accuracy. In a geographical database, objects are commonly tagged with meta-information. This class assesses how correct those values are.
+  \item Logical consistency - This assesses the internal consistency of the dataset.
+For every dataset there may be internal rules and relationships that objects and features must follow, and this class assesses the degree to which these are adhered to.
+  \item Completeness - This assesses the lack of data in a dataset, and the coverage of real-world objects. Objects or features may be missing from a dataset, which reduces its quality.
+  \item Semantic accuracy - This links the way in which an object or feature is recorded and represented to how it should be interpreted.
+  \item Usage, purpose and constraints - This concerns the validity of the dataset in relation to its purpose and how it is used.
+  \item Temporal quality - This assesses the validity of the dataset in relation to real-world changes over time.
+\end{itemize}
+
+Other than for the purpose of mentioning various sources or error, this project will focus exclusively on the Positional accuracy, Completeness and Semantic accuracy classes.
+The comparison of the OSM and SLU maps will focus on these accuracy classes to determine the error tolerance when manipulating features in the 3D map presentation and the road network generating phase.
+
+\subsection{The accuracy classes that will not be focused on}
+
+The lineage aspect of the OSM data is available through the OSM History Viewer, an online debugging tool that lets anyone freely view the change history of individual fetures in a commit history-like fashion.
+There is an option for editors to include a personal note with their changesets to include additional details or motivate why a change was made, but there is no guaratee that the data includes any information on the aquisition method used (OSM 2020).
+The history aspect of the SLU dataset is not freely available online, but the acquistion method is included in the file metadata on a per-object basis.
+Lantmäteriet uses internal codes to specifify the acquistion method, and these codes can be referred to in the product description the accompanies the map files.
+As the scope of this project is limited to making a broad comparison of the positional accuracy of the two datasets, only the required positional accuracy as described by the SLU product description will be used.
+TODO: Dig deeper in the SLU dataset, how can one find the method of aquisition for Stockholm buildings, and its precision?
+
+The attribute accuracy measure is interesting, as both maps use encoded metadata on a per feature basis, the attribute accuracy of the two datasets could be assessed by doing a simple feature comparison and creating a translation table between the two attribute maps.
+To the knowledge of the author, this has never been done with specifically these two datasets, at the time of writing.
+
+The temporal accuracy will not be assessed as this projects focuses on comparing a single snapshot of the two datasets at the same point in time.
+
+The logical accuracy will not be assessed because the underlying logical relations between features in both maps serve no purpose to this project.
 
@@
 \bibitem{osm-19}
-(OSM 19) Converting to WGS84, accessed 13-03-2020\\ 
-\href{https://wiki.openstreetmap.org/wiki/Converting_to_WGS84}.
+(OSM 19) Converting to WGS84, accessed 13-03-2020 \\ 
+\href{https://wiki.openstreetmap.org/wiki/Converting\_to\_WGS84}.
+
+\bibitem{osm-20}
+(OSM 20) OSM History Viewer\\
+\href{https://wiki.openstreetmap.org/wiki/OSM\_History\_Viewer}.
+
+\bibitem{haklay-10}
+(Haklay, 2010), How good is volunteered geographical information?A comparative study of OpenStreetMap and OrdnanceSurvey datasets \\
+\href{https://kfrichter.org/crowdsourcing-material/day1/haklay10.pdf}.
+
+\bitiem{fan-14}
+(Fan et al, 2014), Quality assessment for building footprints data on OpenStreetMap \\
+\href{https://www.researchgate.net/publication/262163378\_Quality\_assessment\_for\_building\_footprints\_data\_on\_OpenStreetMap}.
 

### 2020-03-17 paper/paper.tex
Add building correspondence function + details

@@
 
-The SLU map contains 
+The SLU map contains building features and property limits, but no road data.
 The SLU map is maintained by The Swedish National Land Survey (Swedish: Lantmäteriet).
@@
 
+\section{Dataset quality evaluation}
+
+//TODO: Feature mapping, how to, first of all
+
+\subsection{Building correspondence}
+
+The building footprint correspondence will be evaluated using a similarity function that assesses the turning funciton and relative area overlap.
+The similarity function between any building in the OSM data set and any building in the SLU datasets will be defined as a function of their respective footprints ($foot_{OSM}, foot_{SLU}$), as follows:
+$S(foot_{OSM}, foot_{SLU}) = S_{T}(foot_{OSM}, foot_{SLU}) + S_{RO}(foot_{OSM}, foot_{SLU})$, where $S_{T}$ measures the difference in turning function and $S_{RO}$ measures the relative overlap between the footprints.
+
+\subsection{Turning function definition}
+
+The turning function was first defined by Arkin et al.
+(1991), as a method for measuring the similarity of two polygons.
+The Turning funciton $T_c(l)$ measures the cumulative angle of the polygon's counter-clockwise tangent, as a function of the cumulative normalized length l.
+This project uses the turning function as it is defined by (Fan et al., 2014).
+For a polygon with vertices ${v_1 ...
+v_n}$ and line segments ${e_1 ...
+e_n}$ It is defined as follows.
+Fix a starting vertex $v_1$.
+The tangent angle at $v_1$ is $\theta_{n,1}$. This is the angle between the neighbouring line segments $e_n$ and $e_1$.
+For any $i$ such that $i>1$ and $n<i$, the tangent angle at $v_i$ is recursively defined as: $\theta_{i, i-1} = \theta_{n,1} + \sum^{i}_{k=1} \theta_{k, k-1}$
+The turning function has some nice geometric properties, in that it is invariant to both rotation and scaling of the polygon.
+The function contains no information of the orientation of the polygon, only of the relative angle between successive line segments, thus it does not change under roation.
+It also measures only the normalized cumulative length, which does not change under scaling.
+The similarity of two polygons A and B in terms of their turning function is defined as their distance of their cumulative turning functions: $S_{T}(A, B) = 1 - (\int^{1}_{0} T_{C,A}(l) - T_{C,B}(l) dl)^{1/2}$.
+The value range of $S(A, B)$ will be 0 to 1, where $S(A,B) = 1$ if the polygons are identical.
+
+//TODO: Completeness by building area, quotation
+
+\subsection{Correspondence by building area}
+
+(Fan et al., 2014) further discussed how the relative overlap in building area can be used to determine building correspondence between datasets.
+The relative building overlap between the OSM and SLU datasets will be defined as folows, given the footprints of any building in the OSM set ($foot_{OSM}$) and the SLU set ($foot_{SLU}$):
+
+$S_{RO}(foot_{OSM}, foot_{SLU}) = \frac{A_{Overlap}}{min(A_{foot_{OSM}}, A_{foot_{SLU}}))}$
+
+For this project the relative overlap will be used both to find possible matching building candidates, and as a term in their similarity function.
+
+//TODO: Finding identical points
+
+//TODO: Point proximity for positional accuracy
+
+//TODO: Cite biljecki
+When presenting CityGML geometric errors, Biljecki recommends to obtain a measure of the distribution of error, introduce a tolerance and calculate the number of features or segments that fall within this tolerance.
+We can used this as a complementary statistic to the average error and error variance.
+
 \section{Implementation}
@@
 
+\bibitem{arkin-91}
+Arkin E.M., Chew L.P., Huttenlocher D.P., Kedem K. and Mitchell J.S.B. 1991. An Efficiently Computable \\
+Metric for Comparing Polygonal Shapes. In: IEEE Transaction on Pattern Analysis and Machine \\
+Intelligence, Vol. 13, No. 3, March 1991. \\
+
 \end{thebibliography}

### 2020-03-17 paper/paper.tex
Add section about dataset completeness

@@
 
-//TODO: Feature mapping, how to, first of all
+\subsection{Assesment of dataset completeness}
+
+(Fan et al., 2014) used the total building area in their reference dataset and the OpenStreetMap dataset to assess the completeness of OpenStreetMap data.
+The motivation for this is to eliminate semantic differences between both datasets.
+A very common example of a semantic error when comparing geographic data is that a large building in one dataset may be segmented into several smaller ones in the other, making one-to-one object mapping impossible.
+Thus using building area as an estimator for completeness solves this issue much better than i.e.
+the number of buildings or other objects in each dataset.
+This project will use the building area in the OSM dataset versus that of the SLU dataset to determine the completeness of OpenStreetMap building data in the Stockholm area, which will then be used as an argument for how accurate the building correspondence and point proximity measures can be.
 
@@
 
+Feature mapping
 The building footprint correspondence will be evaluated using a similarity function that assesses the turning funciton and relative area overlap.
@@
 
-//TODO: Completeness by building area, quotation
-
 \subsection{Correspondence by building area}
@@
 For this project the relative overlap will be used both to find possible matching building candidates, and as a term in their similarity function.
+
+\subsection{Closest point and point proximity}
 

### 2020-03-17 paper/paper.tex
A statement on positional accuracy

@@
 
-//TODO: Point proximity for positional accuracy
-
-//TODO: Cite biljecki
-When presenting CityGML geometric errors, Biljecki recommends to obtain a measure of the distribution of error, introduce a tolerance and calculate the number of features or segments that fall within this tolerance.
-We can used this as a complementary statistic to the average error and error variance.
+Finally, when two building footprints and their vertices have been mapped, we will use the offset of matching vertices as a measure of positional error.
+The average and variance in vertex spread will be calculated and used to make an assessment of the geometric precision of the OpenStreetMap dataset, and this precision will further be used to form an upper boundary on geometrical corrections that can be made to the map once it has been imported into the application.
 

### 2020-03-17 paper/paper.tex
Add a section about relative overlap and semantic accuracy

@@
 
+The relative overlap may also be used to determine building matching relations even when the semantic accuracy is low.
+If a large building is represented by a single footprint in one dataset but by several smaller footprints in the other, XXX found that if $S_{RO}(foot_{A}, foot_{B}) < 30\%$ for two buildings $A$ and $B$ from different sets, then $A$ and $B$ are highly likely to be separate, neighbouring buildings and not in fact identical.
+Therefore we will here assume that two buildings are matching candidates if their relative area overlap is greater than 30\%.
+If a one-to-many object matching is found, the compound perimeter of the footprints in the many-set will be used when calculating the turning function and finding closest vertices between the polygons.
+
 \subsection{Closest point and point proximity}

### 2020-03-17 paper/paper.tex
Change description of S_T value range
Fix report do that it compiles

@@
 (1991), as a method for measuring the similarity of two polygons.
-The Turning funciton $T_c(l)$ measures the cumulative angle of the polygon's counter-clockwise tangent, as a function of the cumulative normalized length l.
+The Turning function $T_c(l)$ measures the cumulative angle of the polygon's counter-clockwise tangent, as a function of the cumulative normalized length l.
 This project uses the turning function as it is defined by (Fan et al., 2014).
@@
 The similarity of two polygons A and B in terms of their turning function is defined as their distance of their cumulative turning functions: $S_{T}(A, B) = 1 - (\int^{1}_{0} T_{C,A}(l) - T_{C,B}(l) dl)^{1/2}$.
-The value range of $S(A, B)$ will be 0 to 1, where $S(A,B) = 1$ if the polygons are identical.
+The value range will be $(0 < S_{T} < 1)$, where $S_{T}(A,B) = 1$ if the polygons are identical.
 
@@
 
-The relative overlap may also be used to determine building matching relations even when the semantic accuracy is low.
+The relative overlap may be used to determine building matching relations even when the semantic accuracy is low.
 If a large building is represented by a single footprint in one dataset but by several smaller footprints in the other, XXX found that if $S_{RO}(foot_{A}, foot_{B}) < 30\%$ for two buildings $A$ and $B$ from different sets, then $A$ and $B$ are highly likely to be separate, neighbouring buildings and not in fact identical.
@@
 
-\bitiem{fan-14}
+\bibitem{fan-14}
 (Fan et al, 2014), Quality assessment for building footprints data on OpenStreetMap \\

### 2020-03-18 paper/paper.tex
Motivation behind area overlap, explanation for Douglas-Peucker algorithm

@@
 
-(Fan et al., 2014) further discussed how the relative overlap in building area can be used to determine building correspondence between datasets.
+(Fan et al., 2014) further discussed how the relative overlap in building area can be used to determine building correspondence between datasets, in cases where there is not much displacement between OSM building footprints data and the reference data set.
+Figure X (TODO) shows a typical comparison between the OpenStreetMap and SLU datasets, and by eyesight it was determined that the displacement of features is small.
 The relative building overlap between the OSM and SLU datasets will be defined as folows, given the footprints of any building in the OSM set ($foot_{OSM}$) and the SLU set ($foot_{SLU}$):
@@
 
-//TODO: Finding identical points
+The final problem is that even when building polygons have been matched between the two datasets, they may not have one-to-one vertex relationship.
+Footprints from different datasets may be formed at a different level of detail.
+Problems that arise are e.g.
+that the vertex counts are dissimilar, or that vertex clusters may be found at different parts of the polygon in the two datasets.
+To avoid this effect, key points are extracted using the Douglas-Peucker algorithm (Douglas and Peucker, 1973), to create a simplifed footprint with less points that still retain information about the rough features of the detailed footprint.
+The idea behind Douglas-Peucker is to recursively divide a polyline.
+It initially marks only the start and end points $(v_0, v_n)$ to be kept, and finds the point $v_i$ in between whose distance is the greatest to the line segment between $v_0$ and $v_n$.
+It then recursively refines the line segments $(v_0, v_i)$ and $(v_i, v_n)$, and proceeds to do so until a line segment $(v_j, v_k)$ is found, where every point in between $v_j$ and $v_k$ have a distance to the line segment $(v_j, v_k)$ that is smaller than some resolution $\epsilon$.
+Then $v_j$ are added to the simplified polyline, and all nodes in between them are discarded.
+See Algorithm 1 for a detailed view.
+In either case The minimum bounding rectangle (MBR) is calculated for the two polygons.
+Finally the MBR for the OSM building footprint is shifted so that its centroid aligns with the centroid of the MBR of the SLU building footprint.
+Any edges in the simplified footprints that coinside with the MBR from the same dataset are extracted, and the corresponding points in the original footprints will be matched with each other.
+
+//TODO: Algorithm of Douglas-Peucker
 
@@
 
+\bibitem{douglas-peucker-73}
+Douglas, D.H. and Peucker T.K. 1973. Algorithms for the reduction of the number of points required to \\
+represent a digitalized line or its caricature. In: Cartographica: The International Journal for Geographic \\
+Information and Geovisualization, 10(2), 112-122. \\
+
 \end{thebibliography}

### 2020-03-18 paper/paper.tex
Fix titles
Add Douglas-Peucker algorithm

@@
 \usepackage{multicol}
+\usepackage[ruled,vlined]{algorithm2e}
 \setlength\parindent{0pt}
@@
 
-\section(The geodata used in this project)
+\section{The geodata used in this project}
 
@@
 
-\subsection(OSM Data)
+\subsection{OSM Data}
 
@@
 
-\subsection(SLU Data)
+\subsection{SLU Data}
 
@@
 
-\subsection(A word on coordinate systems)
+\subsection{A word on coordinate systems}
 
@@
 
-//TODO: Algorithm of Douglas-Peucker
+\begin{algorithm}[H]
+\SetAlgoLined
+\KwResult{Write here the result }
+    // Find the point with the maximum distance\;
+    $d_{max}$ = 0\;
+    $index$ = 0\;
+    $end$ = length(PointList)\;
+
+    \For{$i$=2 to ($end$-1)}{
+        $d$ = perpendicularDistance(PointList[$i$], Line(PointList[1], PointList[$end$]))\;
+        \If{$d$ $>$ $d_{max}$} {
+            $index$ = $i$\;
+            $d_{max}$ = $d$\;
+        }
+    }
+
+    ResultList = []\;
+
+    // If max distance is greater than epsilon, recursively simplify\;
+    \eIf{$d_{max}$ $>$ $\epsilon$} {
+        // Recursive call\;
+        recResults1[] = Douglas-Peucker(PointList[1...$index$], $\epsilon$)\;
+        recResults2[] = Douglas-Peucker(PointList[$index$...$end$], $\epsilon$)\;
+
+        // Build the result list\;
+        ResultList[] = {recResults1[1...length(recResults1) - 1], recResults2[1...length(recResults2)]}\;
+    }{
+        ResultList[] = {PointList[1], PointList[$end$]}\;
+    }
+    \Return{ResultList[]}\;
+
+    \caption{Douglas-Peucker}
+\end{algorithm}
 

### 2020-03-18 paper/paper.tex
Add reference to Rutzinger et al

@@
 The relative overlap may be used to determine building matching relations even when the semantic accuracy is low.
-If a large building is represented by a single footprint in one dataset but by several smaller footprints in the other, XXX found that if $S_{RO}(foot_{A}, foot_{B}) < 30\%$ for two buildings $A$ and $B$ from different sets, then $A$ and $B$ are highly likely to be separate, neighbouring buildings and not in fact identical.
+If a large building is represented by a single footprint in one dataset but by several smaller footprints in the other, (Rutzinger et al, 2009) found that if $S_{RO}(foot_{A}, foot_{B}) < 30\%$ for two buildings $A$ and $B$ from different sets, then $A$ and $B$ are highly likely to be separate, neighbouring buildings and not in fact identical.
 Therefore we will here assume that two buildings are matching candidates if their relative area overlap is greater than 30\%.
@@
 
+\bibitem{rutzinger-09}
+Rutzinger, M., Rottensteiner, F. and Pfeifer, N. 2009. A Comparison of Evaluation Techniques for Building \\
+Extraction From Airborne Laser Scanning. IEEE Journal of Selected Topics in Applied Earth \\
+Observations and Remote Sensing 2(1): 11-20. \\
+
 \end{thebibliography}

### 2020-03-18 paper/paper.tex
Include query coordinates in the report
Include the date that the geodata was collected

@@
 Both map segments were obtained around the Stockholm greater metropolitan area.
-The exact map segments that were extracted were in both cases rectangles with the following coordinates:.
-TODO: Get coordinates
+The exact map segments that were extracted were in both cases rectangles with the following coordinates in WGS84: (N: 59.42, E: 18.15, S: 59.23, W: 17.79).
+Both datasets were collcted on March 18th, 2020.
 

### 2020-03-20 osm-importer/building-cropper.py
Update execution order again

@@ -1,7 +1,8 @@
 # Processes data with good vertex structure. Removes OSM buildings that interect the edge
 # of an OSM query, and the corresponding overlapping buildings in the SLU set
 # REQUIREMENTS:
-# - Run building-vertex-formatter.py
+# - Run building-data-formatter-osm.py
+# - Run building-data-formatter-slu.py
 
 from geojson import Point, Feature, FeatureCollection, load, dump
 from pyproj import Proj, transform

### 2020-03-20 osm-importer/building-cropper.py
Fix polygon detection, building-cropper.py

@@ -14,12 +14,11 @@ q_bbox_s = 59.328128796834925
 q_bbox_w = 18.03809503228281
 
 def polygon_line_intersection(polygon, line_point_1, line_point_2):
-    #print(polygon)
-    shapely_poly = shapely.geometry.Polygon(polygon[0])
+    shapely_poly = shapely.geometry.Polygon(polygon)
     shapely_line = shapely.geometry.LineString([line_point_1, line_point_2])
-    intersection_line = list(shapely_poly.intersection(shapely_line).coords)
-    print(intersection_line)
-    if len(intersection_line) == 0:
+    intersection_line = shapely_poly.intersection(shapely_line)
+
+    if intersection_line.length == 0.0:
         return False
     return True
 
@@ -37,6 +36,15 @@ def polygon_intersects_query_bbox(polygon):
 def polygon_relative_overlap(polygon_1, polygon_2):
     return 0.0
 
+def extract_polygon_from_feature(feature):
+    feature_type = feature['geometry']['type']
+
+    if feature_type == 'Polygon':
+        # Even for multipolygons, the big surrounding polygon is always the first one
+        return feature_osm['geometry']['coordinates'][0]
+
+    print("extract_polygon_from_feature: feature type not implemented: " + feature)
+
 # read files
 with open('raw_data/buildings-osm.geojson', 'r') as f:
     OSM_data = load(f)
@@ -48,7 +56,8 @@ for feature_osm in OSM_data['features']:
     print("Cropping, progess: " + str(int(100*progress/len(OSM_data['features']))) + '%')
     progress+=1.0
 
-    polygon_osm = feature_osm['geometry']['coordinates']
+    # Build polygon
+    polygon_osm = extract_polygon_from_feature(feature_osm)
     if polygon_intersects_query_bbox(polygon_osm):
 
         for feature_slu in SLU_data['features']:

### 2020-03-20 paper/paper.tex
Added notes about building data refinement to report

@@
 
+\section{Notes about data processing}
+
+Once data from the both sets had been extracted, they looked dissimilar.
+(Show picture) Had to go over all OSM buildings that intersected with the query box and remove them, along with any SLU buildings that intersected with them.
+
+When calculating the area overlap we used the fact that multipolygons in geojson are represented with the main, bounding polygon first.
+When calculating the area of a multipolygon building, we simply took the area of the first polygon and subtracted the subsequent ones.
+
+However, when calculating the relative area overlap, this becomes unesseccary since any semantically incorrect subdivision of the polygon will always be inside the bounding polygon.
+
 \begin{thebibliography}{9}

### 2020-03-20 osm-importer/building-convert-to-sweref.py
New interim files for storing cropped building data

@@ -17,9 +17,9 @@ def wgs84_to_epsg3006(point):
     return (y_new, x_new)
 
 # read files
-with open('raw_data/buildings-osm.geojson', 'r') as f:
+with open('raw_data/buildings-osm-cropped.geojson', 'r') as f:
     OSM_data = load(f)
-with open('raw_data/buildings-slu.geojson', 'r') as f:
+with open('raw_data/buildings-slu-cropped.geojson', 'r') as f:
     SLU_data = load(f)
 
 progress = 0.0

### 2020-03-20 osm-importer/building-cropper.py

@@ -69,7 +69,7 @@ for feature_osm in OSM_data['features']:
         OSM_data['features'].remove(feature_osm)
 
 # Write all building features to files
-with open('raw_data/buildings-osm.geojson', 'w') as f:
+with open('raw_data/buildings-osm-cropped.geojson', 'w') as f:
     dump(FeatureCollection(OSM_data), f)
-with open('raw_data/buildings-slu.geojson', 'w') as f:
+with open('raw_data/buildings-slu-cropped.geojson', 'w') as f:
     dump(FeatureCollection(SLU_data), f)

### 2020-03-20 osm-importer/building-cropper.py
Build cropped data additively

@@ -51,6 +51,9 @@ with open('raw_data/buildings-osm.geojson', 'r') as f:
 with open('raw_data/buildings-slu.geojson', 'r') as f:
     SLU_data = load(f)
 
+OSM_data_out = []
+SLU_data_out = []
+
 progress = 0.0
 for feature_osm in OSM_data['features']:
     print("Cropping, progess: " + str(int(100*progress/len(OSM_data['features']))) + '%')
@@ -59,17 +62,17 @@ for feature_osm in OSM_data['features']:
     # Build polygon
     polygon_osm = extract_polygon_from_feature(feature_osm)
     if polygon_intersects_query_bbox(polygon_osm):
-
         for feature_slu in SLU_data['features']:
             polygon_slu = feature_slu['geometry']['coordinates']
 
             if polygon_relative_overlap(polygon_osm, polygon_slu) > 0.3:
                 SLU_data['features'].remove(feature_slu)
 
-        OSM_data['features'].remove(feature_osm)
+    else:
+        OSM_data_out.append(feature_osm)
 
 # Write all building features to files
 with open('raw_data/buildings-osm-cropped.geojson', 'w') as f:
-    dump(FeatureCollection(OSM_data), f)
+    dump(FeatureCollection(OSM_data_out), f)
 with open('raw_data/buildings-slu-cropped.geojson', 'w') as f:
-    dump(FeatureCollection(SLU_data), f)
+    dump(FeatureCollection(SLU_data_out), f)

### 2020-03-25 osm-importer/plot_utils.py
Add utils library for plotting. Add methods for plotting polygons

@@ -1,0 +1,55 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib
+import geometry_utils
+from matplotlib.patches import Polygon
+from matplotlib.collections import PatchCollection
+
+def plot_polygons(polygons):
+    fig, ax = plt.subplots()
+    ax.set_xlim(0)
+    patches = []
+
+    min_x = 999999999999.0
+    min_y = 999999999999.0
+    max_x = -999999999999.0
+    max_y = -999999999999.0
+
+    for p in polygons:
+        patches.append(Polygon(p, True))
+
+        minmax_coords = geometry_utils.minmax_points_of_polygon(p)
+        if minmax_coords[0][1] > max_y:
+            max_y = minmax_coords[0][1]
+        if minmax_coords[1][0] > max_x:
+            max_x = minmax_coords[1][0]
+        if minmax_coords[2][1] < min_y:
+            min_y = minmax_coords[2][1]
+        if minmax_coords[3][0] < min_x:
+            min_x = minmax_coords[3][0]
+
+
+    p = PatchCollection(patches, cmap=matplotlib.cm.jet, alpha=0.4)
+
+    colors = 100*np.random.rand(len(patches))
+    p.set_array(np.array(colors))
+
+    ax.add_collection(p)
+
+    # Set zoom limits
+    r_x = max_x-min_x
+    r_y = max_y-min_y
+    if r_x > r_y:
+        r = r_x-r_y
+        min_y -= r/2
+        max_y += r/2
+    else:
+        r = r_y-r_x
+        min_x -= r/2
+        max_x += r/2
+
+    plt.axis([min_x,max_x,min_y,max_y])
+    plt.show()
+
+def plot_polygon(polygon):
+    plot_polygons([polygon])

### 2020-03-25 osm-importer/building-metrics.py
Fix SLU multipolygon representation

@@ -66,7 +66,8 @@ for feature_osm in OSM_data['features']:
         polygon_osm = geometry_utils.extract_polygon_from_feature(feature_osm)
         polygon_slu = []
         for feature_slu in features_slu:
-            polygon_slu += geometry_utils.extract_polygon_from_feature(feature_slu)
+            polygon_slu.append(geometry_utils.extract_polygon_from_feature(feature_slu))
+
 
         # TODO: Find out how Fan et al did for situations where a building is made up out of many small footprints
         # Compute convex hulls

### 2020-03-25 osm-importer/geometry_utils.py
Fix Douglas-Peucker implementation

@@ -117,44 +117,6 @@ def perp_distance_point_to_line(point, line_point_1, line_point_2):
     return nom/denom
 
 def douglas_peucker(polygon, e):
-    # Find starting points, use northernmost point and the furthest point from that
-    #Pick northernmost point to ensure it is the same in both OSM and SLU cases
-    sp_1 = None
-    xy_max = 999999999999.0
-    for i in range(0, len(polygon)):
-        p = polygon[i]
-        x = p[0]
-        y = p[1]
-        if x+y < xy_max:
-            xy_max = x+y
-            sp_1 = p
-
-    sp_2 = None
-    r_max = 0.0
-    for i in range(0, len(polygon)):
-        p = polygon[i]
-        r = point_distance(sp_1, p)
-        if r > r_max:
-            r_max = r
-            sp_2 = p
-
-    sp_1_index = polygon.index(sp_1)
-    sp_2_index = polygon.index(sp_2)
-
-    if sp_1_index > sp_2_index:
-        temp = sp_1_index
-        sp_1_index = sp_2_index
-        sp_2_index = temp
-
-    slice_1 = polygon[:sp_1_index] + polygon[sp_2_index:]
-    slice_2 = polygon[sp_1_index:sp_2_index]
-
-    res_polygon_1 = douglas_peucker_helper(slice_1, e)
-    res_polygon_2 = douglas_peucker_helper(slice_2, e)
-
-    return res_polygon_1 + res_polygon_2
-
-def douglas_peucker_helper(polygon, e):
     # Find the point with the maximum distance to the line: polygon[0], polygon[end]
     d_max = 0
     index = 0
@@ -169,8 +131,8 @@ def douglas_peucker_helper(polygon, e):
     #If max distance is greater than epsilon, recursively simplify
     if d_max > e:
         # Recursive call
-        rec_polygon_1 = douglas_peucker_helper(polygon[:index], e)
-        rec_polygon_2 = douglas_peucker_helper(polygon[index:], e)
+        rec_polygon_1 = douglas_peucker(polygon[:index], e)
+        rec_polygon_2 = douglas_peucker(polygon[index:], e)
         # Build the result list
         polygon_out = rec_polygon_1 + rec_polygon_2
     else:

### 2020-03-25 osm-importer/geometry_utils.py
Douglas-Peucker: Find common starting point in the outer call of the algorithm

@@ -115,8 +115,23 @@ def perp_distance_point_to_line(point, line_point_1, line_point_2):
     nom = abs((y2-y1)*x0 - (x2-x1)*y0 + x2*y1 - y2*x1)
     denom = math.sqrt((y2-y1)**2 + (x2-x1)**2)
     return nom/denom
-
 def douglas_peucker(polygon, e):
+    # Find starting point, use northernmost point
+    #Pick northernmost point to ensure it is the same in both OSM and SLU cases
+    sp_index = 0
+    y_max = 999999999999.0
+    for i in range(0, len(polygon)):
+        p = polygon[i]
+        if p[1] < y_max:
+            y_max = p[1]
+            sp_index = i
+
+    # Rotate the polygon until the starting point is first
+    polygon_rotated = polygon[sp_index:] + polygon[:sp_index]
+
+    return douglas_peucker_helper(polygon_rotated, e)
+
+def douglas_peucker_helper(polygon, e):
     # Find the point with the maximum distance to the line: polygon[0], polygon[end]
     d_max = 0
     index = 0
@@ -131,8 +146,8 @@ def douglas_peucker(polygon, e):
     #If max distance is greater than epsilon, recursively simplify
     if d_max > e:
         # Recursive call
-        rec_polygon_1 = douglas_peucker(polygon[:index], e)
-        rec_polygon_2 = douglas_peucker(polygon[index:], e)
+        rec_polygon_1 = douglas_peucker_helper(polygon[:index], e)
+        rec_polygon_2 = douglas_peucker_helper(polygon[index:], e)
         # Build the result list
         polygon_out = rec_polygon_1 + rec_polygon_2
     else:

### 2020-03-25 osm-importer/building-metrics.py
Reduce todo-list, add commented-out plot statements for debugging

@@ -4,6 +4,7 @@
 
 from geojson import Point, Feature, FeatureCollection, load
 import geometry_utils
+import plot_utils
 
 OSM_data = {}
 SLU_data = {}
@@ -74,9 +75,16 @@ for feature_osm in OSM_data['features']:
         cv_osm = geometry_utils.convex_hull(polygon_osm)
         cv_slu = geometry_utils.convex_hull(polygon_slu)
 
+        #plot_utils.plot_polygon(polygon_osm)
+        #plot_utils.plot_polygons(polygon_slu)
+
         # Turning funcitons
         tc_osm = geometry_utils.turning_function(cv_osm)
         tc_slu = geometry_utils.turning_function(cv_slu)
+
+        #plot_utils.plot_polygons([cv_osm, polygon_osm])
+        #plot_utils.plot_polygons([cv_slu] + polygon_slu)
+        #plot_utils.plot_polygons([cv_slu, cv_osm])
 
         # Remove last points to prepare for Douglas-Peucker
         cv_osm.pop(-1)
@@ -86,6 +94,10 @@ for feature_osm in OSM_data['features']:
         # TODO: Adjust tolerance
         cv_osm_dp = geometry_utils.douglas_peucker(cv_osm, 5)
         cv_slu_dp = geometry_utils.douglas_peucker(cv_slu, 5)
+
+        #plot_utils.plot_polygons([cv_osm, cv_osm_dp])
+        #plot_utils.plot_polygons([cv_slu, cv_slu_dp])
+        #plot_utils.plot_polygons([cv_osm_dp, cv_slu_dp])
 
         bounding_points_osm = geometry_utils.minmax_points_of_polygon(cv_osm_dp)
         bounding_points_slu = geometry_utils.minmax_points_of_polygon(cv_slu_dp)
@@ -101,12 +113,4 @@ avg_pos_error /= counted_points
 print("Average position error: " + str(avg_pos_error))
 
 
-
-
-
-# TODO: Check if there is source code for Fan's project
-# TODO: What is the turning function used for in Fan's project?
-# TODO: Plot: Check if the convex hull is calculated properly
-# TODO: Plot: Check if the DP-reduction is correct
-# TODO: Plot: Check that the found points on the OSM and SLU convex hulls are calculated properly
-# TODO: Plot: Check that the DP-reduction + MBR finds the correct matching points
+#TODO: Read through Fan et al, make a list of metrics that we will obtain

### 2020-03-25 osm-importer/geometry_utils.py
Geometry helper methods

@@ -1,5 +1,6 @@
 import shapely.geometry
 import math
+import numpy as np
 from pyproj import Proj, transform
 
 def wgs84_to_epsg3006(point):
@@ -48,8 +49,33 @@ def polygon_line_intersection(polygon, line_point_1, line_point_2):
         return False
     return True
 
+def polygon_centroid(polygon):
+    centroid = [0.0, 0.0]
+    for i in range(0, len(polygon)):
+        point = polygon[i]
+        if i == len(polygon)-1 and point[0] == polygon[0][0] and point[1] == polygon[0][1]:
+            # Ensure that the starting point is not counted twice
+            continue
+        centroid[0] += point[0]
+        centroid[1] += point[1]
+    centroid[0] /= len(polygon)
+    centroid[1] /= len(polygon)
+    return centroid
+
 def point_distance(p1, p2):
     return math.sqrt((p1[0]-p2[0])**2 + (p1[1]-p2[1])**2)
+
+def point_len(p):
+    return point_distance(p, [0.0, 0.0])
+
+def normalize(a):
+    d = point_len(a)
+    return [a[0]/d, a[1]/d]
+
+def project(a, b):
+    # Project a onto b
+    proj_factor = np.dot(a, b)/point_len(a)**2
+    return [proj_factor*a[0], proj_factor*a[1]]
 
 def polygon_perimeter(polygon):
     perimeter_out = 0.0
@@ -64,6 +90,12 @@ def signed_vector_angle(vector1, vector2):
     len1 = math.hypot(x1, y1)
     len2 = math.hypot(x2, y2)
     return math.acos(inner_product/(len1*len2))
+
+def perpendicular(a):
+    b = np.empty_like(a)
+    b[0] = -a[1]
+    b[1] = a[0]
+    return b
 
 def add_areas_recursively(c):
     area_out = 0.0

### 2020-03-25 osm-importer/geometry_utils.py
Oriented MBR function

@@ -263,6 +263,83 @@ def convex_hull(polygons):
         point = far_point
     return hull_points
 
+def oriented_mbr(points):
+    cv = convex_hull(points)
+    center = polygon_centroid(cv)
+
+    best_area = 0.0
+    best_f_1_min = None
+    best_f_1_max = None
+    best_f_2_min = None
+    best_f_2_max = None
+
+    for i in range(0, len(cv)-1):
+        a = cv[i]
+        b = cv[i+1]
+        ev_1 = [b[0]-a[0], b[1]-a[1]]
+        ev_1 = normalize(ev_1)
+        ev_2 = perpendicular(ev_1)
+
+        # Projected edge points from the center
+        f_1_min = [0.0, 0.0]
+        f_1_max = [0.0, 0.0]
+        f_2_min = [0.0, 0.0]
+        f_2_max = [0.0, 0.0]
+
+        for j in range(0, len(cv)-1):
+            # p = cv[j] - center
+            p = [cv[j][0] - center[0], cv[j][1] - center[1]]
+            print("%%%")
+            print(cv[j])
+            print(center)
+            print(p)
+            print("%%%")
+
+            p_1 = project(p, ev_1)
+            p_2 = project(p, ev_2)
+
+            print("***")
+            print(np.dot(p_1, ev_1))
+            print(np.dot(f_1_max, ev_1))
+            print("***")
+
+            if np.dot(p_1, ev_1) > np.dot(f_1_max, ev_1):
+                f_1_max = p_1
+            if np.dot(p_1, ev_1) < np.dot(f_1_min, ev_1):
+                f_1_min = p_1
+            if np.dot(p_2, ev_2) > np.dot(f_2_max, ev_2):
+                f_2_max = p_2
+            if np.dot(p_2, ev_2) > np.dot(f_2_min, ev_2):
+                f_2_min = p_2
+
+        print(f_1_max)
+        print(f_1_min)
+        print(f_2_max)
+        print(f_2_min)
+
+        area = polygon_area([f_1_min, f_2_min, f_1_max, f_2_max])
+        print(area)
+        if area > best_area:
+            best_f_1_min = f_1_min
+            best_f_1_max = f_1_max
+            best_f_2_min = f_2_min
+            best_f_2_max = f_2_max
+
[…]

### 2020-03-25 osm-importer/building-metrics.py
Calculate MBR in building metrics (WIP)

@@ -78,6 +78,10 @@ for feature_osm in OSM_data['features']:
         #plot_utils.plot_polygon(polygon_osm)
         #plot_utils.plot_polygons(polygon_slu)
 
+        mbr_osm = geometry_utils.oriented_mbr(polygon_osm)
+        print(mbr_osm)
+        plot_utils.plot_polygons([polygon_osm, mbr_osm])
+
         # Turning funcitons
         tc_osm = geometry_utils.turning_function(cv_osm)
         tc_slu = geometry_utils.turning_function(cv_slu)

### 2020-03-26 paper/paper.tex
Reduce todo-list, add commented-out plot statements for debugging
Mention Oriented MBR in the report
Update todo list, report updates

@@
 
+\section{Related work}
+Fan et al, 2015
+Kunze, 2012 (reference in Fan et al)
+
 \section{The geodata used in this project}
@@
 Both map segments were obtained around the Stockholm greater metropolitan area.
-The exact map segments that were extracted were in both cases rectangles with the following coordinates in WGS84: (N: 59.42, E: 18.15, S: 59.23, W: 17.79).
+The exact map segments that were extracted were in both cases rectangular boundaries with the following coordinates in WGS84: (N: 59.42, E: 18.15, S: 59.23, W: 17.79).
 Both datasets were collcted on March 18th, 2020.
@@
 The motivation for this is to eliminate semantic differences between both datasets.
-A very common example of a semantic error when comparing geographic data is that a large building in one dataset may be segmented into several smaller ones in the other, making one-to-one object mapping impossible.
+A very common example of a semantic error when comparing geographic data is that a large building in one dataset may be segmented into several smaller ones in the other, that is to say a building in one set is represented as an aggreation of mutliple buildings in the other set.
+This makes one-to-one object mapping impossible.
 Thus using building area as an estimator for completeness solves this issue much better than i.e.
@@
 
+TODO: Include a distribution of the shap similarity of matched building footprints, see Fan et al, page 14
+
 \subsection{Correspondence by building area}
@@
 See Algorithm 1 for a detailed view.
-In either case The minimum bounding rectangle (MBR) is calculated for the two polygons.
+In either case The Oriented Minimum Bounding Rectangle (MBR) is calculated for the two polygons.
 Finally the MBR for the OSM building footprint is shifted so that its centroid aligns with the centroid of the MBR of the SLU building footprint.

### 2020-03-26 osm-importer/building-metrics.py

@@ -156,4 +156,21 @@ print("Average position error: " + str(avg_pos_error_cp) + " (Counting Points me
 avg_pos_error_mbr /= counted_points_mbr
 print("Average position error: " + str(avg_pos_error_mbr) + " (MBR method, reasonable)")
 
-#TODO: Read through Fan et al, make a list of metrics that we will obtain
+#TODO: Line plot function in plot-utils
+#TODO: Line plot of turning functions
+#TODO: Function for evaluating shape similarity, using turning function
+#TODO: MBR method: Record which quadrant (MBR edge) That the edge was found on. Only match edges on the same sides
+#TODO: Find the SLU MBR that minimizes area overlap between the OSM and SLU MBRs, not the one with least area
+#TODO: Write in the report about why we had to cut buildings close to the boundary
+#TODO: Position accuracy per building by taking the average of the distance between the corresponding points
+#TODO: Fan et al, only buildings with a 1:1 semantic relationship are involved in the measurement of of positional and shape accuracy. Should we do this too?
+    # In the case of turning function, this is the only thing we can do!
+#TODO: Try switching to the area-minimizing MBR implementation that chris sent
+#TODO: Fan et al, normalize shape similarity by the rectangularity of the footprints (area divided by the area of the oriented MBR)
+#TODO: Metric: Number of buildings in each dataset
+#TODO: Metric: Statistic of the matching result using area overlap (Fan et al, page 9)
+#TODO: Metric: Max, min and std deviation of position offsets (Fan et al, page 12)
+#TODO: Metric: Bar diagram of position offsets (Fan et al, page 12)
+#TODO: Metric: Bar diagram of footprint shape similarity (Fan et al, page 12)
+#TODO: Mention as a source of error: Fan et al claimed that the main source of error was the limited resolution of satellite photos
+

### 2020-03-26 osm-importer/plot_utils.py
Add line plot functions to plot utils

@@ -65,6 +65,20 @@ def plot_scatter(points):
     plt.ylabel('y')
     plt.show()
 
+def plot_line(points):
+    plot_lines([points])
+
+def plot_lines(lines):
+    for line in lines:
+        data = np.array(line)
+        x, y = data.T
+        plt.plot(x, y)
+
+    plt.title('Line plot')
+    plt.xlabel('x')
+    plt.ylabel('y')
+    plt.show()
+
 def plot_bar(data_values, bar_width):
     min_val = min(data_values)
     max_val = max(data_values)

### 2020-03-26 osm-importer/building-metrics.py
Turning function is now a step function, plot turning functions

@@ -81,6 +81,8 @@ for feature_osm in OSM_data['features']:
         tc_osm = geometry_utils.turning_function(cv_osm)
         tc_slu = geometry_utils.turning_function(cv_slu)
 
+        plot_utils.plot_lines([tc_osm, tc_slu])
+
         #plot_utils.plot_polygons([cv_osm, polygon_osm])
         #plot_utils.plot_polygons([cv_slu] + polygon_slu)
         #plot_utils.plot_polygons([cv_slu, cv_osm])
@@ -156,8 +158,6 @@ print("Average position error: " + str(avg_pos_error_cp) + " (Counting Points me
 avg_pos_error_mbr /= counted_points_mbr
 print("Average position error: " + str(avg_pos_error_mbr) + " (MBR method, reasonable)")
 
-#TODO: Line plot function in plot-utils
-#TODO: Line plot of turning functions
 #TODO: Function for evaluating shape similarity, using turning function
 #TODO: MBR method: Record which quadrant (MBR edge) That the edge was found on. Only match edges on the same sides
 #TODO: Find the SLU MBR that minimizes area overlap between the OSM and SLU MBRs, not the one with least area

### 2020-03-26 osm-importer/geometry_utils.py

@@ -142,7 +142,6 @@ def add_areas_recursively(c):
     return area_out
 
 def minmax_points_of_polygon(polygon):
-    # TODO: Using this instead of MBR for now, since I have not figured out how to do rotated MBR
     min_point_n = [0.0, -999999999.0]
     min_point_e = [-999999999.0, 0.0]
     min_point_s = [0.0, 999999999.0]
@@ -222,6 +221,8 @@ def turning_function(polygon):
 
         edge_len = point_distance(p2, p1)
         acc_len += edge_len
+
+        turnpoints_out.append([acc_len-0.001, acc_angle])
 
         angle = signed_vector_angle([p2[0]-p1[0], p2[1]-p1[1]], [p3[0]-p2[0], p3[1]-p2[1]])
         acc_angle += angle

### 2020-03-26 osm-importer/building-metrics.py
Print the total number of buildings in each dataset

@@ -148,6 +148,9 @@ for feature_osm in OSM_data['features']:
 
 print("Results...")
 
+print("#Buildings, OSM: " + str(len(OSM_data['features'])))
+print("#Buildings, SLU: " + str(len(SLU_data['features'])))
+
 print("Total area, OSM: " + str(total_area_OSM))
 print("Total area, SLU: " + str(total_area_SLU))
 print("Total area, fraction: " + str(total_area_OSM / total_area_SLU))
@@ -167,7 +170,6 @@ print("Average position error: " + str(avg_pos_error_mbr) + " (MBR method, reaso
     # In the case of turning function, this is the only thing we can do!
 #TODO: Try switching to the area-minimizing MBR implementation that chris sent
 #TODO: Fan et al, normalize shape similarity by the rectangularity of the footprints (area divided by the area of the oriented MBR)
-#TODO: Metric: Number of buildings in each dataset
 #TODO: Metric: Statistic of the matching result using area overlap (Fan et al, page 9)
 #TODO: Metric: Max, min and std deviation of position offsets (Fan et al, page 12)
 #TODO: Metric: Bar diagram of position offsets (Fan et al, page 12)

### 2020-03-27 osm-importer/building-metrics.py
Add source of error section to report

@@ -161,6 +161,7 @@ print("Average position error: " + str(avg_pos_error_cp) + " (Counting Points me
 avg_pos_error_mbr /= counted_points_mbr
 print("Average position error: " + str(avg_pos_error_mbr) + " (MBR method, reasonable)")
 
+#TODO: Turning function fixes
 #TODO: Function for evaluating shape similarity, using turning function
 #TODO: MBR method: Record which quadrant (MBR edge) That the edge was found on. Only match edges on the same sides
 #TODO: Find the SLU MBR that minimizes area overlap between the OSM and SLU MBRs, not the one with least area

### 2020-03-27 osm-importer/building-metrics.py
Normalized shape dissimilarity implementation

@@ -68,8 +68,6 @@ for feature_osm in OSM_data['features']:
         for feature_slu in features_slu:
             polygon_slu.append(geometry_utils.extract_polygon_from_feature(feature_slu))
 
-
-        # TODO: Find out how Fan et al did for situations where a building is made up out of many small footprints
         # Compute convex hulls
         cv_osm = geometry_utils.convex_hull(polygon_osm)
         cv_slu = geometry_utils.convex_hull(polygon_slu)
@@ -82,6 +80,13 @@ for feature_osm in OSM_data['features']:
         tc_slu = geometry_utils.turning_function(cv_slu)
 
         plot_utils.plot_lines([tc_osm, tc_slu])
+
+        # Normalized shape dissimilarities, only if there is a 1:1 match
+        if len(polygon_slu) == 1:
+            #TODO: Metric: Bar diagram of footprint shape similarity (Fan et al, page 12)
+            # Only buildings with a 1:1 semantic relationship are involved in the measurement of of positional and shape accuracy. Should we do this too?
+            shape_dissimilarity_normalized = geometry_utils.normalized_shape_dissimilarity(polygon_osm, polygon_slu)
+            print("1:1 MATCH MATCH MATCH")
 
         #plot_utils.plot_polygons([cv_osm, polygon_osm])
         #plot_utils.plot_polygons([cv_slu] + polygon_slu)
@@ -161,19 +166,16 @@ print("Average position error: " + str(avg_pos_error_cp) + " (Counting Points me
 avg_pos_error_mbr /= counted_points_mbr
 print("Average position error: " + str(avg_pos_error_mbr) + " (MBR method, reasonable)")
 
+
+#TODO: Position accuracy per building by taking the average of the distance between the corresponding points
+
 #TODO: Turning function fixes
 #TODO: Function for evaluating shape similarity, using turning function
 #TODO: MBR method: Record which quadrant (MBR edge) That the edge was found on. Only match edges on the same sides
 #TODO: Find the SLU MBR that minimizes area overlap between the OSM and SLU MBRs, not the one with least area
 #TODO: Write in the report about why we had to cut buildings close to the boundary
-#TODO: Position accuracy per building by taking the average of the distance between the corresponding points
-#TODO: Fan et al, only buildings with a 1:1 semantic relationship are involved in the measurement of of positional and shape accuracy. Should we do this too?
-    # In the case of turning function, this is the only thing we can do!
[…]

### 2020-03-27 osm-importer/geometry_utils.py

@@ -222,7 +222,7 @@ def turning_function(polygon):
         edge_len = point_distance(p2, p1)
         acc_len += edge_len
 
-        turnpoints_out.append([acc_len-0.001, acc_angle])
+        turnpoints_out.append([acc_len-0.0000001, acc_angle])
 
         angle = signed_vector_angle([p2[0]-p1[0], p2[1]-p1[1]], [p3[0]-p2[0], p3[1]-p2[1]])
         acc_angle += angle
@@ -348,6 +348,37 @@ def oriented_mbr(points):
             best_area = area
 
     return [best_corner_1, best_corner_2, best_corner_3, best_corner_4]
+
+def step_funtion_area(points):
+    # Area under a step function by piecewise summation
+    area_out = 0.0
+    for i in range(0, len(points)-1):
+        p1 = points[i]
+        p2 = points[i+1]
+        x_diff = p2[0]-p1[0]
+        y = p1[1]
+        area_out += y*x_diff
+    return area_out
+
+def polygon_rectangularity(polygon):
+    mbr = oriented_mbr(polygon)
+    a_poly = polygon_area(polygon)
+    a_rect = polygon_area(mbr)
+    return a_poly/a_rect
+
+def shape_dissimilarity(polygon_1, polygon_2):
+    # Implementation of eqn 1, Fan et al
+    tc_1 = turning_function(polygon_1)
+    tc_2 = turning_function(polygon_2)
+    area_diff = step_funtion_area(tc_1) - step_funtion_area(tc_2)
+    return math.sqrt(abs(area_diff))
+
+def normalized_shape_dissimilarity(polygon_1, polygon_2):
+    # Implementation of eqn 4, Fan et al
+    # Normalize shape similarity by the rectangularity of the footprints (area divided by the area of the oriented MBR)
+    rectangularity = polygon_rectangularity(polygon_1)
+    mbr = oriented_mbr(polygon_1)
+    return 1 - shape_dissimilarity(polygon_1, polygon_2)*(1-rectangularity)/shape_dissimilarity(polygon_1, mbr)
 
 def extract_polygon_from_feature(feature):
     feature_type = feature['geometry']['type']

### 2020-03-27 osm-importer/building-metrics.py
Fix indentation of position error method #2

@@ -137,17 +137,17 @@ for feature_osm in OSM_data['features']:
             edges_on_perimeter_osm.remove(edge_osm)
             edges_on_perimeter_slu.remove(best_edge_slu)
 
-            for i in range(0, len(mbr_osm)):
-                p1 = mbr_osm[i]
-                best_r = 9999999999.0
-                p2_best = None
-                for p2 in mbr_slu:
-                    r = geometry_utils.point_distance(p1, p2)
-                    if r < best_r:
-                        p2_best = p2
-                        best_r = r
-                avg_pos_error_cp += best_r
-                counted_points_cp += 1
+        for i in range(0, len(mbr_osm)):
+            p1 = mbr_osm[i]
+            best_r = 9999999999.0
+            p2_best = None
+            for p2 in mbr_slu:
+                r = geometry_utils.point_distance(p1, p2)
+                if r < best_r:
+                    p2_best = p2
+                    best_r = r
+            avg_pos_error_cp += best_r
+            counted_points_cp += 1
 
 ### OUTPUT ###
 

### 2020-03-27 osm-importer/building-metrics.py
Update todos

@@ -26,6 +26,7 @@ for feature in SLU_data['features']:
     total_area_SLU += geometry_utils.add_areas_recursively(feature['geometry']['coordinates'])
 
 # Feature mapping
+#TODO: Metric: Statistic of the matching result using area overlap (Fan et al, page 9)
 overlapping_buildings_osm_bigger = {}
 overlapping_buildings_slu_bigger = {}
 progress = 0.0
@@ -121,6 +122,7 @@ for feature_osm in OSM_data['features']:
         #print(edges_on_perimeter_slu)
 
         # While both edge sets are non-emptyh
+        #TODO: MBR method: Record which quadrant (MBR edge) That the edge was found on. Only match edges on the same sides
         while edges_on_perimeter_osm and edges_on_perimeter_slu:
             edge_osm = edges_on_perimeter_osm[0]
             best_edge_slu = None
@@ -136,6 +138,9 @@ for feature_osm in OSM_data['features']:
                 #print(min_edge_distance)
             edges_on_perimeter_osm.remove(edge_osm)
             edges_on_perimeter_slu.remove(best_edge_slu)
+            #TODO: Position accuracy per building by taking the average of the distance between the corresponding points
+            #TODO: Metric: Bar diagram of position offsets (Fan et al, page 12)
+            #TODO: Metric: Max, min and std deviation of position offsets (Fan et al, page 12)
 
         for i in range(0, len(mbr_osm)):
             p1 = mbr_osm[i]
@@ -165,17 +170,4 @@ print("Average position error: " + str(avg_pos_error_cp) + " (Counting Points me
 
 avg_pos_error_mbr /= counted_points_mbr
 print("Average position error: " + str(avg_pos_error_mbr) + " (MBR method, reasonable)")
-
-
-#TODO: Position accuracy per building by taking the average of the distance between the corresponding points
-
-#TODO: Turning function fixes
-#TODO: Function for evaluating shape similarity, using turning function
-#TODO: MBR method: Record which quadrant (MBR edge) That the edge was found on. Only match edges on the same sides
-#TODO: Find the SLU MBR that minimizes area overlap between the OSM and SLU MBRs, not the one with least area
-#TODO: Write in the report about why we had to cut buildings close to the boundary
-#TODO: Try switching to the area-minimizing MBR implementation that chris sent
-#TODO: Metric: Statistic of the matching result using area overlap (Fan et al, page 9)
[…]

### 2020-03-27 osm-importer/geometry_utils.py

@@ -209,6 +209,7 @@ def douglas_peucker_helper(polygon, e):
         polygon_out = [polygon[0], polygon[end]]
     return polygon_out
 
+#TODO: Turning function fixes
 def turning_function(polygon):
     turnpoints_out = []
     acc_len = 0.0 # Accumluated length that has been stepped through
@@ -289,6 +290,7 @@ def convex_hull(polygons):
     return hull_points
 
 def oriented_mbr(points):
+    #TODO: Try switching to the area-minimizing MBR implementation that chris sent
     cv = convex_hull(points)
     center = polygon_centroid(cv)
 
@@ -351,6 +353,7 @@ def oriented_mbr(points):
 
 def step_funtion_area(points):
     # Area under a step function by piecewise summation
+    # TODO: Incorrect at the moment, need to actually integrate the difference between the curves
     area_out = 0.0
     for i in range(0, len(points)-1):
         p1 = points[i]
@@ -383,12 +386,11 @@ def normalized_shape_dissimilarity(polygon_1, polygon_2):
 def extract_polygon_from_feature(feature):
     feature_type = feature['geometry']['type']
 
+    # Even for multipolygons, the big surrounding polygon is always the first one
     if feature_type == 'Polygon':
-        # Even for multipolygons, the big surrounding polygon is always the first one
         return feature['geometry']['coordinates'][0]
 
     if feature_type == 'MultiPolygon':
-        #TODO: Return the very first polygon for now. Is this correct?
         return feature['geometry']['coordinates'][0][0]
 
     print("extract_polygon_from_feature: feature type not implemented: " + feature_type)

### 2020-03-27 paper/paper.tex
Add source of error section to report
Add seciton about OSM-SLU-specific data preprocessing

@@
 The coordinate conversions for this paper were obtained using proj: a Linux commandline application for geospatial coordinate conversion.
+
+\subsection{Specific preprocessing for SLU-OSM geodata}
+
+The SLU map is delivered with cropped buildings to exactly matched the query coordinates. This means that area comparsion will not work.
+Therefore it was neccessary to crop both datasets to ensure that all buildings were complete and had matching candidates in the other set.
+(Show a picture)
 
@@
 
+\section{Analysis}
+
+(Fan et al, 2014) claimed that the positional error is largely due to the limited resolution in satellite imagery.
+
 \begin{thebibliography}{9}

### 2020-03-27 osm-importer/building-metrics.py
Calculate average offset per building. Finally plot the data points

@@ -57,7 +57,7 @@ for feature_osm in OSM_data['features']:
 avg_pos_error_cp = 0.0
 counted_points_cp = 0
 avg_pos_error_mbr = 0.0
-counted_points_mbr = 0
+counted_data_points_mbr = []
 for feature_osm in OSM_data['features']:
     # Footprints to a single OSM footprint
     if feature_osm['id'] in overlapping_buildings_osm_bigger:
@@ -80,7 +80,7 @@ for feature_osm in OSM_data['features']:
         tc_osm = geometry_utils.turning_function(cv_osm)
         tc_slu = geometry_utils.turning_function(cv_slu)
 
-        plot_utils.plot_lines([tc_osm, tc_slu])
+        #plot_utils.plot_lines([tc_osm, tc_slu])
 
         # Normalized shape dissimilarities, only if there is a 1:1 match
         if len(polygon_slu) == 1:
@@ -121,7 +121,12 @@ for feature_osm in OSM_data['features']:
         #print(edges_on_perimeter_osm)
         #print(edges_on_perimeter_slu)
 
+        # Calculate average positional offset
+        # Position accuracy per building by taking the average of the distance between the corresponding points
+        avg_point_distance = 0.0
+        counted_points = 0
         # While both edge sets are non-emptyh
+        #TODO: Try to only match poisition error for 1:1 building matches, and see what that does for the spread of the results
         #TODO: MBR method: Record which quadrant (MBR edge) That the edge was found on. Only match edges on the same sides
         while edges_on_perimeter_osm and edges_on_perimeter_slu:
             edge_osm = edges_on_perimeter_osm[0]
@@ -132,15 +137,19 @@ for feature_osm in OSM_data['features']:
                 if edge_distance < min_edge_distance:
                     best_edge_slu = edge_slu
                     min_edge_distance = edge_distance
-            if(min_edge_distance < 10.0):   # Cutoff point to remove points that were inducted due to mismatching of edges
-                avg_pos_error_mbr += min_edge_distance
-                counted_points_mbr += 2
+            #if(min_edge_distance < 10.0):   # Cutoff point to remove points that were inducted due to mismatching of edges
+            avg_point_distance += min_edge_distance
+            counted_points += 2
                 #print(min_edge_distance)
             edges_on_perimeter_osm.remove(edge_osm)
             edges_on_perimeter_slu.remove(best_edge_slu)
[…]

### 2020-03-27 osm-importer/building-metrics.py
Collect and plot shape dissimilarity data

@@ -58,6 +58,7 @@ avg_pos_error_cp = 0.0
 counted_points_cp = 0
 avg_pos_error_mbr = 0.0
 counted_data_points_mbr = []
+shape_dissimilarity_data = []
 for feature_osm in OSM_data['features']:
     # Footprints to a single OSM footprint
     if feature_osm['id'] in overlapping_buildings_osm_bigger:
@@ -84,10 +85,9 @@ for feature_osm in OSM_data['features']:
 
         # Normalized shape dissimilarities, only if there is a 1:1 match
         if len(polygon_slu) == 1:
-            #TODO: Metric: Bar diagram of footprint shape similarity (Fan et al, page 12)
             # Only buildings with a 1:1 semantic relationship are involved in the measurement of of positional and shape accuracy. Should we do this too?
             shape_dissimilarity_normalized = geometry_utils.normalized_shape_dissimilarity(polygon_osm, polygon_slu)
-            print("1:1 MATCH MATCH MATCH")
+            shape_dissimilarity_data.append(shape_dissimilarity_normalized)
 
         #plot_utils.plot_polygons([cv_osm, polygon_osm])
         #plot_utils.plot_polygons([cv_slu] + polygon_slu)
@@ -144,9 +144,7 @@ for feature_osm in OSM_data['features']:
             edges_on_perimeter_osm.remove(edge_osm)
             edges_on_perimeter_slu.remove(best_edge_slu)
 
-        print(counted_points)
         if counted_points > 0:
-            print(avg_point_distance/counted_points)
             avg_pos_error_mbr += avg_point_distance/counted_points
             counted_data_points_mbr.append(avg_point_distance/counted_points)
         #TODO: Metric: Max, min and std deviation of position offsets (Fan et al, page 12)
@@ -183,3 +181,5 @@ print("Average position error: " + str(avg_pos_error_mbr) + " (MBR method, reaso
 # Metric: Bar diagram of position offsets (Fan et al, page 12)
 plot_utils.plot_bar(counted_data_points_mbr, 1.0)
 
+# Metric: Bar diagram of footprint shape similarity (Fan et al, page 12)
+plot_utils.plot_bar(shape_dissimilarity_data, 0.1)

### 2020-03-27 osm-importer/building-metrics.py
Mean, max, min and stdev of position offsets

@@ -5,6 +5,7 @@
 from geojson import Point, Feature, FeatureCollection, load
 import geometry_utils
 import plot_utils
+import statistics
 
 OSM_data = {}
 SLU_data = {}
@@ -56,7 +57,6 @@ for feature_osm in OSM_data['features']:
 # Geometrical distance between the two sets
 avg_pos_error_cp = 0.0
 counted_points_cp = 0
-avg_pos_error_mbr = 0.0
 counted_data_points_mbr = []
 shape_dissimilarity_data = []
 for feature_osm in OSM_data['features']:
@@ -145,9 +145,7 @@ for feature_osm in OSM_data['features']:
             edges_on_perimeter_slu.remove(best_edge_slu)
 
         if counted_points > 0:
-            avg_pos_error_mbr += avg_point_distance/counted_points
             counted_data_points_mbr.append(avg_point_distance/counted_points)
-        #TODO: Metric: Max, min and std deviation of position offsets (Fan et al, page 12)
 
         for i in range(0, len(mbr_osm)):
             p1 = mbr_osm[i]
@@ -175,8 +173,11 @@ print("Total area, fraction: " + str(total_area_OSM / total_area_SLU))
 avg_pos_error_cp /= counted_points_cp
 print("Average position error: " + str(avg_pos_error_cp) + " (Counting Points method, upper threshold)")
 
-avg_pos_error_mbr /= len(counted_data_points_mbr)
-print("Average position error: " + str(avg_pos_error_mbr) + " (MBR method, reasonable)")
+# Metric: Max, min and std deviation of position offsets (Fan et al, page 12)
+print("Average position error: " + str(statistics.mean(counted_data_points_mbr)) + " (MBR method, reasonable)")
+print("Position error, max: " + str(max(counted_data_points_mbr)))
+print("Position error, min: " + str(min(counted_data_points_mbr)))
+print("Position error, stdev: " + str(statistics.stdev(counted_data_points_mbr)))
 
 # Metric: Bar diagram of position offsets (Fan et al, page 12)
 plot_utils.plot_bar(counted_data_points_mbr, 1.0)

### 2020-04-08 osm-importer/building-metrics.py
Building metrics script, save statistics with pickle

@@ -213,6 +213,24 @@ with open('metric-data-pos-acc', 'wb') as fp:
 with open('metric-data-shape-diss-norm', 'wb') as fp:
     pickle.dump(shape_dissimilarity_data, fp)
 
+# Write important statistical features to a dictionary
+statistics_dict = {}
+
+statistics_dict["building_count_osm"] = len(OSM_data['features'])
+statistics_dict["building_count_slu"] = len(SLU_data['features'])
+
+statistics_dict["total_area_osm"] = total_area_OSM
+statistics_dict["total_area_slu"] = total_area_SLU
+
+statistics_dict["total_matches_count"] = total_matches_count
+statistics_dict["one_to_one_matches_count"] = one_to_one_matches_count
+statistics_dict["one_to_many_matches_count"] = one_to_many_matches_count
+
+print(statistics_dict)
+
+with open('metric-data-statistics', 'wb') as fp:
+    pickle.dump(statistics_dict, fp)
+
 print("Results...")
 
 print("#Buildings, OSM: " + str(len(OSM_data['features'])))
@@ -239,8 +257,3 @@ plot_utils.plot_bar(counted_data_points_mbr, 1.0)
 # Metric: Bar diagram of footprint shape similarity (Fan et al, page 12)
 plot_utils.plot_bar(shape_dissimilarity_data, 0.1)
 
-
-#TODO: Report outline + add space for the metrics and images (think about what metrics and images to include)
-#TODO: Crunch metrics for the report (Remember to remove cutoff point)
-
-#TODO: Pictures for the report

### 2020-04-08 osm-importer/building-statistics-recap.py
Add statistics recap script

@@ -1,0 +1,57 @@
+import pickle
+import statistics
+import plot_utils
+
+counted_data_points_mbr = []
+shape_dissimilarity_data = []
+
+with open('metric-data-pos-acc', 'rb') as fp:
+    counted_data_points_mbr = pickle.load(fp)
+
+with open('metric-data-shape-diss-norm', 'rb') as fp:
+    shape_dissimilarity_data = pickle.load(fp)
+
+statistics_dict = {}
+
+with open('metric-data-statistics', 'rb') as fp:
+    statistics_dict = pickle.load(fp)
+
+building_count_osm = statistics_dict["building_count_osm"]
+building_count_slu = statistics_dict["building_count_slu"]
+
+total_area_OSM = statistics_dict["total_area_osm"]
+total_area_SLU = statistics_dict["total_area_slu"]
+building_count_osm = statistics_dict["building_count_osm"]
+
+total_matches_count = statistics_dict["total_matches_count"]
+one_to_one_matches_count = statistics_dict["one_to_one_matches_count"]
+one_to_many_matches_count = statistics_dict["one_to_many_matches_count"]
+
+print(statistics_dict)
+
+
+print("Results...")
+
+print("#Buildings, OSM: " + str(building_count_osm))
+print("#Buildings, SLU: " + str(building_count_slu))
+
+print("Total area, OSM: " + str(total_area_OSM))
+print("Total area, SLU: " + str(total_area_SLU))
+print("Total area, fraction: " + str(total_area_OSM / total_area_SLU))
+
+# Metric: Statistic of the matching result using area overlap (Fan et al, page 9)
+print("Number of 1:1 matches: " + str(one_to_one_matches_count))
+print("Number of 1:N matches: " + str(one_to_many_matches_count))
+print("Number of 1:0 matches: " + str(total_matches_count - one_to_one_matches_count - one_to_many_matches_count))
+
+# Metric: Max, min and std deviation of position offsets (Fan et al, page 12)
+print("Average position error: " + str(statistics.mean(counted_data_points_mbr)) + " (MBR method, reasonable)")
+print("Position error, max: " + str(max(counted_data_points_mbr)))
+print("Position error, min: " + str(min(counted_data_points_mbr)))
+print("Position error, stdev: " + str(statistics.stdev(counted_data_points_mbr)))
+
+# Metric: Bar diagram of position offsets (Fan et al, page 12)
[…]

### 2020-04-08 paper/paper.tex
Report todo
Add metric tables with data

@@
 
+% TODO: Figure: Overlapping buildings, OSM-SLU, 1:1
+% TODO: Figure: Overlapping buildings, OSM-SLU, 1:n
+
 \subsection{Building correspondence}
@@
 \subsection{Turning function definition}
+
+% TODO: Turning function example figure for a polygon
 
@@
 
-TODO: Include a distribution of the shap similarity of matched building footprints, see Fan et al, page 14
+% TODO: Turning function overlap, example figure with two polygos
 
@@
 Any edges in the simplified footprints that coinside with the MBR from the same dataset are extracted, and the corresponding points in the original footprints will be matched with each other.
+
+\section{Results, Metrics}
+
+% TODO: Table of positional offset statistics
+% TODO: Table of completeness statistics
+% TODO: Statement on completeness
+
+% TODO: Bar plot positional accuracy
+% TODO: Bar plot turning functions
 
@@
 
+\section{Results, Metrics}
+
+\subsection{Completeness by building area}
+
+\begin{table}[]
+\begin{tabular}{lllll}
+    & Total number of buildings & Area cover & Number of buildings (\%) & Area cover (\%) \\
+    OSM & 102755 & 33105023.32... $m^2$ & 60.17\% & 98.18\% \\
+    SLU & 170783 & 33718628.97... $m^2$ & 100.00\% & 100.00\%
+\end{tabular}
+\end{table}
+
+The area cover of the OSM dataset in relation to the SLU dataset was estimated to 98.18\%.
+
+\subsection{Positional offset statistics}
+
+Building matching statistics:
+
+\begin{table}[]
+\begin{tabular}{lllll}
+    & 1:0 & 1:1 & 1:n & Total \\
+    Match count & 9812 & 81722 & 11221 & 102755
+\end{tabular}
+\end{table}
+
+% TODO: Bar plot positional accuracy
+% TODO: Bar plot turning functions
+
 \section{Analysis}

### 2020-04-08 paper/paper.tex
Creat non-floating tables for metrics in report

@@
 
-\section{Results, Metrics}
-
-% TODO: Table of positional offset statistics
-% TODO: Table of completeness statistics
-% TODO: Statement on completeness
-
-% TODO: Bar plot positional accuracy
-% TODO: Bar plot turning functions
-
 \begin{algorithm}[H]
@@
 
-\begin{table}[]
+\begin{table}[H]
 \begin{tabular}{lllll}
-    & Total number of buildings & Area cover & Number of buildings (\%) & Area cover (\%) \\
-    OSM & 102755 & 33105023.32... $m^2$ & 60.17\% & 98.18\% \\
-    SLU & 170783 & 33718628.97... $m^2$ & 100.00\% & 100.00\%
+    & Total number of buildings & Area cover \\
+    OSM & 102755 & 33105023.32... $m^2$ \\
+    SLU & 170783 & 33718628.97... $m^2$
+\end{tabular}
+\end{table}
+
+\begin{table}[H]
+\begin{tabular}{lllll}
+    & Number of buildings (\%) & Area cover (\%) \\
+    OSM & 60.17\% & 98.18\% \\
+    SLU & 100.00\% & 100.00\%
 \end{tabular}
@@
 
-\begin{table}[]
+\begin{table}[H]
 \begin{tabular}{lllll}

### 2020-04-08 osm-importer/building-statistics-recap.py
Table axis labels and max/min values

@@ -51,7 +51,7 @@ print("Position error, min: " + str(min(counted_data_points_mbr)))
 print("Position error, stdev: " + str(statistics.stdev(counted_data_points_mbr)))
 
 # Metric: Bar diagram of position offsets (Fan et al, page 12)
-plot_utils.plot_bar(counted_data_points_mbr, 1.0)
+plot_utils.plot_bar(counted_data_points_mbr, 0.5, max_value=10.0, min_value=0.0, x_label="Average positional error (m)", y_label="Number of buildings")
 
 # Metric: Bar diagram of footprint shape similarity (Fan et al, page 12)
-plot_utils.plot_bar(shape_dissimilarity_data, 0.1)
+plot_utils.plot_bar(shape_dissimilarity_data, 0.05, max_value=1.1, min_value=0.0, x_label="Turning function distance", y_label="Number of building pairs")

### 2020-04-08 osm-importer/plot_utils.py

@@ -94,26 +94,35 @@ def plot_lines(lines):
     plt.ylabel('y')
     plt.show()
 
-def plot_bar(data_values, bar_width):
+def plot_bar(data_values, bar_width, max_value = None, min_value = None, x_label = None, y_label = None):
+    #TODO: Axis labels
     min_val = min(data_values)
     max_val = max(data_values)
-    counter = min_val
+
+    if max_value is not None:
+        max_val = max_value
+    if min_value is not None:
+        min_val = min_value
+
+    counter = min_val + bar_width/2.0
     bars = []
-    while counter < max_val:
+    while counter <= max_val:
         bars.append(counter)
         counter += bar_width
     values = []
     for i in range(0, len(bars)):
         values.append(0)
     for val in data_values:
-        added = False
         for i in range(0, len(bars)-1):
             if val > bars[i] and val < bars[i+1]:
                 values[i] += 1
-                added = True
                 break
-        if not added:
-            values[-1] += 1
+
+    ax2 = plt.axes()
+    if x_label is not None:
+        ax2.set_xlabel(x_label)
+    if y_label is not None:
+        ax2.set_ylabel(y_label)
 
     plt.bar(bars, values, width=bar_width*0.8)
     plt.show()

### 2020-04-09 paper/paper.tex
Include plots in report
Include table with positional error statistics
Remove test image from report
Add building match example images

@@
 \section{Abstract}
-
-\begin{figure}[H]
-\centering
-\includegraphics[height=0.5\textheight]{test_image}
-\caption{En typisk rymdhiss enligt Wikipedia. Bildkälla (\cite{sampleitem1}).}
-\label{fig:space}
-\end{figure}
 
@@
 
-% TODO: Figure: Overlapping buildings, OSM-SLU, 1:1
-% TODO: Figure: Overlapping buildings, OSM-SLU, 1:n
+\begin{figure}[H]
+\centering
+\includegraphics[height=0.5\textheight]{img_match_1_1}
+\caption{An example of a 1:1 building match, where the sought-after building was found both in the OSM (red) and SLU (blue) datasets.}
+\label{fig:space}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[height=0.5\textheight]{img_match_1_n}
+\caption{An example of a 1:n building match, where the sought-after building was represented as one polygon in the OSM (red) dataset, but divided into many smaller lots in the SLU (multicolored) dataset.}
+\label{fig:space}
+\end{figure}
 
@@
 
-\subsection{Positional offset statistics}
+\subsection{Positional accuracy statistics}
+
+\begin{table}[H]
+\begin{tabular}{lllll}
+    $e_{average}$ & $e_{max}$ & $e_{min}$ & $e_{standard deviation}$ \\
+    2.0579... & 9.9878... & 0.0 & 1.2464...
+\end{tabular}
+\end{table}
 
@@
 
-% TODO: Bar plot positional accuracy
-% TODO: Bar plot turning functions
+% Bar plot positional accuracy
+
+\begin{figure}[H]
+\centering
+\includegraphics[height=0.5\textheight]{img_pos_error_plot}
+\caption{The average positional accuracies per building as a frequency diagram.}
+\label{fig:space}
+\end{figure}
+
+\subsection{Shape accuracy statistics}
+
+% Bar plot turning functions
+
+\begin{figure}[H]
+\centering
+\includegraphics[height=0.5\textheight]{img_turning_function_plot}
+\caption{The turning function distances between 1:1 building pairs as a frequency diagram.}
+\label{fig:space}
+\end{figure}
 

### 2020-04-09 paper/paper.tex
Include turning function plots in report
Report todo
Report abstract

@@
 
-This report covers... \cite{sampleitem2}
+This project assesses the feasibility of creating a 3D representation of geodata with the inclusion of pedestrian footpaths.
+\\
+Procedural modeling of road networks and even entire cities is of interest to the fields of urban planning and urban architecture.
+In recent years there has been an advent of interactive tools that aid urban planners in placing or generating features of an urban plan, particularly roads and road networks.
+These tools are created with the purpose of aiding urban planners by giving them increased visualization capabilities, and the ability to quickly perform editing both in the large scale and small scale, and quickly view the results.
+Specifically, there have been an advent of tools that exract geodata from OpenStreetMap (OSM) and generate 3D representations of real world locations by placing terrain, buildings, roads and other features.
+However, freely available such tools seldom include smaller roads and pedestrian footpaths such as walkways, sidewalks or trails.
+\\
+This project assesses the feasibility of creating such a service, by focusing on OSM road and building data in the Stockholm metropolitan area.
+An estimate of the positional accuracy of OSM data has been obtained by comparing OSM positional data to a dataset with a known accuracy, which was obtained from the Swedish National Land Survey.
+This estimate of the positional error was used to assess the integrity of the OSM Stockholm road network, by locating critical points where supplying the roads and paths with indivdual widths would cause collision with other features.
+An algortitm then took the road network and made adjustments that were within the found margin of error for the OSM dataset, to eliminate as many collisions as possible.
+As secondary contributions, the completeness and shape accuracy of the OSM data set have also been estimated and commented.
+\\
+It was found that the average positional error in the OSM dataset was about 2m, and approximately 98\% of all buildings in the reference dataset were represented in the OSM dataset.
+The report concludes with a discussion of the findings of this study, and an assessment of the feasibility of generating city models with pedestrian footpaths in a number of different applications from urban planning tools to videogames.
 
 \section{Introduction}
+
+This report covers... \cite{sampleitem2}
 
@@
 
-% TODO: Turning function overlap, example figure with two polygos
+% TODO: building match figures in a horisontal grid
+% TODO: figure with buildings-slu-cropped
+% TODO: mention utility buildings inside compounds, source of marginal errors
+
+% TODO: Scrape road data with same coords as the large data run
+% TODO: Organize large building and road datasets
+% TODO: Script that imports both building and road data
+% TODO: Reconstruct geometrical hashmap to take both polygons and lines
+% TODO: Implement new geometrical hashmap in new script
+% TODO: Assign default path widths (min)
+% TODO: Per road, go over narby features and infer max width, log collisions (max width - min width)
+
+\begin{figure}[H]
+\centering
+\includegraphics[height=0.5\textheight]{img_turn_function_diff_filled}
+\caption{An example of the turning functions of two corresponding buildings, with the area between them highlighted.}
+\label{fig:space}
+\end{figure}
 

### 2020-04-13 paper/paper.tex
Fix report figures

@@
     \centering
-\includegraphics[height=0.5\textheight]{img_match_1_1}
-\caption{An example of a 1:1 building match, where the sought-after building was found both in the OSM (red) and SLU (blue) datasets.}
-\label{fig:space}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-\includegraphics[height=0.5\textheight]{img_match_1_n}
-\caption{An example of a 1:n building match, where the sought-after building was represented as one polygon in the OSM (red) dataset, but divided into many smaller lots in the SLU (multicolored) dataset.}
+    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_building_match}
+    \caption{Examples of a 1:1 building match (left), where the sought-after building was found both in the OSM (red) and SLU (blue) datasets, and of a 1:n building match (right), where the sought-after building was represented as one polygon in the OSM (red) dataset, but divided into many smaller lots in the SLU (multicolored) dataset.}
     \label{fig:space}
@@
 \subsection{Turning function definition}
-
-% TODO: Turning function example figure for a polygon
 
@@
 
-% TODO: building match figures in a horisontal grid
 % TODO: figure with buildings-slu-cropped
 % TODO: mention utility buildings inside compounds, source of marginal errors
+% TODO: Turning function example figure for a polygon
+
+% TODO: Update project plan
 
@@
     \centering
-\includegraphics[height=0.5\textheight]{img_turn_function_diff_filled}
+    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_turn_function_diff_filled}
     \caption{An example of the turning functions of two corresponding buildings, with the area between them highlighted.}
@@
     \centering
-\includegraphics[height=0.5\textheight]{img_pos_error_plot}
+    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_pos_error_plot}
     \caption{The average positional accuracies per building as a frequency diagram.}
@@
     \centering
-\includegraphics[height=0.5\textheight]{img_turning_function_plot}
+    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_turning_function_plot}
     \caption{The turning function distances between 1:1 building pairs as a frequency diagram.}

### 2020-04-13 paper/paper.tex
Mention utility buildings as a source of area overlap error

@@
 % TODO: figure with buildings-slu-cropped
-% TODO: mention utility buildings inside compounds, source of marginal errors
 % TODO: Turning function example figure for a polygon
@@
 
+The difference in area cover can partially be explained by small utility buildings that are often not represented in the OpenStreetMap dataset.
+Particularily the residential areas in downtown, Stockholm, are characteristic in that each city block consists of a street front and an open, communal compound in the middle, inside of which there are usually a number of small utility buildings such as bicycle garages, toolsheds and such.
+
 (Fan et al, 2014) claimed that the positional error is largely due to the limited resolution in satellite imagery.

### 2020-04-13 paper/paper.tex
Add SLU vs OSM map images

--- 2020-04-13 paper/paper.tex
@@
 
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_map_compare}
+    \caption{Example segments of the Stockholm metropolitan area with unsystematic errors in precision.
+The SLU building dataset (blue) in superimposed on the OSM dataset (dark gray).
+The top picture is centered on a city block which shows a semantic mismatching between the two datasets.
+The whole block consists of only two buildings in the OSM dataset, but is divided into 15 smaller building lots in the SLU dataset.
+The bottom picture shows 4 smaller buildings whose the positional accuracy is low, due to the buildings being incorrectly scaled, skewed and rotated.}
+    \label{fig:space}
+\end{figure}
+
 \subsection{The accuracy classes that will not be focused on}
@@
 
-% TODO: figure with buildings-slu-cropped
-% TODO: mention utility buildings inside compounds, source of marginal errors
 % TODO: Turning function example figure for a polygon
@@
 
+The difference in area cover can partially be explained by small utility buildings that are often not represented in the OpenStreetMap dataset.
+Particularily the residential areas in downtown, Stockholm, are characteristic in that each city block consists of a street front and an open, communal compound in the middle, inside of which there are usually a number of small utility buildings such as bicycle garages, toolsheds, refuse rooms and such.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_map_utility_buildings}
+    \caption{A map segment which shows an area with many small utility buildings that are represented in the SLU dataset but not in the OSM dataset.}
+    \label{fig:space}
+\end{figure}
+
 (Fan et al, 2014) claimed that the positional error is largely due to the limited resolution in satellite imagery.

### 2020-04-14 osm-importer/building-metrics.py
Logging

@@ -126,6 +126,10 @@ for feature_osm in OSM_data['features']:
         #plot_utils.plot_polygon(polygon_osm)
         #plot_utils.plot_polygons(polygon_slu)
 
+        #if len(polygon_slu) == 1:
+        #    plot_utils.plot_polygons([polygon_osm, polygon_slu[0]])
+
+
         # Normalized shape dissimilarities, only if there is a 1:1 match
         if len(polygon_slu) == 1:
             # Only buildings with a 1:1 semantic relationship are involved in the measurement of of positional and shape accuracy. Should we do this too?
@@ -134,9 +138,12 @@ for feature_osm in OSM_data['features']:
             #print("SDN: " + str(shape_dissimilarity_normalized))
 
             # Turning funcitons
-            #tc_osm, tc_slu = geometry_utils.turning_function_2(polygon_osm, polygon_slu[0])
-            #plot_utils.plot_step_functions([tc_osm, tc_slu])
+            tc_osm, tc_slu = geometry_utils.turning_function_2(polygon_osm, polygon_slu[0])
             #plot_utils.plot_polygons([polygon_slu[0], polygon_osm])
+            #plot_utils.plot_step_functions([tc_osm, tc_slu])
+
+            plot_utils.plot_polygon(polygon_osm)
+            plot_utils.plot_step_function(tc_osm)
 
             #print("Shape dissimilarity:" + str(shape_dissimilarity_normalized))
 

### 2020-04-14 paper/paper.tex
Include image with example turning funciton in report
Remove interactive constraint from report

@@
 \item It will be assumed that the primary, secondary and tertiatry roads all have a fixed width. Primary roads have a set width, and secondary roads and footpaths as well.
-\item It will be assumed that the use case of the application is for generating areas with the size of a neighbourhood or a few city blocks.
-The area that the user will sketch will typically be around 0.1 km2.
 \item The program will not handle terrain features and altitude. It will be assume that the road mesh is projected on a flat 2D surface.
@@
 
-% TODO: Turning function example figure for a polygon
-
 % TODO: Update project plan
@@
 % TODO: Per road, go over narby features and infer max width, log collisions (max width - min width)
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_turn_function}
+    \caption{An example of the turning function of a polygon.}
+    \label{fig:space}
+\end{figure}
 

### 2020-04-14 osm-importer/way-collisions.py
New way scripts and updated execution order

@@ -1,7 +1,8 @@
-# This script reads path data by highway type, and outputs a file with the width of every street by id
-# It requires that way-data-formatter.py has been run first to generate the neccessary road files.
+# This script reads path data by highway type, infers weights and outputs collisions
+# It requires that both OSM way data and OSM building data have been generated in SWEREF
 # REQUIREMENTS:
-# - Run way-data-formatter.py
+# - Run way-convert-to-sweref.py
+# - Run building-convert-to-sweref.py
 
 from geojson import Point, Feature, FeatureCollection, load
 from json import dump

### 2020-04-14 osm-importer/way-collisions.py
Way collisions script: import both way and building data

@@ -21,19 +21,24 @@ standard_widths = {
         'primary': 1.0
         }
 
-path_data = {}
+way_data = {}
+building_data = []
 
 # read files
 for hwy in highway_categories:
     with open('raw_data/' + hwy + '.geojson', 'r') as f:
-        path_data[hwy] = load(f)
+        way_data[hwy] = load(f)
+
+with open('raw_data/buildings-osm-sweref.geojson', 'r') as f:
+    building_data = load(f)
+
 
 # Store all found paths by street name
 paths_by_name = {}
 widths_by_id = {}
 
-for hwy in path_data.keys():
-    for feature in path_data[hwy]['features'] :
+for hwy in way_data.keys():
+    for feature in way_data[hwy]['features'] :
         properties = feature['properties']
 
         # Set width to default

### 2020-04-14 osm-importer/way-convert-to-sweref.py
Way coordinate conversion script

@@ -1,0 +1,53 @@
+# Coverts processed OSM way data to sweref coordinates
+# REQUIREMENTS:
+# - Run way-data-formatter.py
+
+from geojson import Point, Feature, FeatureCollection, load, dump
+import geometry_utils
+
+def step_recursively(c):
+    if type(c) == list:
+        if type(c[0]) == float and len(c) == 2:
+            # Found a float pair
+            c_new = geometry_utils.wgs84_to_epsg3006(c)
+            c[0] = c_new[0]
+            c[1] = c_new[1]
+        else:
+            for c_sub in c:
+                step_recursively(c_sub)
+    else:
+        print(type(c))
+
+highway_categories = ['footpath', 'residential', 'secondary', 'primary']
+
+way_data = {}
+
+print("Reading files...")
+
+# Read files
+for hwy in highway_categories:
+    with open('raw_data/' + hwy + '.geojson', 'r') as f:
+        way_data[hwy] = load(f)
+
+total_len = 0
+total_progress = 0
+for hwy in highway_categories:
+    total_len += len(way_data[hwy]['features'])
+
+print("Total number of features to convert: " + str(total_len))
+
+# Conversion run
+for hwy in highway_categories:
+    print("Number of " + hwy + " features to convert: " + str(len(way_data[hwy]['features'])))
+    progress = 0
+    for feature in way_data[hwy]['features']:
+        step_recursively(feature['geometry']['coordinates'])
+        # Progress
+        print("Converting " + hwy + " features, progess: " + str(progress) + "/" + str(len(way_data[hwy]['features'])) + " | Total progress: " + str(100*total_progress/total_len) + "%")
+        progress+=1
+        total_progress+=1
+
+# Write all building features to files
+for highway in highway_categories:
+    with open('raw_data/' + highway + '-converted.geojson', 'w') as f:
+        dump(FeatureCollection(way_data[highway]), f)

### 2020-04-14 osm-importer/way-collisions.py
Standard path widths

@@ -15,10 +15,10 @@ highway_categories = [
         ]
 
 standard_widths = {
-        'footpath': 0.25,
-        'residential': 0.5,
-        'secondary': 0.75,
-        'primary': 1.0
+        'footpath': 2.0,        # Measured range: 2.0m - 5.0m
+        'residential': 6.0,     # Measured range: 6.0m - 9.0m
+        'secondary': 6.5,       # Measured range: 7.0m - 16.0m
+        'primary': 8.0          # Measured range: 8.0m - 10.0m
         }
 
 way_data = {}

### 2020-04-14 osm-importer/way-collisions.py
Read converted way data

@@ -26,7 +26,7 @@ building_data = []
 
 # read files
 for hwy in highway_categories:
-    with open('raw_data/' + hwy + '.geojson', 'r') as f:
+    with open('raw_data/' + hwy + '-converted.geojson', 'r') as f:
         way_data[hwy] = load(f)
 
 with open('raw_data/buildings-osm-sweref.geojson', 'r') as f:
@@ -54,7 +54,3 @@ for hwy in way_data.keys():
         else:
             print("[WARNING] The following way did not have a name:\n" + str(feature))
 
-# Write all paths widths to file
-with open('raw_data/path_widths.json', 'w') as f:
-    dump(widths_by_id, f)
-        

### 2020-04-14 osm-importer/geometry_hashtable.py
Optimize hash table for storing both ways and buildings

@@ -35,9 +35,15 @@ class GeometryHashtable:
             for j in range(0, cell_count_y):
                 hashtable[i].append([])
 
+        # Add the feature to every hash bucket that its points belong to
         for feature in features:
-            x, y = self._get_hash_keys_of_feature(feature)
-            hashtable[x][y].append(feature)
+            hashbuckets_added_to = []
+            polygon = geometry_utils.extract_polygon_from_feature(feature)
+            for point in polygon:
+                x, y = self._get_hash_keys_of_feature(point)
+                if not hashtable[x][y] in hashbuckets_added_to:
+                    hashtable[x][y].append(feature)
+                    hashbuckets_added_to.append(hashtable[x][y])
 
         self.hashtable = hashtable
 
@@ -49,17 +55,19 @@ class GeometryHashtable:
         features_out = []
         for x_i in range(max(x-1, 0), min(x+2, len(self.hashtable))):
             for y_i in range(max(y-1, 0), min(y+2, len(self.hashtable[0]))):
-                features_out += self._get_features_in_bucket(x_i, y_i)
+                # Don't add duplicate features
+                features_got = self._get_features_in_bucket(x_i, y_i)
+                for feature in features_got:
+                    if not feature in features_out:
+                        features_out.append(feature)
         return features_out
 
     def _get_features_in_bucket(self, x, y):
         return self.hashtable[x][y]
 
-    def _get_hash_keys_of_feature(self, feature):
-        polygon = geometry_utils.extract_polygon_from_feature(feature)
-        center = geometry_utils.polygon_centroid(polygon)
-        dx = center[0] - self.top_left_x
-        dy = center[1] - self.top_left_y
+    def _get_hash_keys_of_point(self, point):
+        dx = point[0] - self.top_left_x
+        dy = point[1] - self.top_left_y
 
         px = dx - (dx % self.cell_size)
         py = dy - (dy % self.cell_size)

### 2020-04-14 osm-importer/way-collisions.py
Way collisions script, make hashtables

@@ -32,6 +32,16 @@ for hwy in highway_categories:
 with open('raw_data/buildings-osm-sweref.geojson', 'r') as f:
     building_data = load(f)
 
+all_features_list = []
+all_features_list += building_data['features']
+for hwy in highway_categories:
+    all_features_list.append(way_data[hwy]['features'])
+
+# Build polygon hashtables
+print("Building hash tables...")
+hashtable = GeometryHashtable("building-way-hashtable", 230)
+hashtable.create_from_features_list(all_features_list)
+print("Hash tables complete!")
 
 # Store all found paths by street name
 paths_by_name = {}

### 2020-04-15 paper/paper.tex
Update todo
Update TODO
Add note about standard road widths from Trafikverket to report

@@
 As the scope of this project is limited to making a broad comparison of the positional accuracy of the two datasets, only the required positional accuracy as described by the SLU product description will be used.
-TODO: Dig deeper in the SLU dataset, how can one find the method of aquisition for Stockholm buildings, and its precision?
+%TODO: Dig deeper in the SLU dataset, how can one find the method of aquisition for Stockholm buildings, and its precision?
 
@@
 
-% TODO: Update project plan
-
-% TODO: Scrape road data with same coords as the large data run
-% TODO: Organize large building and road datasets
-% TODO: Script that imports both building and road data
-% TODO: Reconstruct geometrical hashmap to take both polygons and lines
-% TODO: Implement new geometrical hashmap in new script
-% TODO: Assign default path widths (min)
-% TODO: Per road, go over narby features and infer max width, log collisions (max width - min width)
+% TODO: Assign default road widths
+% TODO: Implement closest edge distance (project end points on the other edge, get closest perpendicular distance) 
+% TODO: Per road, go over nearby features and infer max width, log collisions (max width - min width)
 
@@
 However, when calculating the relative area overlap, this becomes unesseccary since any semantically incorrect subdivision of the polygon will always be inside the bounding polygon.
+
+\section{Collision checks}
+
+The default minimum widths were obtained from (Lantmäteriet, 20). The standard minimum width for a two-lane road in Sweden is 6.5 meters.
 
@@
 
+\bibitem{trafikverket-20}
+https://trafikverket.ineko.se/Files/sv-SE/71830/Ineko.Product.RelatedFiles/2020\_029\_vagar\_och\_gators\_utformning\_krav.pdf \\
+
 \end{thebibliography}

### 2020-04-16 osm-importer/geometry_utils.py
Implement closest edge distance, WIP

@@ -259,6 +259,45 @@ def perp_distance_point_to_line(point, line_point_1, line_point_2):
     denom = math.sqrt((y2-y1)**2 + (x2-x1)**2)
     return nom/denom
 
+def shortest_distance_between_edges_projected(edge_1, edge_2):
+    a1 = edge_1[0]
+    a2 = edge_1[1]
+    b1 = edge_2[0]
+    b2 = edge_2[1]
+    a = [a2[0]-a1[0], a2[1]-a1[1]]
+    b = [b2[0]-b1[0], b2[1]-b1[1]]
+
+    b_proj_a = project(b, a)
+    arc_theta = point_len(b_proj_a)/point_len(b)
+
+    b1_proj_a = project(b1, a)
+    b1_proj_a = [b1_proj_a[0]-a1[0], b1_proj_a[1]-a1[1]]
+    b2_proj_a = project(b2, a)
+    b2_proj_a = [b2_proj_a[0]-a2[0], b2_proj_a[1]-a2[1]]
+
+    if np.dot(b1_proj_a, a) < 0 and np.dot(b2_proj_a, a) < 0:
+        return None
+    if np.dot(b1_proj_a, a) > np.dot(a, a) and np.dot(b2_proj_a, a) > np.dot(a, a):
+        return None
+
+    if np.dot(b1_proj_a, a) >= 0:
+        ba1 = [a1[0]-b1[0], a1[1]-b1[1]]
+        bp1 = [ba1[0]/arc_theta, ba1[1]/arc_theta]
+        p1 = [bp1[0]-b1[0], bp1[1]-b1[1]]
+        line_dist_1 = perp_distance_point_to_line(p1, a1, a2)
+    else:
+        line_dist_1 = perp_distance_point_to_line(b1, a1, a2)
+    
+    if np.dot(b2_proj_a, a) <= 0:
+        ba2 = [a2[0]-b2[0], a2[1]-b2[1]]
+        bp2 = [ba2[0]/arc_theta, ba2[1]/arc_theta]
+        p2 = [bp2[0]-b2[0], bp2[1]-b2[1]]
+        line_dist_2 = perp_distance_point_to_line(p2, a1, a2)
+    else:
+        line_dist_2 = perp_distance_point_to_line(b2, a1, a2)
+
+    return min(line_dist_1, line_dist_2)
+
 def douglas_peucker(polygon, e):
     # Find starting point, use northernmost point
     #Pick northernmost point to ensure it is the same in both OSM and SLU cases

### 2020-04-16 osm-importer/way-collisions.py
Fix closest edge distance call and debugging

@@ -5,6 +5,7 @@
 # - Run building-convert-to-sweref.py
 
 import geometry_utils
+import plot_utils
 from geojson import Point, Feature, FeatureCollection, load
 from geometry_hashtable import GeometryHashtable
 from json import dump
@@ -56,6 +57,9 @@ for hwy in way_data.keys():
         feature.max_way_width = 2*feature.min_way_width
 
         for feature_2 in hashtable.get_collision_canditates(feature):
+            if feature is feature_2:
+                continue
+
             # Get polygons
             polygon_1 = geometry_utils.extract_polygon_from_feature(feature)
             polygon_2 = geometry_utils.extract_polygon_from_feature(feature_2)
@@ -63,17 +67,25 @@ for hwy in way_data.keys():
             # Get edges
             for i in range(0, len(polygon_1)-1):
                 edge_1 = [polygon_1[i], polygon_1[i+1]]
-                for j in range(0, len(polygon_2)):
+                for j in range(0, len(polygon_2)-1):
                     edge_2 = [polygon_2[j], polygon_2[j+1]]
 
-                    shortest_dist = geometry_utils.shortest_distance_between_edges(edge_1, edge_2)
-                    if shortest_dist < feature.way_width:
+                    shortest_dist = geometry_utils.shortest_distance_between_edges_projected(edge_1, edge_2)
+                    if shortest_dist == None:
+                        continue
+                    print("Shortest dist was not None! It was: " + str(shortest_dist))
+                    plot_utils.plot_polygons([polygon_1, polygon_2])
+                    if shortest_dist < feature.min_way_width:
                         print("Features collision!")
                         print(feature)
                         print(feature_2)
-                        print("min dist: " + feature.min_way_width)
+                        print("min dist: " + str(feature.min_way_width))
                         print("************")
+                        plot_utils.plot_polygons([polygon_1, polygon_2])
                     else:
                         feature.max_way_width = min(feature.max_way_width, shortest_dist)
 
+            break
+        break
+    break
 

### 2020-04-19 osm-importer/geometry_utils.py
Fix angle error by checking when edge points are the same

@@ -310,6 +310,9 @@ def shortest_distance_between_edges_projected(edge_1, edge_2):
 
     x = line_line_intersection([a1, a2], [b1, b2])
 
+    if a1 == b1 or a1 == b2 or a2 == b1 or a2 == b2:
+        return 0.0
+
     # Case 1, no intersection, lines are parallel
     if x == None:
         return perp_distance_point_to_line(a1, b1, b2)

### 2020-04-19 osm-importer/way-collisions.py
Do not handle cases where two roads are intersecting or joining

@@ -85,7 +85,13 @@ for hwy in way_data.keys():
                     shortest_dist = geometry_utils.shortest_distance_between_edges_projected(edge_1, edge_2)
                     if shortest_dist == None:
                         continue
-                    print("Shortest dist was not None! It was: " + str(shortest_dist))
+
+                    # For two roads, move on if the distance is 0 (meaning adjoining roads or intersections)
+                    if 'highway' in feature['properties'] and 'highway' in feature_2['properties']:
+                        if shortest_dist == 0.0:
+                            continue
+                    
+                    #print("Shortest dist was not None! It was: " + str(shortest_dist))
                     #plot_utils.plot_polygons([polygon_1, polygon_2])
                     if shortest_dist < feature.min_way_width:
                         print("Features collision!")
@@ -93,7 +99,7 @@ for hwy in way_data.keys():
                         print(feature_2)
                         print("min dist: " + str(feature.min_way_width))
                         print("************")
-                        plot_utils.plot_polygons([polygon_1, polygon_2])
+                        #plot_utils.plot_polygons([polygon_1, polygon_2])
                     else:
                         feature.max_way_width = min(feature.max_way_width, shortest_dist)
 

### 2020-04-19 osm-importer/way-collisions.py
Update Todo

@@ -107,3 +107,5 @@ for hwy in way_data.keys():
 
     break
 
+#TODO: Printa statistik: antal kolliderande features, antal kolliderande noder (#/%)
+#TODO: Printa statistik: längd av kolliderande väg, total längd

### 2020-04-19 osm-importer/geometry_utils.py
Import plot utils

@@ -2,6 +2,7 @@ import shapely.geometry
 import math
 import numpy as np
 import copy
+import plot_utils
 from pyproj import Proj, transform
 
 def wgs84_to_epsg3006(point):

### 2020-04-19 osm-importer/geometry_utils.py
Obtain closest node to intersection point when going over features

@@ -311,12 +311,14 @@ def shortest_distance_between_edges_projected(edge_1, edge_2):
 
     x = line_line_intersection([a1, a2], [b1, b2])
 
-    if a1 == b1 or a1 == b2 or a2 == b1 or a2 == b2:
-        return 0.0
+    if a1 == b1 or a1 == b2:
+        return 0.0, a1
+    if a2 == b1 or a2 == b2:
+        return 0.0, a2
 
     # Case 1, no intersection, lines are parallel
     if x == None:
-        return perp_distance_point_to_line(a1, b1, b2)
+        return perp_distance_point_to_line(a1, b1, b2), a1
 
     # First, check if both the projections of b1 and b2 on a are outside of the a domain. If so, return None
     b1_x = [b1[0]-x[0], b1[1]-x[1]]
@@ -330,22 +332,21 @@ def shortest_distance_between_edges_projected(edge_1, edge_2):
     a2_d1 = [x_d1[0]-a2[0], x_d1[1]-a2[1]]
     a2_d2 = [x_d2[0]-a2[0], x_d2[1]-a2[1]]
     if np.dot(a1_d1, a2_d1) > 0 or np.dot(a1_d2, a2_d2) > 0:
-        return None
-
-    #points = [a1, a2, b1, b2, x, x_d1, x_d2]
-    #edges = [[0,1],[2,3]]
-    #plot_utils.plot_edges_and_points(points, edges)
+        return None, None
 
     # Case 2, x between a1 and a2
     x_a1 = [x[0]-a1[0], x[1]-a1[1]]
     a2_x = [a2[0]-x[0], a2[1]-x[1]]
     if np.dot(x_a1, a2_x) >= 0.0:
         if line_segment_line_segment_intersection(edge_1, edge_2):
-            return 0.0
+            return 0.0, a1
         else:
             perp_dist_1 = perp_distance_point_to_line(b1, a1, a2)
             perp_dist_2 = perp_distance_point_to_line(b2, a1, a2)
-            return min(perp_dist_1, perp_dist_2)
+            if perp_dist_1 <= perp_dist_2:
+                return perp_dist_1, a1
+            else:
+                return perp_dist_2, a2
 
     # Case 3, x < a1
     if point_distance(x, a1) < point_distance(x, a2):
@@ -375,7 +376,7 @@ def shortest_distance_between_edges_projected(edge_1, edge_2):
     theta = math.acos(point_len(d)/point_distance(b, x))
[…]

### 2020-04-19 osm-importer/way-collisions.py

@@ -60,6 +60,8 @@ for hwy in way_data.keys():
         feature.min_way_width = standard_widths[hwy]
         feature.max_way_width = 2*feature.min_way_width
 
+        feature_collided = False
+
         for feature_2 in hashtable.get_collision_canditates(feature):
             if feature is feature_2:
                 continue
@@ -82,7 +84,7 @@ for hwy in way_data.keys():
                 for j in range(0, len(polygon_2)-1):
                     edge_2 = [polygon_2[j], polygon_2[j+1]]
 
-                    shortest_dist = geometry_utils.shortest_distance_between_edges_projected(edge_1, edge_2)
+                    shortest_dist, closest_node = geometry_utils.shortest_distance_between_edges_projected(edge_1, edge_2)
                     if shortest_dist == None:
                         continue
 
@@ -104,6 +106,8 @@ for hwy in way_data.keys():
                         feature.max_way_width = min(feature.max_way_width, shortest_dist)
 
         feature.handled = True
+        if feature_collided:
+            stat_collision_feature_count += 1
 
     break
 

### 2020-04-19 osm-importer/way-collisions.py
Statistics from way collisions run, print and save

@@ -6,6 +6,7 @@
 
 import geometry_utils
 import plot_utils
+import pickle
 from geojson import Point, Feature, FeatureCollection, load
 from geometry_hashtable import GeometryHashtable
 from json import dump
@@ -50,6 +51,16 @@ print("Hash tables complete!")
 paths_by_name = {}
 widths_by_id = {}
 
+stat_collision_feature_count = 0
+stat_collision_node_count = 0
+stat_collision_edge_len = 0.0
+stat_total_features_count = 0
+stat_total_node_count = 0
+stat_total_edge_len = 0.0
+
+for hwy in highway_categories:
+    stat_total_features_count += len(way_data[hwy]['features'])
+
 for hwy in way_data.keys():
     for feature in way_data[hwy]['features'] :
         # Skip tunnels
@@ -81,6 +92,8 @@ for hwy in way_data.keys():
             # Get edges
             for i in range(0, len(polygon_1)-1):
                 edge_1 = [polygon_1[i], polygon_1[i+1]]
+                stat_total_node_count += 1
+                stat_total_edge_len += geometry_utils.point_distance(edge_1[0], edge_1[1])
                 for j in range(0, len(polygon_2)-1):
                     edge_2 = [polygon_2[j], polygon_2[j+1]]
 
@@ -96,6 +109,18 @@ for hwy in way_data.keys():
                     #print("Shortest dist was not None! It was: " + str(shortest_dist))
                     #plot_utils.plot_polygons([polygon_1, polygon_2])
                     if shortest_dist < feature.min_way_width:
+                        stat_collision_node_count += 1
+                        if closest_node == polygon_1[0]:
+                            stat_collision_edge_len += geometry_utils.point_distance(polygon_1[0], polygon_1[1]) / 2
+                        elif closest_node == polygon_1[-1]:
+                            stat_collision_edge_len += geometry_utils.point_distance(polygon_1[-2], polygon_1[-1]) / 2
+                        else:
+                            index = polygon_1.index(closest_node)
+                            stat_collision_edge_len += geometry_utils.point_distance(polygon_1[index-1], polygon_1[index]) / 2
+                            stat_collision_edge_len += geometry_utils.point_distance(polygon_1[index], polygon_1[index+1]) / 2
+
+                        feature_collided = True
+
                         print("Features collision!")
                         print(feature)
                         print(feature_2)
[…]

### 2020-04-19 osm-importer/geometry_utils.py
Logging

@@ -373,7 +373,11 @@ def shortest_distance_between_edges_projected(edge_1, edge_2):
     #print("d: " + str(d))
     #print("len(d): " + str(point_len(d)))
     #print("len(b): " + str(point_distance(b, x)))
+    #points = [a1, a2, b1, b2, x]
+    #edges = [[0,1],[2,3]]
+    #plot_utils.plot_edges_and_points(points, edges)
     theta = math.acos(point_len(d)/point_distance(b, x))
+
     #print(math.tan(theta)*point_len(k))
 
     return math.tan(theta)*point_len(k), a

### 2020-04-19 paper/paper.tex
Update todo
Update Todo
Add way collision statistics table to report
Report: equation centering

@@
 The lab is particularly interested in the generation of footpaths, since most commercial software that import road maps only import arterial and secondary car roads, and do not include pedestrian walkways, cycle paths and such smaller roads.
-% TODO: any official description of ESAL that can be used here?
 
@@
 As the scope of this project is limited to making a broad comparison of the positional accuracy of the two datasets, only the required positional accuracy as described by the SLU product description will be used.
-%TODO: Dig deeper in the SLU dataset, how can one find the method of aquisition for Stockholm buildings, and its precision?
 
@@
 The similarity function between any building in the OSM data set and any building in the SLU datasets will be defined as a function of their respective footprints ($foot_{OSM}, foot_{SLU}$), as follows:
-$S(foot_{OSM}, foot_{SLU}) = S_{T}(foot_{OSM}, foot_{SLU}) + S_{RO}(foot_{OSM}, foot_{SLU})$, where $S_{T}$ measures the difference in turning function and $S_{RO}$ measures the relative overlap between the footprints.
+\begin{center}
+    $S(foot_{OSM}, foot_{SLU}) = S_{T}(foot_{OSM}, foot_{SLU}) + S_{RO}(foot_{OSM}, foot_{SLU})$
+\end{center}
+where $S_{T}$ measures the difference in turning function and $S_{RO}$ measures the relative overlap between the footprints.
 
@@
 The tangent angle at $v_1$ is $\theta_{n,1}$. This is the angle between the neighbouring line segments $e_n$ and $e_1$.
-For any $i$ such that $i>1$ and $n<i$, the tangent angle at $v_i$ is recursively defined as: $\theta_{i, i-1} = \theta_{n,1} + \sum^{i}_{k=1} \theta_{k, k-1}$
+For any $i$ such that $i>1$ and $n<i$, the tangent angle at $v_i$ is recursively defined as:
+\begin{center}
+    $\theta_{i, i-1} = \theta_{n,1} + \sum^{i}_{k=1} \theta_{k, k-1}$
+\end{center}
 The turning function has some nice geometric properties, in that it is invariant to both rotation and scaling of the polygon.
@@
 It also measures only the normalized cumulative length, which does not change under scaling.
-The similarity of two polygons A and B in terms of their turning function is defined as their distance of their cumulative turning functions: $S_{T}(A, B) = 1 - (\int^{1}_{0} T_{C,A}(l) - T_{C,B}(l) dl)^{1/2}$.
+The similarity of two polygons A and B in terms of their turning function is defined as their distance of their cumulative turning functions:
+\begin{center}
+    $S_{T}(A, B) = 1 - (\int^{1}_{0} T_{C,A}(l) - T_{C,B}(l) dl)^{1/2}$.
+\end{center}
 The value range will be $(0 < S_{T} < 1)$, where $S_{T}(A,B) = 1$ if the polygons are identical.
 
-% TODO: Assign default road widths
-% TODO: Implement closest edge distance (project end points on the other edge, get closest perpendicular distance) 
-% TODO: Per road, go over nearby features and infer max width, log collisions (max width - min width)
+% REPORT
+% TODO: RELATED WORK, mention current sources (Haklay, Fan)
+% TODO: ANALYSIS SECTION, figure out what to write
+% TODO: Figure of street, before and after correction, with collision visible in the first half
+
+% TODO: Format formulas
+% TODO: Quote Todor's paper for geodata import services and current state of the field
+
+% TODO: Dig deeper in the SLU dataset, how can one find the method of aquisition for Stockholm buildings, and its precision?
 
@@
 
+\begin{center}
     $S_{RO}(foot_{OSM}, foot_{SLU}) = \frac{A_{Overlap}}{min(A_{foot_{OSM}}, A_{foot_{SLU}}))}$
+\end{center}
 
@@
 
+\subsection{Way collision statistics}
+
+\begin{table}[H]
+\begin{tabular}{lllll}
+                    & Colliding     & Total        & Percent \\
+    Features        & 186           & 1951         & 9.53\% \\
+    Nodes           & 458           & 264179       & 0.17\% \\
+    Edge length     & 14445.66 m    & 5004997.66 m & 0.29\%
+\end{tabular}
+\end{table}
+
 \section{Analysis}

### 2020-04-19 osm-importer/way-collisions.py
way-collisions.py progress bar

@@ -61,11 +61,15 @@ stat_total_edge_len = 0.0
 for hwy in highway_categories:
     stat_total_features_count += len(way_data[hwy]['features'])
 
+progress = 0.0
 for hwy in way_data.keys():
     for feature in way_data[hwy]['features'] :
         # Skip tunnels
         if 'tunnel' in feature['properties'] and feature['properties']['tunnel'] == 'yes':
             continue
+
+        print("Feature collision check, progess: " + str(100*progress/stat_total_features_count) + '%')
+        progress+=1.0
 
         # Set width to default
         feature.min_way_width = standard_widths[hwy]

### 2020-04-19 osm-importer/way-collisions.py
Rework way-collisions.py to print statistics for all highway categories

@@ -51,15 +51,23 @@ print("Hash tables complete!")
 paths_by_name = {}
 widths_by_id = {}
 
-stat_collision_feature_count = 0
-stat_collision_node_count = 0
-stat_collision_edge_len = 0.0
-stat_total_features_count = 0
-stat_total_node_count = 0
-stat_total_edge_len = 0.0
+stat_collision_feature_count = {}
+stat_collision_node_count = {}
+stat_collision_edge_len = {}
+stat_total_features_count = {}
+stat_total_node_count = {}
+stat_total_edge_len = {}
+total_total_features_count = 0
+
 
 for hwy in highway_categories:
-    stat_total_features_count += len(way_data[hwy]['features'])
+    total_total_features_count += len(way_data[hwy]['features'])
+    stat_collision_feature_count[hwy] = 0
+    stat_collision_node_count[hwy] = 0
+    stat_collision_edge_len[hwy] = 0.0
+    stat_total_features_count[hwy] = 0
+    stat_total_node_count[hwy] = 0
+    stat_total_edge_len[hwy] = 0.0
 
 progress = 0.0
 for hwy in way_data.keys():
@@ -68,8 +76,10 @@ for hwy in way_data.keys():
         if 'tunnel' in feature['properties'] and feature['properties']['tunnel'] == 'yes':
             continue
 
-        print("Feature collision check, progess: " + str(100*progress/stat_total_features_count) + '%')
+        print("Feature collision check, progess: " + str(100*progress/total_total_features_count) + '%')
         progress+=1.0
+
+        stat_total_features_count[hwy] += 1
 
         # Set width to default
         feature.min_way_width = standard_widths[hwy]
@@ -96,8 +106,8 @@ for hwy in way_data.keys():
             # Get edges
             for i in range(0, len(polygon_1)-1):
                 edge_1 = [polygon_1[i], polygon_1[i+1]]
-                stat_total_node_count += 1
-                stat_total_edge_len += geometry_utils.point_distance(edge_1[0], edge_1[1])
+                stat_total_node_count[hwy] += 1
+                stat_total_edge_len[hwy] += geometry_utils.point_distance(edge_1[0], edge_1[1])
                 for j in range(0, len(polygon_2)-1):
                     edge_2 = [polygon_2[j], polygon_2[j+1]]
 
@@ -113,15 +123,15 @@ for hwy in way_data.keys():
                     #print("Shortest dist was not None! It was: " + str(shortest_dist))
[…]

### 2020-04-19 report/report.tex
Update way collision statistics in the report

--- 2020-04-19 paper/paper.tex
@@
 
-% REPORT
+% TODAY
+% TODO: way-collisions.py: Progress bar
+% TODO: Quote Todor's paper for geodata import services and current state of the field
 % TODO: RELATED WORK, mention current sources (Haklay, Fan)
 % TODO: ANALYSIS SECTION, figure out what to write
-% TODO: Figure of street, before and after correction, with collision visible in the first half
+% TODO: Figure of street, before correction
 
-% TODO: Format formulas
-% TODO: Quote Todor's paper for geodata import services and current state of the field
+% TOMORROW
+% TODO: Proof read entire report
+% TODO: Plan collision correction
+% TODO: Collision correction run
+% TODO: Table for collision correction data in report
 
+% TUESDAY
+% TODO: Figure of street, after correction
+% TODO: Write analysis section
 % TODO: Dig deeper in the SLU dataset, how can one find the method of aquisition for Stockholm buildings, and its precision?
+% TODO: Start large data run, way collision
 
@@
 \begin{tabular}{lllll}
-                    & Colliding     & Total        & Percent \\
-    Features        & 186           & 1951         & 9.53\% \\
-    Nodes           & 458           & 264179       & 0.17\% \\
-    Edge length     & 14445.66 m    & 5004997.66 m & 0.29\%
+                            & Footpath      & Residential  & Secondary   & Primary     \\
+    Features, total         & 1391          & 256          & 107         & 148         \\
+    Nodes, total            & 264179        & 26771        & 6997        & 10686       \\
+    Edge length, total      & 5004997.34 m  & 814349.47 m  & 169129.19 m & 307461.81 m \\
+    Features, colliding     & 186           & 65           & 20          & 29          \\
+    Nodes, colliding        & 458           & 239          & 45          & 59          \\
+    Edge length, colliding  & 14445.66 m    & 8887.32 m    & 1467.54 m   & 1930.88 m   \\
+    Features, percent       & 13.37 \%      & 25.39 \%     & 18.69 \%    & 19.59 \%    \\
+    Nodes, percent          & 0.17 \%       & 0.89 \%      & 0.64 \%     & 0.55 \%     \\
+    Edge length, percent    & 0.29 \%       & 1.09 \%      & 0.87 \%     & 0.63 \%
+
 \end{tabular}

### 2020-04-19 osm-importer/plot_utils.py
Helper functions for plotting edges and polygons

@@ -5,6 +5,15 @@ import geometry_utils
 from matplotlib.patches import Polygon
 from matplotlib.collections import PatchCollection
 
+def plot_polygons_and_edges(polygons, points):
+    plot_polygons(polygons,  False)
+    edges = []
+
+    print(len(points))
+    for i in range(0, len(points)-1):
+        edges.append([i, i+1])
+    plot_edges_and_points(points, edges)
+
 def plot_edges_and_points(p, e):
     #points = np.array(p)
     #edges = np.array(e)
@@ -23,8 +32,26 @@ def plot_edges_and_points(p, e):
     plt.plot(x, y, 'ro') # Points
     plt.show()
 
+def plot_edges(e):
+    points = []
+    edges = []
+    t = 0
+    for i in range(0, len(e)):
+        edge_collection = e[i]
+        points += edge_collection
+        for j in range(0, len(edge_collection)-1):
+            edges.append([t, t+1])
+            t += 1
+        t += 1
 
-def plot_polygons(polygons):
+    print(points)
+    print(e)
+    print(edges)
+
+    plot_edges_and_points(points, edges)
+    
+
+def plot_polygons(polygons, plot=True):
     fig, ax = plt.subplots()
     ax.set_xlim(0)
     patches = []
@@ -68,7 +95,8 @@ def plot_polygons(polygons):
         max_x += r/2
 
     plt.axis([min_x,max_x,min_y,max_y])
-    plt.show()
+    if plot:
+        plt.show()
 
 def plot_polygon(polygon):
     plot_polygons([polygon])

### 2020-04-19 osm-importer/way-collisions.py
Plot roads as edges

@@ -140,7 +140,12 @@ for hwy in way_data.keys():
                         print(feature_2)
                         print("min dist: " + str(feature.min_way_width))
                         print("************")
-                        #plot_utils.plot_polygons([polygon_1, polygon_2])
+
+                        if 'highway' in feature_2['properties']:
+                            print("PLOT EDGES")
+                            plot_utils.plot_edges([polygon_2, polygon_1])
+                        else:
+                            plot_utils.plot_polygons_and_edges([polygon_2], polygon_1)
                     else:
                         feature.max_way_width = min(feature.max_way_width, shortest_dist)
 

### 2020-04-19 osm-importer/plot_utils.py
Reduce logging

@@ -9,7 +9,6 @@ def plot_polygons_and_edges(polygons, points):
     plot_polygons(polygons,  False)
     edges = []
 
-    print(len(points))
     for i in range(0, len(points)-1):
         edges.append([i, i+1])
     plot_edges_and_points(points, edges)
@@ -43,11 +42,6 @@ def plot_edges(e):
             edges.append([t, t+1])
             t += 1
         t += 1
-
-    print(points)
-    print(e)
-    print(edges)
-
     plot_edges_and_points(points, edges)
     
 

### 2020-04-19 osm-importer/way-collisions.py

@@ -142,7 +142,6 @@ for hwy in way_data.keys():
                         print("************")
 
                         if 'highway' in feature_2['properties']:
-                            print("PLOT EDGES")
                             plot_utils.plot_edges([polygon_2, polygon_1])
                         else:
                             plot_utils.plot_polygons_and_edges([polygon_2], polygon_1)

### 2020-04-20 osm-importer/way-collisions.py
Queue tentative colliding features

@@ -7,6 +7,7 @@
 import geometry_utils
 import plot_utils
 import pickle
+from queue import Queue
 from geojson import Point, Feature, FeatureCollection, load
 from geometry_hashtable import GeometryHashtable
 from json import dump
@@ -58,7 +59,7 @@ stat_total_features_count = {}
 stat_total_node_count = {}
 stat_total_edge_len = {}
 total_total_features_count = 0
-
+colliding_features_tentative = {}
 
 for hwy in highway_categories:
     total_total_features_count += len(way_data[hwy]['features'])
@@ -68,6 +69,9 @@ for hwy in highway_categories:
     stat_total_features_count[hwy] = 0
     stat_total_node_count[hwy] = 0
     stat_total_edge_len[hwy] = 0.0
+    colliding_features_tentative[hwy] = Queue()
+
+# Collision detection
 
 progress = 0.0
 for hwy in way_data.keys():
@@ -124,6 +128,12 @@ for hwy in way_data.keys():
                     #plot_utils.plot_polygons([polygon_1, polygon_2])
                     if shortest_dist < feature.min_way_width:
                         stat_collision_node_count[hwy] += 1
+
+                        if not feature in colliding_features_tentative[hwy]:
+                            colliding_features_tentative[hwy].enqueue(feature)
+                        if not feature_2 in colliding_features_tentative[hwy]:
+                            colliding_features_tentative[hwy].enqueue(feature_2)
+
                         if closest_node == polygon_1[0]:
                             stat_collision_edge_len[hwy] += geometry_utils.point_distance(polygon_1[0], polygon_1[1]) / 2
                         elif closest_node == polygon_1[-1]:

### 2020-04-20 osm-importer/way-collisions.py
Collision correction, skeleton code

@@ -162,6 +162,96 @@ for hwy in way_data.keys():
         if feature_collided:
             stat_collision_feature_count[hwy] += 1
 
+
+# Collision correction
+# Collision correction: Store collided features in a tentative map. When a collision is fixed, add both feature_1 and feature_2 to the tentative map
+reached_stable = False
+while not reached_stable:
+    reached_stable = True
+
+    feature_1 = colliding_features_tentative.dequeue()
+    polygon_1 = geometry_utils.extract_polygon_from_feature(feature_1)
+    for feature_2 in hashtable.get_collision_canditates(feature):
+        if feature is feature_2:
+            continue
+        polygon_2 = geometry_utils.extract_polygon_from_feature(feature_2)
+    
+        for i in range(0, len(polygon_1)-1):
+            edge_1 = [polygon_1[i], polygon_1[i+1]]
+            for j in range(0, len(polygon_2)-1):
+                edge_2 = [polygon_2[j], polygon_2[j+1]]
+
+                shortest_dist, closest_node = geometry_utils.shortest_distance_between_edges_projected(edge_1, edge_2)
+                if shortest_dist < feature_1.min_way_width:
+
+                    # If a collision was fixed, queue this feature again and continue lates
+                    colliding_features_tentative.enqueue(feature_1)
+                    reached_stable = False
+
+# Collect statistics after correction run
+
+stat_corrected_collision_feature_count = {}
+stat_corrected_collision_node_count = {}
+stat_corrected_collision_edge_len = {}
+stat_corrected_total_features_count = {}
+stat_corrected_total_node_count = {}
+stat_corrected_total_edge_len = {}
+
+for hwy in highway_categories:
+    stat_corrected_collision_feature_count[hwy] = 0
+    stat_corrected_collision_node_count[hwy] = 0
+    stat_corrected_collision_edge_len[hwy] = 0.0
+    stat_corrected_total_features_count[hwy] = 0
+    stat_corrected_total_node_count[hwy] = 0
+    stat_corrected_total_edge_len[hwy] = 0.0
+
+for hwy in way_data.keys():
+    for feature in way_data[hwy]['features'] :
+        for feature_2 in hashtable.get_collision_canditates(feature):
+            if feature is feature_2:
+                continue
+
+            # Get polygons
+            polygon_1 = geometry_utils.extract_polygon_from_feature(feature)
+            polygon_2 = geometry_utils.extract_polygon_from_feature(feature_2)
+
+            # Get edges
+            for i in range(0, len(polygon_1)-1):
[…]

### 2020-04-20 osm-importer/way-collisions.py
Print post-correction statistics

@@ -260,16 +260,24 @@ statistics_dict['stat_collision_node_count'] = stat_collision_node_count
 statistics_dict['stat_total_node_count'] = stat_total_node_count
 statistics_dict['stat_collision_edge_len'] = stat_collision_edge_len
 statistics_dict['stat_total_edge_len'] = stat_total_edge_len
+statistics_dict['stat_corrected_collision_feature_count'] = stat_corrected_collision_feature_count
+statistics_dict['stat_corrected_total_feature_count'] = stat_corrected_total_features_count
+statistics_dict['stat_corrected_collision_node_count'] = stat_corrected_collision_node_count
+statistics_dict['stat_corrected_total_node_count'] = stat_corrected_total_node_count
+statistics_dict['stat_corrected_collision_edge_len'] = stat_corrected_collision_edge_len
+statistics_dict['stat_corrected_total_edge_len'] = stat_corrected_total_edge_len
 
 with open('way-collision-statistics', 'wb') as fp:
     pickle.dump(statistics_dict, fp)
 
 # Results
-    print("Results...")
+print("Results...")
 for hwy in highway_categories:
     print("****************")
     print(hwy)
     print("****************")
+    print("Pre correction")
+    print("----------------")
     print("Colliding features count: " + str(stat_collision_feature_count[hwy]))
     print("Colliding nodes count: " + str(stat_collision_node_count[hwy]))
     print("Colliding edges cumulative length: " + str(stat_collision_edge_len[hwy]))
@@ -281,4 +289,19 @@ for hwy in highway_categories:
     print("% features count: " + str(stat_collision_feature_count[hwy]/stat_total_features_count[hwy]))
     print("% nodes count: " + str(stat_collision_node_count[hwy]/stat_total_node_count[hwy]))
     print("% edges cumulative length: " + str(stat_collision_edge_len[hwy]/stat_total_edge_len[hwy]))
+
+    print("----------------")
+    print("Post correction")
+    print("----------------")
+    print("Colliding features count: " + str(stat_corrected_collision_feature_count[hwy]))
+    print("Colliding nodes count: " + str(stat_corrected_collision_node_count[hwy]))
+    print("Colliding edges cumulative length: " + str(stat_corrected_collision_edge_len[hwy]))
+
+    print("Total features count: " + str(stat_corrected_total_features_count[hwy]))
+    print("Total nodes count: " + str(stat_corrected_total_node_count[hwy]))
[…]

### 2020-04-20 report/report.tex
Update todo
Include way widths in report
Update Todo
Add table for post-correction data

@@
 
-% TODAY
-% TODO: way-collisions.py: Progress bar
+% MONDAY
 % TODO: Quote Todor's paper for geodata import services and current state of the field
@@
 % TODO: ANALYSIS SECTION, figure out what to write
-% TODO: Figure of street, before correction
 
-% TOMORROW
-% TODO: Proof read entire report
-% TODO: Plan collision correction
+% TODO: Collision correction: Fix collisions by relaxing both feature_1 and feature_2
+
+% TUESDAY
+% TODO: Collision correction: Edit edge distance function to factor in width of the other path
 % TODO: Collision correction run
+% TODO: Figure of street, before correction, after correction
 % TODO: Table for collision correction data in report
 
-% TUESDAY
-% TODO: Figure of street, after correction
 % TODO: Write analysis section
 % TODO: Dig deeper in the SLU dataset, how can one find the method of aquisition for Stockholm buildings, and its precision?
+
+% WEDNESDAY
+% TODO: Proof read entire report
 % TODO: Start large data run, way collision
@@
 
+Subsequent visual measurements of a few key locations in the Stockholm area revealed the following minimum way widths:
+\begin{table}[H]
+\begin{tabular}{lllll}
+                  & Footpath & Residential & Secondary & Primary \\
+    Minimum width & 2.0 m    & 6.0 m       & 6.5 m     & 8.0 m   \\
+    Maximum width & 5.0 m    & 9.0 m       & 10.0 m    & 16.0 m
+\end{tabular}
+\end{table}
+
 \section{Results, Metrics}
@@
 
+After collision correction:
+
+\begin{table}[H]
+\begin{tabular}{lllll}
+                            & Footpath      & Residential  & Secondary   & Primary     \\
+    Features, total         & & & & \\
+    Nodes, total            & & & & \\
+    Edge length, total      & m & m & m & m \\
+    Features, colliding     & & & & \\
+    Nodes, colliding        & & & & \\
+    Edge length, colliding  & m & m & m & m \\
+    Features, percent       & \% & \% & \% & \% \\
+    Nodes, percent          & \% & \% & \% & \% \\
+    Edge length, percent    & \% & \% & \% & \%
+
+\end{tabular}
+\end{table}
+
 \section{Analysis}

### 2020-04-21 report/report.tex
Cite Todor
Summary of Todor's viewpoint in Background

@@
 In recent years there has been an advent of interactive tools that aid urban planners in placing or generating features of an urban plan, particularly roads and road networks.
+Procedural and AI solutions that aid the arcitect in a form of human-machine symbiosis.
+
+The last decade has seen multiple startups in the area of software that aid urban arcitecture.
+The software uses procedural models to render a city grid based on real-world locations, or generate new user-specified parameters, and allow for varying degrees of manual editing on grid, road, neighbourhood, city block or individual feature level.
+Some of the most prominent examples are CityEngine (made by the startup Procedural, bought by ESRI), Urban Canvas (made by the startup Synthicity, bought by Autodesk) and ArcGIS (***).
+
+But such tools are as of today not commonly applied in urbanist practises, due to failures in reflecting the unique workspace of arcitects and their design needs.
+Urban arcitecture is a social factors, stakeholders.
+Each arcitect is an actor in a continously developing environment, and affecting that environment in a complex manner through their action.
+
+Procedural city-generation models combine architectural elemets, such as streets, lots, buildings, building facades.
+
+Urban design practises with conventional two-dimenstional deliverables range from delivering neighbourhood designs, urban zoning and building plans, to writing design codes and building ordinances.
+
+(Stojanivski, 2020)
+
 Additionally, the topic is of interest for 3D content creators in for example the fields of video game design or 3D animation.
@@
 Studios who create large worlds are often interested in solutions that generate large road networks without any hand modeling.
-
-Motivation
-
 Navigation agents, they might fail navigating when faced with a too narrow path or intersecting obstacles
@@
 Kunze, 2012 (reference in Fan et al)
+Haklay
 
@@
 
+\bibitem{stojanovski-20}
+Stojanivski T. 2020. Viewpoint: City Information Modeling (CIM) And Digitizing Urban Design Practises.
+Under review
+
 \end{thebibliography}

### 2020-04-21 report/report.tex
Update todo
Reorganize Background

@@
 
+Procedural city-generation models combine architectural elemets, such as streets, lots, buildings, building facades.
+Urban design practises with conventional two-dimenstional deliverables range from delivering neighbourhood designs, urban zoning and building plans, to writing design codes and building ordinances.
 The last decade has seen multiple startups in the area of software that aid urban arcitecture.
 The software uses procedural models to render a city grid based on real-world locations, or generate new user-specified parameters, and allow for varying degrees of manual editing on grid, road, neighbourhood, city block or individual feature level.
-Some of the most prominent examples are CityEngine (made by the startup Procedural, bought by ESRI), Urban Canvas (made by the startup Synthicity, bought by Autodesk) and ArcGIS (***).
+Some of the most prominent examples are CityEngine (made by the startup Procedural, bought by ESRI), Urban Canvas (made by the startup Synthicity, bought by Autodesk, now defunct) and ArcGIS (***).
 
@@
 Each arcitect is an actor in a continously developing environment, and affecting that environment in a complex manner through their action.
-
-Procedural city-generation models combine architectural elemets, such as streets, lots, buildings, building facades.
-
-Urban design practises with conventional two-dimenstional deliverables range from delivering neighbourhood designs, urban zoning and building plans, to writing design codes and building ordinances.
 
@@
 
-% MONDAY
-% TODO: Quote Todor's paper for geodata import services and current state of the field
+% TUESDAY
 % TODO: RELATED WORK, mention current sources (Haklay, Fan)
@@
 % TODO: Collision correction: Fix collisions by relaxing both feature_1 and feature_2
+% TODO: Collision correction: Edit edge distance function to factor in width of the other path
 
-% TUESDAY
-% TODO: Collision correction: Edit edge distance function to factor in width of the other path
+% WEDNESDAY
 % TODO: Collision correction run
@@
 
-% WEDNESDAY
+% THURSDAY
 % TODO: Proof read entire report

### 2020-04-21 report/report.tex
Mention van Oorts quality criteria

@@
 \section{Related work}
+Van Oorts criteria for evaluating the quality of geographical information, which at the time of publication received attention from surveyors and cartographers. (van Oort, 2006)
+
+Haklay
+
 Fan et al, 2015
 Kunze, 2012 (reference in Fan et al)
-Haklay
 
@@
 
-(Haklay 2010) specifices a comprehesive list of 8 accuracy classes when it comes to evaluating the quality of geographic information.
+(Haklay 2010) presents a comprehesive list of 8 accuracy classes when it comes to evaluating the quality of geographic information (van Oorts criteria)
 
@@
 
+\bibitem{van-oort-06}
+van Oort P A J. 2006. Spatial Data Quality: From Description to Application. PhD Thesis, Wageningen University.
+
 \end{thebibliography}

### 2020-04-21 report/report.tex
Write related work section

@@
 
-Haklay
+Since then a number of publications have examined slices of geodata from van Oort's criteria.
+In 2008, (Haklay, 2010) conducted such a study to estimate of the quality of OpenStreetMap positional geodata in London and England.
+Since the OpenStreetMap (OSM) project started in London it was thought that OSM data in the London metropolitan area would be representative of the highest quality data available, and therefore a good indicator for the whole global OSM dataset.
+The study was conducted by comparing OSM data to Ordnance Survey datasets.
+The study showed that the OSM dataset had a rough positional accuracy of 6 meters from the reference dataset.
+However, at the time of the study, the OSM project had captured roughly 29\% of the area of England, meaning that the data completeness was fairly low.
 
-Fan et al, 2015
-Kunze, 2012 (reference in Fan et al)
+The next notable study to assess the quality of OpenStreetMap data was published by (Kunze, 2012).
+The study applied different methods to assess the completeness of OSM data in two federal states in Germany, mainly by analysing the area difference between the OSM dataset and an administrative dataset.
+
+Finally, a notable study in 2013-2014, conducted by researchers from various universities in Norway and Germany (Fan et al, 2014), examined the quality and accuracy in building footprint data of the Munich area, at a time and place where the completeness of the OSM dataset was significantly higher (the results of their study showed that the OSM had captured close to 100\% of all building footprints).
+At the time Munich was one of the most developed cities in OSM.
+Their study included an insight in the geometrical calculations neccessary to match features and points between two datasets, and how to reliably calculate the metrics needed for the most important and assessible of van Oorts criteria.
+Their study revealed that although the feature completeness of OSM was high, some architectural details were missing.
+The results reveal that the positional accuracy of OSM data at the time was about 4 meters.
 
@@
 
+Volunteered Geographical Information.
 OpenStreetMap road data and building footprints are commonly obtained by manually tracing features in commercially available satellite images.
@@
 % TUESDAY
-% TODO: RELATED WORK, mention current sources (Haklay, Fan)
 % TODO: ANALYSIS SECTION, figure out what to write
@@
 
+\bibitem{kunze-12}
+Kunze, C. 2012. Vergleichsanalyse des Gebäudedatenbestandes aus OpenStreetMap mit amtlichen Datenquellen. \\
+Student research project at the Technical University of Dresden, online avaiable: http://nbn- \\
+resolving.de/urn:nbn:de:bsz:14-qucosa-88141
+
 \end{thebibliography}

### 2020-04-21 report/report.tex
Analysis section, quality of OSM data

@@
 
-\section{Analysis}
+\section{Analysis and Future Work}
+
+\subsection{Quality of the OSM dataset}
+
+Accuracy: we can assume that Stockholm, being one of the more developed cities in one of the most technologically connected country in the world, would be fairly well mapped in OpenStreetMap.
+The positional accuracy has improved by roughly 3.5 meters since the study by (Haklay, 2010), and by roughly 1.5 meters since the study by (Fan et al, 2014), if the SLU reference dataset is to be believed as absolute.
+
+The completeness analysis by building area shows that the OSM Stockholm dataset has high completeness in terms of both building area and counted features, in the same level as the study of Munich by (Fan et al, 2014).
 
@@
 Particularily the residential areas in downtown, Stockholm, are characteristic in that each city block consists of a street front and an open, communal compound in the middle, inside of which there are usually a number of small utility buildings such as bicycle garages, toolsheds, refuse rooms and such.
+These buildings are likely hard to make out and trace on publicly available satellite image data.
+They are, according to (Fan et al, 2014) occluded by their surroundings due to shadows or forestry.
+This appears to be a large consensus in the field.
+
+The low semantic accuracy has made analysis hard.
+The available SLU dataset from Lantmäteriet uses a plot-based subdivision of buildings, whereas OSM uses a subdivision that more closely resembles the actual building footprint as seen from above.
+In many cases whole city blocks are represented as single buildings, which means that many less meaningful feature matches can be found.
+As seen in table ***, about 80\% of all OSM features were found to have a 1:1 mapping to the SLU dataset, but upon further visual comparison it can be seen that many of the 1:n matches are found in a very specified area, namely the residential areas in downtown, Stockholm.
+This is problematic since this typical area with the same level of accessibility was likely measured with the same measuring technique, and the unsystematic positional error will likely be similar in these areas.
+The question is wether it comforms well to the positional error in the rest of the dataset or not.
+This analysis will not be carried out in this project, but could be a good subject for a future master's thesis project.
+
+\subsection{Utility in real world applications}
+
+TODO: Interview with Todor
 

### 2020-04-21 osm-importer/way-collisions.py
If feature 2 is a way, factor in its width

@@ -234,7 +234,14 @@ for hwy in way_data.keys():
                     
                     #print("Shortest dist was not None! It was: " + str(shortest_dist))
                     #plot_utils.plot_polygons([polygon_1, polygon_2])
-                    if shortest_dist < feature.min_way_width:
+                    collision = False
+                    if 'highway' in feature_2['properties']:
+                        # If the other feature is a path, factor in its road width
+                        collision = shortest_dist < feature.min_way_width + feature_2.min_way_width
+                    else:
+                        collision = shortest_dist < feature.min_way_width
+
+                    if collision:
                         stat_correced_collision_node_count[hwy] += 1
 
                         if closest_node == polygon_1[0]:

### 2020-04-22 osm-importer/way-collisions.py
Find translation vector and perform correction by relaxation

@@ -7,6 +7,7 @@
 import geometry_utils
 import plot_utils
 import pickle
+import numpy as np
 from geojson import Point, Feature, FeatureCollection, load
 from geometry_hashtable import GeometryHashtable
 from json import dump
@@ -192,16 +193,45 @@ while not reached_stable:
 
                 if collision:
                     # Feature correction by relaxing both edges in the opposite direction
-                    # TODO: Figure out a way to do dis
+                    perp_1 = geometry_utils.perpendicular([edge_1[1][0]-edge_1[0][0], edge_1[1][1]-edge_1[0][1]])
+                    perp_2 = geometry_utils.perpendicular([edge_2[1][0]-edge_2[0][0], edge_2[1][1]-edge_2[0][1]])
+                    perp_1 = geometry_utils.normalize(perp_1)
+                    perp_2 = geometry_utils.normalize(perp_2)
 
-                    translation_vec = 
+                    translation_vec = None
+                    if np.dot(perp_1, perp_2) > np.dot(perp_1, -perp_2):
+                        translation_vec = [perp_1[0]+perp_2[0], perp_1[1]+perp_2[1]]
+                    else:
+                        translation_vec = [perp_1[0]-perp_2[0], perp_1[1]-perp_2[1]]
+
                     translation_vec = geometry_utils.normalize(translation_vec)
                     translation_dist = shortest_dist/2
 
-                    edge_1[0] = edge_1[0] + translation_vec[0]*translation_dist
-                    edge_1[1] = edge_1[1] + translation_vec[1]*translation_dist
-                    edge_2[0] = edge_2[0] - translation_vec[0]*translation_dist
-                    edge_2[1] = edge_2[1] - translation_vec[1]*translation_dist
+                    edge_1_tr_1 = []
+                    edge_2_tr_1 = []
+                    edge_1_tr_2 = []
+                    edge_2_tr_2 = []
+
+                    edge_1_tr_1[0] = edge_1[0] + translation_vec[0]*translation_dist
+                    edge_1_tr_1[1] = edge_1[1] + translation_vec[1]*translation_dist
+                    edge_2_tr_1[0] = edge_2[0] - translation_vec[0]*translation_dist
+                    edge_2_tr_1[1] = edge_2[1] - translation_vec[1]*translation_dist
+
+                    edge_1_tr_2[0] = edge_1[0] - translation_vec[0]*translation_dist
[…]

### 2020-04-22 osm-importer/way-collisions.py
Add exceptions to checked roads when correcting features

@@ -184,6 +184,15 @@ while not reached_stable:
                 edge_2 = [polygon_2[j], polygon_2[j+1]]
 
                 shortest_dist, closest_node = geometry_utils.shortest_distance_between_edges_projected(edge_1, edge_2)
+
+                if shortest_dist == None:
+                    continue
+
+                # For two roads, move on if the distance is 0 (meaning adjoining roads or intersections)
+                if 'highway' in feature['properties'] and 'highway' in feature_2['properties']:
+                    if shortest_dist == 0.0:
+                        continue
+
                 collision = False
                 if 'highway' in feature_2['properties']:
                     # If the other feature is a path, factor in its road width

### 2020-04-22 osm-importer/way-collisions.py
Skip tunnels everywhere

@@ -188,6 +188,10 @@ while not reached_stable:
                 if shortest_dist == None:
                     continue
 
+                # Skip tunnels
+                if 'tunnel' in feature_2['properties'] and feature_2['properties']['tunnel'] == 'yes':
+                    continue
+
                 # For two roads, move on if the distance is 0 (meaning adjoining roads or intersections)
                 if 'highway' in feature['properties'] and 'highway' in feature_2['properties']:
                     if shortest_dist == 0.0:
@@ -284,6 +288,10 @@ for hwy in way_data.keys():
 
                     shortest_dist, closest_node = geometry_utils.shortest_distance_between_edges_projected(edge_1, edge_2)
                     if shortest_dist == None:
+                        continue
+
+                    # Skip tunnels
+                    if 'tunnel' in feature_2['properties'] and feature_2['properties']['tunnel'] == 'yes':
                         continue
 
                     # For two roads, move on if the distance is 0 (meaning adjoining roads or intersections)

### 2020-04-22 osm-importer/way-collisions.py
Fix spelling errors

@@ -309,16 +309,16 @@ for hwy in way_data.keys():
                         collision = shortest_dist < feature.min_way_width
 
                     if collision:
-                        stat_correced_collision_node_count[hwy] += 1
+                        stat_corrected_collision_node_count[hwy] += 1
 
                         if closest_node == polygon_1[0]:
-                            stat_correced_collision_edge_len[hwy] += geometry_utils.point_distance(polygon_1[0], polygon_1[1]) / 2
+                            stat_corrected_collision_edge_len[hwy] += geometry_utils.point_distance(polygon_1[0], polygon_1[1]) / 2
                         elif closest_node == polygon_1[-1]:
-                            stat_correced_collision_edge_len[hwy] += geometry_utils.point_distance(polygon_1[-2], polygon_1[-1]) / 2
+                            stat_corrected_collision_edge_len[hwy] += geometry_utils.point_distance(polygon_1[-2], polygon_1[-1]) / 2
                         else:
                             index = polygon_1.index(closest_node)
-                            stat_correced_collision_edge_len[hwy] += geometry_utils.point_distance(polygon_1[index-1], polygon_1[index]) / 2
-                            stat_correced_collision_edge_len[hwy] += geometry_utils.point_distance(polygon_1[index], polygon_1[index+1]) / 2
+                            stat_corrected_collision_edge_len[hwy] += geometry_utils.point_distance(polygon_1[index-1], polygon_1[index]) / 2
+                            stat_corrected_collision_edge_len[hwy] += geometry_utils.point_distance(polygon_1[index], polygon_1[index+1]) / 2
 
                         feature_collided = True
 

### 2020-04-22 osm-importer/way-collisions.py
Make sure to skip all tunnels

@@ -270,6 +270,11 @@ for hwy in highway_categories:
 
 for hwy in way_data.keys():
     for feature in way_data[hwy]['features'] :
+
+        # Skip tunnels
+        if 'tunnel' in feature['properties'] and feature['properties']['tunnel'] == 'yes':
+            continue
+
         for feature_2 in hashtable.get_collision_canditates(feature):
             if feature is feature_2:
                 continue

### 2020-04-22 osm-importer/way-collisions.py
Don't count total features, nodes and edge length twice

@@ -256,17 +256,11 @@ print("Statistics collection...")
 stat_corrected_collision_feature_count = {}
 stat_corrected_collision_node_count = {}
 stat_corrected_collision_edge_len = {}
-stat_corrected_total_features_count = {}
-stat_corrected_total_node_count = {}
-stat_corrected_total_edge_len = {}
 
 for hwy in highway_categories:
     stat_corrected_collision_feature_count[hwy] = 0
     stat_corrected_collision_node_count[hwy] = 0
     stat_corrected_collision_edge_len[hwy] = 0.0
-    stat_corrected_total_features_count[hwy] = 0
-    stat_corrected_total_node_count[hwy] = 0
-    stat_corrected_total_edge_len[hwy] = 0.0
 
 for hwy in way_data.keys():
     for feature in way_data[hwy]['features'] :
@@ -286,8 +280,6 @@ for hwy in way_data.keys():
             # Get edges
             for i in range(0, len(polygon_1)-1):
                 edge_1 = [polygon_1[i], polygon_1[i+1]]
-                stat_corrected_total_node_count[hwy] += 1
-                stat_corrected_total_edge_len[hwy] += geometry_utils.point_distance(edge_1[0], edge_1[1])
                 for j in range(0, len(polygon_2)-1):
                     edge_2 = [polygon_2[j], polygon_2[j+1]]
 
@@ -340,11 +332,8 @@ statistics_dict['stat_total_node_count'] = stat_total_node_count
 statistics_dict['stat_collision_edge_len'] = stat_collision_edge_len
 statistics_dict['stat_total_edge_len'] = stat_total_edge_len
 statistics_dict['stat_corrected_collision_feature_count'] = stat_corrected_collision_feature_count
-statistics_dict['stat_corrected_total_feature_count'] = stat_corrected_total_features_count
 statistics_dict['stat_corrected_collision_node_count'] = stat_corrected_collision_node_count
-statistics_dict['stat_corrected_total_node_count'] = stat_corrected_total_node_count
 statistics_dict['stat_corrected_collision_edge_len'] = stat_corrected_collision_edge_len
-statistics_dict['stat_corrected_total_edge_len'] = stat_corrected_total_edge_len
 
 with open('way-collision-statistics', 'wb') as fp:
     pickle.dump(statistics_dict, fp)
@@ -376,11 +365,7 @@ for hwy in highway_categories:
     print("Colliding nodes count: " + str(stat_corrected_collision_node_count[hwy]))
[…]

### 2020-04-23 osm-importer/way-collisions.py
Ensure both statistic collection runs are exactly the same

@@ -70,6 +70,12 @@ for hwy in highway_categories:
     stat_total_node_count[hwy] = 0
     stat_total_edge_len[hwy] = 0.0
 
+# Set way default widths per features
+for hwy in way_data.keys():
+    for feature in way_data[hwy]['features'] :
+        feature.min_way_width = standard_widths[hwy]
+        feature.max_way_width = 2*feature.min_way_width
+
 # Collision detection
 
 progress = 0.0
@@ -83,10 +89,6 @@ for hwy in way_data.keys():
         progress+=1.0
 
         stat_total_features_count[hwy] += 1
-
-        # Set width to default
-        feature.min_way_width = standard_widths[hwy]
-        feature.max_way_width = 2*feature.min_way_width
 
         feature_collided = False
 
@@ -123,9 +125,15 @@ for hwy in way_data.keys():
                         if shortest_dist == 0.0:
                             continue
                     
-                    #print("Shortest dist was not None! It was: " + str(shortest_dist))
                     #plot_utils.plot_polygons([polygon_1, polygon_2])
-                    if shortest_dist < feature.min_way_width:
+                    collision = False
+                    if 'highway' in feature_2['properties']:
+                        # If the other feature is a path, factor in its road width
+                        collision = shortest_dist < feature.min_way_width + feature_2.min_way_width
+                    else:
+                        collision = shortest_dist < feature.min_way_width
+
+                    if collision:
                         stat_collision_node_count[hwy] += 1
 
                         if not feature in colliding_features_tentative:
@@ -143,12 +151,6 @@ for hwy in way_data.keys():
                             stat_collision_edge_len[hwy] += geometry_utils.point_distance(polygon_1[index], polygon_1[index+1]) / 2
 
                         feature_collided = True
-
-                        print("Features collision!")
-                        print(feature)
-                        print(feature_2)
-                        print("min dist: " + str(feature.min_way_width))
-                        print("************")
 
                         """
                         if 'highway' in feature_2['properties']:
@@ -262,15 +264,36 @@ for hwy in highway_categories:
     stat_corrected_collision_node_count[hwy] = 0
     stat_corrected_collision_edge_len[hwy] = 0.0
 
+# Mark all features as unhandled
+for hwy in way_data.keys():
+    for feature in way_data[hwy]['features']:
+        feature.handled = False
+
+for feature in building_data['features']:
+    feature.handled = False
+
+progress = 0.0
 for hwy in way_data.keys():
     for feature in way_data[hwy]['features'] :
-
         # Skip tunnels
[…]

### 2020-04-25 osm-importer/geometry_utils.py
Fix list equals check in shortest_distance_between_edges

@@ -311,9 +311,10 @@ def shortest_distance_between_edges_projected(edge_1, edge_2):
 
     x = line_line_intersection([a1, a2], [b1, b2])
 
-    if a1 == b1 or a1 == b2:
+    # If any of the endpoints are the same, the distance is 0
+    if (a1[0] == b1[0] and a1[1] == b1[1]) or (a1[0] == b2[0] and a1[1] == b2[1]):
         return 0.0, a1
-    if a2 == b1 or a2 == b2:
+    if (a2[0] == b1[0] and a2[1] == b1[1]) or (a2[0] == b2[0] and a2[1] == b2[1]):
         return 0.0, a2
 
     # Case 1, no intersection, lines are parallel

### 2020-04-25 osm-importer/way-collisions.py
Naming fix: feature -> feature_1

@@ -175,8 +175,8 @@ while not reached_stable:
 
     feature_1 = colliding_features_tentative.pop(0)
     polygon_1 = geometry_utils.extract_polygon_from_feature(feature_1)
-    for feature_2 in hashtable.get_collision_canditates(feature):
-        if feature is feature_2:
+    for feature_2 in hashtable.get_collision_canditates(feature_1):
+        if feature_1 is feature_2:
             continue
         polygon_2 = geometry_utils.extract_polygon_from_feature(feature_2)
     
@@ -195,16 +195,16 @@ while not reached_stable:
                     continue
 
                 # For two roads, move on if the distance is 0 (meaning adjoining roads or intersections)
-                if 'highway' in feature['properties'] and 'highway' in feature_2['properties']:
+                if 'highway' in feature_1['properties'] and 'highway' in feature_2['properties']:
                     if shortest_dist == 0.0:
                         continue
 
                 collision = False
                 if 'highway' in feature_2['properties']:
                     # If the other feature is a path, factor in its road width
-                    collision = shortest_dist < feature.min_way_width + feature_2.min_way_width
+                    collision = shortest_dist < feature_1.min_way_width + feature_2.min_way_width
                 else:
-                    collision = shortest_dist < feature.min_way_width
+                    collision = shortest_dist < feature_1.min_way_width
 
                 if collision:
                     # Feature correction by relaxing both edges in the opposite direction
@@ -250,6 +250,7 @@ while not reached_stable:
 
                     # If a collision was fixed, queue this feature again and continue lates
                     colliding_features_tentative.append(feature_1)
+                    colliding_features_tentative.append(feature_2)
                     reached_stable = False
 
 # Collect statistics after correction run

### 2020-04-25 osm-importer/way-collisions.py
Vector dot check fix

@@ -214,7 +214,7 @@ while not reached_stable:
                     perp_2 = geometry_utils.normalize(perp_2)
 
                     translation_vec = None
-                    if np.dot(perp_1, perp_2) > np.dot(perp_1, -perp_2):
+                    if np.dot(perp_1, perp_2) > 0.0:
                         translation_vec = [perp_1[0]+perp_2[0], perp_1[1]+perp_2[1]]
                     else:
                         translation_vec = [perp_1[0]-perp_2[0], perp_1[1]-perp_2[1]]

### 2020-04-25 osm-importer/way-collisions.py
Initialize transformed edges

@@ -222,10 +222,10 @@ while not reached_stable:
                     translation_vec = geometry_utils.normalize(translation_vec)
                     translation_dist = shortest_dist/2
 
-                    edge_1_tr_1 = []
-                    edge_2_tr_1 = []
-                    edge_1_tr_2 = []
-                    edge_2_tr_2 = []
+                    edge_1_tr_1 = [0.0, 0.0]
+                    edge_2_tr_1 = [0.0, 0.0]
+                    edge_1_tr_2 = [0.0, 0.0]
+                    edge_2_tr_2 = [0.0, 0.0]
 
                     edge_1_tr_1[0] = edge_1[0] + translation_vec[0]*translation_dist
                     edge_1_tr_1[1] = edge_1[1] + translation_vec[1]*translation_dist

### 2020-04-25 osm-importer/geometry_utils.py
Move dot function to geometry utils to avoid numpy import

@@ -27,6 +27,9 @@ def epsg3006_to_wgs84(point):
     x_new, y_new = transform(SWEREF, WGS84, y, x)
     return (y_new, x_new)
 
+def dot(vec_1, vec_2):
+    return np.dot(vec_1, vec_2)
+
 def polygon_intersection_area(polygon_1, polygon_2):
     shapely_poly_1 = shapely.geometry.Polygon(polygon_1)
     shapely_poly_2 = shapely.geometry.Polygon(polygon_2)

### 2020-04-25 osm-importer/way-collisions.py

@@ -7,7 +7,6 @@
 import geometry_utils
 import plot_utils
 import pickle
-import numpy as np
 from geojson import Point, Feature, FeatureCollection, load
 from geometry_hashtable import GeometryHashtable
 from json import dump
@@ -214,7 +213,7 @@ while not reached_stable:
                     perp_2 = geometry_utils.normalize(perp_2)
 
                     translation_vec = None
-                    if np.dot(perp_1, perp_2) > 0.0:
+                    if geometry_utils.dot(perp_1, perp_2) > 0.0:
                         translation_vec = [perp_1[0]+perp_2[0], perp_1[1]+perp_2[1]]
                     else:
                         translation_vec = [perp_1[0]-perp_2[0], perp_1[1]-perp_2[1]]

### 2020-04-25 osm-importer/way-collisions.py
Count iterations

@@ -165,108 +165,122 @@ for hwy in way_data.keys():
             stat_collision_feature_count[hwy] += 1
 
 print("Collision correction...")
+iter_count = 1
 
 # Collision correction
 # Collision correction: Store collided features in a tentative map. When a collision is fixed, add both feature_1 and feature_2 to the tentative map
 reached_stable = False
 while not reached_stable:
+    print("Iteration #" + str(iter_count) + ", remaining features: " + str(len(colliding_features_tentative)))
+    iter_count += 1
+
     reached_stable = True
 
-    feature_1 = colliding_features_tentative.pop(0)
-    polygon_1 = geometry_utils.extract_polygon_from_feature(feature_1)
-    for feature_2 in hashtable.get_collision_canditates(feature_1):
-        if feature_1 is feature_2:
-            continue
-        polygon_2 = geometry_utils.extract_polygon_from_feature(feature_2)
-    
-        for i in range(0, len(polygon_1)-1):
-            edge_1 = [polygon_1[i], polygon_1[i+1]]
-            for j in range(0, len(polygon_2)-1):
-                edge_2 = [polygon_2[j], polygon_2[j+1]]
+    feature_count = len(colliding_features_tentative)
+    for f in range(0, feature_count):
+        feature_fixed = False
+        feature_1 = colliding_features_tentative.pop(0)
+        polygon_1 = geometry_utils.extract_polygon_from_feature(feature_1)
+        for feature_2 in hashtable.get_collision_canditates(feature_1):
+            if feature_1 is feature_2:
+                continue
+            polygon_2 = geometry_utils.extract_polygon_from_feature(feature_2)
+        
+            for i in range(0, len(polygon_1)-1):
+                edge_1 = [polygon_1[i], polygon_1[i+1]]
+                for j in range(0, len(polygon_2)-1):
+                    edge_2 = [polygon_2[j], polygon_2[j+1]]
 
-                shortest_dist, closest_node = geometry_utils.shortest_distance_between_edges_projected(edge_1, edge_2)
+                    shortest_dist, closest_node = geometry_utils.shortest_distance_between_edges_projected(edge_1, edge_2)
 
-                if shortest_dist == None:
-                    continue
-
-                # Skip tunnels
-                if 'tunnel' in feature_2['properties'] and feature_2['properties']['tunnel'] == 'yes':
-                    continue
-
[…]

### 2020-04-25 osm-importer/way-collisions.py
Quit loop if the number of rounds with same feature count is < 40

@@ -170,9 +170,12 @@ iter_count = 1
 # Collision correction
 # Collision correction: Store collided features in a tentative map. When a collision is fixed, add both feature_1 and feature_2 to the tentative map
 reached_stable = False
-while not reached_stable:
+rounds_with_same_feature_count = 0
+
+while not reached_stable and rounds_with_same_feature_count < 40:
     print("Iteration #" + str(iter_count) + ", remaining features: " + str(len(colliding_features_tentative)))
     iter_count += 1
+    feature_count = len(colliding_features_tentative)
 
     reached_stable = True
 
@@ -281,6 +284,10 @@ while not reached_stable:
                     break
             if feature_fixed:
                 break
+
+    if len(colliding_features_tentative) == feature_count:
+        rounds_with_same_feature_count += 1
+
 
 # Collect statistics after correction run
 print("Statistics collection...")

### 2020-04-25 osm-importer/way-collisions.py
Include reference to maximum positional error

@@ -28,6 +28,8 @@ standard_widths = {
 way_data = {}
 building_data = []
 
+max_feature_distance = 2.0579
+
 # read files
 for hwy in highway_categories:
     with open('raw_data/' + hwy + '-converted.geojson', 'r') as f:

### 2020-04-25 osm-importer/way-collisions.py
Don't pass by feature by feature, but by edge by edge

@@ -183,7 +183,6 @@ while not reached_stable and rounds_with_same_feature_count < 40:
 
     feature_count = len(colliding_features_tentative)
     for f in range(0, feature_count):
-        feature_fixed = False
         feature_1 = colliding_features_tentative.pop(0)
         polygon_1 = geometry_utils.extract_polygon_from_feature(feature_1)
         for feature_2 in hashtable.get_collision_canditates(feature_1):
@@ -276,16 +275,11 @@ while not reached_stable and rounds_with_same_feature_count < 40:
                             polygon_2[j+1] = edge_2_tr_2[1]
 
                         # If a collision was fixed, queue this feature again and continue lates
-                        colliding_features_tentative.append(feature_1)
-                        #colliding_features_tentative.append(feature_2)
+                        if not feature_1 in colliding_features_tentative:
+                            colliding_features_tentative.append(feature_1)
+                        #if not feature_2 in colliding_features_tentative:
+                            #colliding_features_tentative.append(feature_2)
                         reached_stable = False
-                        feature_fixed = True
-                        break
-
-                if feature_fixed:
-                    break
-            if feature_fixed:
-                break
 
     if len(colliding_features_tentative) == feature_count:
         rounds_with_same_feature_count += 1

### 2020-04-25 osm-importer/way-collisions.py
Use half the minimum road width when calculating collisions

@@ -130,9 +130,9 @@ for hwy in way_data.keys():
                     collision = False
                     if 'highway' in feature_2['properties']:
                         # If the other feature is a path, factor in its road width
-                        collision = shortest_dist < feature.min_way_width + feature_2.min_way_width
+                        collision = shortest_dist < feature.min_way_width/2 + feature_2.min_way_width/2
                     else:
-                        collision = shortest_dist < feature.min_way_width
+                        collision = shortest_dist < feature.min_way_width/2
 
                     if collision:
                         stat_collision_node_count[hwy] += 1
@@ -212,9 +212,9 @@ while not reached_stable and rounds_with_same_feature_count < 40:
                     collision = False
                     if 'highway' in feature_2['properties']:
                         # If the other feature is a path, factor in its road width
-                        collision = shortest_dist < feature_1.min_way_width + feature_2.min_way_width
+                        collision = shortest_dist < feature_1.min_way_width/2 + feature_2.min_way_width/2
                     else:
-                        collision = shortest_dist < feature_1.min_way_width
+                        collision = shortest_dist < feature_1.min_way_width/2
 
                     if collision:
 
@@ -231,7 +231,10 @@ while not reached_stable and rounds_with_same_feature_count < 40:
                             translation_vec = [perp_1[0]-perp_2[0], perp_1[1]-perp_2[1]]
 
                         translation_vec = geometry_utils.normalize(translation_vec)
-                        translation_dist = shortest_dist/2
+                        if 'highway' in feature_2['properties']:
+                            translation_dist = feature_1.min_way_width/2 + feature_2.min_way_width/2 - shortest_dist
+                        else:
+                            translation_dist = feature_1.min_way_width/2 - shortest_dist
 
                         edge_1_tr_1 = [[0,0], [0,0]]
                         edge_2_tr_1 = [[0,0], [0,0]]
@@ -352,9 +355,9 @@ for hwy in way_data.keys():
                     collision = False
                     if 'highway' in feature_2['properties']:
                         # If the other feature is a path, factor in its road width
-                        collision = shortest_dist < feature.min_way_width + feature_2.min_way_width
+                        collision = shortest_dist < feature.min_way_width/2 + feature_2.min_way_width/2
                     else:
-                        collision = shortest_dist < feature.min_way_width
+                        collision = shortest_dist < feature.min_way_width/2
 
                     if collision:
                         stat_corrected_collision_node_count[hwy] += 1

### 2020-04-25 osm-importer/way-collisions.py
Count the number of rounds since the best solution was found. Use this as a loop terminal

@@ -169,15 +169,20 @@ for hwy in way_data.keys():
 print("Collision correction...")
 iter_count = 1
 
+original_polygons = {}
+#TODO: build original polygons
+
 # Collision correction
 # Collision correction: Store collided features in a tentative map. When a collision is fixed, add both feature_1 and feature_2 to the tentative map
 reached_stable = False
-rounds_with_same_feature_count = 0
+best_solution = 0
+rounds_with_same_best_solution = 0
+unfixable_features = []
+# TODO: Add and compare unfixable features. First round only
 
-while not reached_stable and rounds_with_same_feature_count < 40:
+while not reached_stable and rounds_with_same_best_solution < 20:
     print("Iteration #" + str(iter_count) + ", remaining features: " + str(len(colliding_features_tentative)))
     iter_count += 1
-    feature_count = len(colliding_features_tentative)
 
     reached_stable = True
 
@@ -265,7 +270,6 @@ while not reached_stable and rounds_with_same_feature_count < 40:
                         avg_dist_2 = geometry_utils.point_distance(matching_2[0][0], matching_2[0][1]) + geometry_utils.point_distance(matching_2[1][0], matching_2[1][1])
 
                         # vertex swap
-                        # Convert to list to avoid conversion to ndarray
                         if avg_dist_1 > avg_dist_2:
                             polygon_1[i] = edge_1_tr_1[0]
                             polygon_1[i+1] = edge_1_tr_1[1]
@@ -280,13 +284,15 @@ while not reached_stable and rounds_with_same_feature_count < 40:
                         # If a collision was fixed, queue this feature again and continue lates
                         if not feature_1 in colliding_features_tentative:
                             colliding_features_tentative.append(feature_1)
-                        #if not feature_2 in colliding_features_tentative:
-                            #colliding_features_tentative.append(feature_2)
+                        if 'highway' in feature_2['properties']:
+                            if not feature_2 in colliding_features_tentative:
+                                colliding_features_tentative.append(feature_2)
                         reached_stable = False
 
-    if len(colliding_features_tentative) == feature_count:
-        rounds_with_same_feature_count += 1
-
+    rounds_with_same_best_solution += 1
+    if len(colliding_features_tentative) < best_solution:
+        rounds_with_same_best_solution = 0
+        best_solution = len(colliding_features_tentative)
 
 # Collect statistics after correction run
 print("Statistics collection...")

### 2020-04-25 osm-importer/way-collisions.py
Fix start value for best solution

@@ -175,7 +175,7 @@ original_polygons = {}
 # Collision correction
 # Collision correction: Store collided features in a tentative map. When a collision is fixed, add both feature_1 and feature_2 to the tentative map
 reached_stable = False
-best_solution = 0
+best_solution = 9999999999
 rounds_with_same_best_solution = 0
 unfixable_features = []
 # TODO: Add and compare unfixable features. First round only

### 2020-04-25 osm-importer/way-collisions.py
Update todo

@@ -170,7 +170,7 @@ print("Collision correction...")
 iter_count = 1
 
 original_polygons = {}
-#TODO: build original polygons
+#TODO: build original polygons, compare the distance to the original points and only make adjustments if the distance from the target point to the original points is smaller than 2.05
 
 # Collision correction
 # Collision correction: Store collided features in a tentative map. When a collision is fixed, add both feature_1 and feature_2 to the tentative map
@@ -178,7 +178,7 @@ reached_stable = False
 best_solution = 9999999999
 rounds_with_same_best_solution = 0
 unfixable_features = []
-# TODO: Add and compare unfixable features. First round only
+# TODO: Add and compare unfixable features. First round only, to optimize
 
 while not reached_stable and rounds_with_same_best_solution < 20:
     print("Iteration #" + str(iter_count) + ", remaining features: " + str(len(colliding_features_tentative)))

### 2020-04-25 osm-importer/way-collisions.py
Incorporate maximum feature width in collision checks

@@ -125,14 +125,14 @@ for hwy in way_data.keys():
                     if 'highway' in feature['properties'] and 'highway' in feature_2['properties']:
                         if shortest_dist == 0.0:
                             continue
-                    
+                     
                     #plot_utils.plot_polygons([polygon_1, polygon_2])
                     collision = False
                     if 'highway' in feature_2['properties']:
                         # If the other feature is a path, factor in its road width
-                        collision = shortest_dist < feature.min_way_width/2 + feature_2.min_way_width/2
+                        collision = shortest_dist + 2*max_feature_distance < feature.min_way_width/2 + feature_2.min_way_width/2 + 2*max_feature_distance
                     else:
-                        collision = shortest_dist < feature.min_way_width/2
+                        collision = shortest_dist + 2*max_feature_distance < feature.min_way_width/2 + 2*max_feature_distance
 
                     if collision:
                         stat_collision_node_count[hwy] += 1
@@ -217,9 +217,9 @@ while not reached_stable and rounds_with_same_best_solution < 20:
                     collision = False
                     if 'highway' in feature_2['properties']:
                         # If the other feature is a path, factor in its road width
-                        collision = shortest_dist < feature_1.min_way_width/2 + feature_2.min_way_width/2
+                        collision = shortest_dist + 2*max_feature_distance < feature_1.min_way_width/2 + feature_2.min_way_width/2
                     else:
-                        collision = shortest_dist < feature_1.min_way_width/2
+                        collision = shortest_dist + 2*max_feature_distance < feature_1.min_way_width/2
 
                     if collision:
 
@@ -361,9 +361,9 @@ for hwy in way_data.keys():
                     collision = False
                     if 'highway' in feature_2['properties']:
                         # If the other feature is a path, factor in its road width
-                        collision = shortest_dist < feature.min_way_width/2 + feature_2.min_way_width/2
+                        collision = shortest_dist + 2*max_feature_distance < feature.min_way_width/2 + feature_2.min_way_width/2
                     else:
-                        collision = shortest_dist < feature.min_way_width/2
+                        collision = shortest_dist + 2*max_feature_distance < feature.min_way_width/2
 
                     if collision:
                         stat_corrected_collision_node_count[hwy] += 1

### 2020-04-25 report/report.tex
Update todo
Add percentages to building matching statistics table
Update SLU building accuracy
Update data from collision run

@@
 The map is updated continously by Lantmäteriet, in conjoncture with the forming or reforming of property.
-Property limits and building features have a position accuracy requirement of 5 meters.
+Building features have a position accuracy requirement of 2 meters.
 (Lantmäteriet, 2019)
@@
 
-% TUESDAY
-% TODO: ANALYSIS SECTION, figure out what to write
+% TODO: Collision correction run (Enter data in report)
+% TODO: Figure of street, before correction, after correction
 
-% TODO: Collision correction: Fix collisions by relaxing both feature_1 and feature_2
-% TODO: Collision correction: Edit edge distance function to factor in width of the other path
-
-% WEDNESDAY
-% TODO: Collision correction run
-% TODO: Figure of street, before correction, after correction
-% TODO: Table for collision correction data in report
-
-% TODO: Write analysis section
-% TODO: Dig deeper in the SLU dataset, how can one find the method of aquisition for Stockholm buildings, and its precision?
-
-% THURSDAY
-% TODO: Proof read entire report
+% LARGE DATA RUN
 % TODO: Start large data run, way collision
+% TODO: Enter for correction data in report
 
@@
 
+% TODO: Write disposition for implementation of the collision checker
+
 \section{Notes about data processing}
@@
     & 1:0 & 1:1 & 1:n & Total \\
-    Match count & 9812 & 81722 & 11221 & 102755
+    Match count & 9812   & 81722   & 11221   & 102755 \\
+    Percent     & 9.55\% & 79.53\% & 10.92\% & 100\%
 \end{tabular}
@@
     Edge length, total      & 5004997.34 m  & 814349.47 m  & 169129.19 m & 307461.81 m \\
-    Features, colliding     & 186           & 65           & 20          & 29          \\
-    Nodes, colliding        & 458           & 239          & 45          & 59          \\
-    Edge length, colliding  & 14445.66 m    & 8887.32 m    & 1467.54 m   & 1930.88 m   \\
-    Features, percent       & 13.37 \%      & 25.39 \%     & 18.69 \%    & 19.59 \%    \\
-    Nodes, percent          & 0.17 \%       & 0.89 \%      & 0.64 \%     & 0.55 \%     \\
-    Edge length, percent    & 0.29 \%       & 1.09 \%      & 0.87 \%     & 0.63 \%
+    Features, colliding     & 141           & 24           & 20          & 22          \\
+    Nodes, colliding        & 302           & 48           & 38          & 42          \\
+    Edge length, colliding  & 7779.66 m     & 1197.50 m    & 1074.88 m   & 1291.26 m   \\
+    Features, percent       & 10.13 \%      & 9.38 \%      & 18.69 \%    & 14.86 \%    \\
+    Nodes, percent          & 0.11 \%       & 0.18 \%      & 0.54 \%     & 0.39 \%     \\
+    Edge length, percent    & 0.015 \%      & 0.15 \%      & 0.64 \%     & 0.42 \%
 
@@
                             & Footpath      & Residential  & Secondary   & Primary     \\
-    Features, total         & & & & \\
-    Nodes, total            & & & & \\
-    Edge length, total      & m & m & m & m \\
-    Features, colliding     & & & & \\
-    Nodes, colliding        & & & & \\
-    Edge length, colliding  & m & m & m & m \\
-    Features, percent       & \% & \% & \% & \% \\
-    Nodes, percent          & \% & \% & \% & \% \\
-    Edge length, percent    & \% & \% & \% & \%
+    Features, colliding     & 1             & 0            & 2           & 0           \\
+    Nodes, colliding        & 1             & 0            & 2           & 0           \\
+    Edge length, colliding  & 4.20 m        & 0.00 m       & 17.67 m     & 0.00 m      \\
+    Features, percent       & 0.07 \%       & 0.00 \%      & 1.87 \%     & 0.00 \%     \\
+    Nodes, percent          & 0.00 \%       & 0.00 \%      & 0.02 \%     & 0.00 \%     \\
+    Edge length, percent    & 0.00 \%       & 0.00 \%      & 0.01 \%     & 0.00 \%
 
@@
 
-TODO: Interview with Todor
+% TODO: Interview with Todor
 

### 2020-04-25 osm-importer/way-collisions.py
Update todo

@@ -169,16 +169,12 @@ for hwy in way_data.keys():
 print("Collision correction...")
 iter_count = 1
 
-original_polygons = {}
-#TODO: build original polygons, compare the distance to the original points and only make adjustments if the distance from the target point to the original points is smaller than 2.05
-
 # Collision correction
 # Collision correction: Store collided features in a tentative map. When a collision is fixed, add both feature_1 and feature_2 to the tentative map
 reached_stable = False
 best_solution = 9999999999
 rounds_with_same_best_solution = 0
 unfixable_features = []
-# TODO: Add and compare unfixable features. First round only, to optimize
 
 while not reached_stable and rounds_with_same_best_solution < 20:
     print("Iteration #" + str(iter_count) + ", remaining features: " + str(len(colliding_features_tentative)))

### 2020-04-26 osm-importer/geometry_utils.py
Deep copy polygon functions in geometry utils

@@ -30,6 +30,12 @@ def epsg3006_to_wgs84(point):
 def dot(vec_1, vec_2):
     return np.dot(vec_1, vec_2)
 
+def deepcopy_polygon(polygon):
+    polygon_out = []
+    for point in polygon:
+        polygon_out.append[point[0], point[1]]
+    return points_out
+
 def polygon_intersection_area(polygon_1, polygon_2):
     shapely_poly_1 = shapely.geometry.Polygon(polygon_1)
     shapely_poly_2 = shapely.geometry.Polygon(polygon_2)

### 2020-04-26 osm-importer/way-collisions.py
Log colliding features before and after correction run

@@ -59,6 +59,7 @@ stat_collision_edge_len = {}
 stat_total_features_count = {}
 stat_total_node_count = {}
 stat_total_edge_len = {}
+stat_colliding_features_before_correction = {}
 total_total_features_count = 0
 colliding_features_tentative = []
 
@@ -150,6 +151,11 @@ for hwy in way_data.keys():
                             index = polygon_1.index(closest_node)
                             stat_collision_edge_len[hwy] += geometry_utils.point_distance(polygon_1[index-1], polygon_1[index]) / 2
                             stat_collision_edge_len[hwy] += geometry_utils.point_distance(polygon_1[index], polygon_1[index+1]) / 2
+
+                        if not feature in stat_colliding_features_before_correction:
+                            stat_colliding_features_before_correction[feature] = {}
+                        if not feature_2 in stat_colliding_features_before_correction[feature]:
+                            stat_colliding_features_before_correction[feature][feature_2] = [geometry_utils.deepcopy_polygon(polygon_1), geometry_utils.deepcopy_polygon(polygon_2)]
 
                         feature_collided = True
 
@@ -296,6 +302,7 @@ print("Statistics collection...")
 stat_corrected_collision_feature_count = {}
 stat_corrected_collision_node_count = {}
 stat_corrected_collision_edge_len = {}
+stat_colliding_features_after_correction = {}
 
 for hwy in highway_categories:
     stat_corrected_collision_feature_count[hwy] = 0
@@ -373,6 +380,11 @@ for hwy in way_data.keys():
                             stat_corrected_collision_edge_len[hwy] += geometry_utils.point_distance(polygon_1[index-1], polygon_1[index]) / 2
                             stat_corrected_collision_edge_len[hwy] += geometry_utils.point_distance(polygon_1[index], polygon_1[index+1]) / 2
 
+                        if not feature_1 in stat_colliding_features_after_correction:
+                            stat_colliding_features_after_correction[feature_1] = {}
+                        if not feature_2 in stat_colliding_features_after_correction[feature_1]:
+                            stat_colliding_features_after_correction[feature_1][feature_2] = [geometry_utils.deepcopy_polygon(polygon_1), geometry_utils.deepcopy_polygon(polygon_2)]
+
                         feature_collided = True
 
         feature.handled = True
@@ -390,6 +402,8 @@ statistics_dict['stat_total_edge_len'] = stat_total_edge_len
[…]

### 2020-04-26 osm-importer/geometry_utils.py
Fix typo in polygon deep copy

@@ -33,8 +33,8 @@ def dot(vec_1, vec_2):
 def deepcopy_polygon(polygon):
     polygon_out = []
     for point in polygon:
-        polygon_out.append[point[0], point[1]]
-    return points_out
+        polygon_out.append([point[0], point[1]])
+    return polygon_out
 
 def polygon_intersection_area(polygon_1, polygon_2):
     shapely_poly_1 = shapely.geometry.Polygon(polygon_1)

### 2020-04-26 osm-importer/way-collisions.py
Use hashable feature ids as keys, include features in the colliding features statistics

@@ -152,10 +152,14 @@ for hwy in way_data.keys():
                             stat_collision_edge_len[hwy] += geometry_utils.point_distance(polygon_1[index-1], polygon_1[index]) / 2
                             stat_collision_edge_len[hwy] += geometry_utils.point_distance(polygon_1[index], polygon_1[index+1]) / 2
 
-                        if not feature in stat_colliding_features_before_correction:
-                            stat_colliding_features_before_correction[feature] = {}
-                        if not feature_2 in stat_colliding_features_before_correction[feature]:
-                            stat_colliding_features_before_correction[feature][feature_2] = [geometry_utils.deepcopy_polygon(polygon_1), geometry_utils.deepcopy_polygon(polygon_2)]
+                        if not feature['id'] in stat_colliding_features_before_correction:
+                            stat_colliding_features_before_correction[feature['id']] = {}
+                        if not feature_2['id'] in stat_colliding_features_before_correction[feature['id']]:
+                            stat_colliding_features_before_correction[feature['id']][feature_2['id']] = [
+                                    feature,
+                                    feature_2,
+                                    geometry_utils.deepcopy_polygon(polygon_1),
+                                    geometry_utils.deepcopy_polygon(polygon_2)]
 
                         feature_collided = True
 
@@ -380,10 +384,14 @@ for hwy in way_data.keys():
                             stat_corrected_collision_edge_len[hwy] += geometry_utils.point_distance(polygon_1[index-1], polygon_1[index]) / 2
                             stat_corrected_collision_edge_len[hwy] += geometry_utils.point_distance(polygon_1[index], polygon_1[index+1]) / 2
 
-                        if not feature_1 in stat_colliding_features_after_correction:
-                            stat_colliding_features_after_correction[feature_1] = {}
-                        if not feature_2 in stat_colliding_features_after_correction[feature_1]:
-                            stat_colliding_features_after_correction[feature_1][feature_2] = [geometry_utils.deepcopy_polygon(polygon_1), geometry_utils.deepcopy_polygon(polygon_2)]
+                        if not feature_1['id'] in stat_colliding_features_after_correction:
+                            stat_colliding_features_after_correction[feature_1['id']] = {}
+                        if not feature_2 in stat_colliding_features_after_correction[feature_1['id']]:
+                            stat_colliding_features_after_correction[feature_1['id']][feature_2['id']] = [
+                                    feature,
+                                    feature_2,
[…]

### 2020-04-26 osm-importer/way-collisions.py
Fix typo: missing id

@@ -386,7 +386,7 @@ for hwy in way_data.keys():
 
                         if not feature_1['id'] in stat_colliding_features_after_correction:
                             stat_colliding_features_after_correction[feature_1['id']] = {}
-                        if not feature_2 in stat_colliding_features_after_correction[feature_1['id']]:
+                        if not feature_2['id'] in stat_colliding_features_after_correction[feature_1['id']]:
                             stat_colliding_features_after_correction[feature_1['id']][feature_2['id']] = [
                                     feature,
                                     feature_2,

### 2020-04-26 osm-importer/way-collisions.py
Fix typo: stat_total_features_count -> stat_total_feature_count

@@ -56,7 +56,7 @@ widths_by_id = {}
 stat_collision_feature_count = {}
 stat_collision_node_count = {}
 stat_collision_edge_len = {}
-stat_total_features_count = {}
+stat_total_feature_count = {}
 stat_total_node_count = {}
 stat_total_edge_len = {}
 stat_colliding_features_before_correction = {}
@@ -68,7 +68,7 @@ for hwy in highway_categories:
     stat_collision_feature_count[hwy] = 0
     stat_collision_node_count[hwy] = 0
     stat_collision_edge_len[hwy] = 0.0
-    stat_total_features_count[hwy] = 0
+    stat_total_feature_count[hwy] = 0
     stat_total_node_count[hwy] = 0
     stat_total_edge_len[hwy] = 0.0
 
@@ -90,7 +90,7 @@ for hwy in way_data.keys():
         print("Feature collision check, progess: " + str(100*progress/total_total_features_count) + '%')
         progress+=1.0
 
-        stat_total_features_count[hwy] += 1
+        stat_total_feature_count[hwy] += 1
 
         feature_collided = False
 
@@ -328,7 +328,7 @@ for hwy in way_data.keys():
         if 'tunnel' in feature['properties'] and feature['properties']['tunnel'] == 'yes':
             continue
 
-        print("Feature collision check, progess: " + str(100*progress/total_total_features_count) + '%')
+        #print("Feature collision check, progess: " + str(100*progress/total_total_features_count) + '%')
         progress+=1.0
 
         feature_collided = False
@@ -402,7 +402,7 @@ for hwy in way_data.keys():
 statistics_dict = {}
 
 statistics_dict['stat_collision_feature_count'] = stat_collision_feature_count
-statistics_dict['stat_total_feature_count'] = stat_total_features_count
+statistics_dict['stat_total_feature_count'] = stat_total_feature_count
 statistics_dict['stat_collision_node_count'] = stat_collision_node_count
 statistics_dict['stat_total_node_count'] = stat_total_node_count
 statistics_dict['stat_collision_edge_len'] = stat_collision_edge_len
@@ -428,11 +428,11 @@ for hwy in highway_categories:
     print("Colliding nodes count: " + str(stat_collision_node_count[hwy]))
     print("Colliding edges cumulative length: " + str(stat_collision_edge_len[hwy]))
 
-    print("Total features count: " + str(stat_total_features_count[hwy]))
[…]

### 2020-04-26 osm-importer/way-statistics-recap.py
Statistics recap script for ways

@@ -1,0 +1,59 @@
+import pickle
+import statistics
+import plot_utils
+
+highway_categories = [
+        'footpath',
+        'residential',
+        'secondary',
+        'primary'
+        ]
+
+statistics_dict = {}
+
+with open('way-collision-statistics', 'rb') as fp:
+    statistics_dict = pickle.load(fp)
+
+stat_collision_feature_count = statistics_dict['stat_collision_feature_count']
+stat_total_feature_count = statistics_dict['stat_total_feature_count']
+stat_collision_node_count = statistics_dict['stat_collision_node_count']
+stat_total_node_count = statistics_dict['stat_total_node_count']
+stat_collision_edge_len = statistics_dict['stat_collision_edge_len']
+stat_total_edge_len = statistics_dict['stat_total_edge_len']
+stat_corrected_collision_feature_count = statistics_dict['stat_corrected_collision_feature_count']
+stat_corrected_collision_node_count = statistics_dict['stat_corrected_collision_node_count']
+stat_corrected_collision_edge_len = statistics_dict['stat_corrected_collision_edge_len']
+stat_colliding_features_before_correction = statistics_dict['stat_colliding_features_before_correction']
+stat_colliding_features_after_correction = statistics_dict['stat_colliding_features_after_correction']
+
+# Results
+print("Results...")
+for hwy in highway_categories:
+    print("****************")
+    print(hwy)
+    print("****************")
+    print("Pre correction")
+    print("----------------")
+    print("Colliding features count: " + str(stat_collision_feature_count[hwy]))
+    print("Colliding nodes count: " + str(stat_collision_node_count[hwy]))
+    print("Colliding edges cumulative length: " + str(stat_collision_edge_len[hwy]))
+
+    print("Total features count: " + str(stat_total_feature_count[hwy]))
+    print("Total nodes count: " + str(stat_total_node_count[hwy]))
+    print("Total edges cumulative length: " + str(stat_total_edge_len[hwy]))
+
+    print("% features count: " + str(stat_collision_feature_count[hwy]/stat_total_feature_count[hwy]))
+    print("% nodes count: " + str(stat_collision_node_count[hwy]/stat_total_node_count[hwy]))
+    print("% edges cumulative length: " + str(stat_collision_edge_len[hwy]/stat_total_edge_len[hwy]))
+
+    print("----------------")
+    print("Post correction")
+    print("----------------")
[…]

### 2020-04-26 osm-importer/plot_utils.py
Fix paths+polygon plot function to take multiple paths

@@ -5,15 +5,20 @@ import geometry_utils
 from matplotlib.patches import Polygon
 from matplotlib.collections import PatchCollection
 
-def plot_polygons_and_edges(polygons, points):
+def plot_polygons_and_paths(polygons, paths):
     plot_polygons(polygons,  False)
-    edges = []
 
-    for i in range(0, len(points)-1):
-        edges.append([i, i+1])
-    plot_edges_and_points(points, edges)
+    for i in range(0, len(paths)):
+        path = paths[i]
+        edges = []
+        for i in range(0, len(path)-1):
+            edges.append([i, i+1])
 
-def plot_edges_and_points(p, e):
+        plot_edges_and_points(path, edges, False)
+
+    plt.show()
+
+def plot_edges_and_points(p, e, plot=True):
     #points = np.array(p)
     #edges = np.array(e)
 
@@ -29,7 +34,9 @@ def plot_edges_and_points(p, e):
     y = points[:,1].flatten()
     plt.plot(x[edges.T], y[edges.T], 'y-') # Edges
     plt.plot(x, y, 'ro') # Points
-    plt.show()
+
+    if plot:
+        plt.show()
 
 def plot_edges(e):
     points = []

### 2020-04-26 osm-importer/way-collisions.py
Fix distance checks

@@ -131,9 +131,9 @@ for hwy in way_data.keys():
                     collision = False
                     if 'highway' in feature_2['properties']:
                         # If the other feature is a path, factor in its road width
-                        collision = shortest_dist + 2*max_feature_distance < feature.min_way_width/2 + feature_2.min_way_width/2 + 2*max_feature_distance
+                        collision = shortest_dist < feature.min_way_width/2 + feature_2.min_way_width/2
                     else:
-                        collision = shortest_dist + 2*max_feature_distance < feature.min_way_width/2 + 2*max_feature_distance
+                        collision = shortest_dist < feature.min_way_width/2
 
                     if collision:
                         stat_collision_node_count[hwy] += 1
@@ -368,9 +368,9 @@ for hwy in way_data.keys():
                     collision = False
                     if 'highway' in feature_2['properties']:
                         # If the other feature is a path, factor in its road width
-                        collision = shortest_dist + 2*max_feature_distance < feature.min_way_width/2 + feature_2.min_way_width/2
+                        collision = shortest_dist < feature.min_way_width/2 + feature_2.min_way_width/2
                     else:
-                        collision = shortest_dist + 2*max_feature_distance < feature.min_way_width/2
+                        collision = shortest_dist < feature.min_way_width/2
 
                     if collision:
                         stat_corrected_collision_node_count[hwy] += 1

### 2020-04-26 osm-importer/way-statistics-recap.py
Plot and print polygons when recaping statistics

@@ -1,6 +1,7 @@
 import pickle
 import statistics
 import plot_utils
+import geometry_utils
 
 highway_categories = [
         'footpath',
@@ -57,3 +58,33 @@ for hwy in highway_categories:
     print("% nodes count: " + str(stat_corrected_collision_node_count[hwy]/stat_total_node_count[hwy]))
     print("% edges cumulative length: " + str(stat_corrected_collision_edge_len[hwy]/stat_total_edge_len[hwy]))
 
+print(len(stat_colliding_features_before_correction))
+print(len(stat_colliding_features_after_correction))
+
+for id_1 in stat_colliding_features_before_correction.keys():
+    for id_2 in stat_colliding_features_before_correction[id_1].keys():
+        if id_1 in stat_colliding_features_after_correction.keys():
+            if id_2 in stat_colliding_features_after_correction[id_1].keys():
+                entry_before = stat_colliding_features_before_correction[id_1][id_2]
+                entry_after = stat_colliding_features_after_correction[id_1][id_2]
+                feature_1 = entry_before[0]
+                feature_2 = entry_before[1]
+                polygon_1_before = entry_before[2]
+                polygon_2_before = entry_before[3]
+                polygon_1_after = entry_after[2]
+                polygon_2_after = entry_after[3]
+
+                print("*****")
+                print("Polygon 1")
+                print(polygon_1_before)
+                print(polygon_1_after)
+                print("Polygon 2")
+                print(polygon_2_before)
+                print(polygon_2_after)
+
+
+                # Plot colliding features
+                if 'highway' in feature_2['properties']:
+                    plot_utils.plot_edges([polygon_2_before, polygon_2_after, polygon_1_before, polygon_1_after])
+                else:
+                    plot_utils.plot_polygons_and_paths([polygon_2_before, polygon_2_after], [polygon_1_before, polygon_1_after])

### 2020-04-28 osm-importer/geometry_utils.py
Import script fixes for faulty data from large-data-run

@@ -386,7 +386,12 @@ def shortest_distance_between_edges_projected(edge_1, edge_2):
     #points = [a1, a2, b1, b2, x]
     #edges = [[0,1],[2,3]]
     #plot_utils.plot_edges_and_points(points, edges)
-    theta = math.acos(point_len(d)/point_distance(b, x))
+
+    # Protection against math domain errors
+    try:
+        theta = math.acos(point_len(d)/point_distance(b, x))
+    except:
+        return None, None
 
     #print(math.tan(theta)*point_len(k))
 

### 2020-04-28 osm-importer/way-collisions.py

@@ -94,6 +94,14 @@ for hwy in way_data.keys():
         if 'tunnel' in feature_1['properties'] and feature_1['properties']['tunnel'] == 'yes':
             continue
 
+        # Skip roads that are polygons
+        if 'highway' in feature_1['properties'] and 'area' in feature_1['properties'] and feature_1['properties']['area'] == 'yes':
+            continue
+
+        # Skip roads that are also buildings
+        if 'highway' in feature_1['properties'] and 'building' in feature_1['properties']:
+            continue
+
         print("Feature collision check, progess: " + str(100*progress/total_total_features_count) + '%')
         progress+=1.0
 
@@ -108,6 +116,14 @@ for hwy in way_data.keys():
 
             # Skip tunnels
             if 'tunnel' in feature_2['properties'] and feature_2['properties']['tunnel'] == 'yes':
+                continue
+
+            # Skip roads that are polygons
+            if 'highway' in feature_2['properties'] and 'area' in feature_2['properties'] and feature_2['properties']['area'] == 'yes':
+                continue
+
+            # Skip roads that are also buildings
+            if 'highway' in feature_2['properties'] and 'building' in feature_2['properties']:
                 continue
             
             # Skip features that have already been handled
@@ -125,6 +141,10 @@ for hwy in way_data.keys():
                 stat_total_edge_len[hwy] += geometry_utils.point_distance(edge_1[0], edge_1[1])
                 for j in range(0, len(polygon_2)-1):
                     edge_2 = [polygon_2[j], polygon_2[j+1]]
+
+                    # Prevention against double points, happens in the dataset
+                    if edge_1[0] == edge_1[1] or edge_2[0] == edge_2[1]:
+                        continue
 
                     shortest_dist, closest_node = geometry_utils.shortest_distance_between_edges_projected(edge_1, edge_2)
                     if shortest_dist == None:

### 2020-04-28 report/report.tex
Update todo
Update collision statistics table

@@
 
-% TODO: Collision correction run (Enter data in report)
-% TODO: Figure of street, before correction, after correction
-
-% LARGE DATA RUN
-% TODO: Start large data run, way collision
-% TODO: Enter for correction data in report
+% REPORT
+% TODO: Enter correction data in report
+% TODO: Analysis, write about the building-road case
+% TODO: Analysis, write about the road-road case
+% TODO: Plot features before collision, building-road
+% TODO: Plot features before collision, road-road
+% TODO: Proposed fixing algorithm: building-road
+% TODO: Proposed fixing algorithm: road-road
+% TODO: Analysis, mention that footpaths were nor worse than other categories, which are included in many unity packages already
+% TODO: Figure, suggested morphing, building-road case
+% TODO: Figure, suggested morphing, road-road case
+% TODO: Proof read report
 
@@
                             & Footpath      & Residential  & Secondary   & Primary     \\
-    Features, total         & 1391          & 256          & 107         & 148         \\
-    Nodes, total            & 264179        & 26771        & 6997        & 10686       \\
-    Edge length, total      & 5004997.34 m  & 814349.47 m  & 169129.19 m & 307461.81 m \\
-    Features, colliding     & 141           & 24           & 20          & 22          \\
-    Nodes, colliding        & 302           & 48           & 38          & 42          \\
-    Edge length, colliding  & 7779.66 m     & 1197.50 m    & 1074.88 m   & 1291.26 m   \\
-    Features, percent       & 10.13 \%      & 9.38 \%      & 18.69 \%    & 14.86 \%    \\
-    Nodes, percent          & 0.11 \%       & 0.18 \%      & 0.54 \%     & 0.39 \%     \\
-    Edge length, percent    & 0.015 \%      & 0.15 \%      & 0.64 \%     & 0.42 \%
+    Features, total         & 26473         & 10523        & 3044        & 2756        \\
+    Nodes, total            & 6382579       & 6612383      & 968420      & 420810      \\
+    Edge length, total      & 129984.79 km  & 186089.55 km & 29304.57 km & 16162.95 km \\
+    Features, colliding     & 3189          & 1032         & 412         & 456         \\
+    Nodes, colliding        & 7330          & 2362         & 643         & 824         \\
+    Edge length, colliding  & 110.67 km     & 66.18 km     & 9389.30 m   & 20.41 km    \\
+    Features, percent       & 12.05 \%      & 9.81 \%      & 13.53 \%    & 16.55 \%    \\
+    Nodes, percent          & 0.11 \%       & 0.04 \%      & 0.07 \%     & 0.20 \%     \\
+    Edge length, percent    & 0.09 \%       & 0.04 \%      & 0.03 \%     & 0133 \%
 
@@
                             & Footpath      & Residential  & Secondary   & Primary     \\
-    Features, colliding     & 1             & 0            & 2           & 0           \\
-    Nodes, colliding        & 1             & 0            & 2           & 0           \\
-    Edge length, colliding  & 4.20 m        & 0.00 m       & 17.67 m     & 0.00 m      \\
-    Features, percent       & 0.07 \%       & 0.00 \%      & 1.87 \%     & 0.00 \%     \\
-    Nodes, percent          & 0.00 \%       & 0.00 \%      & 0.02 \%     & 0.00 \%     \\
-    Edge length, percent    & 0.00 \%       & 0.00 \%      & 0.01 \%     & 0.00 \%
+    Features, colliding     & 134           & 336          & 246         & 277         \\
+    Nodes, colliding        & 157           & 530          & 306         & 407         \\
+    Edge length, colliding  & 3248.73       & 6475.99 m    & 3840.00 m   & 8039.54 m   \\
+    Features, percent       & 0.51 \%       & 3.19 \%      & 8.08 \%     & 10.51 \%    \\
+    Nodes, percent          & 0.00 \%       & 0.00 \%      & 0.03 \%     & 0.09 \%     \\
+    Edge length, percent    & 0.00 \%       & 0.00 \%      & 0.01 \%     & 0.05 \%
 

### 2020-04-28 report/report.tex
Add figure with feature overlap cases

@@
 
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_feature_overlap_cases}
+    \caption{The 5 identified cases over feature overlap.
+Top left: A building lies in paralel with a road, and the distance between the two is smaller than the minimum width of the road.
+Top right: A road shares a node with a building.
+Middle: A road is intersecting with certain building features.
+Bottom left: Two roads have a minimum distance that is smaller than the sum of the minimum widths of both roads.
+Bottom right: Two roads share a node, and a subsequent node in one road lies within the minimum width of the other road.}
+    \label{fig:space}
+\end{figure}
+
+All these cases will create issues when generating the 3D mesh representation of the city.
+
 (Fan et al, 2014) claimed that the positional error is largely due to the limited resolution in satellite imagery.

### 2020-04-29 report/report.tex
Update todo
Add footpath vs other highway categories conclusion
Include images showin feature overlap fix algorithms

@@
 % REPORT
-% TODO: Enter correction data in report
-% TODO: Analysis, write about the building-road case
-% TODO: Analysis, write about the road-road case
-% TODO: Plot features before collision, building-road
-% TODO: Plot features before collision, road-road
 % TODO: Proposed fixing algorithm: building-road
 % TODO: Proposed fixing algorithm: road-road
-% TODO: Analysis, mention that footpaths were nor worse than other categories, which are included in many unity packages already
-% TODO: Figure, suggested morphing, building-road case
-% TODO: Figure, suggested morphing, road-road case
+% TODO: Mention propagation of error, have to include all the edited features in the next pass
 % TODO: Proof read report
@@
 
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_feature_overlap_fix_1}
+    \caption{}
+    \label{fig:space}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_feature_overlap_fix_2}
+    \caption{}
+    \label{fig:space}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_feature_overlap_fix_3}
+    \caption{}
+    \label{fig:space}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_feature_overlap_fix_4}
+    \caption{}
+    \label{fig:space}
+\end{figure}
+
 All these cases will create issues when generating the 3D mesh representation of the city.
+Conslusion: footpaths were not worse than other categories, which are included in many unity packages already
 

### 2020-04-29 report/report.tex
Analysis conclusions. Restructure collision cases presentation

@@
 % TODO: Proposed fixing algorithm: road-road
-% TODO: Mention propagation of error, have to include all the edited features in the next pass
 % TODO: Proof read report
@@
 
-\subsection{Utility in real world applications}
-
-% TODO: Interview with Todor
-
 \begin{figure}[H]
@@
 
+\subsection{Utility in real world applications}
+
+% TODO: Interview with Todor
+
+\subsection{Feature overlap}
+
+After visual inspection of the cases where features collide, it was determined that all collision occurences can be almost exclusively classified into one of five categories:
+
+\begin{enumerate}
+\item Case 1: A building lies in paralel with a road, and the distance between the two is smaller than the minimum width of the road.
+\item Case 2: A road shares a node with a building.
+\item Case 3: A road is intersecting with certain building features.
+\item Case 4: Two roads have a minimum distance that is smaller than the sum of the minimum widths of both roads. 
+\item Case 5: Two roads share a node, and a subsequent node in one road lies within the minimum width of the other road.
+\end{enumerate}
+
+Each case is exemplified in figure X. %{TODO}
+
 \begin{figure}[H]
@@
     \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_feature_overlap_cases}
-    \caption{The 5 identified cases over feature overlap.
-Top left: A building lies in paralel with a road, and the distance between the two is smaller than the minimum width of the road.
-Top right: A road shares a node with a building.
-Middle: A road is intersecting with certain building features.
-Bottom left: Two roads have a minimum distance that is smaller than the sum of the minimum widths of both roads.
-Bottom right: Two roads share a node, and a subsequent node in one road lies within the minimum width of the other road.}
+    \caption{The 5 identified cases over feature overlap. Top left: Case 1. Top right: Case 2. Middle: Case 3. Bottom left: Case 4. Bottom right: Case 5.}
     \label{fig:space}
 \end{figure}
+
+All these cases will create issues when generating the 3D mesh representation of the city.
+Conslusion: footpaths were not worse than other categories, which are included in many unity packages already
+
+Propagation of error, have to include all the edited features in the next pass
+
+(Fan et al, 2014) claimed that the positional error is largely due to the limited resolution in satellite imagery.
 
@@
 \end{figure}
-
-All these cases will create issues when generating the 3D mesh representation of the city.
-Conslusion: footpaths were not worse than other categories, which are included in many unity packages already
-
-(Fan et al, 2014) claimed that the positional error is largely due to the limited resolution in satellite imagery.
 

### 2020-04-29 report/report.tex
Add descriptions of collision case algorithms

@@
 
+\subsection{Suggested algorithms for correcting feature overlap}
+
+All collision cases can be solved with relatively simply geometric algorithms.
+This section will present suggestions on algorithms that will resolve collisions by identifying individualcolliding nodes and translating them, after which some degree of smoothing (such as linear interpolation) can be applied to preserve visual features.
+Cases 1 and 3-5 will be handled, while 2 will be left since case 2 collisions do not inherently cause feature overlap when extruding a 3D road mesh from the way edges.
+Thus it will me up to the individual developer to decide what to do with Case 2 collisions.
+
 \begin{figure}[H]
@@
     \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_feature_overlap_fix_1}
-    \caption{}
+    \caption{Case 1 collisions can be solved by translating colliding way points away from the building, along the normal of the nearest surface. Smoothing can then be applied as seen fit.}
     \label{fig:space}
@@
     \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_feature_overlap_fix_2}
-    \caption{}
+    \caption{Case 3 collisions can be solved by eliminating the building features that intersect with the way, and then applying the same algorithm as in case 1.}
     \label{fig:space}
@@
     \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_feature_overlap_fix_3}
-    \caption{}
+    \caption{Case 4 collisions can be solved by calculating the average normal over all colliding edges for both ways, and translating colliding points along the normal, away from the other way.
+Smoothing can then be applied as seen fit.}
     \label{fig:space}
@@
     \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_feature_overlap_fix_4}
-    \caption{}
+    \caption{Case 5 collisions can be solved simply by translating the colliding point along the tangent of the closest edge, and then applying smoothing as seen fit.}
     \label{fig:space}

### 2020-05-06 report/report.tex
Update todo
Update abstract and research questions

@@
 Procedural modeling of road networks and even entire cities is of interest to the fields of urban planning and urban architecture.
-In recent years there has been an advent of interactive tools that aid urban planners in placing or generating features of an urban plan, particularly roads and road networks.
-These tools are created with the purpose of aiding urban planners by giving them increased visualization capabilities, and the ability to quickly perform editing both in the large scale and small scale, and quickly view the results.
-Specifically, there have been an advent of tools that exract geodata from OpenStreetMap (OSM) and generate 3D representations of real world locations by placing terrain, buildings, roads and other features.
+In recent years, there have been an advent of tools that extract geodata from OpenStreetMap (OSM) and generate 3D representations of real world locations by placing terrain, buildings, roads and other features.
 However, freely available such tools seldom include smaller roads and pedestrian footpaths such as walkways, sidewalks or trails.
@@
 \\
-It was found that the average positional error in the OSM dataset was about 2m, and approximately 98\% of all buildings in the reference dataset were represented in the OSM dataset.
+It was found that the average positional error in the OSM dataset was about 2 meters, and approximately 98\% of all buildings in the reference dataset were represented in the OSM dataset.
+%TODO: Include a summary of the collision run
 The report concludes with a discussion of the findings of this study, and an assessment of the feasibility of generating city models with pedestrian footpaths in a number of different applications from urban planning tools to videogames.
@@
 
-This research question deals with the procedural generation of road networks from example data sets, and the edge cases that arise when dealing with footpath planning.
-A common problem is that footpaths are in their nature placed densely.
-How can the program identify critical areas where the footpath width is so large that it collides with existing roads?
-Can this be done so efficiently that the example-driven program can still run in real time?
-An additional question that arises is how much road data will be necessary to create believable procedural road networks for new zones in the Stockholm metropolitan area.
-How can this data be sampled and processed efficiently?
+This project has dealt with the feasibility of including footpath data when generating 3D representations of OpenStreetMap data.
 
-The previous work by Nishida et.
-al.
-has shown that it is possible to create believable procedural road maps from examples.
-Their program has created a network of 35 km road in less than a second, so for the application of zoning new city areas my hypothesis is that such a program can be made to run in real time with such minimal waiting time that it can be used in a real application by a city planner.
+The primary contribution is an investigation into the rate of collision between map features in OpenStreetMap.
+In order to create a 3D representation of geodata, individual features will be extruded to give them width, which may create situations where meshes are intersecting.
+The question is: how can a program identify critical areas where the footpath width is so large that it collides with existing roads?
+Can these collisions be easily fixed by making adjustments that are within the positional accuracy of the OpenStreetMap data?
+The features will be categorized according to type, so that the rate of collisions between footpaths and other features can be compared to that of other way types (primary, secondary, residential).
 
-The challenge lies in how to extract the footpath data and tie it together with the existing road map.
-the work by Nishida et.
-al.
-only deals with arterial and secondary roads, and I will be adding a tertiary layer of roads, wich is highly possible, but may increase the data consumption and program running time.
-The same applies for the statistical model that identifies colliding roads, which I will be adding on top of that.
+The secondary contribution will be an assesment of the geometrical precision of OpenStreetMap data around the Stockholm area.
+The positional accuracy will be needed to assess where road collisions occur and whether they can easily be corrected.
 
@@
 The value range will be $(0 < S_{T} < 1)$, where $S_{T}(A,B) = 1$ if the polygons are identical.
-
-% REPORT
-% TODO: Proposed fixing algorithm: building-road
-% TODO: Proposed fixing algorithm: road-road
-% TODO: Proof read report
 

### 2020-05-07 report/report.tex
New title
Update abstract and background. Include results from collision run

@@
 
-\title{Footpath Planning for Example-driven Procedural Road Networks}
+\title{Quality Assesment of OpenStreetMap Footpath Data for 3D City Modelling}
 
@@
 
-This project assesses the feasibility of creating a 3D representation of geodata with the inclusion of pedestrian footpaths.
+This project assesses the feasibility of creating a 3D representation of OpenStreetMap (OSM) geodata with the inclusion of pedestrian footpaths, by focusing on OSM road and building data in the Stockholm metropolitan area.
 \\
-Procedural modeling of road networks and even entire cities is of interest to the fields of urban planning and urban architecture.
-In recent years, there have been an advent of tools that extract geodata from OpenStreetMap (OSM) and generate 3D representations of real world locations by placing terrain, buildings, roads and other features.
-However, freely available such tools seldom include smaller roads and pedestrian footpaths such as walkways, sidewalks or trails.
+An estimate of the positional accuracy of OSM data has been obtained by comparing OSM positional data to a dataset with a known accuracy, which was obtained from the Swedish National Land Survey.
+It was found that the average positional error in the OSM dataset was about 2 meters, and approximately 98\% of all buildings in the reference dataset were represented in the OSM dataset.
 \\
-This project assesses the feasibility of creating such a service, by focusing on OSM road and building data in the Stockholm metropolitan area.
-An estimate of the positional accuracy of OSM data has been obtained by comparing OSM positional data to a dataset with a known accuracy, which was obtained from the Swedish National Land Survey.
-This estimate of the positional error was used to assess the integrity of the OSM Stockholm road network, by locating critical points where supplying the roads and paths with indivdual widths would cause collision with other features.
+This estimate of the positional error was used to assess the integrity of the OSM Stockholm road network, by locating critical points where extending the roads and paths with indivdual widths would cause collision with other features.
 An algortitm then took the road network and made adjustments that were within the found margin of error for the OSM dataset, to eliminate as many collisions as possible.
+It was found that extending all roads with a standard width led to collisions between roughly 12\% of all roads in the dataset.
+The total length of road that intersected with other features was less than 0.1\% of the total road length.
+It was measured that roughly 80\% of these colliding features could be corrected by simple geometrical translation by a distance smaller than the posiional accuracy of the OSM dataset.
+Similarily, the amount of colliding road by length that could be corrected was calculated to around 99\%.
+\\
+The report concludes with a discussion of the findings of this study, and an assessment of the feasibility of generating city models with pedestrian footpaths in a number of different applications from urban planning tools to videogames.
 As secondary contributions, the completeness and shape accuracy of the OSM data set have also been estimated and commented.
-\\
-It was found that the average positional error in the OSM dataset was about 2 meters, and approximately 98\% of all buildings in the reference dataset were represented in the OSM dataset.
-%TODO: Include a summary of the collision run
-The report concludes with a discussion of the findings of this study, and an assessment of the feasibility of generating city models with pedestrian footpaths in a number of different applications from urban planning tools to videogames.
 
@@
 (Stojanivski, 2020)
+
+In recent years, there have been an advent of tools that extract geodata from OpenStreetMap (OSM) and generate 3D representations of real world locations by placing terrain, buildings, roads and other features.
+However, freely available such tools seldom include smaller roads and pedestrian footpaths such as walkways, sidewalks or trails.
 
@@
     Nodes, percent          & 0.11 \%       & 0.04 \%      & 0.07 \%     & 0.20 \%     \\
-    Edge length, percent    & 0.09 \%       & 0.04 \%      & 0.03 \%     & 0133 \%
+    Edge length, percent    & 0.09 \%       & 0.04 \%      & 0.03 \%     & 0.13 \%
 

### 2020-05-07 report/report.tex
Restructure introduction
Rename OpenStreetMap -> OSM

@@
 Similarily, the amount of colliding road by length that could be corrected was calculated to around 99\%.
+It was found that footpaths were much less problematic than other road types (primary, secondary, residential), in terms of feature collisions.
 \\
@@
 
-This report covers... \cite{sampleitem2}
-
 \subsection{Research Question}
 
-This project has dealt with the feasibility of including footpath data when generating 3D representations of OpenStreetMap data.
-
-The primary contribution is an investigation into the rate of collision between map features in OpenStreetMap.
+This project has dealt with the feasibility of including footpath data when generating 3D representations of OpenStreetMap (OSM) data.
+\\
+The primary contribution is an investigation into the rate of collision between map features in OSM.
 In order to create a 3D representation of geodata, individual features will be extruded to give them width, which may create situations where meshes are intersecting.
 The question is: how can a program identify critical areas where the footpath width is so large that it collides with existing roads?
-Can these collisions be easily fixed by making adjustments that are within the positional accuracy of the OpenStreetMap data?
+Can these collisions be easily fixed by making adjustments that are within the positional accuracy of the OSM data?
 The features will be categorized according to type, so that the rate of collisions between footpaths and other features can be compared to that of other way types (primary, secondary, residential).
-
-The secondary contribution will be an assesment of the geometrical precision of OpenStreetMap data around the Stockholm area.
+\\
+The secondary contribution will be an assesment of the geometrical precision of OSM data around the Stockholm area.
 The positional accuracy will be needed to assess where road collisions occur and whether they can easily be corrected.
-
-\subsection{Problem Constraints}
-
-\begin{enumerate}
-\item The road network is assumed to be a 2D simple graph.
-This means that any overlapping roads, such as tunnels and bridges that cross over street-level roads, will be eliminated from the training set.
-Any self-connected nodes will also be eliminated from the training set.
-\item It will be assumed that the city map consists of only arterial (primary) roads, secondary roads and footpaths.
-\item It will be assumed that the primary, secondary and tertiatry roads all have a fixed width. Primary roads have a set width, and secondary roads and footpaths as well.
-\item The program will not handle terrain features and altitude. It will be assume that the road mesh is projected on a flat 2D surface.
-\end{enumerate}
-
-With that said, Nishida et. al. present an excellent solution for generating new road patches that comform to an underlying terrain function.
 
@@
 
-In recent years, there have been an advent of tools that extract geodata from OpenStreetMap (OSM) and generate 3D representations of real world locations by placing terrain, buildings, roads and other features.
+In recent years, there have been an advent of tools that extract geodata from OSM and generate 3D representations of real world locations by placing terrain, buildings, roads and other features.
 However, freely available such tools seldom include smaller roads and pedestrian footpaths such as walkways, sidewalks or trails.
@@
 Since then a number of publications have examined slices of geodata from van Oort's criteria.
-In 2008, (Haklay, 2010) conducted such a study to estimate of the quality of OpenStreetMap positional geodata in London and England.
-Since the OpenStreetMap (OSM) project started in London it was thought that OSM data in the London metropolitan area would be representative of the highest quality data available, and therefore a good indicator for the whole global OSM dataset.
+In 2008, (Haklay, 2010) conducted such a study to estimate of the quality of OSM positional geodata in London and England.
+Since the OSM project started in London it was thought that OSM data in the London metropolitan area would be representative of the highest quality data available, and therefore a good indicator for the whole global OSM dataset.
 The study was conducted by comparing OSM data to Ordnance Survey datasets.
@@
 
-The next notable study to assess the quality of OpenStreetMap data was published by (Kunze, 2012).
+The next notable study to assess the quality of OSM data was published by (Kunze, 2012).
 The study applied different methods to assess the completeness of OSM data in two federal states in Germany, mainly by analysing the area difference between the OSM dataset and an administrative dataset.
@@
 
+\subsection{Problem Constraints}
+
+\begin{enumerate}
+\item The road network is assumed to be a 2D simple graph.
+This means that any overlapping roads, such as tunnels and bridges that cross over street-level roads, will be eliminated from the training set.
+Any self-connected nodes will also be eliminated from the training set.
+\item It will be assumed that the city map consists of only arterial (primary) roads, secondary roads and footpaths.
+\item It will be assumed that the primary, secondary and tertiatry roads all have a fixed width. Primary roads have a set width, and secondary roads and footpaths as well.
+\item The program will not handle terrain features and altitude. It will be assume that the road mesh is projected on a flat 2D surface.
+\end{enumerate}
+
 \section{The geodata used in this project}
 
-For the purpose of this project, the OpenStreetMap road dataset has been used in the presentation of the road network around Stockholm, and in the generative process of new road networks.
-To assess and correct path widths and path node locations, the OpenStreetMap building footprint dataset has also been used.
-To make an assessment on the quality and precision of the OpenStreetMap data, the SLU map from the Swedish National Land Survey was used in comparison with the OpenStreetMap map.
+For the purpose of this project, the OOSMroad dataset has been used in the presentation of the road network around Stockholm, and in the generative process of new road networks.
+To assess and correct path widths and path node locations, the OSM building footprint dataset has also been used.
+To make an assessment on the quality and precision of the OSM data, the SLU map from the Swedish National Land Survey was used in comparison with the OSM map.
 The SLU map conatins only building footprints and property limits.
@@
 Volunteered Geographical Information.
-OpenStreetMap road data and building footprints are commonly obtained by manually tracing features in commercially available satellite images.
+OSM road data and building footprints are commonly obtained by manually tracing features in commercially available satellite images.
 Such images have a limit on their resolution which puts a theoretical limit on the accuracy of map features compared to their real-world equivalents (Haklay 2010).
-Particularily the high resolution imagery from Bing in 2010 led to an increase in building information in OpenStreetMap (Fan et al, 2014).
+Particularily the high resolution imagery from Bing in 2010 led to an increase in building information in OSM (Fan et al, 2014).
 
@@
 The SLU dataset is delivered in the SWEREF 99 TM (Lantmäteriet, 2019).
-SWEREF 99 TM is a projected coordinate system and there is no linear transformation to WGS 48, which OpenStreetMap uses (OSM 19).
+SWEREF 99 TM is a projected coordinate system and there is no linear transformation to WGS 48, which OSM uses (OSM 19).
 The coordinate conversions for this paper were obtained using proj: a Linux commandline application for geospatial coordinate conversion.
@@
 
-(Fan et al., 2014) used the total building area in their reference dataset and the OpenStreetMap dataset to assess the completeness of OpenStreetMap data.
+(Fan et al., 2014) used the total building area in their reference dataset and the OSM dataset to assess the completeness of OSM data.
 The motivation for this is to eliminate semantic differences between both datasets.
@@
 the number of buildings or other objects in each dataset.
-This project will use the building area in the OSM dataset versus that of the SLU dataset to determine the completeness of OpenStreetMap building data in the Stockholm area, which will then be used as an argument for how accurate the building correspondence and point proximity measures can be.
+This project will use the building area in the OSM dataset versus that of the SLU dataset to determine the completeness of OSM building data in the Stockholm area, which will then be used as an argument for how accurate the building correspondence and point proximity measures can be.
 
@@
 (Fan et al., 2014) further discussed how the relative overlap in building area can be used to determine building correspondence between datasets, in cases where there is not much displacement between OSM building footprints data and the reference data set.
-Figure X (TODO) shows a typical comparison between the OpenStreetMap and SLU datasets, and by eyesight it was determined that the displacement of features is small.
+Figure X (TODO) shows a typical comparison between the OSM and SLU datasets, and by eyesight it was determined that the displacement of features is small.
 The relative building overlap between the OSM and SLU datasets will be defined as folows, given the footprints of any building in the OSM set ($foot_{OSM}$) and the SLU set ($foot_{SLU}$):
@@
 Finally, when two building footprints and their vertices have been mapped, we will use the offset of matching vertices as a measure of positional error.
-The average and variance in vertex spread will be calculated and used to make an assessment of the geometric precision of the OpenStreetMap dataset, and this precision will further be used to form an upper boundary on geometrical corrections that can be made to the map once it has been imported into the application.
+The average and variance in vertex spread will be calculated and used to make an assessment of the geometric precision of the OSM dataset, and this precision will further be used to form an upper boundary on geometrical corrections that can be made to the map once it has been imported into the application.
 
@@
 
-Accuracy: we can assume that Stockholm, being one of the more developed cities in one of the most technologically connected country in the world, would be fairly well mapped in OpenStreetMap.
+Accuracy: we can assume that Stockholm, being one of the more developed cities in one of the most technologically connected country in the world, would be fairly well mapped in OSM.
 The positional accuracy has improved by roughly 3.5 meters since the study by (Haklay, 2010), and by roughly 1.5 meters since the study by (Fan et al, 2014), if the SLU reference dataset is to be believed as absolute.
@@
 
-The difference in area cover can partially be explained by small utility buildings that are often not represented in the OpenStreetMap dataset.
+The difference in area cover can partially be explained by small utility buildings that are often not represented in the OSM dataset.
 Particularily the residential areas in downtown, Stockholm, are characteristic in that each city block consists of a street front and an open, communal compound in the middle, inside of which there are usually a number of small utility buildings such as bicycle garages, toolsheds, refuse rooms and such.

### 2020-05-07 report/report.tex
Update research questions section

@@
 This project has dealt with the feasibility of including footpath data when generating 3D representations of OpenStreetMap (OSM) data.
+A key issue in creating 3D representations of geodata is that since individual features will be extruded to give them width, situations may arise where the generated 3D meshes are intersecting.
 \\
 The primary contribution is an investigation into the rate of collision between map features in OSM.
-In order to create a 3D representation of geodata, individual features will be extruded to give them width, which may create situations where meshes are intersecting.
-The question is: how can a program identify critical areas where the footpath width is so large that it collides with existing roads?
-Can these collisions be easily fixed by making adjustments that are within the positional accuracy of the OSM data?
+A program has identified critical areas in the OSM dataset, where the paths are so wide that they collide with existing map features.
+This has been done both using the linear OSM road representation and on a dataset where each road has been assigned a standard width that is in line with Swedish regulations.
+The program has also calculated how many of these colliding features that can be corrected by making adjustments that are within the positional accuracy of the OSM data (See the following paragraph for the secondary contribution).
 The features will be categorized according to type, so that the rate of collisions between footpaths and other features can be compared to that of other way types (primary, secondary, residential).
+Here, features are defined as the geometrical components that comprise the OSM map.
+For the purpose of this report, features will refer mainly to roads and building footprints.
 \\
 The secondary contribution will be an assesment of the geometrical precision of OSM data around the Stockholm area.
+The positional accuracy will be estimated by comparing feature points between the OSM dataset and a reference dataset.
 The positional accuracy will be needed to assess where road collisions occur and whether they can easily be corrected.

### 2020-05-07 report/report.tex
Reorganize implementation
Shorten abstract

@@
 \\
-This estimate of the positional error was used to assess the integrity of the OSM Stockholm road network, by locating critical points where extending the roads and paths with indivdual widths would cause collision with other features.
+This estimate of the positional error was used to assess the integrity of the OSM Stockholm road network, by locating critical points where assigning widths to the roads would cause collision with other features.
 An algortitm then took the road network and made adjustments that were within the found margin of error for the OSM dataset, to eliminate as many collisions as possible.
@@
 The total length of road that intersected with other features was less than 0.1\% of the total road length.
-It was measured that roughly 80\% of these colliding features could be corrected by simple geometrical translation by a distance smaller than the posiional accuracy of the OSM dataset.
-Similarily, the amount of colliding road by length that could be corrected was calculated to around 99\%.
+It was measured that roughly 80\% of these colliding features, and 99\% of the colliding road by length, could be corrected by simple geometrical translation by a distance smaller than the posiional accuracy of the OSM dataset.
 It was found that footpaths were much less problematic than other road types (primary, secondary, residential), in terms of feature collisions.
@@
 
-\section{Dataset quality evaluation}
-
 \subsection{Assesment of dataset completeness}
@@
 
-\section{Implementation}
-
-In this implementation, roads of different types were allowed to merge into one another to form common intersections.
-To facilitate this, it was neccessary to work with a path-by-path representation of the point data.
-When two path share a node with common coordinates, they form an intersection which can be of multiple different types.
-
-% TODO: Write disposition for implementation of the collision checker
-
-\section{Notes about data processing}
+\subsection{Notes about data processing}
 
@@
 \section{Collision checks}
+
+% TODO: Write disposition for implementation of the collision checker
+In this implementation, roads of different types were allowed to merge into one another to form common intersections.
+To facilitate this, it was neccessary to work with a path-by-path representation of the point data.
+When two path share a node with common coordinates, they form an intersection which can be of multiple different types.
 

### 2020-05-07 report/report.tex
Restructure background
Implementation overview section

@@
 A program has identified critical areas in the OSM dataset, where the paths are so wide that they collide with existing map features.
-This has been done both using the linear OSM road representation and on a dataset where each road has been assigned a standard width that is in line with Swedish regulations.
-The program has also calculated how many of these colliding features that can be corrected by making adjustments that are within the positional accuracy of the OSM data (See the following paragraph for the secondary contribution).
-The features will be categorized according to type, so that the rate of collisions between footpaths and other features can be compared to that of other way types (primary, secondary, residential).
 Here, features are defined as the geometrical components that comprise the OSM map.
@@
 The positional accuracy will be needed to assess where road collisions occur and whether they can easily be corrected.
+
+\subsection{Implementation overview}
+
+\subsubsection{Assesment of Geodata Precision}
+
+The geometrical precision study has been carried out in a number of steps that are all detailed below.
+First a feature map was found, which mapped individual features between the OSM dataset and a reference dataset.
+A program then took matching building pairs and extracted matching geometrical points after first simplifying the polygons using the Douglas-Peucker algorithm.
+Once a set of matching points has been found across the whole dataset, the average distance between them was taken as a measure of the positional accuracy.
+\\
+Additionally, a number of auxillary calculations were neccessary for this study.
+The completeness of the OSM dataset were first calculated by comparing how many features are represented in both the OSM dataset and the reference dataset.
+The completeness is needed to estimate the error boundary of the positional accuracy.
+Additionally, the shape accuracy of the OSM dataset has also been calculated and graphed as an extra verification of the similarity between the datasets.
+Shape accuracy is a measure of polygon similarity that was obtained by comparing polygon turning function between matching features, feature by feature.
+See section XX for more details about the shape accuracy measure. %TODO
+
+\subsubsection{Collision study}
+
+For the purpose of the collision study, each road has been assigned a standard width that is in line with Swedish regulations, since in the OSM dataset, roads are represented as simple polylines, lacking width.
+In a 3D representation however, it is crucial that each road object has a spacial width in order to be rendered and seen.
+After the width assignment, program has iterated over all features and calculated how many of them that overlap with any other feature, as well as the total length of the road polylines that is intersecting some other feature.
+The program has then calculated how many of these colliding features that can be corrected by making adjustments that are within the positional accuracy of the OSM data (As obtained by the previous study in this report: Assessment of Geodata Precision).
+The features will be categorized according to type, so that the rate of collisions between footpaths and other features can be compared to that of other road types.
+OSM makes the distinction between primary, secondary, residential and footpath roads, so this classification will also be used in this report.
 
@@
 
-\subsection{Background}
+\section{Background}
+
+\section{Background of 3D City Representation}
 
@@
 
-\subsection{ESAL, KTH}
-The thesis work will be carried out at the Embodied Social Agents Lab (ESAL) at the Department of Computational Science and Technology (CST).
-The lab has before been working on systems for generating procedural urban environments, and are interested in the possibility of generating more detailed maps than have been done before, using example-based methods to preserve the aesthetic look of a greater metropolitan area.
-The lab is particularly interested in the generation of footpaths, since most commercial software that import road maps only import arterial and secondary car roads, and do not include pedestrian walkways, cycle paths and such smaller roads.
-
 \section{Related work}
@@
 The results reveal that the positional accuracy of OSM data at the time was about 4 meters.
+
+\subsection{ESAL, KTH}
+The thesis work will be carried out at the Embodied Social Agents Lab (ESAL) at the Department of Computational Science and Technology (CST).
+The lab has before been working on systems for generating procedural urban environments, and are interested in the possibility of generating more detailed maps than have been done before, using example-based methods to preserve the aesthetic look of a greater metropolitan area.
+The lab is particularly interested in the generation of footpaths, since most commercial software that import road maps only import arterial and secondary car roads, and do not include pedestrian walkways, cycle paths and such smaller roads.
 

### 2020-05-07 report/report.tex
Add a note on why 2*positional accuracy was chosen as max distance for the collision study

@@
 The software uses procedural models to render a city grid based on real-world locations, or generate new user-specified parameters, and allow for varying degrees of manual editing on grid, road, neighbourhood, city block or individual feature level.
-Some of the most prominent examples are CityEngine (made by the startup Procedural, bought by ESRI), Urban Canvas (made by the startup Synthicity, bought by Autodesk, now defunct) and ArcGIS (***).
+Some of the most prominent examples are CityEngine (made by the startup Procedural, bought by ESRI), Urban Canvas (made by the startup Synthicity, bought by Autodesk, now defunct) and ArcGIS (***) %TODO.
 
@@
 % TODO: Write disposition for implementation of the collision checker
+The maximum distance between two features was set to 2 times the positional accuracy, since then there will always be ample room to make changes to any features that would collide with the newly noved nodes, and thus we do not need to worry about feature collisions propagating.
+
 In this implementation, roads of different types were allowed to merge into one another to form common intersections.

### 2020-05-07 report/report.tex
Divide the background into subsections

@@
 
-\section{Background of 3D City Representation}
+\subsection{Background of 3D City Representation}
 
@@
 
-\section{Related work}
+\subsection{Related work}
 Van Oorts criteria for evaluating the quality of geographical information, which at the time of publication received attention from surveyors and cartographers. (van Oort, 2006)

### 2020-05-08 report/report.tex
Update todo
Add table of contents
Paragraph spacing

@@
 \usepackage[ruled,vlined]{algorithm2e}
+
 \setlength\parindent{0pt}
+\setlength{\parskip}{1em}
+\setcounter{tocdepth}{2}
 
@@
 This project assesses the feasibility of creating a 3D representation of OpenStreetMap (OSM) geodata with the inclusion of pedestrian footpaths, by focusing on OSM road and building data in the Stockholm metropolitan area.
-\\
+
 An estimate of the positional accuracy of OSM data has been obtained by comparing OSM positional data to a dataset with a known accuracy, which was obtained from the Swedish National Land Survey.
 It was found that the average positional error in the OSM dataset was about 2 meters, and approximately 98\% of all buildings in the reference dataset were represented in the OSM dataset.
-\\
+
 This estimate of the positional error was used to assess the integrity of the OSM Stockholm road network, by locating critical points where assigning widths to the roads would cause collision with other features.
@@
 It was found that footpaths were much less problematic than other road types (primary, secondary, residential), in terms of feature collisions.
-\\
+
 The report concludes with a discussion of the findings of this study, and an assessment of the feasibility of generating city models with pedestrian footpaths in a number of different applications from urban planning tools to videogames.
 As secondary contributions, the completeness and shape accuracy of the OSM data set have also been estimated and commented.
+
+\newpage
+\tableofcontents
+\newpage
+
+%TODO: Import research questions from google doc
 
@@
 A key issue in creating 3D representations of geodata is that since individual features will be extruded to give them width, situations may arise where the generated 3D meshes are intersecting.
-\\
+
 The primary contribution is an investigation into the rate of collision between map features in OSM.
@@
 For the purpose of this report, features will refer mainly to roads and building footprints.
-\\
+
 The secondary contribution will be an assesment of the geometrical precision of OSM data around the Stockholm area.
@@
 Once a set of matching points has been found across the whole dataset, the average distance between them was taken as a measure of the positional accuracy.
-\\
+
 Additionally, a number of auxillary calculations were neccessary for this study.

### 2020-05-08 report/report.tex
No paragraph skip in table of contents
Update evaluation section

@@
 \newpage
+\setlength{\parskip}{0em}
 \tableofcontents
+\setlength{\parskip}{1em}
 \newpage
-
-%TODO: Import research questions from google doc
 
@@
 
+\subsection{Evaluation}
+
+The primary study of the rate of collision between roads and other features, will serve to obtain the following three metrics:
+
+\begin{enumerate}
+    \item The number of road features that collide with any other feature (road or building).
+    \item The number of road edges that intersect with any other building polygon or road polyline.
+    \item The total length in meters of these edges will also be calculated as a measure of the total length of road.
+\end{enumerate}
+
+The road features will be categorized according to type, so that the rate of collisions between footpaths and other features can be compared to that of other road types.
+OSM mainly makes the distinction between primary, secondary, residential and footpath roads, so this classification will also be used in this report.
+The above features will be calculated for each road type separately by summing over each road feature of a certain type.
+
+Following this, an algorithm will find the exact same metrics with regards to how many road features that can in theory be corrected by simple geometrical translation of their vertices.
+This will also be preseneted in terms of the number of road features, the number of edges and the total length in meters of remaining colliding road.
+
+The secondary study about geodata precision will primarily serve to obtain an estimate of the positional accuracy of the OSM dataset, in meters.
+A number of auxillary metrics will be needed to estimate the domain of error on the positional accuracy, namely the feature completeness and shape accuracy of the OSM dataset.
+The completeness of the OSM dataset will be taken as a the number of features are represented in both datasets, as a perentage of the total numer of features in the reference dataset.
+Shape accuracy is a measure of polygon similarity that was obtained by comparing polygon turning function between matching features, feature by feature.
+See section XX for more details about the shape accuracy measure. %TODO
+
 \subsection{Implementation overview}
@@
 Additionally, the shape accuracy of the OSM dataset has also been calculated and graphed as an extra verification of the similarity between the datasets.
-Shape accuracy is a measure of polygon similarity that was obtained by comparing polygon turning function between matching features, feature by feature.
-See section XX for more details about the shape accuracy measure. %TODO
 
@@
 The program has then calculated how many of these colliding features that can be corrected by making adjustments that are within the positional accuracy of the OSM data (As obtained by the previous study in this report: Assessment of Geodata Precision).
-The features will be categorized according to type, so that the rate of collisions between footpaths and other features can be compared to that of other road types.
-OSM makes the distinction between primary, secondary, residential and footpath roads, so this classification will also be used in this report.
-
-\subsection{Evaluation}
-
-The question will be examined primarily by creating a piece of software that can generate urban maps with primary and secondary roads, as well as footpaths, from example.
-I will experiment with different methods of collecting road data as well as different sample amounts and densities.
-
-The evaluation plan will consist of programmatic evaluations and quantifyable metrics.
-My main idea is to sample metrics about intersections only (number of connecting roads, the angles between them, distance to connecting intersections, curviness of roads, etc).
-If I can sample all that and create a distribution for a few different cities, I have the two following hypothesises:
-  1. There should be selection of features that produce an observable difference in the distributions between cities that look vastly different.
-  2. If I let my program generate a large number of city parcels from a dataset, the resulting feature distribution should look like to the source dataset, and unlike datasets of other cities.
-These distributions could be created for primary, secondary roads and footpaths separately.
 

### 2020-05-08 report/report.tex
Change nodes to edges in result tables

@@
     Features, total         & 26473         & 10523        & 3044        & 2756        \\
-    Nodes, total            & 6382579       & 6612383      & 968420      & 420810      \\
+    Edges, total            & 6382579       & 6612383      & 968420      & 420810      \\
     Edge length, total      & 129984.79 km  & 186089.55 km & 29304.57 km & 16162.95 km \\
     Features, colliding     & 3189          & 1032         & 412         & 456         \\
-    Nodes, colliding        & 7330          & 2362         & 643         & 824         \\
+    Edges, colliding        & 7330          & 2362         & 643         & 824         \\
     Edge length, colliding  & 110.67 km     & 66.18 km     & 9389.30 m   & 20.41 km    \\
     Features, percent       & 12.05 \%      & 9.81 \%      & 13.53 \%    & 16.55 \%    \\
-    Nodes, percent          & 0.11 \%       & 0.04 \%      & 0.07 \%     & 0.20 \%     \\
+    Edges, percent          & 0.11 \%       & 0.04 \%      & 0.07 \%     & 0.20 \%     \\
     Edge length, percent    & 0.09 \%       & 0.04 \%      & 0.03 \%     & 0.13 \%
@@
     Features, colliding     & 134           & 336          & 246         & 277         \\
-    Nodes, colliding        & 157           & 530          & 306         & 407         \\
+    Edges, colliding        & 157           & 530          & 306         & 407         \\
     Edge length, colliding  & 3248.73       & 6475.99 m    & 3840.00 m   & 8039.54 m   \\
     Features, percent       & 0.51 \%       & 3.19 \%      & 8.08 \%     & 10.51 \%    \\
-    Nodes, percent          & 0.00 \%       & 0.00 \%      & 0.03 \%     & 0.09 \%     \\
+    Edges, percent          & 0.00 \%       & 0.00 \%      & 0.03 \%     & 0.09 \%     \\
     Edge length, percent    & 0.00 \%       & 0.00 \%      & 0.01 \%     & 0.05 \%

### 2020-05-08 report/report.tex
Fix section and figure references

@@
 Shape accuracy is a measure of polygon similarity that was obtained by comparing polygon turning function between matching features, feature by feature.
-See section XX for more details about the shape accuracy measure. %TODO
+See section 5.3 for more details about the shape accuracy measure.
 
@@
 (Fan et al., 2014) further discussed how the relative overlap in building area can be used to determine building correspondence between datasets, in cases where there is not much displacement between OSM building footprints data and the reference data set.
-Figure X (TODO) shows a typical comparison between the OSM and SLU datasets, and by eyesight it was determined that the displacement of features is small.
+Figure 1 shows a typical comparison between the OSM and SLU datasets, and by eyesight it was determined that the displacement of features is small.
 The relative building overlap between the OSM and SLU datasets will be defined as folows, given the footprints of any building in the OSM set ($foot_{OSM}$) and the SLU set ($foot_{SLU}$):
@@
 
-Each case is exemplified in figure X. %{TODO}
+Each case is exemplified in figure 8.
 

### 2020-05-08 report/report.tex
Building correspondence section -> Shape accuracy section

@@
 
-\subsection{Building correspondence}
+\subsection{Shape accuracy definition}
 
-Feature mapping
-The building footprint correspondence will be evaluated using a similarity function that assesses the turning funciton and relative area overlap.
-The similarity function between any building in the OSM data set and any building in the SLU datasets will be defined as a function of their respective footprints ($foot_{OSM}, foot_{SLU}$), as follows:
-\begin{center}
-    $S(foot_{OSM}, foot_{SLU}) = S_{T}(foot_{OSM}, foot_{SLU}) + S_{RO}(foot_{OSM}, foot_{SLU})$
-\end{center}
-where $S_{T}$ measures the difference in turning function and $S_{RO}$ measures the relative overlap between the footprints.
-
-\subsection{Turning function definition}
-
+The building footprint correspondence will be evaluated using a similarity function that difference between the turning funcitons of matching building footprints in both datasets.
 The turning function was first defined by Arkin et al.

### 2020-05-09 report/report.tex
Specify what sections are implementation
Fix obvious quirks over the entire report

@@
 The primary contribution is an investigation into the rate of collision between map features in OSM.
-A program has identified critical areas in the OSM dataset, where the paths are so wide that they collide with existing map features.
+A program has taken an OSM dataset of the Stockholm area and assigned each road with standard withds according to swedish regulation.
+It has then identified critical areas in the OSM dataset, where the paths are so wide that they collide with existing map features.
+It has also identified how many of those areas can in theory be corrected by simple translation of the road vertices, without propagation of collision to other features.
 Here, features are defined as the geometrical components that comprise the OSM map.
@@
 Additionally, a number of auxillary calculations were neccessary for this study.
-The completeness of the OSM dataset were first calculated by comparing how many features are represented in both the OSM dataset and the reference dataset.
+The estimated completeness of the OSM dataset (a measure of how many features are represented in both the OSM dataset and the reference dataset) was calculated as the relative difference in area of all building footprints, between the two datasets.
 The completeness is needed to estimate the error boundary of the positional accuracy.
@@
 In a 3D representation however, it is crucial that each road object has a spacial width in order to be rendered and seen.
-After the width assignment, program has iterated over all features and calculated how many of them that overlap with any other feature, as well as the total length of the road polylines that is intersecting some other feature.
+After the width assignment, a program has iterated over all features and calculated how many of them that overlap with any other feature, as well as the total length of the road polylines that is intersecting some other feature.
 The program has then calculated how many of these colliding features that can be corrected by making adjustments that are within the positional accuracy of the OSM data (As obtained by the previous study in this report: Assessment of Geodata Precision).
+
+\subsection{Problem Constraints}
+
+In this study, when working with OSM geodata, the following assumptions have been made about the dataset to reduce the size of the problem domain:
+
+\begin{enumerate}
+    \item The program will not handle terrain features and altitude. It will be assumed that the road mesh can be projected on a flat 2D surface without feature intersection.
+    \item It will be assumed that the city map consists of only arterial (primary) roads, secondary roads and footpaths.
+    \item It will be assumed that the OSM dataset represents all roads as simple polylines, and building footprints as simple connected polygons.
+Any other features (such as squares, which are roads represented as polygons) will be omitted from the dataset.
+    \item The road network is assumed to be a 2D simple graph.
+This means that any overlapping roads, such as tunnels and bridges that cross over street-level roads, will be eliminated from the dataset.
+Any self-connected nodes will also be eliminated from the dataset.
+\end{enumerate}
 
@@
 \subsection{Related work}
-Van Oorts criteria for evaluating the quality of geographical information, which at the time of publication received attention from surveyors and cartographers. (van Oort, 2006)
-
-Since then a number of publications have examined slices of geodata from van Oort's criteria.
+When assessing the completeness of geodata, a very common scale is Van Oorts criteria for evaluating the quality of geographical information, which at the time of publication received attention from surveyors and cartographers.
+(van Oort, 2006) Since then a number of publications have examined slices of geodata from van Oort's criteria.
 In 2008, (Haklay, 2010) conducted such a study to estimate of the quality of OSM positional geodata in London and England.
@@
 The study applied different methods to assess the completeness of OSM data in two federal states in Germany, mainly by analysing the area difference between the OSM dataset and an administrative dataset.
-
 Finally, a notable study in 2013-2014, conducted by researchers from various universities in Norway and Germany (Fan et al, 2014), examined the quality and accuracy in building footprint data of the Munich area, at a time and place where the completeness of the OSM dataset was significantly higher (the results of their study showed that the OSM had captured close to 100\% of all building footprints).
@@
 
-\subsection{Problem Constraints}
-
-\begin{enumerate}
-\item The road network is assumed to be a 2D simple graph.
-This means that any overlapping roads, such as tunnels and bridges that cross over street-level roads, will be eliminated from the training set.
-Any self-connected nodes will also be eliminated from the training set.
-\item It will be assumed that the city map consists of only arterial (primary) roads, secondary roads and footpaths.
-\item It will be assumed that the primary, secondary and tertiatry roads all have a fixed width. Primary roads have a set width, and secondary roads and footpaths as well.
-\item The program will not handle terrain features and altitude. It will be assume that the road mesh is projected on a flat 2D surface.
-\end{enumerate}
-
 \section{The geodata used in this project}
 
-For the purpose of this project, the OOSMroad dataset has been used in the presentation of the road network around Stockholm, and in the generative process of new road networks.
-To assess and correct path widths and path node locations, the OSM building footprint dataset has also been used.
-To make an assessment on the quality and precision of the OSM data, the SLU map from the Swedish National Land Survey was used in comparison with the OSM map.
-The SLU map conatins only building footprints and property limits.
-Both map segments were obtained around the Stockholm greater metropolitan area.
+For the purpose of this project, the OSM road dataset has been sampled around Stockholm metropolitan area.
+To assess and correct path widths and path node locations, the OSM building footprint dataset areound the same area has also been used.
+As a reference dataset, the SLU map from the Swedish National Land Survey was used in comparison with the OSM map.
+%TODO: What does SLU stand for?
+The SLU map conatins only building and property footprints.
 The exact map segments that were extracted were in both cases rectangular boundaries with the following coordinates in WGS84: (N: 59.42, E: 18.15, S: 59.23, W: 17.79).
@@
 
-Volunteered Geographical Information.
+The OSM project is based on Volunteered Geographical Information (VGI).
 OSM road data and building footprints are commonly obtained by manually tracing features in commercially available satellite images.
@@
 Particularily the high resolution imagery from Bing in 2010 led to an increase in building information in OSM (Fan et al, 2014).
-
-To separate sidewalks from detached footpaths, I used the fact that OSM generally provides sidewalks with the same name as the adjacent motorway.
-Thus they can be paired together and measured against each other.
 
@@
 The map is updated continously by Lantmäteriet, in conjoncture with the forming or reforming of property.
-Building features have a position accuracy requirement of 2 meters.
+Building features in the SLU dataset have a position accuracy requirement of 2 meters.
 (Lantmäteriet, 2019)
@@
 
-The SLU dataset is delivered in the SWEREF 99 TM (Lantmäteriet, 2019).
-SWEREF 99 TM is a projected coordinate system and there is no linear transformation to WGS 48, which OSM uses (OSM 19).
+The SLU dataset is delivered in the SWEREF 99 TM coordinate system (Lantmäteriet, 2019).
+SWEREF 99 TM is a projected coordinate system and there is no linear transformation to the WGS 48 system, which OSM uses (OSM 19).
 The coordinate conversions for this paper were obtained using proj: a Linux commandline application for geospatial coordinate conversion.
 
-\subsection{Specific preprocessing for SLU-OSM geodata}
+\subsection{Specific preprocessing for SLU and OSM geodata}
 
-The SLU map is delivered with cropped buildings to exactly matched the query coordinates. This means that area comparsion will not work.
-Therefore it was neccessary to crop both datasets to ensure that all buildings were complete and had matching candidates in the other set.
-(Show a picture)
+Upon delivery of the OSM dataset, all features that intersect with the user-specified domain are included, with their full geometry.
+The SLU map however, is delivered with building footprints cropped to exactly match the query coordinates, meaning that buildings at the edge of the user-specified domain can have cropped geometries.
+Any building area outside of the query domain will be excluded.
+Since the principles behind what features are delivered and how differ between the datasets, it is neccessary to crop both datasets to ensure that all buildings are complete and have matching candidates in the other set.
+Any features that intersect the edge of the rectagular boundary, in both datasets, were therefore excluded from the study.
 
-\section{Evaluation of the quality of geodata}
+\section{Implemenation: Evaluation of the quality of geodata}
 
@@
 
-(Fan et al., 2014) used the total building area in their reference dataset and the OSM dataset to assess the completeness of OSM data.
+(Fan et al., 2014) used the fraction of the total building area in their reference dataset and the OSM dataset to assess the completeness of OSM data.
 The motivation for this is to eliminate semantic differences between both datasets.
@@
 
+\subsection{Correspondence by building area}
+
+(Fan et al., 2014) further discussed how the relative overlap in building area can be used as an estimate of the building correspondence between datasets.
+In this project, relative area overlap was used in order to pair buildings from the two datasets together and create a feature map.
+This can only be done in cases where there is not much displacement between OSM building footprints data and the reference data set.
+Figure 1 under Section 5 shows a typical comparison between the OSM and SLU datasets.
+By eyesight it was determined that the displacement of features is small.
+
+The relative building overlap between the OSM and SLU datasets will be defined as folows, given the footprints of any building in the OSM set ($foot_{OSM}$) and the SLU set ($foot_{SLU}$):
+\begin{center}
+    $S_{RO}(foot_{OSM}, foot_{SLU}) = \frac{A_{Overlap}}{min(A_{foot_{OSM}}, A_{foot_{SLU}}))}$
+\end{center}
+
+The relative overlap may be used to determine building matching relations even when the semantic accuracy is low.
+If a large building is represented by a single footprint in one dataset but by several smaller footprints in the other, (Rutzinger et al, 2009) found that if $S_{RO}(foot_{A}, foot_{B}) < 30\%$ for two buildings $A$ and $B$ from different sets, then $A$ and $B$ are highly likely to be separate, neighbouring buildings and not in fact identical.
+Therefore it will here be assumed that two buildings are matching candidates if their relative area overlap is greater than 30\%.
+If a one-to-many object matching is found, the compound perimeter of the footprints in the many-set will be used when calculating the shape accuracy and finding closest vertices between the polygons.
+
+When calculating the area overlap, it was utilized the fact that multipolygons in geojson are represented with the main, bounding polygon first.
+Any subsequent polygons always represent cutouts or crop-outs in the footprint.
+When calculating the area of a multipolygon building, the area of the first polygon was therefore taken and subtracted with those of the subsequent polygons.
+
 \subsection{Shape accuracy definition}
 
-The building footprint correspondence will be evaluated using a similarity function that difference between the turning funcitons of matching building footprints in both datasets.
+The building footprint correspondence will be evaluated using a similarity function that depends on the difference between the turning funcitons of matching building footprints in both datasets.
 The turning function was first defined by Arkin et al.
@@
 
-\subsection{Correspondence by building area}
-
-(Fan et al., 2014) further discussed how the relative overlap in building area can be used to determine building correspondence between datasets, in cases where there is not much displacement between OSM building footprints data and the reference data set.
-Figure 1 shows a typical comparison between the OSM and SLU datasets, and by eyesight it was determined that the displacement of features is small.
-The relative building overlap between the OSM and SLU datasets will be defined as folows, given the footprints of any building in the OSM set ($foot_{OSM}$) and the SLU set ($foot_{SLU}$):
-
-\begin{center}
-    $S_{RO}(foot_{OSM}, foot_{SLU}) = \frac{A_{Overlap}}{min(A_{foot_{OSM}}, A_{foot_{SLU}}))}$
-\end{center}
-
-For this project the relative overlap will be used both to find possible matching building candidates, and as a term in their similarity function.
-
-The relative overlap may be used to determine building matching relations even when the semantic accuracy is low.
-If a large building is represented by a single footprint in one dataset but by several smaller footprints in the other, (Rutzinger et al, 2009) found that if $S_{RO}(foot_{A}, foot_{B}) < 30\%$ for two buildings $A$ and $B$ from different sets, then $A$ and $B$ are highly likely to be separate, neighbouring buildings and not in fact identical.
-Therefore we will here assume that two buildings are matching candidates if their relative area overlap is greater than 30\%.
-If a one-to-many object matching is found, the compound perimeter of the footprints in the many-set will be used when calculating the turning function and finding closest vertices between the polygons.
-
 \subsection{Closest point and point proximity}
@@
 See Algorithm 1 for a detailed view.
-In either case The Oriented Minimum Bounding Rectangle (MBR) is calculated for the two polygons.
-Finally the MBR for the OSM building footprint is shifted so that its centroid aligns with the centroid of the MBR of the SLU building footprint.
-Any edges in the simplified footprints that coinside with the MBR from the same dataset are extracted, and the corresponding points in the original footprints will be matched with each other.
+In either case The Oriented Minimum Bounding Rectangle (OMBR) is calculated for the two polygons.
+Finally the OMBR for the OSM building footprint is shifted so that its centroid aligns with the centroid of the OMBR of the SLU building footprint.
+Any edges in the simplified footprints that coinside with the OMBR from the same dataset are extracted, and the corresponding points in the original footprints will be matched with each other.
 
@@
 
-\subsection{Notes about data processing}
+\section{Implementation: Collision study}
 
-Once data from the both sets had been extracted, they looked dissimilar.
-(Show picture) Had to go over all OSM buildings that intersected with the query box and remove them, along with any SLU buildings that intersected with them.
-
-When calculating the area overlap we used the fact that multipolygons in geojson are represented with the main, bounding polygon first.
-When calculating the area of a multipolygon building, we simply took the area of the first polygon and subtracted the subsequent ones.
-
-However, when calculating the relative area overlap, this becomes unesseccary since any semantically incorrect subdivision of the polygon will always be inside the bounding polygon.
-
-\section{Collision checks}
-
-% TODO: Write disposition for implementation of the collision checker
-The maximum distance between two features was set to 2 times the positional accuracy, since then there will always be ample room to make changes to any features that would collide with the newly noved nodes, and thus we do not need to worry about feature collisions propagating.
-
+The maximum distance between two features was set to 1 times the positional accuracy, since then there will always be ample room to make changes to any features that would collide with the newly moved nodes, and thus we do not need to worry about feature collisions propagating.
 In this implementation, roads of different types were allowed to merge into one another to form common intersections.
@@
 The default minimum widths were obtained from (Lantmäteriet, 20). The standard minimum width for a two-lane road in Sweden is 6.5 meters.
+Subsequent visual measurements of a few key locations in the Stockholm area revealed the following minimum way widths:
 
-Subsequent visual measurements of a few key locations in the Stockholm area revealed the following minimum way widths:
 \begin{table}[H]
@@
 
-Accuracy: we can assume that Stockholm, being one of the more developed cities in one of the most technologically connected country in the world, would be fairly well mapped in OSM.
+We can assume that Stockholm, being one of the more developed cities in one of the most technologically connected country in the world, would be fairly well mapped in OSM.
 The positional accuracy has improved by roughly 3.5 meters since the study by (Haklay, 2010), and by roughly 1.5 meters since the study by (Fan et al, 2014), if the SLU reference dataset is to be believed as absolute.
@@
 In many cases whole city blocks are represented as single buildings, which means that many less meaningful feature matches can be found.
+% TODO: Table reference
 As seen in table ***, about 80\% of all OSM features were found to have a 1:1 mapping to the SLU dataset, but upon further visual comparison it can be seen that many of the 1:n matches are found in a very specified area, namely the residential areas in downtown, Stockholm.
@@
 
-\subsection{Utility in real world applications}
-
 % TODO: Interview with Todor
+% TODO: Figure showing how turning function is calculated from edges and angles
+% TODO: Figure refs, all figures
+% TODO: Table numberings + references
 
@@
 
-Propagation of error, have to include all the edited features in the next pass
-
 (Fan et al, 2014) claimed that the positional error is largely due to the limited resolution in satellite imagery.
@@
 All collision cases can be solved with relatively simply geometric algorithms.
-This section will present suggestions on algorithms that will resolve collisions by identifying individualcolliding nodes and translating them, after which some degree of smoothing (such as linear interpolation) can be applied to preserve visual features.
+This section will present suggestions on algorithms that will resolve collisions by identifying individual colliding nodes and translating them, after which some degree of smoothing (such as linear interpolation) can be applied to preserve visual features.
 Cases 1 and 3-5 will be handled, while 2 will be left since case 2 collisions do not inherently cause feature overlap when extruding a 3D road mesh from the way edges.
@@
     \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_feature_overlap_fix_3}
-    \caption{Case 4 collisions can be solved by calculating the average normal over all colliding edges for both ways, and translating colliding points along the normal, away from the other way.
+    \caption{Case 4 collisions can be solved by calculating the average normal over all colliding edges for both roads, and translating colliding points along the normal, away from the other road.
 Smoothing can then be applied as seen fit.}
@@
 \end{figure}
+
+\subsection{Utility in real world applications}
 

### 2020-05-09 report/report.tex
Table captions

@@
 \usepackage[ruled,vlined]{algorithm2e}
+\usepackage{caption}
 
@@
 Each arcitect is an actor in a continously developing environment, and affecting that environment in a complex manner through their action.
-
 (Stojanivski, 2020)
@@
 \end{tabular}
+\captionof{table}{Standard road widths used for the collision study}\label{road-widths-table}
 \end{table}
@@
 \end{tabular}
+\captionof{table}{Total number of buildings and their cumulative area}\label{num-of-buildings-table}
 \end{table}
@@
 \end{tabular}
+\captionof{table}{Total number of buildings and cumulative area as percentages}\label{num-of-buildings-percent-table}
 \end{table}
@@
 \end{tabular}
+\captionof{table}{The estimated positional accuracy of the OSM dataset: average, max, min and standard deviation}\label{positional-accuracy-table}
 \end{table}
@@
 \end{tabular}
+\captionof{table}{The number and percentages of matching buildings between both datasets, ordered by type of correlation (1:0, 1:1 or 1:n)}\label{building-matching-table}
 \end{table}
@@
 
-\subsection{Way collision statistics}
+\subsection{Road collision statistics}
 
@@
 \end{tabular}
+\captionof{table}{The first result table of the collision study.
+This table shows the amount of features and edges, as well as the cumulative edge length, in total and which are intersecting with any other feature.}\label{road-collision-table}
 \end{table}
@@
 \end{tabular}
+\captionof{table}{The second result table of the collision study.
+This table shows the amount of features and edges, as well as the cumulative edge length, which can be corrected by simple geometric translation.}\label{road-collision-corrected-table}
 \end{table}

### 2020-05-09 report/report.tex
Table indents
Update table references

--- 2020-05-09 report/report.tex
@@
     \end{tabular}
+    \captionof{table}{Standard road widths used for the collision study}
+    \label{table:road-widths}
 \end{table}
@@
     \end{tabular}
+    \captionof{table}{Total number of buildings and their cumulative area}
+    \label{table:num-of-buildings}
 \end{table}
@@
     \end{tabular}
+    \captionof{table}{Total number of buildings and cumulative area as percentages}
+    \label{table:num-of-buildings-percent}
 \end{table}
@@
     \end{tabular}
+    \captionof{table}{The estimated positional accuracy of the OSM dataset: average, max, min and standard deviation}
+    \label{table:positional-accuracy}
 \end{table}
@@
     \end{tabular}
+    \captionof{table}{The number and percentages of matching buildings between both datasets, ordered by type of correlation (1:0, 1:1 or 1:n)}
+    \label{table:building-matching}
 \end{table}
@@
     \end{tabular}
+    \captionof{table}{The first result table of the collision study.
+This table shows the amount of features and edges, as well as the cumulative edge length, in total and which are intersecting with any other feature.}
+    \label{table:road-collision}
 \end{table}
@@
     \end{tabular}
+    \captionof{table}{The second result table of the collision study.
+This table shows the amount of features and edges, as well as the cumulative edge length, which can be corrected by simple geometric translation.}
+    \label{table:road-collision-corrected}
 \end{table}

### 2020-05-09 report/report.tex
Update figure references
In-text figure references

@@
 The comparison of the OSM and SLU maps will focus on these accuracy classes to determine the error tolerance when manipulating features in the 3D map presentation and the road network generating phase.
+Figure \ref{fig:osm-slu-map} shows two segments of the Stockholm map where the OSM and SLU datasets show lower semantic and positional accuracy respectively.
 
@@
 The bottom picture shows 4 smaller buildings whose the positional accuracy is low, due to the buildings being incorrectly scaled, skewed and rotated.}
-    \label{fig:space}
+    \label{fig:osm-slu-map}
 \end{figure}
@@
 A very common example of a semantic error when comparing geographic data is that a large building in one dataset may be segmented into several smaller ones in the other, that is to say a building in one set is represented as an aggreation of mutliple buildings in the other set.
-This makes one-to-one object mapping impossible.
-Thus using building area as an estimator for completeness solves this issue much better than i.e.
-the number of buildings or other objects in each dataset.
+This makes a complete one-to-one object mapping impossible.
+Rather, buildings have to be identified as 1:1 (one to one), 1:n (one to many) or 1:0 (one to none) cases, depending on if a building is fully represented in the other dataset, represented as an aggregation of smaller plots, or not found at all.
+Figure \ref{fig:building-match-types} shows examples of two buildings that were found in the OSM and SLU datasets as 1:1 and 1:n matches respectively.
+
+Since it is impossible to find a one-to-one feature map for building footprints, using building area as an estimator for completeness is much better than i.e.
+using the number of buildings or other objects in each dataset.
 This project will use the building area in the OSM dataset versus that of the SLU dataset to determine the completeness of OSM building data in the Stockholm area, which will then be used as an argument for how accurate the building correspondence and point proximity measures can be.
@@
     \caption{Examples of a 1:1 building match (left), where the sought-after building was found both in the OSM (red) and SLU (blue) datasets, and of a 1:n building match (right), where the sought-after building was represented as one polygon in the OSM (red) dataset, but divided into many smaller lots in the SLU (multicolored) dataset.}
-    \label{fig:space}
+    \label{fig:building-match-types}
 \end{figure}
@@
 This can only be done in cases where there is not much displacement between OSM building footprints data and the reference data set.
-Figure 1 under Section 5 shows a typical comparison between the OSM and SLU datasets.
+Figure \ref{fig:osm-slu-map} under Section 5 shows a typical comparison between the OSM and SLU datasets.
 By eyesight it was determined that the displacement of features is small.
@@
 The building footprint correspondence will be evaluated using a similarity function that depends on the difference between the turning funcitons of matching building footprints in both datasets.
-The turning function was first defined by Arkin et al.
-(1991), as a method for measuring the similarity of two polygons.
+The turning function was first defined by Arkin et al. (1991), as a method for measuring the similarity of two polygons.
 The Turning function $T_c(l)$ measures the cumulative angle of the polygon's counter-clockwise tangent, as a function of the cumulative normalized length l.
 This project uses the turning function as it is defined by (Fan et al., 2014).
-For a polygon with vertices ${v_1 ...
-v_n}$ and line segments ${e_1 ...
-e_n}$ It is defined as follows.
+See figure \ref{fig:turning-function} for a side-by-side comparison of a polygon and its turning function.
+For a polygon with vertices ${v_1 ... v_n}$ and line segments ${e_1 ... e_n}$ It is defined as follows.
 Fix a starting vertex $v_1$.
@@
 The value range will be $(0 < S_{T} < 1)$, where $S_{T}(A,B) = 1$ if the polygons are identical. 
+See figure \ref{fig:turning-function-compare} for a side-by-side comparison of two similar polygons and their turning functions.
 
@@
     \caption{An example of the turning function of a polygon.}
-    \label{fig:space}
+    \label{fig:turning-function}
 \end{figure}
@@
     \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_turn_function_diff_filled}
-    \caption{An example of the turning functions of two corresponding buildings, with the area between them highlighted.}
-    \label{fig:space}
+    \caption{An example of the turning functions of two corresponding buildings from the OSM and SLU datasets, with the area between them highlighted.}
+    \label{fig:turning-function-compare}
 \end{figure}
@@
 Footprints from different datasets may be formed at a different level of detail.
-Problems that arise are e.g.
-that the vertex counts are dissimilar, or that vertex clusters may be found at different parts of the polygon in the two datasets.
+Problems that arise are e.g. that the vertex counts are dissimilar, or that vertex clusters may be found at different parts of the polygon in the two datasets.
 To avoid this effect, key points are extracted using the Douglas-Peucker algorithm (Douglas and Peucker, 1973), to create a simplifed footprint with less points that still retain information about the rough features of the detailed footprint.
@@
     \caption{The average positional accuracies per building as a frequency diagram.}
-    \label{fig:space}
+    \label{fig:bar-plot-positional-accuracy}
 \end{figure}
@@
     \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_turning_function_plot}
-    \caption{The turning function distances between 1:1 building pairs as a frequency diagram.}
-    \label{fig:space}
+    \caption{The shape similarities between 1:1 building pairs as a frequency diagram.}
+    \label{fig:bar-plot-similarity}
 \end{figure}
@@
 In many cases whole city blocks are represented as single buildings, which means that many less meaningful feature matches can be found.
-% TODO: Table reference
-As seen in table ***, about 80\% of all OSM features were found to have a 1:1 mapping to the SLU dataset, but upon further visual comparison it can be seen that many of the 1:n matches are found in a very specified area, namely the residential areas in downtown, Stockholm.
+As seen in table \ref{table:building-matching}, about 80\% of all OSM features were found to have a 1:1 mapping to the SLU dataset, but upon further visual comparison it can be seen that many of the 1:n matches are found in a very specified area, namely the residential areas in downtown, Stockholm.
 This is problematic since this typical area with the same level of accessibility was likely measured with the same measuring technique, and the unsystematic positional error will likely be similar in these areas.
@@
     \caption{A map segment which shows an area with many small utility buildings that are represented in the SLU dataset but not in the OSM dataset.}
-    \label{fig:space}
+    \label{fig:osm-slu-map-utility-buildings}
 \end{figure}
@@
 
-Each case is exemplified in figure 8.
+Each case is exemplified in figure \ref{fig:collision-cases}.
 
@@
     \caption{The 5 identified cases over feature overlap. Top left: Case 1. Top right: Case 2. Middle: Case 3. Bottom left: Case 4. Bottom right: Case 5.}
-    \label{fig:space}
+    \label{fig:collision-cases}
 \end{figure}
@@
     \caption{Case 1 collisions can be solved by translating colliding way points away from the building, along the normal of the nearest surface. Smoothing can then be applied as seen fit.}
-    \label{fig:space}
+    \label{fig:collision-case-1}
 \end{figure}
@@
     \caption{Case 3 collisions can be solved by eliminating the building features that intersect with the way, and then applying the same algorithm as in case 1.}
-    \label{fig:space}
+    \label{fig:collision-case-3}
 \end{figure}
@@
 Smoothing can then be applied as seen fit.}
-    \label{fig:space}
+    \label{fig:collision-case-4}
 \end{figure}
@@
     \caption{Case 5 collisions can be solved simply by translating the colliding point along the tangent of the closest edge, and then applying smoothing as seen fit.}
-    \label{fig:space}
+    \label{fig:collision-case-5}
 \end{figure}

### 2020-05-09 report/report.tex
Table in-text references

@@
 
-The default minimum widths were obtained from (Lantmäteriet, 20). The standard minimum width for a two-lane road in Sweden is 6.5 meters.
+The default minimum widths were obtained from (Lantmäteriet, 20). The standard recommended minimum width for a two-lane road in Sweden is 6.5 meters.
 Subsequent visual measurements of a few key locations in the Stockholm area revealed the following minimum way widths:
+Table \ref{table:road-widths} shows the minimum and maximum road widths that were found during these measurements, and that will be used in the collision study.
 
@@
 % TODO: Figure showing how turning function is calculated from edges and angles
-% TODO: Figure refs, all figures
-% TODO: Table numberings + references
 

### 2020-05-09 report/report.tex
Add note on the origin of SLU

@@
 To assess and correct path widths and path node locations, the OSM building footprint dataset areound the same area has also been used.
-As a reference dataset, the SLU map from the Swedish National Land Survey was used in comparison with the OSM map.
-%TODO: What does SLU stand for?
+As a reference dataset, the property map from the Swedish National Land Survey was used in comparison with the OSM map.
+The reference map is provided by the Swedish University of Agricultural Sciences (SLU), and will be referred to in this report as the SLU map or the SLU dataset.
 The SLU map conatins only building and property footprints.

### 2020-05-09 report/report.tex
Add note about ArcGIS and Esri

@@
 The software uses procedural models to render a city grid based on real-world locations, or generate new user-specified parameters, and allow for varying degrees of manual editing on grid, road, neighbourhood, city block or individual feature level.
-Some of the most prominent examples are CityEngine (made by the startup Procedural, bought by ESRI), Urban Canvas (made by the startup Synthicity, bought by Autodesk, now defunct) and ArcGIS (***) %TODO.
+Some of the most prominent examples are ArcGIS Pro (made by Esri), CityEngine (made by the startup Procedural, bought by ESRI) and Urban Canvas (made by the startup Synthicity, bought by Autodesk, now defunct).
 

### 2020-05-10 report/report.tex
Update todo
Figure ref 7
Fix intro to section 8.1

@@
 
-We can assume that Stockholm, being one of the more developed cities in one of the most technologically connected country in the world, would be fairly well mapped in OSM.
-The positional accuracy has improved by roughly 3.5 meters since the study by (Haklay, 2010), and by roughly 1.5 meters since the study by (Fan et al, 2014), if the SLU reference dataset is to be believed as absolute.
+One could reasonably assume that Stockholm, being one of the more developed cities in one of the most technologically connected country in the world, would be fairly well mapped in OSM.
+Let us take the positional accuracy of the Stockholm area, as obtained by this study, as reprentative of that of major cities in developed countries in the OSM dataset.
+Then the positional accuracy has improved by roughly 3.5 meters since the study by (Haklay, 2010), and by roughly 1.5 meters since the study by (Fan et al, 2014).
+This of course assumes that the SLU reference dataset is taken as absolute.
 
 The completeness analysis by building area shows that the OSM Stockholm dataset has high completeness in terms of both building area and counted features, in the same level as the study of Munich by (Fan et al, 2014).
-
 The difference in area cover can partially be explained by small utility buildings that are often not represented in the OSM dataset.
 Particularily the residential areas in downtown, Stockholm, are characteristic in that each city block consists of a street front and an open, communal compound in the middle, inside of which there are usually a number of small utility buildings such as bicycle garages, toolsheds, refuse rooms and such.
+See figure \ref{fig:osm-slu-map-utility-buildings} for an example of an area in stockholm with such characteristics.
 These buildings are likely hard to make out and trace on publicly available satellite image data.
@@
 
-% TODO: Interview with Todor
+% TODO: Reformat 8.3 to show the suggested algorithms in a list, with figure references
+% TODO: Complementary text in the results section
+% TODO: Proper citations (Fix links)
+
 % TODO: Figure showing how turning function is calculated from edges and angles
+% TODO: Figure detailing the steps in the point-matching algorithm
+% TODO: Algorithm showing the edge projected distance algorithm for the collision study
+% TODO: Expand on analysis about Feature overlap
+
+% After response from Todor:
+% TODO: Write section: Utility in real world applications
+% TODO: Spell check
+% TODO: Go over report one final time
 

### 2020-05-10 report/report.tex
Update analysis with conclusion about the percentages of roads and edges

@@
 
-\subsection{Feature overlap}
+\subsection{Collision study}
+
+The biggest reason is thought to be that footpaths in their nature are narrow, so narrow that they can be displaced by more than half of their full width before the positional accuracy is reached.
+
+Conslusion: footpaths were not worse than other categories, which are included in many unity packages already
+(Fan et al, 2014) claimed that the positional error is largely due to the limited resolution in satellite imagery.
+
+***\% of feautures collide and ***\% of edges.
+This means that for each colliding road, on average ***\% of its edges will be intersecting some other feature, thus bearing responsibility for the collision.
+In total there are ***\% feautures ***\% edges, meaning that each feature on average is contructed out of *** edges.
+Thus the percentage of edges colliding per road is on average ***\%.
+The conclusion is that most often, a small part of a road is colliding, and will need correction.
+Visual inspection of the cases where roads collide with other features has supported this.
 
@@
 All these cases will create issues when generating the 3D mesh representation of the city.
-Conslusion: footpaths were not worse than other categories, which are included in many unity packages already
-
-(Fan et al, 2014) claimed that the positional error is largely due to the limited resolution in satellite imagery.
 

### 2020-05-10 report/report.tex
Update table and figure references in results

@@
 
-\section{Results, Metrics}
+\section{Results and metrics}
+
+\subsection{Positional accuracy statistics}
+
+For the secondary study, the average positional accuracy per building was calculated to roughly 2.06 meters, and this will be taken as an estaimate of the positional accuracy of the OSM dataset.
+Table \ref{table:positional-accuracy} shows the estimate in positional accuracy that was obtained from the study.
+Figure \ref{fig:bar-plot-positional-accuracy} shows the positional accuracies measured per building as a frequency diagram.
+
+\begin{table}[H]
+    \begin{tabular}{lllll}
+        $e_{average}$ & $e_{max}$ & $e_{min}$ & $e_{standard deviation}$ \\
+        2.0579... & 9.9878... & 0.0 & 1.2464...
+    \end{tabular}
+    \captionof{table}{The estimated positional accuracy of the OSM dataset: average, max, min and standard deviation}
+    \label{table:positional-accuracy}
+\end{table}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_pos_error_plot}
+    \caption{The average positional accuracies per building as a frequency diagram.}
+    \label{fig:bar-plot-positional-accuracy}
+\end{figure}
+
+\subsection{Road collision study}
+
+For the primary collision study.
+
+Table \ref{table:road-collision} shows the statistics of the collision study, and the number of roads, number of edges and total edge length that showed intersections with other features before road extrusion, by road category.
+For footpaths in particular, it was shown that 12.05\% of all roads collide with some other feature after extrusion to the minimum width.
+The same was shown to be true for 0.11\% of the road polyline edges and 0.09\% of the total road length.
+Table \ref{table:road-collision-corrected} shows the remaining roads, edges and total edges length whose intersections with other features could not be resolved by geometrical translation of a distance smaller than the positional accuracy.
+
+\begin{table}[H]
+    \begin{tabular}{lllll}
+                                & Footpath      & Residential  & Secondary   & Primary     \\
+        Features, total         & 26473         & 10523        & 3044        & 2756        \\
+        Edges, total            & 6382579       & 6612383      & 968420      & 420810      \\
+        Edge length, total      & 129984.79 km  & 186089.55 km & 29304.57 km & 16162.95 km \\
+        Features, colliding     & 3189          & 1032         & 412         & 456         \\
+        Edges, colliding        & 7330          & 2362         & 643         & 824         \\
+        Edge length, colliding  & 110.67 km     & 66.18 km     & 9389.30 m   & 20.41 km    \\
+        Features, percent       & 12.05 \%      & 9.81 \%      & 13.53 \%    & 16.55 \%    \\
+        Edges, percent          & 0.11 \%       & 0.04 \%      & 0.07 \%     & 0.20 \%     \\
+        Edge length, percent    & 0.09 \%       & 0.04 \%      & 0.03 \%     & 0.13 \%
+
+    \end{tabular}
+    \captionof{table}{The first result table of the collision study.
+This table shows the amount of features and edges, as well as the cumulative edge length, in total and which are intersecting with any other feature.}
+    \label{table:road-collision}
+\end{table}
+
+\begin{table}[H]
+    \begin{tabular}{lllll}
+                                & Footpath      & Residential  & Secondary   & Primary     \\
+        Features, colliding     & 134           & 336          & 246         & 277         \\
+        Edges, colliding        & 157           & 530          & 306         & 407         \\
+        Edge length, colliding  & 3248.73       & 6475.99 m    & 3840.00 m   & 8039.54 m   \\
+        Features, percent       & 0.51 \%       & 3.19 \%      & 8.08 \%     & 10.51 \%    \\
+        Edges, percent          & 0.00 \%       & 0.00 \%      & 0.03 \%     & 0.09 \%     \\
+        Edge length, percent    & 0.00 \%       & 0.00 \%      & 0.01 \%     & 0.05 \%
+
+    \end{tabular}
+    \captionof{table}{The second result table of the collision study.
+This table shows the amount of features and edges, as well as the cumulative edge length, which can be corrected by simple geometric translation.}
+    \label{table:road-collision-corrected}
+\end{table}
 
 \subsection{Completeness by building area}
+
+Table \ref{table:num-of-buildings} shows the building coverage of the OSM and SLU datasets in terms of the number of buildings and their total area.
+Table \ref{table:num-of-buildings-percent} shows the building coverage of both datasets as perentages.
+The area cover of the OSM dataset in relation to the SLU dataset was estimated to 98.18\%. This will be taken as an estimate of the completeness of the OSM dataset.
 
@@
 
-The area cover of the OSM dataset in relation to the SLU dataset was estimated to 98.18\%.
+\subsection{Building matching statistics}
 
-\subsection{Positional accuracy statistics}
-
-\begin{table}[H]
-    \begin{tabular}{lllll}
-        $e_{average}$ & $e_{max}$ & $e_{min}$ & $e_{standard deviation}$ \\
-        2.0579... & 9.9878... & 0.0 & 1.2464...
-    \end{tabular}
-    \captionof{table}{The estimated positional accuracy of the OSM dataset: average, max, min and standard deviation}
-    \label{table:positional-accuracy}
-\end{table}
-
-Building matching statistics:
+Table \ref{table:building-matching} shows how many of the OSM to SLU building matches were found in each category of 1:1, 1:n and 1:0 matches.
 
@@
 
-% Bar plot positional accuracy
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_pos_error_plot}
-    \caption{The average positional accuracies per building as a frequency diagram.}
-    \label{fig:bar-plot-positional-accuracy}
-\end{figure}
-
 \subsection{Shape accuracy statistics}
 
-% Bar plot turning functions
+Figure \ref{fig:bar-plot-similarity} shows the shape similarity measured per building as a frequency diagram.
 
@@
 \end{figure}
-
-\subsection{Road collision statistics}
-
-\begin{table}[H]
-    \begin{tabular}{lllll}
-                                & Footpath      & Residential  & Secondary   & Primary     \\
-        Features, total         & 26473         & 10523        & 3044        & 2756        \\
-        Edges, total            & 6382579       & 6612383      & 968420      & 420810      \\
-        Edge length, total      & 129984.79 km  & 186089.55 km & 29304.57 km & 16162.95 km \\
-        Features, colliding     & 3189          & 1032         & 412         & 456         \\
-        Edges, colliding        & 7330          & 2362         & 643         & 824         \\
-        Edge length, colliding  & 110.67 km     & 66.18 km     & 9389.30 m   & 20.41 km    \\
-        Features, percent       & 12.05 \%      & 9.81 \%      & 13.53 \%    & 16.55 \%    \\
-        Edges, percent          & 0.11 \%       & 0.04 \%      & 0.07 \%     & 0.20 \%     \\
-        Edge length, percent    & 0.09 \%       & 0.04 \%      & 0.03 \%     & 0.13 \%
-
-    \end{tabular}
-    \captionof{table}{The first result table of the collision study.
-This table shows the amount of features and edges, as well as the cumulative edge length, in total and which are intersecting with any other feature.}
-    \label{table:road-collision}
-\end{table}
-
-After collision correction:
-
-\begin{table}[H]
-    \begin{tabular}{lllll}
-                                & Footpath      & Residential  & Secondary   & Primary     \\
-        Features, colliding     & 134           & 336          & 246         & 277         \\
-        Edges, colliding        & 157           & 530          & 306         & 407         \\
-        Edge length, colliding  & 3248.73       & 6475.99 m    & 3840.00 m   & 8039.54 m   \\
-        Features, percent       & 0.51 \%       & 3.19 \%      & 8.08 \%     & 10.51 \%    \\
-        Edges, percent          & 0.00 \%       & 0.00 \%      & 0.03 \%     & 0.09 \%     \\
-        Edge length, percent    & 0.00 \%       & 0.00 \%      & 0.01 \%     & 0.05 \%
-
-    \end{tabular}
-    \captionof{table}{The second result table of the collision study.
-This table shows the amount of features and edges, as well as the cumulative edge length, which can be corrected by simple geometric translation.}
-    \label{table:road-collision-corrected}
-\end{table}
 

### 2020-05-10 report/report.tex
Update todo
Add note about the percentages of road that could not be corrected
Updated turning function example image with key features

@@
     \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_turn_function}
-    \caption{An example of the turning function of a polygon.}
+    \caption{An example of the turning function of a polygon. The edges, vertices and angle of the initial step of the turning funciton are marked.}
     \label{fig:turning-function}
@@
 Table \ref{table:road-collision-corrected} shows the remaining roads, edges and total edges length whose intersections with other features could not be resolved by geometrical translation of a distance smaller than the positional accuracy.
+For footpaths this was shown to hold of 0.51\% of all roads, 
+The number of road edges and percentage of the total road length that could not be corrected by geometric translation was significantly small.
 
@@
 % TODO: Reformat 8.3 to show the suggested algorithms in a list, with figure references
-% TODO: Complementary text in the results section
+% TODO: Figure showing how turning function is calculated from edges and angles
+
 % TODO: Proper citations (Fix links)
+% TODO: Expand on analysis about Feature overlap
 
-% TODO: Figure showing how turning function is calculated from edges and angles
+% TODO: Algorithm showing the edge projected distance algorithm for the collision study
 % TODO: Figure detailing the steps in the point-matching algorithm
-% TODO: Algorithm showing the edge projected distance algorithm for the collision study
-% TODO: Expand on analysis about Feature overlap
 
@@
 
+%TODO: Calculations here
 ***\% of feautures collide and ***\% of edges.

### 2020-05-10 report/report.tex
Update todo
Colliding edges per feature, calculations

@@
 % TODO: Reformat 8.3 to show the suggested algorithms in a list, with figure references
-% TODO: Figure showing how turning function is calculated from edges and angles
 
@@
 
-%TODO: Calculations here
-***\% of feautures collide and ***\% of edges.
-This means that for each colliding road, on average ***\% of its edges will be intersecting some other feature, thus bearing responsibility for the collision.
-In total there are ***\% feautures ***\% edges, meaning that each feature on average is contructed out of *** edges.
-Thus the percentage of edges colliding per road is on average ***\%.
+3189 features collide and 7330 of edges.
+This means that for each colliding road, on average 2.3 of its edges will be intersecting some other feature, thus bearing responsibility for the collision.
+In total there are 26473 feautures 6382579 edges, meaning that each feature on average is contructed out of 241 edges.
+Thus the percentage of edges colliding per road is on average roughly 0.95\%.
+A similar calculation for the number of features that could not be corrected by translation yields that the number of colliding edges per feature is on average 1.2, and that the percentage of colliding edges per feature is 0.49\%.
 The conclusion is that most often, a small part of a road is colliding, and will need correction.

### 2020-05-10 report/report.tex
Reformat Section 8.3 to show the suggested algorithms in a list, with figure references

@@
 
-% TODO: Reformat 8.3 to show the suggested algorithms in a list, with figure references
-
 % TODO: Proper citations (Fix links)
@@
 
+\begin{enumerate}
+\item Case 1 collisions can be solved by translating colliding way points away from the building, along the normal of the nearest surface. Smoothing can then be applied as seen fit.
+\item Case 3 collisions can be solved by eliminating the building features that intersect with the way, and then applying the same algorithm as in case 1.
+\item Case 4 collisions can be solved by calculating the average normal over all colliding edges for both roads, and translating colliding points along the normal, away from the other road.
+Smoothing can then be applied as seen fit.
+\item Case 5 collisions can be solved simply by translating the colliding point along the tangent of the closest edge, and then applying smoothing as seen fit.
+\end{enumerate}
+
 \begin{figure}[H]
@@
     \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_feature_overlap_fix_1}
-    \caption{Case 1 collisions can be solved by translating colliding way points away from the building, along the normal of the nearest surface. Smoothing can then be applied as seen fit.}
     \label{fig:collision-case-1}
+    \caption{Example of the solution algorithm for case 1 collision.
+The yellow line shows the original road placement.
+The blue line shows the updated road placement.
+The second picture illustrates how smoothing can be applied to the translated road if desired.}
 \end{figure}
@@
     \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_feature_overlap_fix_2}
-    \caption{Case 3 collisions can be solved by eliminating the building features that intersect with the way, and then applying the same algorithm as in case 1.}
     \label{fig:collision-case-3}
+    \caption{Example of the solution algorithm for case 3 collision.
+The portruding building feature as seen in \ref{fig:collision-cases} has been eliminated.
+The yellow line shows the original road placement.
+The blue line shows the updated road placement.}
 \end{figure}
@@
     \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_feature_overlap_fix_3}
-    \caption{Case 4 collisions can be solved by calculating the average normal over all colliding edges for both roads, and translating colliding points along the normal, away from the other road.
-Smoothing can then be applied as seen fit.}
+    \caption{Example of the solution algorithm for case 4 collision.
+The yellow lines show the original road placements.
+The blue lines show the updated road placements.
+The second picture illustrates how smoothing can be applied to the translated road if desired.}
     \label{fig:collision-case-4}
@@
     \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_feature_overlap_fix_4}
-    \caption{Case 5 collisions can be solved simply by translating the colliding point along the tangent of the closest edge, and then applying smoothing as seen fit.}
+    \caption{Example of the solution algorithm for case 5 collision.
+The yellow lines show the original road placements.
+The blue line segment with the cyan vertex shows the updated placement of the interesting vertex.
+The remaining blue lines show the boundaries of the original road.}
     \label{fig:collision-case-5}

### 2020-05-10 report/report.tex
Update todo
Fix journal sources

@@
 
+% After response from Todor:
+% TODO: Write section: Utility in real world applications
+
+% After wednesday meeting with Chris
 % TODO: Algorithm showing the edge projected distance algorithm for the collision study
 % TODO: Figure detailing the steps in the point-matching algorithm
-
-% After response from Todor:
-% TODO: Write section: Utility in real world applications
 % TODO: Spell check
@@
 \bibitem{haklay-10}
-(Haklay, 2010), How good is volunteered geographical information?A comparative study of OpenStreetMap and OrdnanceSurvey datasets \\
-\href{https://kfrichter.org/crowdsourcing-material/day1/haklay10.pdf}.
+Haklay, M. (2010). How Good is Volunteered Geographical Information? A Comparative Study of OpenStreetMap and Ordnance Survey Datasets. Environment and Planning B: Planning and Design, 37(4), 682–703.
 
 \bibitem{fan-14}
-(Fan et al, 2014), Quality assessment for building footprints data on OpenStreetMap \\
-\href{https://www.researchgate.net/publication/262163378\_Quality\_assessment\_for\_building\_footprints\_data\_on\_OpenStreetMap}.
+Fan, Hongchao \& Zipf, Alexander \& Fu, Qing \& Neis, Pascal.
+(2014).
+Quality assessment for building footprints data on OpenStreetMap.
+International Journal of Geographical Information Science.
+28.
+700-719.
 
 \bibitem{arkin-91}
-Arkin E.M., Chew L.P., Huttenlocher D.P., Kedem K. and Mitchell J.S.B. 1991. An Efficiently Computable \\
-Metric for Comparing Polygonal Shapes. In: IEEE Transaction on Pattern Analysis and Machine \\
-Intelligence, Vol. 13, No. 3, March 1991. \\
+Arkin, Esther \& Chew, L.
+\& Huttenlocher, Daniel \& Kedem, Klara \& Mitchell, Joseph.
+(1991).
+An Efficiently Computable Metric for Comparing Polygonal Shapes.
+IEEE Transaction Pattern Analysis and Machine Intelligence 13.
+209-216.
 
 \bibitem{douglas-peucker-73}
-Douglas, D.H. and Peucker T.K. 1973. Algorithms for the reduction of the number of points required to \\
-represent a digitalized line or its caricature. In: Cartographica: The International Journal for Geographic \\
-Information and Geovisualization, 10(2), 112-122. \\
+Douglas, D.H.
+\& Peucker T.K.
+(1973).
+Algorithms for the Reduction of the Number of Points Required to Represent a Digitalized Line or Its Caricature.
+Cartographica: The International Journal for Geographic Information and Geovisualization, 10(2), 112-122.
 
 \bibitem{rutzinger-09}
-Rutzinger, M., Rottensteiner, F. and Pfeifer, N. 2009. A Comparison of Evaluation Techniques for Building \\
-Extraction From Airborne Laser Scanning. IEEE Journal of Selected Topics in Applied Earth \\
-Observations and Remote Sensing 2(1): 11-20. \\
+Rutzinger, M., Rottensteiner, F.
+and Pfeifer, N.
+(2009).
+A Comparison of Evaluation Techniques for Building Extraction From Airborne Laser Scanning.
+IEEE Journal of Selected Topics in Applied Earth Observations and Remote Sensing 2(1): 11-20.
 
@@
 \bibitem{van-oort-06}
-van Oort P A J. 2006. Spatial Data Quality: From Description to Application. PhD Thesis, Wageningen University.
+Van Oort P.A.J. (2006). Spatial Data Quality: From Description to Application. PhD Thesis, Wageningen University.
 
 \bibitem{kunze-12}
-Kunze, C. 2012. Vergleichsanalyse des Gebäudedatenbestandes aus OpenStreetMap mit amtlichen Datenquellen. \\
-Student research project at the Technical University of Dresden, online avaiable: http://nbn- \\
-resolving.de/urn:nbn:de:bsz:14-qucosa-88141
+Kunze, C.
+(2012).
+Vergleichsanalyse des Gebäudedatenbestandes aus OpenStreetMap mit amtlichen Datenquellen.
+Student research project, Technical University of Dresden.
+http://nbn-resolving.de/urn:nbn:de:bsz:14-qucosa-88141
 

### 2020-05-10 report/report.tex
Update todo
Update online sources

@@
 % TODO: Figure detailing the steps in the point-matching algorithm
+% TODO: Fix citations so they adhere to the same model
 % TODO: Spell check
@@
 \bibitem{lantmateriet-19}
-(Lantmäteriet, 2019), Produktbeskrivning: GSD-Fastighetskartan vektor \\ 
-\href{https://www.lantmateriet.se/globalassets/kartor-och-geografisk-information/kartor/fastshmi.pdf}.
+Lantmäteriet. (2019). Produktbeskrivning: GSD-Fastighetskartan vektor. Acessed 16-04-2020. \\ 
+https://www.lantmateriet.se/globalassets/kartor-och-geografisk-information/kartor/fastshmi.pdf.
 
 \bibitem{osm-19}
-(OSM 19) Converting to WGS84, accessed 13-03-2020 \\ 
-\href{https://wiki.openstreetmap.org/wiki/Converting\_to\_WGS84}.
+OSM. (2019). Converting to WGS84 - OpenStreetMap Wiki. Accessed 13-03-2020. \\ 
+https://wiki.openstreetmap.org/wiki/Converting\_to\_WGS84.
 
 \bibitem{osm-20}
-(OSM 20) OSM History Viewer\\
-\href{https://wiki.openstreetmap.org/wiki/OSM\_History\_Viewer}.
+(OSM 20) OSM History Viewer - OpenStreetMap Wiki. Accessed 20-03-2020. \\
+https://wiki.openstreetmap.org/wiki/OSM\_History\_Viewer.
 

### 2020-05-10 report/report.tex
Update Stojanovsky reference
Update todo
Expand on analysis of the collision study

@@
 Such images have a limit on their resolution which puts a theoretical limit on the accuracy of map features compared to their real-world equivalents (Haklay 2010).
-Particularily the high resolution imagery from Bing in 2010 led to an increase in building information in OSM (Fan et al, 2014).
+Particularily the high resolution imagery from Bing in 2010 led to an increase in building information in OSM, and the positional error is largely due to the limited resolution in such satellite imagery.
+(Fan et al, 2014).
 
@@
 
-% TODO: Proper citations (Fix links)
-% TODO: Expand on analysis about Feature overlap
-
 % After response from Todor:
@@
 
-The biggest reason is thought to be that footpaths in their nature are narrow, so narrow that they can be displaced by more than half of their full width before the positional accuracy is reached.
+The immediate conclusion of this study is that, in terms of feature collision, footpaths were not significantly worse than other road categories.
+As a result, it can promptly be assumed that city modelling services which already utilize OSM car road data could also use OSM footpath data without a drastic increase in feature collision.
+The biggest reason for this is thought to be that footpaths in their nature are narrow.
+So narrow in fact that they can be displaced by more than half of their full width (minimum 1 meter) before the limit of half the positional accuracy (1.025 meters) is reached.
 
-Conslusion: footpaths were not worse than other categories, which are included in many unity packages already
-(Fan et al, 2014) claimed that the positional error is largely due to the limited resolution in satellite imagery.
-
-3189 features collide and 7330 of edges.
+Looking closer at the individual collision cases, the next conclusion is that most often it is actually a small part of a road that is colliding, and will need correction.
+We make an exemplifying calculation to support this claim.
+Looking in tables \ref{table:road-collision} and \ref{table:road-collision-corrected}, we note that before road extrusion, 3189 features and 7330 of their edges are intersecting some other feature.
 This means that for each colliding road, on average 2.3 of its edges will be intersecting some other feature, thus bearing responsibility for the collision.
-In total there are 26473 feautures 6382579 edges, meaning that each feature on average is contructed out of 241 edges.
+In total there are 26473 feautures and 6382579 edges, meaning that each road feature is on average contructed out of 241 edges.
 Thus the percentage of edges colliding per road is on average roughly 0.95\%.
 A similar calculation for the number of features that could not be corrected by translation yields that the number of colliding edges per feature is on average 1.2, and that the percentage of colliding edges per feature is 0.49\%.
-The conclusion is that most often, a small part of a road is colliding, and will need correction.
 Visual inspection of the cases where roads collide with other features has supported this.
 
-After visual inspection of the cases where features collide, it was determined that all collision occurences can be almost exclusively classified into one of five categories:
+After visual inspection of the cases where features collide, it was determined that all collision occurences can be almost exclusively classified into one of five categories.
+These categories are explained in the list below, and each case is exemplified in figure \ref{fig:collision-cases}.
+All these cases will create issues when generating the 3D mesh representation of the city.
 
@@
 
-Each case is exemplified in figure \ref{fig:collision-cases}.
-
 \begin{figure}[H]
@@
 \end{figure}
-
-All these cases will create issues when generating the 3D mesh representation of the city.
 
@@
 \bibitem{stojanovski-20}
-Stojanivski T. 2020. Viewpoint: City Information Modeling (CIM) And Digitizing Urban Design Practises.
-Under review
+Stojanivski T. (2020). Viewpoint: City Information Modeling (CIM) And Digitizing Urban Design Practises. Article. Under review.
 

### 2020-05-10 report/report.tex
Fix reference to Trafikverket

@@
 
-The default minimum widths were obtained from (Lantmäteriet, 20). The standard recommended minimum width for a two-lane road in Sweden is 6.5 meters.
+The default minimum widths were obtained from (Trafikverket, 20). The standard recommended minimum width for a two-lane road in Sweden is 6.5 meters.
 Subsequent visual measurements of a few key locations in the Stockholm area revealed the following minimum way widths:
@@
 \bibitem{trafikverket-20}
-https://trafikverket.ineko.se/Files/sv-SE/71830/Ineko.Product.RelatedFiles/2020\_029\_vagar\_och\_gators\_utformning\_krav.pdf \\
+Trafikverket. (2020). Vägars och gators utformning. Publikation 2020:029. \\
+https://trafikverket.ineko.se/Files/sv-SE/71830/ \\
+Ineko.Product.RelatedFiles/2020\_029\_vagar\_och\_gators\_utformning\_krav.pdf
 

### 2020-05-10 report/report.tex
Url formatting, references

@@
 \bibitem{lantmateriet-19}
-Lantmäteriet. (2019). Produktbeskrivning: GSD-Fastighetskartan vektor. Acessed 16-04-2020. \\ 
+Lantmäteriet. (2019). Produktbeskrivning: GSD-Fastighetskartan vektor. Acessed 16-04-2020.
 https://www.lantmateriet.se/globalassets/kartor-och-geografisk-information/kartor/fastshmi.pdf.
@@
 \bibitem{osm-19}
-OSM. (2019). Converting to WGS84 - OpenStreetMap Wiki. Accessed 13-03-2020. \\ 
+OSM. (2019). Converting to WGS84 - OpenStreetMap Wiki. Accessed 13-03-2020.
 https://wiki.openstreetmap.org/wiki/Converting\_to\_WGS84.
@@
 \bibitem{osm-20}
-(OSM 20) OSM History Viewer - OpenStreetMap Wiki. Accessed 20-03-2020. \\
+OSM. (2020). OSM History Viewer - OpenStreetMap Wiki. Accessed 20-03-2020
 https://wiki.openstreetmap.org/wiki/OSM\_History\_Viewer.

### 2020-05-12 report/report.tex
References in alphabetical order
Update todo
Correct Stojanovsky citing
Update SLU/OSM descriptions

@@
 
-The OSM project is based on Volunteered Geographical Information (VGI).
+The OSM project is a wiki-like geodatabase based on Volunteered Geographical Information (VGI), meaning anyone can contribute by adding or editing features.
 OSM road data and building footprints are commonly obtained by manually tracing features in commercially available satellite images.
@@
 
+The SLU map is maintained by The Swedish National Land Survey (Swedish: Lantmäteriet).
+The data is collected by geodetic professionals, by land surveying methods such as GPS or DGNSS positioning, or by reproduction of features from ortophoto or stereo mapping from 3D aerial images.
+Building features in the SLU dataset have a position accuracy requirement of 2 meters.
 The SLU map contains building features and property limits, but no road data.
-The SLU map is maintained by The Swedish National Land Survey (Swedish: Lantmäteriet).
-The SLU data is aqqcuired by land surveying methods such as GPS or DGNSS positioning, or by reproduction of features from ortophoto or stereo mapping from 3D aerial images.
 The map is updated continously by Lantmäteriet, in conjoncture with the forming or reforming of property.
-Building features in the SLU dataset have a position accuracy requirement of 2 meters.
+Whereas OSM is a free service, the SLU map is available to paying customers or for free to students and researchers.
 (Lantmäteriet, 2019)
@@
 
+% TODO: Move section 3.3 to preface (Exclude entirely?)
+% TODO: Read through Todor's recommended papers
+
 % After response from Todor:
@@
 
-\bibitem{lantmateriet-19}
-Lantmäteriet. (2019). Produktbeskrivning: GSD-Fastighetskartan vektor. Acessed 16-04-2020.
-https://www.lantmateriet.se/globalassets/kartor-och-geografisk-information/kartor/fastshmi.pdf.
-
-\bibitem{osm-19}
-OSM. (2019). Converting to WGS84 - OpenStreetMap Wiki. Accessed 13-03-2020.
-https://wiki.openstreetmap.org/wiki/Converting\_to\_WGS84.
-
-\bibitem{osm-20}
-OSM. (2020). OSM History Viewer - OpenStreetMap Wiki. Accessed 20-03-2020
-https://wiki.openstreetmap.org/wiki/OSM\_History\_Viewer.
-
-\bibitem{haklay-10}
-Haklay, M. (2010). How Good is Volunteered Geographical Information? A Comparative Study of OpenStreetMap and Ordnance Survey Datasets. Environment and Planning B: Planning and Design, 37(4), 682–703.
-
-\bibitem{fan-14}
-Fan, Hongchao \& Zipf, Alexander \& Fu, Qing \& Neis, Pascal.
-(2014).
-Quality assessment for building footprints data on OpenStreetMap.
-International Journal of Geographical Information Science.
-28.
-700-719.
-
 \bibitem{arkin-91}
@@
 
-\bibitem{rutzinger-09}
-Rutzinger, M., Rottensteiner, F.
-and Pfeifer, N.
-(2009).
-A Comparison of Evaluation Techniques for Building Extraction From Airborne Laser Scanning.
-IEEE Journal of Selected Topics in Applied Earth Observations and Remote Sensing 2(1): 11-20.
+\bibitem{fan-14}
+Fan, Hongchao \& Zipf, Alexander \& Fu, Qing \& Neis, Pascal.
+(2014).
+Quality assessment for building footprints data on OpenStreetMap.
+International Journal of Geographical Information Science.
+28.
+700-719.
 
-\bibitem{trafikverket-20}
-Trafikverket. (2020). Vägars och gators utformning. Publikation 2020:029. \\
-https://trafikverket.ineko.se/Files/sv-SE/71830/ \\
-Ineko.Product.RelatedFiles/2020\_029\_vagar\_och\_gators\_utformning\_krav.pdf
-
-\bibitem{stojanovski-20}
-Stojanivski T. (2020). Viewpoint: City Information Modeling (CIM) And Digitizing Urban Design Practises. Article. Under review.
-
-\bibitem{van-oort-06}
-Van Oort P.A.J. (2006). Spatial Data Quality: From Description to Application. PhD Thesis, Wageningen University.
+\bibitem{haklay-10}
+Haklay, M. (2010). How Good is Volunteered Geographical Information? A Comparative Study of OpenStreetMap and Ordnance Survey Datasets. Environment and Planning B: Planning and Design, 37(4), 682–703.
 
@@
 
+\bibitem{lantmateriet-19}
+Lantmäteriet. (2019). Produktbeskrivning: GSD-Fastighetskartan vektor. Acessed 16-04-2020.
+https://www.lantmateriet.se/globalassets/kartor-och-geografisk-information/kartor/fastshmi.pdf.
+
+\bibitem{osm-19}
+OSM. (2019). Converting to WGS84 - OpenStreetMap Wiki. Accessed 13-03-2020.
+https://wiki.openstreetmap.org/wiki/Converting\_to\_WGS84.
+
+\bibitem{osm-20}
+OSM. (2020). OSM History Viewer - OpenStreetMap Wiki. Accessed 20-03-2020
+https://wiki.openstreetmap.org/wiki/OSM\_History\_Viewer.
+
+\bibitem{rutzinger-09}
+Rutzinger, M., Rottensteiner, F.
+and Pfeifer, N.
+(2009).
+A Comparison of Evaluation Techniques for Building Extraction From Airborne Laser Scanning.
+IEEE Journal of Selected Topics in Applied Earth Observations and Remote Sensing 2(1): 11-20.
+
+\bibitem{stojanovski-20}
+Stojanovski, T. Partanen J., Samuels I., Sanders, P. and Peters C. (2020, forthcoming). City Information Modelling (CIM) and Digitizing Urban Design Practices. Built Environment.
+
+\bibitem{trafikverket-20}
+Trafikverket. (2020). Vägars och gators utformning. Publikation 2020:029. \\
+https://trafikverket.ineko.se/Files/sv-SE/71830/ \\
+Ineko.Product.RelatedFiles/2020\_029\_vagar\_och\_gators\_utformning\_krav.pdf
+
+\bibitem{van-oort-06}
+Van Oort P.A.J. (2006). Spatial Data Quality: From Description to Application. PhD Thesis, Wageningen University.
+
 \end{thebibliography}

### 2020-05-12 report/report.tex
Revome ESAL description

@@
 The results reveal that the positional accuracy of OSM data at the time was about 4 meters.
-
-\subsection{ESAL, KTH}
-The thesis work will be carried out at the Embodied Social Agents Lab (ESAL) at the Department of Computational Science and Technology (CST).
-The lab has before been working on systems for generating procedural urban environments, and are interested in the possibility of generating more detailed maps than have been done before, using example-based methods to preserve the aesthetic look of a greater metropolitan area.
-The lab is particularly interested in the generation of footpaths, since most commercial software that import road maps only import arterial and secondary car roads, and do not include pedestrian walkways, cycle paths and such smaller roads.
 

### 2020-05-13 report/report.tex
Update todo
Include updated project specification
Use kth template

@@
-\documentclass[a4paper]{article}
+\documentclass{kththesis}
+
+\usepackage{blindtext} % This is just to get some nonsense text in this template, can be safely removed
+
+\usepackage{csquotes} % Recommended by biblatex
+\usepackage[style=numeric,sorting=none,backend=biber]{biblatex}
+\addbibresource{references.bib} % The file containing our references, in BibTeX format
+
+\title{Quality Assesment of OpenStreetMap Footpath Data for 3D City Modelling}
+\alttitle{Detta är den svenska översättningen av titeln}
+\author{Alexander Hjelm}
+\email{osquar@kth.se}
+\supervisor{Lotta Larsson}
+\examiner{Lennart Bladgren}
+\hostcompany{Företaget AB} % Remove this line if the project was not done at a host company
+\programme{Master in Computer Science}
+\school{School of Electrical Engineering and Computer Science}
+\date{\today}
 
@@
 
-\title{Quality Assesment of OpenStreetMap Footpath Data for 3D City Modelling}
-
-\author{Alexander Hjelm}
-
-\date{\today}
+% Uncomment the next line to include cover generated at https://intra.kth.se/kth-cover?l=en
+% \kthcover{kth-cover.pdf}
 
 \begin{document}
-\maketitle
 
-\section{Abstract}
+% Frontmatter includes the titlepage, abstracts and table-of-contents
+\frontmatter
 
+\titlepage
+
+\begin{abstract}
   This project assesses the feasibility of creating a 3D representation of OpenStreetMap (OSM) geodata with the inclusion of pedestrian footpaths, by focusing on OSM road and building data in the Stockholm metropolitan area.
@@
   As secondary contributions, the completeness and shape accuracy of the OSM data set have also been estimated and commented.
+
+  \blindtext
+\end{abstract}
+
+\begin{otherlanguage}{swedish}
+  \begin{abstract}
+    Träutensilierna i ett tryckeri äro ingalunda en oviktig faktor,
+    för trevnadens, ordningens och ekonomiens upprätthållande, och
+    dock är det icke sällan som sorgliga erfarenheter göras på grund
+    af det oförstånd med hvilket kaster, formbräden och regaler
+    tillverkas och försäljas Kaster som äro dåligt hopkomna och af
+    otillräckligt.
+  \end{abstract}
+\end{otherlanguage}
+
+\tableofcontents
+
+% Mainmatter is where the actual contents of the thesis goes
+\mainmatter
 
@@
 
-% TODO: Move section 3.3 to preface (Exclude entirely?)
+% Wednesday:
+% TODO: Apply KTH thesis format
+% TODO: Bold research questions
+% TODO: Merge 2.1, 2.2
+% TODO: 2.1: Add a sentence about where the question fits into the big picture
+% TODO: 2.1: Add section refereces to implementation and results of the sub-questions
+% TODO: Reduce long section headlines
+% TODO: Rename analysis -> conclusions
+% TODO: Shorten abstract
+% TODO: Verify that citings look correct (compare to other reports)
+
+% Thursday:
+% TODO: Abstract: Mention that the scientific study is a contribution
+% TODO: Main challenge: How to reconstruct a 3D city. Make this more clear in the abstract
+% TODO: Introduction (See meeting log for suggestions)
+% TODO: Introduction, picture of a virtual city
+
+% Scietific study
+% TODO: Find papers and topics for the scientific study (Building footprints, building heights, LOD levels)
 % TODO: Read through Todor's recommended papers
-
-% After response from Todor:
-% TODO: Write section: Utility in real world applications
+% TODO: Scientific study, show how my work integrates in the 3D city reconstruction process.
+    % TODO: A lot of the metrics I've used applyg for building accuracy as well (Fan)
+    % TODO: Scientific study: Potential application: A method of generating cities on the cheap
+% TODO: Scientific study: A set of recommendations.
 
 % After wednesday meeting with Chris
-% TODO: Algorithm showing the edge projected distance algorithm for the collision study
 % TODO: Figure detailing the steps in the point-matching algorithm
-% TODO: Fix citations so they adhere to the same model
+% TODO: Edge distance algorithm, outline, algorithm + picture
 % TODO: Spell check
@@
 
+\appendix
+
+% Tailmatter inserts the back cover page (if enabled)
+\tailmatter
+
 \end{document}

### 2020-05-13 report/report.tex
Include personal info and fix toc

@@
 
-\title{Quality Assesment of OpenStreetMap Footpath Data for 3D City Modelling}
-\alttitle{Detta är den svenska översättningen av titeln}
+\title{Quality Assesment of OpenStreetMap Footpath Data for 3D City Recreation}
+\alttitle{Kvalitetiv undersökning av OpenStreetMaps gångvägsdata för stadsmodellering i 3D}
 \author{Alexander Hjelm}
-\email{osquar@kth.se}
-\supervisor{Lotta Larsson}
-\examiner{Lennart Bladgren}
-\hostcompany{Företaget AB} % Remove this line if the project was not done at a host company
+\email{aljhelm@kth.se}
+\supervisor{Christopher Peters}
+\examiner{Jonas Beskow}
+%\hostcompany{Företaget AB} % Remove this line if the project was not done at a host company
 \programme{Master in Computer Science}
@@
 
-\newpage
-\setlength{\parskip}{0em}
-\tableofcontents
-\setlength{\parskip}{1em}
-\newpage
+%\newpage
+%\setlength{\parskip}{0em}
+%\tableofcontents
+%\setlength{\parskip}{1em}
+%\newpage
 
-\section{Introduction}
+\chapter{Introduction}
 
@@
 
-% After wednesday meeting with Chris
+% After next meeting with Chris
 % TODO: Figure detailing the steps in the point-matching algorithm
@@
 % TODO: Go over report one final time
+% TODO: Swedish abstract
 

### 2020-05-13 report/report.tex
Remove commented out toc
Fix headlines

@@
 
-%\newpage
-%\setlength{\parskip}{0em}
-%\tableofcontents
-%\setlength{\parskip}{1em}
-%\newpage
-
 \chapter{Introduction}
 
-\subsection{Research Question}
+\section{Research Question}
 
@@
 
-\subsection{Evaluation}
+\section{Evaluation}
 
@@
 
-\subsection{Implementation overview}
+\section{Implementation overview}
 
-\subsubsection{Assesment of Geodata Precision}
+\subsection{Assesment of Geodata Precision}
 
@@
 
-\subsubsection{Collision study}
+\subsection{Collision study}
 
@@
 
-\subsection{Problem Constraints}
+\section{Problem Constraints}
 
@@
 
-\section{Background}
+\chapter{Background}
 
-\subsection{Background of 3D City Representation}
+\section{Background of 3D City Representation}
 
@@
 
-\subsection{Related work}
+\section{Related work}
 When assessing the completeness of geodata, a very common scale is Van Oorts criteria for evaluating the quality of geographical information, which at the time of publication received attention from surveyors and cartographers.
@@
 
-\section{The geodata used in this project}
+\chapter{The geodata used in this project}
 
@@
 
-\subsection{OSM Data}
+\section{OSM Data}
 
@@
 
-\subsection{SLU Data}
+\section{SLU Data}
 
@@
 
-\subsection{A word on coordinate systems}
+\section{A word on coordinate systems}
 
@@
 
-\subsection{Specific preprocessing for SLU and OSM geodata}
+\section{Specific preprocessing for SLU and OSM geodata}
 
@@
 
-\section{Implemenation: Evaluation of the quality of geodata}
+\chapter{Implemenation: Evaluation of the quality of geodata}
 
@@
 
-\subsection{The accuracy classes that will not be focused on}
+\section{The accuracy classes that will not be focused on}
 
@@
 
-\subsection{Assesment of dataset completeness}
+\section{Assesment of dataset completeness}
 
@@
 
-\subsection{Correspondence by building area}
+\section{Correspondence by building area}
 
@@
 
-\subsection{Shape accuracy definition}
+\section{Shape accuracy definition}
 
@@
 
-\subsection{Closest point and point proximity}
+\section{Closest point and point proximity}
 
@@
 
-\section{Implementation: Collision study}
+\chapter{Implementation: Collision study}
 
@@
 
-\section{Results and metrics}
+\chapter{Results and metrics}
 
-\subsection{Positional accuracy statistics}
+\section{Positional accuracy statistics}
 
@@
 
-\subsection{Road collision study}
+\section{Road collision study}
 
@@
 
-\subsection{Completeness by building area}
+\section{Completeness by building area}
 
@@
 
-\subsection{Building matching statistics}
+\section{Building matching statistics}
 
@@
 
-\subsection{Shape accuracy statistics}
+\section{Shape accuracy statistics}
 
@@
 
-\section{Analysis and Future Work}
+\chapter{Analysis and Future Work}
 
-\subsection{Quality of the OSM dataset}
+\section{Quality of the OSM dataset}
 
@@
 % Wednesday:
-% TODO: Apply KTH thesis format
 % TODO: Bold research questions
@@
 
-\subsection{Collision study}
+\section{Collision study}
 
@@
 
-\subsection{Suggested algorithms for correcting feature overlap}
+\section{Suggested algorithms for correcting feature overlap}
 
@@
 
-\subsection{Utility in real world applications}
+\section{Utility in real world applications}
 

### 2020-05-13 report/report.tex
Emphasize research questions

@@
 
-The primary contribution is an investigation into the rate of collision between map features in OSM.
+The primary contribution is an \emph{investigation into the rate of collision between map features in OSM}.
 A program has taken an OSM dataset of the Stockholm area and assigned each road with standard withds according to swedish regulation.
@@
 
-The secondary contribution will be an assesment of the geometrical precision of OSM data around the Stockholm area.
+The secondary contribution will be an \emph{assesment of the geometrical precision of OSM data} around the Stockholm area.
 The positional accuracy will be estimated by comparing feature points between the OSM dataset and a reference dataset.
@@
 % Wednesday:
-% TODO: Bold research questions
 % TODO: Merge 2.1, 2.2

### 2020-05-13 report/report.tex
Fix title capitalization, analysis -> conclusions
Shorten long headlines and fix main text linebreaks

@@
 
-\section{Research Question}
+\section{Research question}
 
@@
 
-\subsection{Assesment of Geodata Precision}
+\subsection{Assesment of geodata precision}
 
@@
 
-\section{Problem Constraints}
+\section{Problem constraints}
 
@@
 
-\section{Background of 3D City Representation}
+\section{Background of 3D city reproduction}
 
@@
 
-\section{OSM Data}
+\section{OSM data}
 
@@
 
-\section{SLU Data}
+\section{SLU data}
 
@@
 
-\section{Specific preprocessing for SLU and OSM geodata}
+\section{Specific geodata preprocessing}
 
@@
 
-\chapter{Implemenation: Evaluation of the quality of geodata}
+\chapter{Implementation}
+
+\section{Geodata precision study}
+
+\subsection{Quality criteria}
 
@@
 
-\section{The accuracy classes that will not be focused on}
+\subsection{Other accuracy criteria}
 
+Let us briefly go over all of van Oorts accuracy criteria that will not be assessed through this study.
+Let us see how they could be obtained and assessed for both the OSM and SLU maps, staring with the data lineage.
 The lineage aspect of the OSM data is available through the OSM History Viewer, an online debugging tool that lets anyone freely view the change history of individual fetures in a commit history-like fashion.
@@
 
-The temporal accuracy will not be assessed as this projects focuses on comparing a single snapshot of the two datasets at the same point in time.
+The temporal accuracy will not be assessed as this projects focuses on comparing a single snapshot of the two datasets at the same point in time, and finally the logical accuracy will not be assessed because the underlying logical relations between features in both maps serve no purpose to this project.
 
-The logical accuracy will not be assessed because the underlying logical relations between features in both maps serve no purpose to this project.
-
-\section{Assesment of dataset completeness}
+\subsection{Assesment of dataset completeness}
 
@@
 
-\section{Correspondence by building area}
+\subsection{Correspondence by building area}
 
@@
 
-\section{Shape accuracy definition}
+\subsection{Shape accuracy definition}
 
@@
 
-\section{Closest point and point proximity}
+\subsection{Closest point and point proximity}
 
@@
 
-\chapter{Implementation: Collision study}
+\section{Collision study}
 
@@
 
-\chapter{Analysis and Future Work}
+\chapter{Conclusions and future work}
 
@@
 
-\section{Suggested algorithms for correcting feature overlap}
+\section{Suggested algorithms for collision correction}
 

### 2020-05-13 report/report.tex
Remove blindtext from abstract
Shorten abstract

@@
 \documentclass{kththesis}
-
-\usepackage{blindtext} % This is just to get some nonsense text in this template, can be safely removed
 
@@
 An estimate of the positional accuracy of OSM data has been obtained by comparing OSM positional data to a dataset with a known accuracy, which was obtained from the Swedish National Land Survey.
-It was found that the average positional error in the OSM dataset was about 2 meters, and approximately 98\% of all buildings in the reference dataset were represented in the OSM dataset.
+It was found that the average positional error in the OSM dataset was about 2 meters.
+This estimate of the positional error was then used to assess the integrity of the OSM Stockholm road network, by locating critical points where assigning standard widths to the roads would cause collision with other features, and evaluating how many of those points could be eliminated by making ajustments within the found margin of error.
 
-  This estimate of the positional error was used to assess the integrity of the OSM Stockholm road network, by locating critical points where assigning widths to the roads would cause collision with other features.
-  An algortitm then took the road network and made adjustments that were within the found margin of error for the OSM dataset, to eliminate as many collisions as possible.
-  It was found that extending all roads with a standard width led to collisions between roughly 12\% of all roads in the dataset.
-  The total length of road that intersected with other features was less than 0.1\% of the total road length.
-  It was measured that roughly 80\% of these colliding features, and 99\% of the colliding road by length, could be corrected by simple geometrical translation by a distance smaller than the posiional accuracy of the OSM dataset.
-  It was found that footpaths were much less problematic than other road types (primary, secondary, residential), in terms of feature collisions.
+It was found that extending all roads with widths led to collisions between roughly 12\% of all roads in the dataset, and the length of road that intersected with other features was less than 0.1\% of the total road length.
+Roughly 80\% of these colliding features, and 99\% of the colliding road by length, could be corrected.
+It was found that footpaths were less problematic than other road types (primary, secondary, residential), in terms of feature collisions, and thus should pose no more problems in 3D city reproduction than other road types.
 
 The report concludes with a discussion of the findings of this study, and an assessment of the feasibility of generating city models with pedestrian footpaths in a number of different applications from urban planning tools to videogames.
-  As secondary contributions, the completeness and shape accuracy of the OSM data set have also been estimated and commented.
-
-  \blindtext
 \end{abstract}

### 2020-05-13 report/report.tex
Update todo
Merge research question + evaluation, RQ references, restructure restults

@@
 
-The secondary contribution will be an \emph{assesment of the geometrical precision of OSM data} around the Stockholm area.
-The positional accuracy will be estimated by comparing feature points between the OSM dataset and a reference dataset.
-The positional accuracy will be needed to assess where road collisions occur and whether they can easily be corrected.
-
-\section{Evaluation}
-
 The primary study of the rate of collision between roads and other features, will serve to obtain the following three metrics:
@@
 The above features will be calculated for each road type separately by summing over each road feature of a certain type.
-
 Following this, an algorithm will find the exact same metrics with regards to how many road features that can in theory be corrected by simple geometrical translation of their vertices.
 This will also be preseneted in terms of the number of road features, the number of edges and the total length in meters of remaining colliding road.
+
+For the collision study, see sections 1.3.1 for an implementation overview, 4.2 for a detailed walkthrough of the program implementation, and 5.1 for the results.
+
+The secondary contribution will be an \emph{assesment of the geometrical precision of OSM data} around the Stockholm area.
+The positional accuracy will be estimated by comparing feature points between the OSM dataset and a reference dataset.
+The positional accuracy will be needed to assess where road collisions occur and whether they can easily be corrected.
 
@@
 See section 5.3 for more details about the shape accuracy measure.
+
+For the positional accuracy study, see sections 1.3.2 for an implementation overview, 4.1 for a walkthrough of the program implementation, and 5.2 for the results.
 
@@
 
-\section{Positional accuracy statistics}
-
-For the secondary study, the average positional accuracy per building was calculated to roughly 2.06 meters, and this will be taken as an estaimate of the positional accuracy of the OSM dataset.
-Table \ref{table:positional-accuracy} shows the estimate in positional accuracy that was obtained from the study.
-Figure \ref{fig:bar-plot-positional-accuracy} shows the positional accuracies measured per building as a frequency diagram.
-
-\begin{table}[H]
-    \begin{tabular}{lllll}
-        $e_{average}$ & $e_{max}$ & $e_{min}$ & $e_{standard deviation}$ \\
-        2.0579... & 9.9878... & 0.0 & 1.2464...
-    \end{tabular}
-    \captionof{table}{The estimated positional accuracy of the OSM dataset: average, max, min and standard deviation}
-    \label{table:positional-accuracy}
-\end{table}
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_pos_error_plot}
-    \caption{The average positional accuracies per building as a frequency diagram.}
-    \label{fig:bar-plot-positional-accuracy}
-\end{figure}
-
 \section{Road collision study}
@@
 For the primary collision study.
-
 Table \ref{table:road-collision} shows the statistics of the collision study, and the number of roads, number of edges and total edge length that showed intersections with other features before road extrusion, by road category.
@@
 
-\section{Completeness by building area}
+\section{Positional accuracy study}
+
+For the secondary study, the average positional accuracy per building was calculated to roughly 2.06 meters, and this will be taken as an estaimate of the positional accuracy of the OSM dataset.
+Table \ref{table:positional-accuracy} shows the estimate in positional accuracy that was obtained from the study.
+Figure \ref{fig:bar-plot-positional-accuracy} shows the positional accuracies measured per building as a frequency diagram.
+
+\begin{table}[H]
+    \begin{tabular}{lllll}
+        $e_{average}$ & $e_{max}$ & $e_{min}$ & $e_{standard deviation}$ \\
+        2.0579... & 9.9878... & 0.0 & 1.2464...
+    \end{tabular}
+    \captionof{table}{The estimated positional accuracy of the OSM dataset: average, max, min and standard deviation}
+    \label{table:positional-accuracy}
+\end{table}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_pos_error_plot}
+    \caption{The average positional accuracies per building as a frequency diagram.}
+    \label{fig:bar-plot-positional-accuracy}
+\end{figure}
+
+\section{Secondary metrics}
+
+\subsection{Completeness by building area}
 
@@
 
-\section{Building matching statistics}
+\subsection{Building matching statistics}
 
@@
 
-\section{Shape accuracy statistics}
+\subsection{Shape accuracy statistics}
 
@@
 % TODO: 2.1: Add section refereces to implementation and results of the sub-questions
-% TODO: Reduce long section headlines
-% TODO: Rename analysis -> conclusions
-% TODO: Shorten abstract
-% TODO: Verify that citings look correct (compare to other reports)
 
 % Thursday:
+% TODO: Verify that citings look correct (compare to other reports)
 % TODO: Abstract: Mention that the scientific study is a contribution

### 2020-05-14 report/report.tex
Update todo
Test citing

@@
 Problems that arise are e.g. that the vertex counts are dissimilar, or that vertex clusters may be found at different parts of the polygon in the two datasets.
-To avoid this effect, key points are extracted using the Douglas-Peucker algorithm (Douglas and Peucker, 1973), to create a simplifed footprint with less points that still retain information about the rough features of the detailed footprint.
+To avoid this effect, key points are extracted using the Douglas-Peucker algorithm \cite{douglas-73}, to create a simplifed footprint with less points that still retain information about the rough features of the detailed footprint.
 The idea behind Douglas-Peucker is to recursively divide a polyline.
@@
 
-% Wednesday:
-% TODO: Merge 2.1, 2.2
-% TODO: 2.1: Add a sentence about where the question fits into the big picture
-% TODO: 2.1: Add section refereces to implementation and results of the sub-questions
+% Thursday:
+% TODO: Reference labels
+% TODO: Fix citings like they appear in the example thesis
+% TODO: Abstract: Mention that the scientific study is a contribution
 
-% Thursday:
-% TODO: Verify that citings look correct (compare to other reports)
-% TODO: Abstract: Mention that the scientific study is a contribution
-% TODO: Main challenge: How to reconstruct a 3D city. Make this more clear in the abstract
 % TODO: Introduction (See meeting log for suggestions)
     % TODO: Introduction, picture of a virtual city
+    % TODO: Main challenge: How to reconstruct a 3D city. Make this more clear in the abstract
+    % TODO: 2.1: Add a sentence about where the question fits into the big picture
+% TODO: Collision study implementation, disposition
+    % TODO: Figure detailing the steps in the point-matching algorithm
+    % TODO: Edge distance algorithm, outline, algorithm + picture
 
@@
 % After next meeting with Chris
-% TODO: Figure detailing the steps in the point-matching algorithm
-% TODO: Edge distance algorithm, outline, algorithm + picture
 % TODO: Spell check
@@
 
-\bibitem{douglas-peucker-73}
+\bibitem{douglas-73}
 Douglas, D.H.

### 2020-05-14 report/report.tex
Update abstract and begin introduction disposition

@@
 \begin{abstract}
-This project assesses the feasibility of creating a 3D representation of OpenStreetMap (OSM) geodata with the inclusion of pedestrian footpaths, by focusing on OSM road and building data in the Stockholm metropolitan area.
+Modeling a city in 3D has long been a topic of interest in field of computer graphics.
+It is often desirable to recreate real-world cities in 3D applications such as urban design software and video games, and the topic of 3D city reproduction comes with its own set of unique challenges.
+So far there have been a lot of research in modelling of 3D road networks, but these projects have often focused on the road networks designated for cars and public transit, and seldom on pedestrian footpaths.
+This project assesses the feasibility of reconstructing a city from OpenStreetMap (OSM) road and building data, and will focus on the feasibility of including of pedestrian footpaths, by focusing on OSM road and building data in the Stockholm metropolitan area.
 
-An estimate of the positional accuracy of OSM data has been obtained by comparing OSM positional data to a dataset with a known accuracy, which was obtained from the Swedish National Land Survey.
+An estimate of the positional accuracy of OSM data has been obtained by comparing OSM data to a dataset with a known accuracy.
 It was found that the average positional error in the OSM dataset was about 2 meters.
 This estimate of the positional error was then used to assess the integrity of the OSM Stockholm road network, by locating critical points where assigning standard widths to the roads would cause collision with other features, and evaluating how many of those points could be eliminated by making ajustments within the found margin of error.
-
-It was found that extending all roads with widths led to collisions between roughly 12\% of all roads in the dataset, and the length of road that intersected with other features was less than 0.1\% of the total road length.
-Roughly 80\% of these colliding features, and 99\% of the colliding road by length, could be corrected.
 It was found that footpaths were less problematic than other road types (primary, secondary, residential), in terms of feature collisions, and thus should pose no more problems in 3D city reproduction than other road types.
-
-The report concludes with a discussion of the findings of this study, and an assessment of the feasibility of generating city models with pedestrian footpaths in a number of different applications from urban planning tools to videogames.
+The report concludes with a scientific study of the field of 3D city reproduction.
+It serves to give an overview of the challenges in the field and place the study and findings of this project in a broader scientific context.
 \end{abstract}
@@
 \chapter{Introduction}
+
+Description of a virtual city and its contents.
+
+Description of different levels of subdivision. What levels of subdivision are we looking at? Picture, ref to Todor's paper.
+
+So far there have been a lot of research in modelling of 3D road networks, but these projects have often focused on the road networks designated for cars and public transit, due to environmental concerns (Todors paper)
+
+Description of the unique character of footpaths vs other types of roads (pictures).
 

### 2020-05-14 report/report.tex
Update todo
Update introduction disposition

@@
 
+Description of the field of city reproduction (Parish Müller).
+
 Description of a virtual city and its contents.
@@
 Description of different levels of subdivision. What levels of subdivision are we looking at? Picture, ref to Todor's paper.
+
+Viewpoints can range from modelling individual roads (find reference) and building facades (todor) to looking at entire city parcels and how they connect with each other (find reference).
+
+Description of procedural modelling of cities (find reference).
 
@@
 % Thursday:
-% TODO: Reference labels
 % TODO: Fix citings like they appear in the example thesis
-% TODO: Abstract: Mention that the scientific study is a contribution
 
-% TODO: Introduction (See meeting log for suggestions)
+% TODO: Introduction disposition (See meeting log for suggestions)
     % TODO: Introduction, picture of a virtual city

### 2020-05-14 report/report.tex
Fix typos in abstract

@@
 \begin{abstract}
-Modeling a city in 3D has long been a topic of interest in field of computer graphics.
+Modeling cities in 3D has long been a topic of interest in field of computer graphics.
 It is often desirable to recreate real-world cities in 3D applications such as urban design software and video games, and the topic of 3D city reproduction comes with its own set of unique challenges.
 So far there have been a lot of research in modelling of 3D road networks, but these projects have often focused on the road networks designated for cars and public transit, and seldom on pedestrian footpaths.
-This project assesses the feasibility of reconstructing a city from OpenStreetMap (OSM) road and building data, and will focus on the feasibility of including of pedestrian footpaths, by focusing on OSM road and building data in the Stockholm metropolitan area.
+This project assesses the feasibility of reconstructing a city from OpenStreetMap (OSM) road and building data, and will focus on the feasibility of including pedestrian footpaths, by focusing on OSM road and building data in the Stockholm metropolitan area.
 

### 2020-05-14 report/report.tex
Research question: include a paragrah about how the RQ fits into the bigger picture

@@
 
+This project assesses the feasibility of reconstructing a city from OpenStreetMap (OSM) road and building data, and will focus on the feasibility of including pedestrian footpaths, by focusing on OSM road and building data in the Stockholm metropolitan area.
+
 \section{Research question}
 
-This project has dealt with the feasibility of including footpath data when generating 3D representations of OpenStreetMap (OSM) data.
-A key issue in creating 3D representations of geodata is that since individual features will be extruded to give them width, situations may arise where the generated 3D meshes are intersecting.
+This project has dealt with the feasibility of including footpath data when generating 3D representations of OpenStreetMap (OSM) data, and preserving the integrity of the 3D road network, avoiding overlap and collision between road meshes.
+The project falls within the broader topic of geodata extraction and modelling, primarily of roads and building parcels.
+Geodata modeling is at the foundation of procedural modelling of real-world cities, since any application which models cities based on geodata will need a robust method of aqquiring and laying out road and building data without compromising the data's integrity.
+(find reference)
 

### 2020-05-14 report/report.tex
Update todo
Include introduction disposition points from meeting log

@@
 
+Picture of a virtual city.
+
 Description of a virtual city and its contents.
+Consisting of road network, buildings, details.
+Roads have different types.
 
@@
 Description of procedural modelling of cities (find reference).
+Importing geodata from online databases and converting it to a 3D city representation.
+
+At the base of this process is the procedural layout of road networks in 3D.
 
@@
 Studios who create large worlds are often interested in solutions that generate large road networks without any hand modeling.
-Navigation agents, they might fail navigating when faced with a too narrow path or intersecting obstacles
-Using a road network to creates an underlying navmesh
+%TODO: (Parish Müller: Procedural modelin of cities)
 
@@
 % TODO: Fix citings like they appear in the example thesis
-
-% TODO: Introduction disposition (See meeting log for suggestions)
-    % TODO: Introduction, picture of a virtual city
-    % TODO: Main challenge: How to reconstruct a 3D city. Make this more clear in the abstract
-    % TODO: 2.1: Add a sentence about where the question fits into the big picture
 % TODO: Collision study implementation, disposition

### 2020-05-14 report/report.tex
Add navigation application with reference

@@
 
-Description of the field of city reproduction (Parish Müller).
+Description of the field of city reproduction (Parish Müller, procedural modeling of ciites).
 
@@
 Roads have different types.
+
+Applications:
+    Navigating agents or pedestrian flows.
+    Using a road network to creates an underlying navmesh
+    This process is analytical and relies on the availablity and accurate representation of city maps.
+    %TODO: (Parish Müller: Procedural modelin of cities)
 
@@
     % TODO: Edge distance algorithm, outline, algorithm + picture
+% TODO: Introduction references
+% TODO: Write introduction
 

### 2020-05-14 report/report.tex
Update introduction disposition

@@
     This process is analytical and relies on the availablity and accurate representation of city maps.
-    %TODO: (Parish Müller: Procedural modelin of cities)
+    (Parish Müller: Procedural modelin of cities)
 
 Description of different levels of subdivision. What levels of subdivision are we looking at? Picture, ref to Todor's paper.
-
 Viewpoints can range from modelling individual roads (find reference) and building facades (todor) to looking at entire city parcels and how they connect with each other (find reference).
 
-Description of procedural modelling of cities (find reference).
+Description of procedural modelling of cities
 Importing geodata from online databases and converting it to a 3D city representation.
+It would be labour intensive to recreate these .
+Procedural models for city reproduction instead rely on programmatic solutions for converting map data to 3D road and building models, and composing them to a 3D world that the designer may continue to fine-tweak.
+(Parish Müller, Procedural modeling of buildings)
 
 At the base of this process is the procedural layout of road networks in 3D.
+(Parish Müller, Procedural modeling of cities)
 

### 2020-05-14 report/report.tex
Update todo
Add beginning of scientific study disposition

@@
 % TODO: Fix citings like they appear in the example thesis
+% TODO: Write introduction section
+    % TODO: Figure showing the unique characteristics of footpaths vs other road types (OSM Screeshot)
+    % TODO: Virtual city picture
+    % TODO: Introduction references
+% TODO: Write Collision study implementation section
     % TODO: Collision study implementation, disposition
@@
     % TODO: Edge distance algorithm, outline, algorithm + picture
-% TODO: Introduction references
-% TODO: Write introduction
 
@@
 
-\section{Utility in real world applications}
+\section{Scientific study: Where does this fit in the general Scietific scope of things}
+
+The same methods can be applied to import building footprint data from OSM or some, and the same metrics can be applied to assess the data accuracy.
+When we have the fundamental
+we can proceed with building modeling (Parish Müller, procedural building modeling).
+Other challenges include infering or guessing building height.
+() have presented a method for inferring building height from the building footprint alone, with surprising accuracy.
 

### 2020-05-14 report/report.tex
Update scientific study disposition

@@
 
-The same methods can be applied to import building footprint data from OSM or some, and the same metrics can be applied to assess the data accuracy.
-When we have the fundamental
-we can proceed with building modeling (Parish Müller, procedural building modeling).
+The same methods can be applied to import building footprint data from OSM or some, and the same metrics can be applied to assess the data accuracy. (Fan et al)
+When we have the fundamental procedure for accuratly importing building footprints without compromising the data integritey, we can proceed with building modeling (Parish Müller, procedural building modeling).
 Other challenges include infering or guessing building height.
 () have presented a method for inferring building height from the building footprint alone, with surprising accuracy.
+
+Depending on the level of data we have access to, we can then choose what level of representation we would like for our application (Todor).
+We may also chose what LOD class(es) to use depending on data availablity and application. (find reference)
 

### 2020-05-16 report/report.tex
Update todo
Fix citings

@@
 
-\usepackage{csquotes} % Recommended by biblatex
-\usepackage[style=numeric,sorting=none,backend=biber]{biblatex}
-\addbibresource{references.bib} % The file containing our references, in BibTeX format
+%\usepackage{csquotes} % Recommended by biblatex
+\usepackage[style=numeric,sorting=none,backend=biber,citestyle=verbose]{biblatex}
+%\addbibresource{references.bib} % The file containing our references, in BibTeX format
 
@@
 The study applied different methods to assess the completeness of OSM data in two federal states in Germany, mainly by analysing the area difference between the OSM dataset and an administrative dataset.
-Finally, a notable study in 2013-2014, conducted by researchers from various universities in Norway and Germany (Fan et al, 2014), examined the quality and accuracy in building footprint data of the Munich area, at a time and place where the completeness of the OSM dataset was significantly higher (the results of their study showed that the OSM had captured close to 100\% of all building footprints).
+Finally, a notable study in 2013-2014, conducted by researchers from various universities in Norway and Germany \textbf{(Fan et al, 2014)}, examined the quality and accuracy in building footprint data of the Munich area, at a time and place where the completeness of the OSM dataset was significantly higher (the results of their study showed that the OSM had captured close to 100\% of all building footprints).
 At the time Munich was one of the most developed cities in OSM.
@@
 Particularily the high resolution imagery from Bing in 2010 led to an increase in building information in OSM, and the positional error is largely due to the limited resolution in such satellite imagery.
-(Fan et al, 2014).
+\textbf{(Fan et al, 2014)}.
 
@@
 
-(Fan et al., 2014) used the fraction of the total building area in their reference dataset and the OSM dataset to assess the completeness of OSM data.
+\textbf{(Fan et al, 2014)} used the fraction of the total building area in their reference dataset and the OSM dataset to assess the completeness of OSM data.
 The motivation for this is to eliminate semantic differences between both datasets.
@@
 Problems that arise are e.g. that the vertex counts are dissimilar, or that vertex clusters may be found at different parts of the polygon in the two datasets.
-To avoid this effect, key points are extracted using the Douglas-Peucker algorithm \cite{douglas-73}, to create a simplifed footprint with less points that still retain information about the rough features of the detailed footprint.
+To avoid this effect, key points are extracted using the Douglas-Peucker algorithm \textbf{(Douglas and Peucker, 1973)}, to create a simplifed footprint with less points that still retain information about the rough features of the detailed footprint.
 The idea behind Douglas-Peucker is to recursively divide a polyline.
@@
 % TODO: Find papers and topics for the scientific study (Building footprints, building heights, LOD levels)
-% TODO: Read through Todor's recommended papers
 % TODO: Scientific study, show how my work integrates in the 3D city reconstruction process.
-    % TODO: A lot of the metrics I've used applyg for building accuracy as well (Fan)
     % TODO: Scientific study: Potential application: A method of generating cities on the cheap

### 2020-05-16 report/report.tex
Emph -> Textbf

@@
 
-The primary contribution is an \emph{investigation into the rate of collision between map features in OSM}.
+The primary contribution is an \textbf{investigation into the rate of collision between map features in OSM}.
 A program has taken an OSM dataset of the Stockholm area and assigned each road with standard withds according to swedish regulation.
@@
 
-The secondary contribution will be an \emph{assesment of the geometrical precision of OSM data} around the Stockholm area.
+The secondary contribution will be an \textbf{assesment of the geometrical precision of OSM data} around the Stockholm area.
 The positional accuracy will be estimated by comparing feature points between the OSM dataset and a reference dataset.

### 2020-05-16 report/report.tex
Fix scientific citings

@@
 First a feature map was found, which mapped individual features between the OSM dataset and a reference dataset.
-A program then took matching building pairs and extracted matching geometrical points after first simplifying the polygons using the Douglas-Peucker algorithm.
+A program then took matching building pairs and extracted matching geometrical points after first simplifying the polygons using the Douglas-Peucker algorithm \textbf{(Douglas and Peucker, 1973)}.
 Once a set of matching points has been found across the whole dataset, the average distance between them was taken as a measure of the positional accuracy.
@@
 Urban arcitecture is a social factors, stakeholders.
-Each arcitect is an actor in a continously developing environment, and affecting that environment in a complex manner through their action.
-(Stojanivski, 2020)
+Each arcitect is an actor in a continously developing environment, and affecting that environment in a complex manner through their action \textbf{(Stojanovski et al, 2020)}.
 
@@
 When assessing the completeness of geodata, a very common scale is Van Oorts criteria for evaluating the quality of geographical information, which at the time of publication received attention from surveyors and cartographers.
-(van Oort, 2006) Since then a number of publications have examined slices of geodata from van Oort's criteria.
-In 2008, (Haklay, 2010) conducted such a study to estimate of the quality of OSM positional geodata in London and England.
+\textbf{(Van Oort, 2006)} Since then a number of publications have examined slices of geodata from van Oort's criteria.
+In 2008, \textbf{(Haklay, 2010)} conducted such a study to estimate of the quality of OSM positional geodata in London and England.
 Since the OSM project started in London it was thought that OSM data in the London metropolitan area would be representative of the highest quality data available, and therefore a good indicator for the whole global OSM dataset.
@@
 
-The next notable study to assess the quality of OSM data was published by (Kunze, 2012).
+The next notable study to assess the quality of OSM data was published by \textbf{(Kunze, 2012)}.
 The study applied different methods to assess the completeness of OSM data in two federal states in Germany, mainly by analysing the area difference between the OSM dataset and an administrative dataset.
@@
 
-(Haklay 2010) presents a comprehesive list of 8 accuracy classes when it comes to evaluating the quality of geographic information (van Oorts criteria)
+\textbf{(Haklay, 2010)} presents a comprehesive list of 8 accuracy classes when it comes to evaluating the quality of geographic information (van Oorts criteria)
 
@@
 
-(Fan et al., 2014) further discussed how the relative overlap in building area can be used as an estimate of the building correspondence between datasets.
+\textbf{(Fan et al, 2014)} further discussed how the relative overlap in building area can be used as an estimate of the building correspondence between datasets.
 In this project, relative area overlap was used in order to pair buildings from the two datasets together and create a feature map.
@@
 The relative overlap may be used to determine building matching relations even when the semantic accuracy is low.
-If a large building is represented by a single footprint in one dataset but by several smaller footprints in the other, (Rutzinger et al, 2009) found that if $S_{RO}(foot_{A}, foot_{B}) < 30\%$ for two buildings $A$ and $B$ from different sets, then $A$ and $B$ are highly likely to be separate, neighbouring buildings and not in fact identical.
+If a large building is represented by a single footprint in one dataset but by several smaller footprints in the other, \textbf{(Rutzinger, Rottensteiner and Pfeifer, 2009)} found that if $S_{RO}(foot_{A}, foot_{B}) < 30\%$ for two buildings $A$ and $B$ from different sets, then $A$ and $B$ are highly likely to be separate, neighbouring buildings and not in fact identical.
 Therefore it will here be assumed that two buildings are matching candidates if their relative area overlap is greater than 30\%.
@@
 The building footprint correspondence will be evaluated using a similarity function that depends on the difference between the turning funcitons of matching building footprints in both datasets.
-The turning function was first defined by Arkin et al. (1991), as a method for measuring the similarity of two polygons.
+The turning function was first defined by \textbf{(Arkin et al, 1991)}, as a method for measuring the similarity of two polygons.
 The Turning function $T_c(l)$ measures the cumulative angle of the polygon's counter-clockwise tangent, as a function of the cumulative normalized length l.
-This project uses the turning function as it is defined by (Fan et al., 2014).
+This project uses the turning function as it is defined by \textbf{(Fan et al, 2014)}.
 See figure \ref{fig:turning-function} for a side-by-side comparison of a polygon and its turning function.
@@
 Let us take the positional accuracy of the Stockholm area, as obtained by this study, as reprentative of that of major cities in developed countries in the OSM dataset.
-Then the positional accuracy has improved by roughly 3.5 meters since the study by (Haklay, 2010), and by roughly 1.5 meters since the study by (Fan et al, 2014).
+Then the positional accuracy has improved by roughly 3.5 meters since the study by \textbf{(Haklay, 2010)}, and by roughly 1.5 meters since the study by \textbf{(Fan et al, 2014)}.
 This of course assumes that the SLU reference dataset is taken as absolute.
 
-The completeness analysis by building area shows that the OSM Stockholm dataset has high completeness in terms of both building area and counted features, in the same level as the study of Munich by (Fan et al, 2014).
+The completeness analysis by building area shows that the OSM Stockholm dataset has high completeness in terms of both building area and counted features, in the same level as the study of Munich by \textbf{(Fan et al, 2014)}.
 The difference in area cover can partially be explained by small utility buildings that are often not represented in the OSM dataset.
@@
 These buildings are likely hard to make out and trace on publicly available satellite image data.
-They are, according to (Fan et al, 2014) occluded by their surroundings due to shadows or forestry.
+They are, according to \textbf{(Fan et al, 2014)} occluded by their surroundings due to shadows or forestry.
 This appears to be a large consensus in the field.
@@
 
-The same methods can be applied to import building footprint data from OSM or some, and the same metrics can be applied to assess the data accuracy. (Fan et al)
+The same methods can be applied to import building footprint data from OSM or some, and the same metrics can be applied to assess the data accuracy. \textbf{(Fan et al, 2014)}
 When we have the fundamental procedure for accuratly importing building footprints without compromising the data integritey, we can proceed with building modeling (Parish Müller, procedural building modeling).

### 2020-05-16 report/report.tex
Fix reference author formatting
Convert non-scientific references to footnotes

@@
 Whereas OSM is a free service, the SLU map is available to paying customers or for free to students and researchers.
-(Lantmäteriet, 2019)
+\footnote{Lantmäteriet.
+Produktbeskrivning: GSD-Fastighetskartan vektor - Lantmäteriet.
+Acessed 16-04-2020.
+https://www.lantmateriet.se/globalassets/kartor-och-geografisk-information/kartor/fastshmi.pdf}
 
@@
 
-The SLU dataset is delivered in the SWEREF 99 TM coordinate system (Lantmäteriet, 2019).
-SWEREF 99 TM is a projected coordinate system and there is no linear transformation to the WGS 48 system, which OSM uses (OSM 19).
+The SLU dataset is delivered in the SWEREF 99 TM coordinate system.
+SWEREF 99 TM is a projected coordinate system and there is no linear transformation to the WGS 48 system, which OSM uses.
+\footnote{Converting to WGS84 - OpenStreetMap Wiki.
+Accessed 13-03-2020.
+https://wiki.openstreetmap.org/wiki/Converting\_to\_WGS84}
 The coordinate conversions for this paper were obtained using proj: a Linux commandline application for geospatial coordinate conversion.
@@
 The lineage aspect of the OSM data is available through the OSM History Viewer, an online debugging tool that lets anyone freely view the change history of individual fetures in a commit history-like fashion.
-There is an option for editors to include a personal note with their changesets to include additional details or motivate why a change was made, but there is no guaratee that the data includes any information on the aquisition method used (OSM 2020).
+There is an option for editors to include a personal note with their changesets to include additional details or motivate why a change was made, but there is no guaratee that the data includes any information on the aquisition method used \footnote{OSM History Viewer - OpenStreetMap Wiki.
+Accessed 20-03-2020.
+https://wiki.openstreetmap.org/wiki/OSM\_History\_Viewer}
 The history aspect of the SLU dataset is not freely available online, but the acquistion method is included in the file metadata on a per-object basis.
@@
 
-The default minimum widths were obtained from (Trafikverket, 20). The standard recommended minimum width for a two-lane road in Sweden is 6.5 meters.
+See the footnote for the publication that the default minimum widths were obtained from.
+\footnote{Vägars och gators utformning - Trafikverket.
+Publikation 2020:029.
+https://trafikverket.ineko.se/Files/sv-SE/71830/\\Ineko.Product.RelatedFiles/2020\_029\_vagar\_och\_gators\_utformning\_krav.pdf}
+The standard recommended minimum width for a two-lane road in Sweden is 6.5 meters.
+
 Subsequent visual measurements of a few key locations in the Stockholm area revealed the following minimum way widths:
@@
 Douglas, D.H.
-\& Peucker T.K.
+and Peucker T.K.
 (1973).
@@
 \bibitem{fan-14}
-Fan, Hongchao \& Zipf, Alexander \& Fu, Qing \& Neis, Pascal.
-(2014).
-Quality assessment for building footprints data on OpenStreetMap.
-International Journal of Geographical Information Science.
-28.
-700-719.
+Fan, H., Zipf, A., Fu, Q. and Neis, P. (2014). Quality assessment for building footprints data on OpenStreetMap. International Journal of Geographical Information Science. 28. 700-719.
 
@@
 
-\bibitem{lantmateriet-19}
-Lantmäteriet. (2019). Produktbeskrivning: GSD-Fastighetskartan vektor. Acessed 16-04-2020.
-https://www.lantmateriet.se/globalassets/kartor-och-geografisk-information/kartor/fastshmi.pdf.
-
-\bibitem{osm-19}
-OSM. (2019). Converting to WGS84 - OpenStreetMap Wiki. Accessed 13-03-2020.
-https://wiki.openstreetmap.org/wiki/Converting\_to\_WGS84.
-
-\bibitem{osm-20}
-OSM. (2020). OSM History Viewer - OpenStreetMap Wiki. Accessed 20-03-2020
-https://wiki.openstreetmap.org/wiki/OSM\_History\_Viewer.
-
 \bibitem{rutzinger-09}
@@
 Stojanovski, T. Partanen J., Samuels I., Sanders, P. and Peters C. (2020, forthcoming). City Information Modelling (CIM) and Digitizing Urban Design Practices. Built Environment.
-
-\bibitem{trafikverket-20}
-Trafikverket. (2020). Vägars och gators utformning. Publikation 2020:029. \\
-https://trafikverket.ineko.se/Files/sv-SE/71830/ \\
-Ineko.Product.RelatedFiles/2020\_029\_vagar\_och\_gators\_utformning\_krav.pdf
 

### 2020-05-16 report/report.tex
Rebuild report
Fix remaining citations

@@
 
-Description of the field of city reproduction (Parish Müller, procedural modeling of ciites).
+Description of the field of city reproduction \textbf{(Parish and Müller, 2001)}.
 
@@
     This process is analytical and relies on the availablity and accurate representation of city maps.
-    (Parish Müller: Procedural modelin of cities)
+    \textbf{(Parish and Müller, 2001)}
 
 Description of different levels of subdivision. What levels of subdivision are we looking at? Picture, ref to Todor's paper.
-Viewpoints can range from modelling individual roads (find reference) and building facades (todor) to looking at entire city parcels and how they connect with each other (find reference).
+Viewpoints can range from modelling individual roads (***find reference) and building facades \textbf{(Stojanovski, 20***)} to looking at entire city parcels and how they connect with each other (***find reference).
 
@@
 Procedural models for city reproduction instead rely on programmatic solutions for converting map data to 3D road and building models, and composing them to a 3D world that the designer may continue to fine-tweak.
-(Parish Müller, Procedural modeling of buildings)
+\textbf{(Müller et al, 2006)}
 
 At the base of this process is the procedural layout of road networks in 3D.
-(Parish Müller, Procedural modeling of cities)
+\textbf{(Müller et al, 2006)}
 
-So far there have been a lot of research in modelling of 3D road networks, but these projects have often focused on the road networks designated for cars and public transit, due to environmental concerns (Todors paper)
+So far there have been a lot of research in modelling of 3D road networks, but these projects have often focused on the road networks designated for cars and public transit, due to environmental concerns.
+\textbf{(Stojanovski, 2020)}
 
@@
 Geodata modeling is at the foundation of procedural modelling of real-world cities, since any application which models cities based on geodata will need a robust method of aqquiring and laying out road and building data without compromising the data's integrity.
-(find reference)
+(***find reference)
 
@@
 Studios who create large worlds are often interested in solutions that generate large road networks without any hand modeling.
-%TODO: (Parish Müller: Procedural modelin of cities)
+\textbf{(Parish and Müller, 2001)}
 
@@
 % Thursday:
-% TODO: Fix citings like they appear in the example thesis
 % TODO: Write introduction section
@@
 The same methods can be applied to import building footprint data from OSM or some, and the same metrics can be applied to assess the data accuracy. \textbf{(Fan et al, 2014)}
-When we have the fundamental procedure for accuratly importing building footprints without compromising the data integritey, we can proceed with building modeling (Parish Müller, procedural building modeling).
-Other challenges include infering or guessing building height.
+When we have the fundamental procedure for accuratly importing building footprints without compromising the data integritey, we can proceed with building modeling.
+\textbf{(Müller et al, 2006)} Other challenges include infering or guessing building height.
 () have presented a method for inferring building height from the building footprint alone, with surprising accuracy.
@@
 Depending on the level of data we have access to, we can then choose what level of representation we would like for our application (Todor).
-We may also chose what LOD class(es) to use depending on data availablity and application. (find reference)
+We may also chose what LOD class(es) to use depending on data availablity and application. (***find reference)
 

### 2020-05-16 report/report.tex
Add Parish, Müller references
Update todo
Rebuild report
References in alphabetical order
Collision study, disposition

@@
 
+\subsection{Road widths and properties}
+
 The maximum distance between two features was set to 1 times the positional accuracy, since then there will always be ample room to make changes to any features that would collide with the newly moved nodes, and thus we do not need to worry about feature collisions propagating.
@@
 \end{table}
+
+\subsection{Feature overlap algorithm}
+
+Outline of the feature overlap algorithm (Go over all features)
+Since only road collisions are of interests, all roads were added to a queue and gone over one by one.
+
+The boundary polygon of the buildings are extracted, as well as the polylines of the roads.
+
+To speed up the collision checking process, a geometric hashing solution was developed.
+This way feature collision lookup could be made without comparing a feature to all other available features, but instead from a selection of the closest candidates.
+The hash bucket size was 230 meters and a feature could be in multiple hash buckets.
+
+Special cases: Where two roads intersect though they share a node.
+Then this is a natural connection in the road network and should not be counted as a problematic case of feature overlap.
+Also ignored collisions between tunnels and bridges and other roads since these are of different heights.
+These were identified by looking at feature attributes in OSM.
+Any roads that were represented as polygons or multipolygons instead of plylines were also excluded from the study.
+
+\subsection{Projected edge distance}
+
+Outline of the minimum edge distance algorithm.
 
@@
     % TODO: Virtual city picture
-    % TODO: Introduction references
+    % TODO: read Biljecki
+    % TODO: read Vanega
 % TODO: Write Collision study implementation section
-    % TODO: Collision study implementation, disposition
     % TODO: Figure detailing the steps in the point-matching algorithm
@@
 \bibitem{arkin-91}
-Arkin, Esther \& Chew, L.
-\& Huttenlocher, Daniel \& Kedem, Klara \& Mitchell, Joseph.
+Arkin, E., Chew, L., Huttenlocher, D., Kedem, K.
+and Mitchell, J.
 (1991).
@@
 
+\bibitem{muller-06}
+Müller, P., Wonka, P., Haegler, S., Ulmer, A. and Van Gool, L. (2006). Procedural modeling of buildings. In ACM SIGGRAPH 2006 Papers, 614-623.
+
+\bibitem{parish-01}
+Parish, Y.
+I., and Müller, P.
+(2001) Procedural modeling of cities.
+In Proceedings of the 28th annual conference on Computer graphics and interactive techniques, SIGGRAPH, 12-17 August 2001, Los Angeles, Calif., 301-308.
+
 \bibitem{rutzinger-09}

### 2020-05-16 report/report.tex
Add note about LOD to scientific study

@@
 Depending on the level of data we have access to, we can then choose what level of representation we would like for our application (Todor).
-We may also chose what LOD class(es) to use depending on data availablity and application. (***find reference)
+
+We may also chose what LOD class(es) to use depending on data availablity and application.
+With the current means of data aquisition being limited to building footprints, using OSM data we can only produce CityGML 2.0 LOD-B-models, that is building blocks with simplifed roof shapes (LOD-A being a rendering of the building footprint itself, or building blocks extruded from the footprint to give the building height).
+\textbf{(Biljecki, Ledoux and Stoter, 2016)}
+If we wish to aqquire building assets at a LOD-C level of detail, including features such as doors, windows, alcoves, building decor, indoor features or vegetation, we would have to complete the data with other sources, or rely on procedural methods to populate.
+Such procedural techniques have long been available.
+\textbf{(Mũller et al, 2006)} present how to accomplish this using shape grammar, a method for describing the possible relations between 
+For simple presentation in e.g. animation and videogames or even urban planning software, this could very well be enough.
+As long as one can be satisified with looking at rough representations of buildings and does not require additional knowledge about minor features.
 
@@
 209-216.
+
+\bibitem{biljecki-14}
+Biljecki, F., Ledoux, H., Stoter, J., Zhao J. (2014): Formalisation of the level of detail in 3D city modelling. Computers, Environment and Urban Systems, vol. 48, pp. 1-15.
 

### 2020-05-16 report/report.tex
Update todo
Reorder scientific study

@@
 % Scietific study
-% TODO: Find papers and topics for the scientific study (Building footprints, building heights, LOD levels)
-% TODO: Scientific study, show how my work integrates in the 3D city reconstruction process.
+% TODO: Find papers and topics for the scientific study (Building footprints, building heights, LOD levels, road rendering)
+% TODO: Scientific study, disposition
     % TODO: Scientific study: Potential application: A method of generating cities on the cheap
@@
 
+Another crucial concept in city modelling is the Level Of Detail (LOD). LOD is a definition of a series of different 3D representation of the same data, at different detail levels.
+The LOD chosen for a particulary application suggests how the data has been aqquired and modelled.
+The CityGML standard divides the LOD representation of 3D models into three classes: A B and C.
+LOD-A models are renderings of the building footprint itself, or building blocks that have been extruded from the footprint to give the building height.
+LOD-B models are building blocks with simplifed roof shapes, and LOD-C models include features such as doors, windows, alcoves, building decor, indoor features or vegetation.
+\textbf{(Biljecki, Ledoux and Stoter, 2016)}
 We may also chose what LOD class(es) to use depending on data availablity and application.
-With the current means of data aquisition being limited to building footprints, using OSM data we can only produce CityGML 2.0 LOD-B-models, that is building blocks with simplifed roof shapes (LOD-A being a rendering of the building footprint itself, or building blocks extruded from the footprint to give the building height).
-\textbf{(Biljecki, Ledoux and Stoter, 2016)}
-If we wish to aqquire building assets at a LOD-C level of detail, including features such as doors, windows, alcoves, building decor, indoor features or vegetation, we would have to complete the data with other sources, or rely on procedural methods to populate.
+With the current means of data aquisition being limited to building footprints, using OSM data we can only produce CityGML 
+If we wish to aqquire building assets at a LOD-C level, we would have to complete the data with other sources, or rely on procedural methods to populate.
 Such procedural techniques have long been available.

### 2020-05-16 report/report.tex
Update procedural modeling section, scientific study

@@
 Such procedural techniques have long been available.
-\textbf{(Mũller et al, 2006)} present how to accomplish this using shape grammar, a method for describing the possible relations between 
+\textbf{(Mũller et al, 2006)} present how to accomplish this using shape grammar: a method for iteratively constructing 3D models based in a hierarchical set of rules.
+By starting with a crude volumetric model of a building and iteratively remodeling it from a set of shape grammars, a procedural model can structure the façade, add texture and add details for windows, doors and ornaments, and so on.
+Shape grammars often work from a databank of 3D models and a set of rules for how these can be pieced together.
+This creates a possible generative space of many different models that can be created from the same data and rule set, with the induction of random fluctuations.
 For simple presentation in e.g. animation and videogames or even urban planning software, this could very well be enough.
+The problem is that these features will rarely match real life.
 As long as one can be satisified with looking at rough representations of buildings and does not require additional knowledge about minor features.

### 2020-05-16 report/report.tex
Update mention of previous activity in the field

@@
 
-So far there have been a lot of research in modelling of 3D road networks, but these projects have often focused on the road networks designated for cars and public transit, due to environmental concerns.
+So far there have been a lot of research in modelling of 3D road networks, but these projects have often focused on the road networks designated for cars and public transit, as a part of the sustainable mobility paradigm and the topic of sustainable cities.
+Also the rise of investigation into self-propelled cars and carpools.
+The study of energy efficient mobility is only in its emerging phase.
 \textbf{(Stojanovski, 2020)}
@@
 
+\bibitem{stojanovski-19}
+%TODO
+Stojanovski, T.
+2019
+Typo-morphology of transportation - Looking at historical development and multimodal futures of Swedish streets and roads
+
 \bibitem{stojanovski-20}

### 2020-05-16 report/report.tex
Update todo
Study on multimodal data, update outlook of the field

@@
 Also the rise of investigation into self-propelled cars and carpools.
-The study of energy efficient mobility is only in its emerging phase.
-\textbf{(Stojanovski, 2020)}
+The study of energy efficient mobility such as walking and cycling is only in its emerging phase. \textbf{(Stojanovski, 2019)}
 
@@
     % TODO: Virtual city picture
-    % TODO: read Biljecki
     % TODO: read Vanega
@@
 % Scietific study
-% TODO: Find papers and topics for the scientific study (Building footprints, building heights, LOD levels, road rendering)
+% TODO: Find papers and topics for the scientific study (Building footprints, building heights, LOD levels, road rendering, terrain (Procedural roads))
 % TODO: Scientific study, disposition
@@
 
+The fact is that much of the morpological information is not present in the OSM dataset.
+Urban architects often work morpological structures where a single road can be comprised of many different sections with different purposes.
+A single inner-city street for example, often contains sidewalks 
+All of a certain placement and dimension that determines what type of traffic it can serve.
+This type of representation is known as a multimodal structures, due to the possibility of incorporating mulitple modes of transportation.
+But may even describe the building facades and distances between curbline and building lot.
+\textbf{Stojanovski, 2019}
+This type of information could in theory be generated procedurally using shape grammars and some sort of wayfinding to create logical routes for different types of traffic.
+But for urban planners requiring an exact morphological representation of some specification, this complementary data will have to be obtained from another dataset.
+
 \begin{thebibliography}{9}

### 2020-05-16 report/report.tex
Update todo
Add main conclusion to the scientific study

@@
 
-% Thursday:
+% Writing:
 % TODO: Write introduction section
+    % TODO: read Vanega
+% TODO: Write Collision study implementation section
+    % TODO: Edge distance algorithm, algorithm
+
+% Pictures
+% Introduction
     % TODO: Figure showing the unique characteristics of footpaths vs other road types (OSM Screeshot)
     % TODO: Virtual city picture
-    % TODO: read Vanega
-% TODO: Write Collision study implementation section
+% Collision study implementation
     % TODO: Figure detailing the steps in the point-matching algorithm
-    % TODO: Edge distance algorithm, outline, algorithm + picture
+    % TODO: Edge distance algorithm, picture
 
@@
     % TODO: Scientific study, disposition
-    % TODO: Scientific study: Potential application: A method of generating cities on the cheap
-% TODO: Scientific study: A set of recommendations.
 
@@
 
+The main conclusion is that generating virtual cities purely from OpenStreetMap data is a perfectly feasibile procedure in the media industry, with the inclusion of procedural methods such as shape grammars to generate detail.
+It will be a cheap method that reuires minimum effort in terms of source data collection. And the shape grammar method allows for a great deal of scalability.
+The level of detail of individual models and the composition as a whole can be fitted to the scope of the project and the team size.
+In the area of urban design however, urban planners often require complementary data to build morphological representations of the data that they are working with.
+The feasibility of creating virtual cities that are of use to urban planners will depend mainly on the availablity of that complementary data.
+
 \begin{thebibliography}{9}

### 2020-05-17 report/report.tex
Add note about GIS quality, biljecki

@@
 
-This project assesses the feasibility of reconstructing a city from OpenStreetMap (OSM) road and building data, and will focus on the feasibility of including pedestrian footpaths, by focusing on OSM road and building data in the Stockholm metropolitan area.
+Currently, when modeling
+3D GIS datasets without errors are rare.
+While previously \textbf{(Biljecki et al, 2016)} note that there are coming more and more sophisticated methods for automatically correcting the most common errors.
+Among the most common errors: Missing or invalid geometry, invalid semantics.
+GIS validation har only been a topis of research lately, and many such datasets were usually obtained years ago, and creating error-free datasets was neither possible due to the lack of validation techniques, nor a high priority.
+
+The idea of gathering geodata from OpenStreetMap (OSM), where the temporal aspect of data accuracy is perhaps mitigated by the crowd contribution nature.
+There have already been a few services (*** Name a few services + links)
+but such services are usually limited in the type of data that they can provide.
+Usually they focus on obtaining either road networks or buildings separately, and in the case of road networks they usually focus on car road networks (why?
+Driving games?***), and roads designated for energy efficient mobility are left out.
+
+This project assesses the feasibility of reconstructing a city from OSM road and building data, and will focus on the feasibility of including pedestrian footpaths, by focusing on OSM road and building data in the Stockholm metropolitan area.
 
@@
 
+\bibitem{biljecki-16}
+%TODO
+Biljecki, F
+2016
+THE MOST COMMON GEOMETRIC AND SEMANTIC ERRORS IN CITYGML DATASETS
+
 \bibitem{douglas-73}

### 2020-05-17 report/report.tex
Add Vanegas reference
Introduction to virtual cities by Vanegas

@@
 
-Description of the field of city reproduction \textbf{(Parish and Müller, 2001)}.
-
-Picture of a virtual city.
-
-Description of a virtual city and its contents.
-Consisting of road network, buildings, details.
-Roads have different types.
+Cities consist of complex collections of buildings, blocks, parcels and neighbourhoods which are all interconnected via a street network, where individual streets can serve different types of traffic.
+Virtual city reproduction has a growing number of applications, including visualization, urban planning and entertainment in the form of video games and 3D animation.
+Traditionally, modelling virtual cities is a labour-intensive and time-consuming task.
+In recent years there has been major research work into the field of procedural content generation, meaning automatic or semi automatic generation of 3D models
+Such methods can indeed be used in the field of virtual city reproduction to reduce the labour required to reproduce accurate 3D representations of cities.
+But the field of virtual city reproduction comes with its own host of challenges.
+Urban environments are typically dense and large, spanning a few to hundreds of square kilometers, and are influenced by variables that are hard to quantify, including land policies, government plans, population changes and transportation infrastructure.
+\textbf{(Vanegas et al, 2010)}.
 
@@
 It would be labour intensive to recreate these .
-Procedural models for city reproduction instead rely on programmatic solutions for converting map data to 3D road and building models, and composing them to a 3D world that the designer may continue to fine-tweak.
+Procedural models for city reproduction instead rely on programmatic solutions for converting map data to 3D road and building models, and composing them to a 3D world that the designer may continue to fine-tweak through interfaces fo interactive editing.
 \textbf{(Müller et al, 2006)}
@@
 % TODO: Write introduction section
-    % TODO: read Vanega
 % TODO: Write Collision study implementation section
@@
 
+\bibitem{vanegas-10}
+%TODO
+Vanegas, C.
+2010
+Modelling the Appearance and Behaviour of Urban Spaces
+
 \bibitem{van-oort-06}

### 2020-05-17 report/report.tex
Update todo
Fix remaining references

@@
 
-% Writing:
-% TODO: Write introduction section
-% TODO: Write Collision study implementation section
-    % TODO: Edge distance algorithm, algorithm
-
+% Sunday
 % Pictures
@@
     % TODO: Virtual city picture
+% Scietific study
+    % TODO: Find papers and topics for the scientific study (building heights, road rendering, terrain (Procedural roads))
+    % TODO: Scientific study, disposition
+% TODO: Edge distance algorithm, algorithm
+
+% Monday
+% TODO: Write introduction section
+% TODO: Write Collision study implementation section
+% TODO: Write Scientific study
 % Collision study implementation
@@
     % TODO: Edge distance algorithm, picture
-
-% Scietific study
-    % TODO: Find papers and topics for the scientific study (Building footprints, building heights, LOD levels, road rendering, terrain (Procedural roads))
-    % TODO: Scientific study, disposition
 
@@
 \bibitem{biljecki-16}
-%TODO
-Biljecki, F
-2016
-THE MOST COMMON GEOMETRIC AND SEMANTIC ERRORS IN CITYGML DATASETS
+Biljecki, F., Ledoux, H., Du, X., Stoter, J., Soon, K., and Khoo, V. (2016). The most common geometric and semantic errors in CityGML datasets. 10.5194/isprs-annals-IV-2-W1-13-2016.
 
@@
 \bibitem{stojanovski-19}
-%TODO
 Stojanovski, T.
-2019
-Typo-morphology of transportation - Looking at historical development and multimodal futures of Swedish streets and roads
+(2019).
+Typo-morphology of transportation -Looking at historical development and multimodal futures of Swedish streets and roads.
+Conference paper.
+ISUF 2019 – ХХVI International Seminar on Urban Form 2019, At Nicosia, Cyprus
 
@@
 \bibitem{vanegas-10}
-%TODO
-Vanegas, C.
-2010
-Modelling the Appearance and Behaviour of Urban Spaces
+Vanegas, C., Aliaga, D., Mueller, P., Waddell, P., Watson, B.
+and Wonka, P.
+(2010).
+Modeling the Appearance and Behavior of Urban Spaces.
+Computer Graphics Forum.
+29.
+25-42.
+10.1111/j.1467-8659.2009.01535.x.
 

### 2020-05-17 report/report.tex
Add figure comparing footpaths and residential streets

@@
 Description of the unique character of footpaths vs other types of roads (pictures).
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_map_residential_footpath_compare}
+    \caption{Example segments of the Stockholm metropolitan area with highlighted residential streets (left) and footpaths (right).
+Here can be seen the unique character of footpaths.
+Walkways in parks appear to grow much more organically, and pavements form rough grid systems with organic branches that bridge over city blocks.}
+    \label{fig:osm-slu-map}
+\end{figure}
 

### 2020-05-17 report/report.tex
Add virtual city image

@@
 
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_virtual_city}
+    \caption{Real-world cities are complex collections of buildings, blocks and road networks that interconnect them all.
+Creating 3D virtual cities using methods of procedual generation is an emerging research field that comes with its own set of challenges.}
+    \label{fig:virtual-city}
+\end{figure}
+
 Cities consist of complex collections of buildings, blocks, parcels and neighbourhoods which are all interconnected via a street network, where individual streets can serve different types of traffic.
@@
 Walkways in parks appear to grow much more organically, and pavements form rough grid systems with organic branches that bridge over city blocks.}
-    \label{fig:osm-slu-map}
+    \label{fig:residential-footpath-map}
 \end{figure}

### 2020-05-17 report/report.tex
Fix non-compliant unicode characters
Add note about building elevation, Biljecki

@@
 
-Depending on the level of data we have access to, we can then choose what level of representation we would like for our application (Todor).
+Building elevation. Occasinally buildings in OSM may be tagged, but the norm is that no elevation data is available.
+\textbf{(Biljecki, Ledoux and Stoter, 2017)} presented a method for inferring building height exclusively from the building footprint and surrounding footprints.
+Surprisingly, their model was shown to be accurace within a meter on average.
 
@@
 LOD-B models are building blocks with simplifed roof shapes, and LOD-C models include features such as doors, windows, alcoves, building decor, indoor features or vegetation.
-\textbf{(Biljecki, Ledoux and Stoter, 2016)}
+\textbf{(Biljecki et al, 2016)}
 We may also chose what LOD class(es) to use depending on data availablity and application.
@@
 
+\bibitem{biljecki-17}
+Biljecki, F., Ledoux, H. and Stoter, J. (2017). Generating 3D city models without elevation data. Computers Environment and Urban Systems. 64. 1-18. 10.1016/j.compenvurbsys.2017.01.001. 
 \bibitem{douglas-73}
@@
 Conference paper.
-ISUF 2019 – ХХVI International Seminar on Urban Form 2019, At Nicosia, Cyprus
+ISUF 2019 -XXVI International Seminar on Urban Form 2019, At Nicosia, Cyprus
 

### 2020-05-17 report/report.tex
Add note about terrain elevation, Galin

@@
 The same methods can be applied to import building footprint data from OSM or some, and the same metrics can be applied to assess the data accuracy. \textbf{(Fan et al, 2014)}
-When we have the fundamental procedure for accuratly importing building footprints without compromising the data integritey, we can proceed with building modeling.
-\textbf{(Müller et al, 2006)} Other challenges include infering or guessing building height.
-() have presented a method for inferring building height from the building footprint alone, with surprising accuracy.
+When we have the fundamental procedure for accuratly importing building footprints without compromising the data integritey, we can proceed with building modeling. \textbf{(Müller et al, 2006)}
 
-Building elevation. Occasinally buildings in OSM may be tagged, but the norm is that no elevation data is available.
+Terrain elevation. Elevation data is available from other sources, but not by OSM.
+\textbf{(Galin et al, 2010)} present a geometrical model for placing a 3d road asset onto a non-flat terrain surface.
+
+Other challenges include infering or guessing building height.
+Occasinally buildings in OSM may be tagged, but the norm is that no elevation data is available.
 \textbf{(Biljecki, Ledoux and Stoter, 2017)} presented a method for inferring building height exclusively from the building footprint and surrounding footprints.
@@
 
+\bibitem{galin-10}
+Galin, E., Peytavie, A., Maréchal, N., and Guérin, E. (2010). Procedural Generation of Roads. Computer Graphics Forum. 29. 429 - 438. 10.1111/j.1467-8659.2009.01612.x.
+
 \bibitem{haklay-10}

### 2020-05-17 report/report.tex
Update todo
Refine introduction section

@@
     \caption{Real-world cities are complex collections of buildings, blocks and road networks that interconnect them all.
-Creating 3D virtual cities using methods of procedual generation is an emerging research field that comes with its own set of challenges.}
+Creating 3D virtual cities using methods of procedual generation is an emerging research field that comes with its own set of challenges.
+This project deals with generating virtual cities from OpenStreetMap data, and the technical challenges therein.}
     \label{fig:virtual-city}
@@
 
-Cities consist of complex collections of buildings, blocks, parcels and neighbourhoods which are all interconnected via a street network, where individual streets can serve different types of traffic.
-Virtual city reproduction has a growing number of applications, including visualization, urban planning and entertainment in the form of video games and 3D animation.
+This project assesses the feasibility of reconstructing a city from OpenStreetMap (OSM) road and building data, and will focus on the feasibility of including pedestrian footpaths, by focusing on OSM road and building data in the Stockholm metropolitan area.
+
+Virtual cities has in recent years been a popular research topic in the field of computer visualization.
+The idea of recreating real world cities in 3D applications has a growing number of applications, including visualization and modelling, urban planning and architecture, as well as entertainment in the form of video games and 3D animation.
 Traditionally, modelling virtual cities is a labour-intensive and time-consuming task.
-In recent years there has been major research work into the field of procedural content generation, meaning automatic or semi automatic generation of 3D models
-Such methods can indeed be used in the field of virtual city reproduction to reduce the labour required to reproduce accurate 3D representations of cities.
-But the field of virtual city reproduction comes with its own host of challenges.
+However, in recent years there has been major research work in the field of procedural content generation, meaning automatic or semi automatic generation of 3D models without the direct involvment of a 3D artist.
+Such methods can indeed be used in the field of virtual city reproduction to reduce the labour required to reproduce accurate 3D representations of cities, but the field of procedual city reproduction comes with its own host of challenges.
 Urban environments are typically dense and large, spanning a few to hundreds of square kilometers, and are influenced by variables that are hard to quantify, including land policies, government plans, population changes and transportation infrastructure.
@@
 
-Applications:
-    Navigating agents or pedestrian flows.
-    Using a road network to creates an underlying navmesh
-    This process is analytical and relies on the availablity and accurate representation of city maps.
-    \textbf{(Parish and Müller, 2001)}
-
-Description of different levels of subdivision. What levels of subdivision are we looking at? Picture, ref to Todor's paper.
-Viewpoints can range from modelling individual roads (***find reference) and building facades \textbf{(Stojanovski, 20***)} to looking at entire city parcels and how they connect with each other (***find reference).
-
-Description of procedural modelling of cities
-Importing geodata from online databases and converting it to a 3D city representation.
-It would be labour intensive to recreate these .
-Procedural models for city reproduction instead rely on programmatic solutions for converting map data to 3D road and building models, and composing them to a 3D world that the designer may continue to fine-tweak through interfaces fo interactive editing.
+Cities consist of complex collections of buildings, blocks, parcels and neighbourhoods which are all interconnected via a street network, where individual streets can serve different types of traffic.
+These features are commonly represented as geodata in map databases, and processed and organized into visualization-ready data by a Geographic Information System (GIS).
+In practise, cities are often made up out of many square kilometers worth of geodata, and it would be highly labour intensive to recreate all the features in 3D.
+Instead, procedural models for city reproduction instead rely on programmatic solutions for converting map data to 3D road and building models, and composing them to a 3D world that the designer may continue to fine-tweak through interfaces fo interactive editing.
+At the very base of this process is the procedural layout of road networks in 3D,
 \textbf{(Müller et al, 2006)}
-
-At the base of this process is the procedural layout of road networks in 3D.
-\textbf{(Müller et al, 2006)}
-
-So far there have been a lot of research in modelling of 3D road networks, but these projects have often focused on the road networks designated for cars and public transit, as a part of the sustainable mobility paradigm and the topic of sustainable cities.
-Also the rise of investigation into self-propelled cars and carpools.
+and so far there have been a lot of research in modelling of 3D road networks, but these projects have often focused on the road networks designated for cars and public transit, as a part of the sustainable mobility paradigm and the topic of sustainable cities, particularily due to the interest in self-propelled cars.
 The study of energy efficient mobility such as walking and cycling is only in its emerging phase. \textbf{(Stojanovski, 2019)}
-
-Description of the unique character of footpaths vs other types of roads (pictures).
 
@@
 
-Currently, when modeling
-3D GIS datasets without errors are rare.
-While previously \textbf{(Biljecki et al, 2016)} note that there are coming more and more sophisticated methods for automatically correcting the most common errors.
-Among the most common errors: Missing or invalid geometry, invalid semantics.
-GIS validation har only been a topis of research lately, and many such datasets were usually obtained years ago, and creating error-free datasets was neither possible due to the lack of validation techniques, nor a high priority.
+When it comes to the quality of available data, it turns out that in practise, 3D GIS datasets without errors are rare.
+Among the most common errors are missing or invalid geometry, and invalid semantics.
+However, \textbf{(Biljecki et al, 2016)} note that there are coming more and more sophisticated methods for automatically correcting the most common errors.
+GIS validation and correction is only an emerging field of research.
+and many such datasets were usually obtained years ago, and creating error-free datasets was neither possible due to the lack of validation techniques, nor a high priority.
 
-The idea of gathering geodata from OpenStreetMap (OSM), where the temporal aspect of data accuracy is perhaps mitigated by the crowd contribution nature.
-There have already been a few services (*** Name a few services + links)
-but such services are usually limited in the type of data that they can provide.
-Usually they focus on obtaining either road networks or buildings separately, and in the case of road networks they usually focus on car road networks (why?
-Driving games?***), and roads designated for energy efficient mobility are left out.
-
-This project assesses the feasibility of reconstructing a city from OSM road and building data, and will focus on the feasibility of including pedestrian footpaths, by focusing on OSM road and building data in the Stockholm metropolitan area.
+A possible way of mitigating this problem is looking towards gathering geodata from volunteer-based geoservices such as OpenStreetMap (OSM), where the temporal aspect of data accuracy is perhaps more assured by the crowd contribution nature.
+There are already a few such services that use OSM data to generate 3D geodata visualizations for data visualization and video game applications.
+A few prominent examples are Mantle
+\footnote{Mantle homepage. Accessed 17-05-2020. https://www.mantle.tech/}
+and Mapbox
+\footnote{Mapbox homepage. Accessed 17-05-2020. https://www.mapbox.com/}.
+Such services are however usually limited in the accuracy and type of data that they can provide.
+While they can provide both builing footprints and road networks, they usually provide road data with no respect to overlap between features.
+Collision between road meshes in 3D will lead to strange-looking mesh intersections and inaccuracies that modelers often want to avoid.
+In particular, streets and roads designated for energy efficient mobility, such as footpaths, are much more organic and dense in their nature than main streets, and thus the risk of feature collision may be higher.
 
@@
 % Sunday
-% Pictures
-% Introduction
-    % TODO: Figure showing the unique characteristics of footpaths vs other road types (OSM Screeshot)
-    % TODO: Virtual city picture
-% Scietific study
-    % TODO: Find papers and topics for the scientific study (building heights, road rendering, terrain (Procedural roads))
-    % TODO: Scientific study, disposition
-% TODO: Edge distance algorithm, algorithm
+% TODO: Write Collision study implementation section
+% TODO: Write Scientific study
 
 % Monday
-% TODO: Write introduction section
-% TODO: Write Collision study implementation section
-% TODO: Write Scientific study
+% TODO: Edge distance algorithm, algorithm
 % Collision study implementation

### 2020-05-17 report/report.tex
Remove redundant OSM shortening references

@@
 
-A possible way of mitigating this problem is looking towards gathering geodata from volunteer-based geoservices such as OpenStreetMap (OSM), where the temporal aspect of data accuracy is perhaps more assured by the crowd contribution nature.
+A possible way of mitigating this problem is looking towards gathering geodata from volunteer-based geoservices such as OSM, where the temporal aspect of data accuracy is perhaps more assured by the crowd contribution nature.
 There are already a few such services that use OSM data to generate 3D geodata visualizations for data visualization and video game applications.
@@
 
-This project has dealt with the feasibility of including footpath data when generating 3D representations of OpenStreetMap (OSM) data, and preserving the integrity of the 3D road network, avoiding overlap and collision between road meshes.
+This project has dealt with the feasibility of including footpath data when generating 3D representations of OSM data, and preserving the integrity of the 3D road network, avoiding overlap and collision between road meshes.
 The project falls within the broader topic of geodata extraction and modelling, primarily of roads and building parcels.

### 2020-05-18 report/report.tex
Update todo
Update scietific study

@@
 
-% Sunday
+% Monday
 % TODO: Write Collision study implementation section
-% TODO: Write Scientific study
-
-% Monday
 % TODO: Edge distance algorithm, algorithm
@@
     % TODO: Figure detailing the steps in the point-matching algorithm
-    % TODO: Edge distance algorithm, picture
 
@@
 
-\section{Scientific study: Where does this fit in the general Scietific scope of things}
+\section{Scientific field study}
 
-The same methods can be applied to import building footprint data from OSM or some, and the same metrics can be applied to assess the data accuracy. \textbf{(Fan et al, 2014)}
-When we have the fundamental procedure for accuratly importing building footprints without compromising the data integritey, we can proceed with building modeling. \textbf{(Müller et al, 2006)}
+We now have a firm idea of the technical feasibility of generating virtual cities from OpenStreetMap-data, including different types of road networks.
+We have however yet to touch upon the importance for road network modelling in the broader context of 3D city reproduction.
+Where exactly in the pipeline of reconstructing a virtual city does this happen, and what are the remaining challenges?
+This will be the goal of this final scientific study: to place everything we have learnt so far in the broader context of GIS and city modeling, and the respective research fields.
 
-Terrain elevation. Elevation data is available from other sources, but not by OSM.
-\textbf{(Galin et al, 2010)} present a geometrical model for placing a 3d road asset onto a non-flat terrain surface.
+As has been mentiond in the introduction to this report, procedural generaion of road networks is a basal requirement for quick and low-effort modeling virtual cities.
+Since the road network ties together the city, and forms the grid that divides the lots, quarters and neighbourhoods, it make sense that we can only continue with the remaining challenges once we can a robust procedure for extracting road geodata and generating the 3D represention of it without compromising the data integrity.
+The same is true for building modelling.
+It turns out that the same methods that have been used in this project to obtain and qualify road network data can be applied to building footprint data from OSM. \textbf{(Fan et al, 2014)}
+And only when we have such a fundamental procedure for accuratly importing building footprints without compromising the data integritey, we can proceed with the remaining challenges in building modeling.
+\textbf{(Müller et al, 2006)}
 
-Other challenges include infering or guessing building height.
+Such challenges, which have not been mentioned up until this point, include infering or guessing building height, since OSM does not provide data about building elevation.
 Occasinally buildings in OSM may be tagged, but the norm is that no elevation data is available.
+There are however very sophisticated methods from guessing the building height after looking at reference datasets.
 \textbf{(Biljecki, Ledoux and Stoter, 2017)} presented a method for inferring building height exclusively from the building footprint and surrounding footprints.
@@
 
-Another crucial concept in city modelling is the Level Of Detail (LOD). LOD is a definition of a series of different 3D representation of the same data, at different detail levels.
-The LOD chosen for a particulary application suggests how the data has been aqquired and modelled.
+Another challenge is that of including terrain elevation when generating a city.
+Even if high resolution terrain elevation data is available from other sources, it is not provided by OSM, and will have to be obtained from a complementary data source.
+Method exist for how to make a city model comform to an underlying terrain, however.
+\textbf{(Galin et al, 2010)} presented such a geometrical model for placing a 3D road asset onto a non-flat terrain surface, in the context of procedurally generated roads.
+
+Another crucial concept in city modelling is the Level Of Detail (LOD) aspect.
+LOD is a definition of a series of different 3D representation of the same data, at different detail levels, and the LOD chosen for a particulary application suggests how the data has been aqquired and modelled.
 The CityGML standard divides the LOD representation of 3D models into three classes: A B and C.
-LOD-A models are renderings of the building footprint itself, or building blocks that have been extruded from the footprint to give the building height.
-LOD-B models are building blocks with simplifed roof shapes, and LOD-C models include features such as doors, windows, alcoves, building decor, indoor features or vegetation.
+\begin{enumerate}
+    \item LOD-A models are renderings of the building footprint itself, or rectangular building blocks that have been extruded from the footprint to give the building height.
+    \item LOD-B models are building blocks with simplifed roof shapes.
+    \item LOD-C models are LOD-B models which include features such as doors, windows, alcoves, building decor, indoor features or vegetation.
+\end{enumerate}
 \textbf{(Biljecki et al, 2016)}
-We may also chose what LOD class(es) to use depending on data availablity and application.
-With the current means of data aquisition being limited to building footprints, using OSM data we can only produce CityGML 
-If we wish to aqquire building assets at a LOD-C level, we would have to complete the data with other sources, or rely on procedural methods to populate.
+We may chose what LOD class(es) to use depending on data availablity and application.
+It turns out that, with the current means of data aquisition from OSM being limited to building footprints, using OSM data we can only produce LOB-B models at best.
+If we wish to aqquire building assets at a LOD-C level, we would have to complete the data with other sources, or rely on procedural methods to populate our crude city model with additional details and building features that give the model more aestethic depth, but do not neccessarily reflect reality.
+
 Such procedural techniques have long been available.
-\textbf{(Mũller et al, 2006)} present how to accomplish this using shape grammar: a method for iteratively constructing 3D models based in a hierarchical set of rules.
+\textbf{(Mũller et al, 2006)} present how to accomplish this using Shape Grammar: a method for automatically and iteratively constructing 3D models based in a hierarchical set of rules.
 By starting with a crude volumetric model of a building and iteratively remodeling it from a set of shape grammars, a procedural model can structure the façade, add texture and add details for windows, doors and ornaments, and so on.
-Shape grammars often work from a databank of 3D models and a set of rules for how these can be pieced together.
-This creates a possible generative space of many different models that can be created from the same data and rule set, with the induction of random fluctuations.
-For simple presentation in e.g. animation and videogames or even urban planning software, this could very well be enough.
-The problem is that these features will rarely match real life.
-As long as one can be satisified with looking at rough representations of buildings and does not require additional knowledge about minor features.
+Shape grammars often work from a databank of 3D models and a set of hierarchical rules for how these can be pieced together.
+This creates a possible generative space of many different models that can be created from the same data and rule set, and with the induction of random fluctuations, many different models can be generated from the same dataset and rules.
+For simple presentation in e.g. animation and videogames or even urban planning software, such procedurally generated cities could very well be enough.
+As said however, the big problem is that these features will rarely match real life, such models will be sufficient as long as one can be satisified with looking at rough representations of buildings and does not require additional knowledge about minor features.
+Many studies urban development have been done before that requred detailed knowledge about a city's geometry, e.g.
+for the purposes of modeling heat development (***find reference) and reducing noise (**find reference).
 
-The fact is that much of the morpological information is not present in the OSM dataset.
-Urban architects often work morpological structures where a single road can be comprised of many different sections with different purposes.
-A single inner-city street for example, often contains sidewalks 
-All of a certain placement and dimension that determines what type of traffic it can serve.
-This type of representation is known as a multimodal structures, due to the possibility of incorporating mulitple modes of transportation.
-But may even describe the building facades and distances between curbline and building lot.
+It even turns out, by looking at such studies, that a lot of research in urban development require detailed morpological information that goes beyond simply knowing the geometry of the city.
+Urban architects often work with morpological structures where a single feature, such as a city neighbourhood, can be examined from many different levels of perspective.
+A neighbourhood can be analysed from the level of its city blocks, individual lots, properties or building limits, building facades or even indoor spaces.
+Simlarily a road network can be analysed by looking at it either at its connectivity by viewing it a a simple 2D graph
+But it might as well be seen as a hierarchical structure where primary roads branch off into sub-streets. 
+The very unfortunate fact is that much of that morpological information is not present in the OSM dataset.
+Furthermore, take the fact that even a single road can be comprised of many different sections with different purposes.
+A single inner-city street for example, often contains sidewalks and lanes reserved for taxis or public transit, all of a certain placement and dimension that determines what type of traffic it can serve.
+This type of representation is known as a multimodal structures, due to the possibility of incorporating mulitple modes of transportation, but it may even describe subtilities like building facades and distances between curbline and building lot.
+OSM, while being an excellent source of city geometry, is simply not a sophisticated enough dataset to cater to all these possible areas of research.
 \textbf{Stojanovski, 2019}
-This type of information could in theory be generated procedurally using shape grammars and some sort of wayfinding to create logical routes for different types of traffic.
+For media applications, such morppological and multimodal information could in theory be generated procedurally, by using shape grammars and some sort of wayfinding to create logical routes for different types of traffic.
 But for urban planners requiring an exact morphological representation of some specification, this complementary data will have to be obtained from another dataset.
 
-The main conclusion is that generating virtual cities purely from OpenStreetMap data is a perfectly feasibile procedure in the media industry, with the inclusion of procedural methods such as shape grammars to generate detail.
-It will be a cheap method that reuires minimum effort in terms of source data collection. And the shape grammar method allows for a great deal of scalability.
+The final conclusion is that generating virtual cities purely from OpenStreetMap data is a perfectly feasibile procedure in the media industry, with the inclusion of procedural methods such as shape grammars to generate neccessary detail.
+This will be a cheap method that reuires minimum effort in terms of source data collection, and the shape grammar method even allows for a great deal of scalability.
 The level of detail of individual models and the composition as a whole can be fitted to the scope of the project and the team size.
-In the area of urban design however, urban planners often require complementary data to build morphological representations of the data that they are working with.
-The feasibility of creating virtual cities that are of use to urban planners will depend mainly on the availablity of that complementary data.
+However, in the area of urban design, urban planners often require complementary data to build morphological representations of the data that they are working with.
+OSM does generally not serve this type of data, so the feasibility of creating virtual cities that are of use to urban planners will not depend upon the accuracy of the OSM dataset, but rather on the availablity of that complementary data elsewhere.
 

### 2020-05-18 report/report.tex
Flesh out collision study implementation

@@
 
-\subsection{Feature overlap algorithm}
+\subsection{Feature Overlap algorithm}
 
-Outline of the feature overlap algorithm (Go over all features)
+The algorithm for determining feature overlap will calculate how many features and how many of their edges are intersecting with some other feature or edge.
+It will also calculate the total length of all edge which are overlapping with some other edge.
+
+The algorithm begins by iterating over all road features.
+Let $A$ be the current road feature.
 Since only road collisions are of interests, all roads were added to a queue and gone over one by one.
-
-The boundary polygon of the buildings are extracted, as well as the polylines of the roads.
+The polyline of the current road feature is extracted.
+Then, all other candidate features for collision are iterated over, and their boundary polygons (for buildings) or polylines (for roads) are extracted.
+Let $B$ be the current feature, road or building, that is being compared to $A$.
+The algorithm then iterates over each edge $e_a$ in the polyline of $A$, and each edge $e_b$ in the polyline or polygon of $B$.
+The Projected Line Segment Distance algorithm (See section 4.2.3) is used to determine the shortest distance between $e_a$ and $e_b$, and the smallest distance between any combination of edges $e_a$ and $e_b$ is taken to be the shortest distance between features $A$ and $B$.
+If this distance is smaller than the minimum width of $A$, then $A$ and $B$ are colliding.
 
 To speed up the collision checking process, a geometric hashing solution was developed.
-This way feature collision lookup could be made without comparing a feature to all other available features, but instead from a selection of the closest candidates.
-The hash bucket size was 230 meters and a feature could be in multiple hash buckets.
+In this way, feature collision lookup could be made without comparing a feature to all other available features, but instead from a selection of the closest candidates.
+The hash bucket size was 230 by 230 meters and a feature could be in multiple hash buckets.
 
-Special cases: Where two roads intersect though they share a node.
-Then this is a natural connection in the road network and should not be counted as a problematic case of feature overlap.
-Also ignored collisions between tunnels and bridges and other roads since these are of different heights.
-These were identified by looking at feature attributes in OSM.
-Any roads that were represented as polygons or multipolygons instead of plylines were also excluded from the study.
+Here a few special cases are noted:
+\begin{enumerate}
+    \item If two roads intersect even though they share a node, it is a natural connection in the road network and should not be counted as a problematic case of feature overlap.
+    \item Collisions between tunnels or bridges and other roads were ignored since these are of different heights. Brdiges and tunnels were identified by looking at feature attributes in OSM.
+    \item Any roads that were represented as polygons or multipolygons instead of ploylines were also excluded from the study. These are typically town squares or areas designated for walking.
+\end{enumerate}
 
-\subsection{Projected edge distance}
+\subsection{Projected Line Segment Distance}
 
-Outline of the minimum edge distance algorithm.
+The Projected Line Segment Distance (PLSD) algorithm is an extented version of the Line Segment Distance (LSD) algorithm.
+Whereas LSD simply measures the distance between the two absolute closest points on two line segments, PLSD will yield the distance of the shortest vector $v$ between line segments $e_a$ and $e_b$, with the condition that $v$ is perpendicular to the tangent of $e_a$.
+In the Feature Overlap algorithm, using the same notation as in section 4.2.2, PLSD takes $e_a$ to be the current edge of road feature $A$, and $e_b$ to be the current edge of feature $B$ (road or building).
+See Algorithm 2 for a detailed view of the PLSD algorithm.
 
@@
 % Monday
-% TODO: Write Collision study implementation section
 % TODO: Edge distance algorithm, algorithm

### 2020-05-18 report/report.tex
Add input field to Algorithm 1
Add first draft of algorithm 2

@@
 \KwResult{Write here the result }
+    \textbf{Input:} PointList\\
     // Find the point with the maximum distance\;
@@
 See Algorithm 2 for a detailed view of the PLSD algorithm.
+
+\begin{algorithm}[H]
+\SetAlgoLined
+\KwResult{A float representing the shortests distance between $A$ and $B$}
+    \textbf{Input:} Line segment $A$: [$\vec{a_1}$, $\vec{a_2}$], Line segment $B$: [$\vec{b_1}$, $\vec{b_2}$]\\
+    // First assume that the points of $A$ and $B$ are ordered so that $\vec{a_1}$ is closer to $\vec{b_1}$ than to $\vec{b_2}$.
+
+    // Check if the line segments $A$ and $B$ intersect, if so, the distance between them is 0.
+    $x_s$ = lineLineIntersection($A$, $B$)\;
+    \If{$x_s$ is not None} {
+        \Return{0}
+    }
+
+    // Get the line-line intersection between the lines spanned by $A$ and $B$. Note that if an intersection exists, it must lie outside of both $A$ and $B$.
+    $x$ = lineLineIntersection($A$, $B$)\;
+
+    // Case 1: No intersection. The lines are parallel.
+    \If{$x$ is None} {
+        \Return{perpendicularLineDistance($A$, $B$)}\;
+    }
+
+    // Get the projections of $\vec{b_1}$ and $\vec{b_2}$ on the line spanned by $A$:
+    $\vec{d_1}$ = linearProjection($\vec{b_1}$, $A$)\;
+    $\vec{d_2}$ = linearProjection($\vec{b_2}$, $A$)\;
+
+    // Case 2: The projection of $B$ onto $A$ is completely disjoint from $A$ with no common points. In this case, no perpendicular distance can be inferred.
+    \If{$\vec{d_1}-\vec{a_1} \cdot \vec{d_1}-\vec{a_2} > 0$ or $\vec{d_2}-\vec{a_1} \cdot \vec{d_2}-\vec{a_2} > 0$} {
+        \Return{None}\;
+    }
+
+    // Case 3: $x$ lies between $\vec{a_1}$ and $\vec{a_2}$.
+    \If{$\vec{x}-\vec{a_1} \cdot \vec{a_2}-\vec{x} > 0$} {
+        // Project the endpoints of $B$ on $A$, to find out which endpoint is closer to the intersection
+        $d_{pl1}$ = pointLineDistance($\vec{b_1}$, $A$)\;
+        $d_{pl2}$ = pointLineDistance($\vec{b_2}$, $A$)\;
+        \eIf{$d_{pl1} < d_{pl2}$} {
+            \Return{$d_{pl1}$}\;
+        } {
+            \Return{$d_{pl2}$}\;
+        }
+    }
+
+    \eIf{|$x$-$\vec{a_1}$| < |$x$-$\vec{a_2}$|} {
+        // Case 4: $x$ is to the left of $\vec{a_1}$
+        $\vec{a}$ = $\vec{a_1}$\;
+        $\vec{b}$ = $\vec{b_1}$\;
+    } {
+        // Case 5: $x$ is to the right of $\vec{a_2}$
+        $\vec{a}$ = $\vec{a_2}$\;
+        $\vec{b}$ = $\vec{b_2}$\;
+    }
+
+    $\vec{p}$ = linearProjection($\vec{b}-\vec{x}$, $\vec{a}-\vec{x}$)\;
+    \eIf{|$\vec{p}$| < |$\vec{a}$-$\vec{x}$|} {
+        $\vec{k}$ = $\vec{d}$\;
+    } {
+        $\vec{k}$ = $\vec{a}-\vec{x}$\;
+    }
+
+    $\theta = \arccos{|\vec{d}|/|\vec{b}-\vec{x}|}$\;
+    \Return{$\tan{\theta} \cdot |\vec{k}|$}
+
+    \caption{Projected Line Segment Distance}
+\end{algorithm}
 

### 2020-05-18 report/report.tex
Newlines for algorithm comments, ensure x is a vector everywhere

--- 2020-05-18 report/report.tex
@@
 \KwResult{Write here the result }
-    // Find the point with the maximum distance\;
+    \textbf{Input:} PointList\\
+    // Find the point with the maximum distance\\
     $d_{max}$ = 0\;
@@
 
-    // If max distance is greater than epsilon, recursively simplify\;
+    // If max distance is greater than epsilon, recursively simplify\\
     \eIf{$d_{max}$ $>$ $\epsilon$} {
-        // Recursive call\;
+        // Recursive call\\
         recResults1[] = Douglas-Peucker(PointList[1...$index$], $\epsilon$)\;
@@
 
-        // Build the result list\;
+        // Build the result list\\
         ResultList[] = {recResults1[1...length(recResults1) - 1], recResults2[1...length(recResults2)]}\;
@@
 See Algorithm 2 for a detailed view of the PLSD algorithm.
+
+\begin{algorithm}[H]
+\SetAlgoLined
+\KwResult{A float representing the shortests distance between $A$ and $B$}
+    \textbf{Input:} Line segment $A$: [$\vec{a_1}$, $\vec{a_2}$], Line segment $B$: [$\vec{b_1}$, $\vec{b_2}$]\\
+    // First assume that the points of $A$ and $B$ are ordered so that $\vec{a_1}$ is closer to $\vec{b_1}$ than to $\vec{b_2}$.\\
+
+    // Check if the line segments $A$ and $B$ intersect, if so, the distance between them is 0.\\
+    $\vec{x_s}$ = lineLineIntersection($A$, $B$)\;
+    \If{$\vec{x_s}$ is not None} {
+        \Return{0}
+    }
+
+    // Get the line-line intersection between the lines spanned by $A$ and $B$. Note that if an intersection exists, it must lie outside of both $A$ and $B$.\\
+    $\vec{x}$ = lineLineIntersection($A$, $B$)\;
+
+    // Case 1: No intersection. The lines are parallel.\\
+    \If{$\vec{x}$ is None} {
+        \Return{perpendicularLineDistance($A$, $B$)}\;
+    }
+
+    // Get the projections of $\vec{b_1}$ and $\vec{b_2}$ on the line spanned by $A$:\\
+    $\vec{d_1}$ = linearProjection($\vec{b_1}$, $A$)\;
+    $\vec{d_2}$ = linearProjection($\vec{b_2}$, $A$)\;
+
+    // Case 2: The projection of $B$ onto $A$ is completely disjoint from $A$ with no common points. In this case, no perpendicular distance can be inferred.\\
+    \If{$\vec{d_1}-\vec{a_1} \cdot \vec{d_1}-\vec{a_2} > 0$ or $\vec{d_2}-\vec{a_1} \cdot \vec{d_2}-\vec{a_2} > 0$} {
+        \Return{None}\;
+    }
+    \newpage
+
+    // Case 3: $\vec{x}$ lies between $\vec{a_1}$ and $\vec{a_2}$.\\
+    \If{$\vec{x}-\vec{a_1} \cdot \vec{a_2}-\vec{x} > 0$} {
+        // Project the endpoints of $B$ on $A$, to find out which endpoint is closer to the intersection\\
+        $d_{pl1}$ = pointLineDistance($\vec{b_1}$, $A$)\;
+        $d_{pl2}$ = pointLineDistance($\vec{b_2}$, $A$)\;
+        \eIf{$d_{pl1} < d_{pl2}$} {
+            \Return{$d_{pl1}$}\;
+        } {
+            \Return{$d_{pl2}$}\;
+        }
+    }
+
+    \eIf{|$\vec{x}$-$\vec{a_1}$| < |$\vec{x}$-$\vec{a_2}$|} {
+        // Case 4: $\vec{x}$ is to the left of $\vec{a_1}$\\
+        $\vec{a}$ = $\vec{a_1}$\;
+        $\vec{b}$ = $\vec{b_1}$\;
+    } {
+        // Case 5: $\vec{x}$ is to the right of $\vec{a_2}$\\
+        $\vec{a}$ = $\vec{a_2}$\;
+        $\vec{b}$ = $\vec{b_2}$\;
+    }
+
+    $\vec{p}$ = linearProjection($\vec{b}-\vec{x}$, $\vec{a}-\vec{x}$)\;
+    \eIf{|$\vec{p}$| < |$\vec{a}$-$\vec{x}$|} {
+        $\vec{k}$ = $\vec{d}$\;
+    } {
+        $\vec{k}$ = $\vec{a}-\vec{x}$\;
+    }
+
+    $\theta = \arccos{|\vec{d}|/|\vec{b}-\vec{x}|}$\;
+    \Return{$\tan{\theta} \cdot |\vec{k}|$}
+
+    \caption{Projected Line Segment Distance}
+\end{algorithm}
 

### 2020-05-18 report/report.tex
Split up algorithm 2

@@
 In the Feature Overlap algorithm, using the same notation as in section 4.2.2, PLSD takes $e_a$ to be the current edge of road feature $A$, and $e_b$ to be the current edge of feature $B$ (road or building).
-See Algorithm 2 for a detailed view of the PLSD algorithm.
+See Algorithms 2 and 3 together for a detailed view of the PLSD algorithm.
 
@@
 
-    // Get the line-line intersection between the lines spanned by $A$ and $B$. Note that if an intersection exists, it must lie outside of both $A$ and $B$.\\
+    // Get the line-line intersection between the lines spanned by $A$ and $B$. If an intersection exists, it must lie outside of both $A$ and $B$.\\
     $\vec{x}$ = lineLineIntersection($A$, $B$)\;
 
+    // Now that we have the intersection point, call the helper method to infer the shortest distance between $A$ and $B$.\\
+    \Return{PLSD-Helper($A$, $B$, $\vec{x}$)}\;
+
+    \caption{Projected Line Segment Distance}
+\end{algorithm}
+
+.\\
+
+\begin{algorithm}[H]
+\SetAlgoLined
+    \textbf{Input:} Line segments $A$, $B$, intersection point $\vec{x}$]\\
     // Case 1: No intersection. The lines are parallel.\\
@@
 
-    // Case 2: The projection of $B$ onto $A$ is completely disjoint from $A$ with no common points. In this case, no perpendicular distance can be inferred.\\
-    \If{$\vec{d_1}-\vec{a_1} \cdot \vec{d_1}-\vec{a_2} > 0$ or $\vec{d_2}-\vec{a_1} \cdot \vec{d_2}-\vec{a_2} > 0$} {
+    // Case 2: The projection of $B$ onto $A$ is completely disjoint from $A$ with no common points. No perpendicular distance can be inferred.\\
+    \If{$(\vec{d_1}-\vec{a_1}) \cdot (\vec{d_1}-\vec{a_2}) > 0$ or $(\vec{d_2}-\vec{a_1}) \cdot (\vec{d_2}-\vec{a_2}) > 0$} {
         \Return{None}\;
     }
-    \newpage
 
     // Case 3: $\vec{x}$ lies between $\vec{a_1}$ and $\vec{a_2}$.\\
-    \If{$\vec{x}-\vec{a_1} \cdot \vec{a_2}-\vec{x} > 0$} {
+    \If{$(\vec{x}-\vec{a_1}) \cdot (\vec{a_2}-\vec{x}) > 0$} {
         // Project the endpoints of $B$ on $A$, to find out which endpoint is closer to the intersection\\
@@
     $\theta = \arccos{|\vec{d}|/|\vec{b}-\vec{x}|}$\;
-    \Return{$\tan{\theta} \cdot |\vec{k}|$}
-
-    \caption{Projected Line Segment Distance}
+    \Return{$\tan{\theta} \cdot |\vec{k}|$}\;
+
+    \caption{PLSD-Helper}
 \end{algorithm}

### 2020-05-18 report/report.tex
Update todo
Add references to Nakata-Osaki and Joon

@@
 % Monday
-% TODO: Edge distance algorithm, algorithm
 % Collision study implementation
@@
 As said however, the big problem is that these features will rarely match real life, such models will be sufficient as long as one can be satisified with looking at rough representations of buildings and does not require additional knowledge about minor features.
-Many studies urban development have been done before that requred detailed knowledge about a city's geometry, e.g.
-for the purposes of modeling heat development (***find reference) and reducing noise (**find reference).
+Many studies urban development have been done before that requred detailed knowledge about a city's geometry, e.g. for the purposes of modeling heat development
+\textbf{(Nakata-Osaki, Souza and Rodrigues, 2018)}
+and reducing noise.
+\textbf{(Joon, Seo and Byung, 2011)}
 
@@
 OSM, while being an excellent source of city geometry, is simply not a sophisticated enough dataset to cater to all these possible areas of research.
-\textbf{Stojanovski, 2019}
+\textbf{(Stojanovski, 2019)}
 For media applications, such morppological and multimodal information could in theory be generated procedurally, by using shape grammars and some sort of wayfinding to create logical routes for different types of traffic.
@@
 
+\bibitem{joon-11}
+Joon, H.
+K., Seo, I.
+C.
+and Byung, C.
+L.
+(2011) Noise impact assessment by utilizing noise map and GIS: A case study in the city of Chungju, Republic of Korea.
+Applied Acoustics, Volume 72, Issue 8.
+544-550.
+ISSN 0003-682X.
+
 \bibitem{haklay-10}
@@
 Müller, P., Wonka, P., Haegler, S., Ulmer, A. and Van Gool, L. (2006). Procedural modeling of buildings. In ACM SIGGRAPH 2006 Papers, 614-623.
+
+\bibitem{nakata-osaki-18}
+Nakata-Osaki, C.
+M., Souza, L.
+C.
+L.
+and Rodrigues, D.
+S.
+(2018).
+THIS – Tool for Heat Island Simulation: A GIS extension model to calculate urban heat island intensity based on urban geometry.
+Computers, Environment and Urban Systems, Volume 67.
+157-168.
+ISSN 0198-9715.
 

### 2020-05-19 report/report.tex
Add reference to Müller et al
Add figure illustrating point matching algorithm

@@
 This project has dealt with the feasibility of including footpath data when generating 3D representations of OSM data, and preserving the integrity of the 3D road network, avoiding overlap and collision between road meshes.
+
 The project falls within the broader topic of geodata extraction and modelling, primarily of roads and building parcels.
 Geodata modeling is at the foundation of procedural modelling of real-world cities, since any application which models cities based on geodata will need a robust method of aqquiring and laying out road and building data without compromising the data's integrity.
-(***find reference)
+\textbf{(Müller et al, 2006)}
 
@@
 Any edges in the simplified footprints that coinside with the OMBR from the same dataset are extracted, and the corresponding points in the original footprints will be matched with each other.
+See figure \ref{fig:vertex-match-algo} for an illustration of the algorithm that detects matching vertices between two similar polygons.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_vertex_match_algo}
+    \caption{An illustration of the matching vertices detection algorithm.
+Starting from two highly similar polygons, the first step is to obtain both the Douglas-Peucker reduced polygons.
+The next step is to obtain the OMBRs of the reduced polygons, and for both polygons find the edges that coinside with the edges of the OMBRs.
+These edges are corresponding, and thus their end vertices form a matching vertex map between the original polygons.}
+    \label{fig:vertex-match-algo}
+\end{figure}
 

### 2020-05-27 report/report.tex
Update todo
Improve text connetivity in abstract

@@
 So far there have been a lot of research in modelling of 3D road networks, but these projects have often focused on the road networks designated for cars and public transit, and seldom on pedestrian footpaths.
-This project assesses the feasibility of reconstructing a city from OpenStreetMap (OSM) road and building data, and will focus on the feasibility of including pedestrian footpaths, by focusing on OSM road and building data in the Stockholm metropolitan area.
-
-An estimate of the positional accuracy of OSM data has been obtained by comparing OSM data to a dataset with a known accuracy.
-It was found that the average positional error in the OSM dataset was about 2 meters.
-This estimate of the positional error was then used to assess the integrity of the OSM Stockholm road network, by locating critical points where assigning standard widths to the roads would cause collision with other features, and evaluating how many of those points could be eliminated by making ajustments within the found margin of error.
+
+This project assesses the feasibility of reconstructing a city from OpenStreetMap (OSM) road and building data, and will focus on the feasibility of including pedestrian footpaths, by examining how common it is that these footpaths collide with other map features.
+In the collision detection process, the margin of error has been taken as the positional error in the OSM dataset, which has been obtained by comparing OSM data to a dataset with a known accuracy.
+
 It was found that footpaths were less problematic than other road types (primary, secondary, residential), in terms of feature collisions, and thus should pose no more problems in 3D city reproduction than other road types.
@@
 
+The estimate of the positional error has was used to assess the integrity of the OSM Stockholm road network, by locating critical points where assigning standard widths to the roads would cause collision with other features, and evaluating how many of those points could be eliminated by making ajustments within the found margin of error.
 For the purpose of the collision study, each road has been assigned a standard width that is in line with Swedish regulations, since in the OSM dataset, roads are represented as simple polylines, lacking width.
@@
 
-% Monday
-% Collision study implementation
-    % TODO: Figure detailing the steps in the point-matching algorithm
+% Report, restructuring
+% TODO: Improve connectivity between two first paragraphs in the abstract
+% TODO: Merge section 3 with the background chapter
+% TODO: Move talk about specific geodata to Implementation section
+% TODO: Move 4.1 -> Evaluation/Methodology (new section)
+% TODO: Overview at the beginning of each chapter
+% TODO: Move conclusions to a new Part 2 of Results and Metrics
+% TODO: New conclusions and future work section
+% TODO: Make research questions even clearer
 

### 2020-05-27 report/report.tex
Move talk about specific geodata to Implementation
New evaluation and methodology section
Make research questions even clearer

@@
 This project has dealt with the feasibility of including footpath data when generating 3D representations of OSM data, and preserving the integrity of the 3D road network, avoiding overlap and collision between road meshes.
-
-The project falls within the broader topic of geodata extraction and modelling, primarily of roads and building parcels.
-Geodata modeling is at the foundation of procedural modelling of real-world cities, since any application which models cities based on geodata will need a robust method of aqquiring and laying out road and building data without compromising the data's integrity.
-\textbf{(Müller et al, 2006)}
-
-The primary contribution is an \textbf{investigation into the rate of collision between map features in OSM}.
-A program has taken an OSM dataset of the Stockholm area and assigned each road with standard withds according to swedish regulation.
+The main contributions are as follows:
+
+\begin{enumerate}
+    \item The primary contribution is an \textbf{investigation into the rate of collision between map features in OSM}.
+How many roads of each type will collide with some other map feature once they have been given a geometrical width?
+    \item The secondary contribution will be an \textbf{assesment of the geometrical precision of OSM data}. What is the positional accuracy of OSM map features compared to a reference map?
+\end{enumerate}
+
+For the primary contribution: the collision study, a program has taken an OSM dataset of the Stockholm area and assigned each road with standard withds according to swedish regulation.
 It has then identified critical areas in the OSM dataset, where the paths are so wide that they collide with existing map features.
@@
 For the purpose of this report, features will refer mainly to roads and building footprints.
+For the collision study, see sections 1.3.1 for an implementation overview, 4.2 for a detailed walkthrough of the program implementation, and 5.1 for the results.
+
+For the secondary contribution: the geodata precision study, the positional accuracy will be estimated by comparing feature points between the OSM dataset and a reference dataset.
+The positional accuracy will be needed to assess where road collisions occur and whether they can easily be corrected.
+For the positional accuracy study, see sections 1.3.2 for an implementation overview, 4.1 for a walkthrough of the program implementation, and 5.2 for the results.
+
+\section{Implementation overview}
+
+\subsection{Assesment of geodata precision}
+
+The geometrical precision study has been carried out in a number of steps that are all detailed below.
+First a feature map was found, which mapped individual features between the OSM dataset and a reference dataset.
+A program then took matching building pairs and extracted matching geometrical points after first simplifying the polygons using the Douglas-Peucker algorithm \textbf{(Douglas and Peucker, 1973)}.
+Once a set of matching points has been found across the whole dataset, the average distance between them was taken as a measure of the positional accuracy.
+
+Additionally, a number of auxillary calculations were neccessary for this study.
+The estimated completeness of the OSM dataset (a measure of how many features are represented in both the OSM dataset and the reference dataset) was calculated as the relative difference in area of all building footprints, between the two datasets.
+The completeness is needed to estimate the error boundary of the positional accuracy.
+Additionally, the shape accuracy of the OSM dataset has also been calculated and graphed as an extra verification of the similarity between the datasets.
+
+\subsection{Collision study}
+
+The estimate of the positional error has was used to assess the integrity of the OSM Stockholm road network, by locating critical points where assigning standard widths to the roads would cause collision with other features, and evaluating how many of those points could be eliminated by making ajustments within the found margin of error.
+For the purpose of the collision study, each road has been assigned a standard width that is in line with Swedish regulations, since in the OSM dataset, roads are represented as simple polylines, lacking width.
+In a 3D representation however, it is crucial that each road object has a spacial width in order to be rendered and seen.
+After the width assignment, a program has iterated over all features and calculated how many of them that overlap with any other feature, as well as the total length of the road polylines that is intersecting some other feature.
+The program has then calculated how many of these colliding features that can be corrected by making adjustments that are within the positional accuracy of the OSM data (As obtained by the previous study in this report: Assessment of Geodata Precision).
+
+\chapter{Methodology and evaluation}
+
+\section{Evaluation metrics}
 
@@
 
-For the collision study, see sections 1.3.1 for an implementation overview, 4.2 for a detailed walkthrough of the program implementation, and 5.1 for the results.
-
-The secondary contribution will be an \textbf{assesment of the geometrical precision of OSM data} around the Stockholm area.
-The positional accuracy will be estimated by comparing feature points between the OSM dataset and a reference dataset.
-The positional accuracy will be needed to assess where road collisions occur and whether they can easily be corrected.
-
 The secondary study about geodata precision will primarily serve to obtain an estimate of the positional accuracy of the OSM dataset, in meters.
@@
 See section 5.3 for more details about the shape accuracy measure.
-
-For the positional accuracy study, see sections 1.3.2 for an implementation overview, 4.1 for a walkthrough of the program implementation, and 5.2 for the results.
-
-\section{Implementation overview}
-
-\subsection{Assesment of geodata precision}
-
-The geometrical precision study has been carried out in a number of steps that are all detailed below.
-First a feature map was found, which mapped individual features between the OSM dataset and a reference dataset.
-A program then took matching building pairs and extracted matching geometrical points after first simplifying the polygons using the Douglas-Peucker algorithm \textbf{(Douglas and Peucker, 1973)}.
-Once a set of matching points has been found across the whole dataset, the average distance between them was taken as a measure of the positional accuracy.
-
-Additionally, a number of auxillary calculations were neccessary for this study.
-The estimated completeness of the OSM dataset (a measure of how many features are represented in both the OSM dataset and the reference dataset) was calculated as the relative difference in area of all building footprints, between the two datasets.
-The completeness is needed to estimate the error boundary of the positional accuracy.
-Additionally, the shape accuracy of the OSM dataset has also been calculated and graphed as an extra verification of the similarity between the datasets.
-
-\subsection{Collision study}
-
-The estimate of the positional error has was used to assess the integrity of the OSM Stockholm road network, by locating critical points where assigning standard widths to the roads would cause collision with other features, and evaluating how many of those points could be eliminated by making ajustments within the found margin of error.
-For the purpose of the collision study, each road has been assigned a standard width that is in line with Swedish regulations, since in the OSM dataset, roads are represented as simple polylines, lacking width.
-In a 3D representation however, it is crucial that each road object has a spacial width in order to be rendered and seen.
-After the width assignment, a program has iterated over all features and calculated how many of them that overlap with any other feature, as well as the total length of the road polylines that is intersecting some other feature.
-The program has then calculated how many of these colliding features that can be corrected by making adjustments that are within the positional accuracy of the OSM data (As obtained by the previous study in this report: Assessment of Geodata Precision).
 
@@
 
-\chapter{Background}
-
-\section{Background of 3D city reproduction}
-
-Procedural modeling of road networks is of interest to the fields of urban planning and urban architecture.
-In recent years there has been an advent of interactive tools that aid urban planners in placing or generating features of an urban plan, particularly roads and road networks.
-Procedural and AI solutions that aid the arcitect in a form of human-machine symbiosis.
-
-Procedural city-generation models combine architectural elemets, such as streets, lots, buildings, building facades.
-Urban design practises with conventional two-dimenstional deliverables range from delivering neighbourhood designs, urban zoning and building plans, to writing design codes and building ordinances.
-The last decade has seen multiple startups in the area of software that aid urban arcitecture.
-The software uses procedural models to render a city grid based on real-world locations, or generate new user-specified parameters, and allow for varying degrees of manual editing on grid, road, neighbourhood, city block or individual feature level.
-Some of the most prominent examples are ArcGIS Pro (made by Esri), CityEngine (made by the startup Procedural, bought by ESRI) and Urban Canvas (made by the startup Synthicity, bought by Autodesk, now defunct).
-
-But such tools are as of today not commonly applied in urbanist practises, due to failures in reflecting the unique workspace of arcitects and their design needs.
-Urban arcitecture is a social factors, stakeholders.
-Each arcitect is an actor in a continously developing environment, and affecting that environment in a complex manner through their action \textbf{(Stojanovski et al, 2020)}.
-
-In recent years, there have been an advent of tools that extract geodata from OSM and generate 3D representations of real world locations by placing terrain, buildings, roads and other features.
-However, freely available such tools seldom include smaller roads and pedestrian footpaths such as walkways, sidewalks or trails.
-
-Additionally, the topic is of interest for 3D content creators in for example the fields of video game design or 3D animation.
-A trend in these fields is using methods of procedural modeling to create large amounts of 3D content quickly and efficiently, while requiring as little hand modeling as possible.
-Studios who create large worlds are often interested in solutions that generate large road networks without any hand modeling.
-\textbf{(Parish and Müller, 2001)}
-
-\section{Related work}
-When assessing the completeness of geodata, a very common scale is Van Oorts criteria for evaluating the quality of geographical information, which at the time of publication received attention from surveyors and cartographers.
-\textbf{(Van Oort, 2006)} Since then a number of publications have examined slices of geodata from van Oort's criteria.
-In 2008, \textbf{(Haklay, 2010)} conducted such a study to estimate of the quality of OSM positional geodata in London and England.
-Since the OSM project started in London it was thought that OSM data in the London metropolitan area would be representative of the highest quality data available, and therefore a good indicator for the whole global OSM dataset.
-The study was conducted by comparing OSM data to Ordnance Survey datasets.
-The study showed that the OSM dataset had a rough positional accuracy of 6 meters from the reference dataset.
-However, at the time of the study, the OSM project had captured roughly 29\% of the area of England, meaning that the data completeness was fairly low.
-
-The next notable study to assess the quality of OSM data was published by \textbf{(Kunze, 2012)}.
-The study applied different methods to assess the completeness of OSM data in two federal states in Germany, mainly by analysing the area difference between the OSM dataset and an administrative dataset.
-Finally, a notable study in 2013-2014, conducted by researchers from various universities in Norway and Germany \textbf{(Fan et al, 2014)}, examined the quality and accuracy in building footprint data of the Munich area, at a time and place where the completeness of the OSM dataset was significantly higher (the results of their study showed that the OSM had captured close to 100\% of all building footprints).
-At the time Munich was one of the most developed cities in OSM.
-Their study included an insight in the geometrical calculations neccessary to match features and points between two datasets, and how to reliably calculate the metrics needed for the most important and assessible of van Oorts criteria.
-Their study revealed that although the feature completeness of OSM was high, some architectural details were missing.
-The results reveal that the positional accuracy of OSM data at the time was about 4 meters.
-
-\chapter{The geodata used in this project}
-
-For the purpose of this project, the OSM road dataset has been sampled around Stockholm metropolitan area.
-To assess and correct path widths and path node locations, the OSM building footprint dataset areound the same area has also been used.
-As a reference dataset, the property map from the Swedish National Land Survey was used in comparison with the OSM map.
-The reference map is provided by the Swedish University of Agricultural Sciences (SLU), and will be referred to in this report as the SLU map or the SLU dataset.
-The SLU map conatins only building and property footprints.
-The exact map segments that were extracted were in both cases rectangular boundaries with the following coordinates in WGS84: (N: 59.42, E: 18.15, S: 59.23, W: 17.79).
-Both datasets were collcted on March 18th, 2020.
-
-\section{OSM data}
-
-The OSM project is a wiki-like geodatabase based on Volunteered Geographical Information (VGI), meaning anyone can contribute by adding or editing features.
-OSM road data and building footprints are commonly obtained by manually tracing features in commercially available satellite images.
-Such images have a limit on their resolution which puts a theoretical limit on the accuracy of map features compared to their real-world equivalents (Haklay 2010).
-Particularily the high resolution imagery from Bing in 2010 led to an increase in building information in OSM, and the positional error is largely due to the limited resolution in such satellite imagery.
-\textbf{(Fan et al, 2014)}.
-
-\section{SLU data}
-
-The SLU map is maintained by The Swedish National Land Survey (Swedish: Lantmäteriet).
-The data is collected by geodetic professionals, by land surveying methods such as GPS or DGNSS positioning, or by reproduction of features from ortophoto or stereo mapping from 3D aerial images.
-Building features in the SLU dataset have a position accuracy requirement of 2 meters.
-The SLU map contains building features and property limits, but no road data.
-The map is updated continously by Lantmäteriet, in conjoncture with the forming or reforming of property.
-Whereas OSM is a free service, the SLU map is available to paying customers or for free to students and researchers.
-\footnote{Lantmäteriet.
-Produktbeskrivning: GSD-Fastighetskartan vektor - Lantmäteriet.
-Acessed 16-04-2020.
-https://www.lantmateriet.se/globalassets/kartor-och-geografisk-information/kartor/fastshmi.pdf}
-
-\section{A word on coordinate systems}
-
-The SLU dataset is delivered in the SWEREF 99 TM coordinate system.
-SWEREF 99 TM is a projected coordinate system and there is no linear transformation to the WGS 48 system, which OSM uses.
-\footnote{Converting to WGS84 - OpenStreetMap Wiki.
-Accessed 13-03-2020.
-https://wiki.openstreetmap.org/wiki/Converting\_to\_WGS84}
-The coordinate conversions for this paper were obtained using proj: a Linux commandline application for geospatial coordinate conversion.
-
-\section{Specific geodata preprocessing}
-
-Upon delivery of the OSM dataset, all features that intersect with the user-specified domain are included, with their full geometry.
-The SLU map however, is delivered with building footprints cropped to exactly match the query coordinates, meaning that buildings at the edge of the user-specified domain can have cropped geometries.
-Any building area outside of the query domain will be excluded.
-Since the principles behind what features are delivered and how differ between the datasets, it is neccessary to crop both datasets to ensure that all buildings are complete and have matching candidates in the other set.
-Any features that intersect the edge of the rectagular boundary, in both datasets, were therefore excluded from the study.
-
-\chapter{Implementation}
-
-\section{Geodata precision study}
-
-\subsection{Quality criteria}
+\section{Quality criteria}
 
@@
 The temporal accuracy will not be assessed as this projects focuses on comparing a single snapshot of the two datasets at the same point in time, and finally the logical accuracy will not be assessed because the underlying logical relations between features in both maps serve no purpose to this project.
+
+\chapter{Background}
+
+\section{Background of 3D city reproduction}
+
+The project falls within the broader topic of geodata extraction and modelling, primarily of roads and building parcels.
+Geodata modeling is at the foundation of procedural modelling of real-world cities, since any application which models cities based on geodata will need a robust method of aqquiring and laying out road and building data without compromising the data's integrity.
+\textbf{(Müller et al, 2006)}
+
+Procedural modeling of road networks is of interest to the fields of urban planning and urban architecture.
+In recent years there has been an advent of interactive tools that aid urban planners in placing or generating features of an urban plan, particularly roads and road networks.
+Procedural and AI solutions that aid the arcitect in a form of human-machine symbiosis.
+
+Procedural city-generation models combine architectural elemets, such as streets, lots, buildings, building facades.
+Urban design practises with conventional two-dimenstional deliverables range from delivering neighbourhood designs, urban zoning and building plans, to writing design codes and building ordinances.
+The last decade has seen multiple startups in the area of software that aid urban arcitecture.
+The software uses procedural models to render a city grid based on real-world locations, or generate new user-specified parameters, and allow for varying degrees of manual editing on grid, road, neighbourhood, city block or individual feature level.
+Some of the most prominent examples are ArcGIS Pro (made by Esri), CityEngine (made by the startup Procedural, bought by ESRI) and Urban Canvas (made by the startup Synthicity, bought by Autodesk, now defunct).
+
+But such tools are as of today not commonly applied in urbanist practises, due to failures in reflecting the unique workspace of arcitects and their design needs.
+Urban arcitecture is a social factors, stakeholders.
+Each arcitect is an actor in a continously developing environment, and affecting that environment in a complex manner through their action \textbf{(Stojanovski et al, 2020)}.
+
+In recent years, there have been an advent of tools that extract geodata from OSM and generate 3D representations of real world locations by placing terrain, buildings, roads and other features.
+However, freely available such tools seldom include smaller roads and pedestrian footpaths such as walkways, sidewalks or trails.
+
+Additionally, the topic is of interest for 3D content creators in for example the fields of video game design or 3D animation.
+A trend in these fields is using methods of procedural modeling to create large amounts of 3D content quickly and efficiently, while requiring as little hand modeling as possible.
+Studios who create large worlds are often interested in solutions that generate large road networks without any hand modeling.
+\textbf{(Parish and Müller, 2001)}
+
+\section{Related work}
+When assessing the completeness of geodata, a very common scale is Van Oorts criteria for evaluating the quality of geographical information, which at the time of publication received attention from surveyors and cartographers.
+\textbf{(Van Oort, 2006)} Since then a number of publications have examined slices of geodata from van Oort's criteria.
+In 2008, \textbf{(Haklay, 2010)} conducted such a study to estimate of the quality of OSM positional geodata in London and England.
+Since the OSM project started in London it was thought that OSM data in the London metropolitan area would be representative of the highest quality data available, and therefore a good indicator for the whole global OSM dataset.
+The study was conducted by comparing OSM data to Ordnance Survey datasets.
+The study showed that the OSM dataset had a rough positional accuracy of 6 meters from the reference dataset.
+However, at the time of the study, the OSM project had captured roughly 29\% of the area of England, meaning that the data completeness was fairly low.
+
+The next notable study to assess the quality of OSM data was published by \textbf{(Kunze, 2012)}.
+The study applied different methods to assess the completeness of OSM data in two federal states in Germany, mainly by analysing the area difference between the OSM dataset and an administrative dataset.
+Finally, a notable study in 2013-2014, conducted by researchers from various universities in Norway and Germany \textbf{(Fan et al, 2014)}, examined the quality and accuracy in building footprint data of the Munich area, at a time and place where the completeness of the OSM dataset was significantly higher (the results of their study showed that the OSM had captured close to 100\% of all building footprints).
+At the time Munich was one of the most developed cities in OSM.
+Their study included an insight in the geometrical calculations neccessary to match features and points between two datasets, and how to reliably calculate the metrics needed for the most important and assessible of van Oorts criteria.
+Their study revealed that although the feature completeness of OSM was high, some architectural details were missing.
+The results reveal that the positional accuracy of OSM data at the time was about 4 meters.
+
+\chapter{Implementation}
+
+\section{Geodata precision study}
 
@@
 
+\section{The geodata used in this project}
+
+For the purpose of this project, the OSM road dataset has been sampled around Stockholm metropolitan area.
+To assess and correct path widths and path node locations, the OSM building footprint dataset areound the same area has also been used.
+As a reference dataset, the property map from the Swedish National Land Survey was used in comparison with the OSM map.
+The reference map is provided by the Swedish University of Agricultural Sciences (SLU), and will be referred to in this report as the SLU map or the SLU dataset.
+The SLU map conatins only building and property footprints.
+The exact map segments that were extracted were in both cases rectangular boundaries with the following coordinates in WGS84: (N: 59.42, E: 18.15, S: 59.23, W: 17.79).
+Both datasets were collcted on March 18th, 2020.
+
+\subsection{OSM data}
+
+The OSM project is a wiki-like geodatabase based on Volunteered Geographical Information (VGI), meaning anyone can contribute by adding or editing features.
+OSM road data and building footprints are commonly obtained by manually tracing features in commercially available satellite images.
+Such images have a limit on their resolution which puts a theoretical limit on the accuracy of map features compared to their real-world equivalents (Haklay 2010).
+Particularily the high resolution imagery from Bing in 2010 led to an increase in building information in OSM, and the positional error is largely due to the limited resolution in such satellite imagery.
+\textbf{(Fan et al, 2014)}.
+
+\subsection{SLU data}
+
+The SLU map is maintained by The Swedish National Land Survey (Swedish: Lantmäteriet).
+The data is collected by geodetic professionals, by land surveying methods such as GPS or DGNSS positioning, or by reproduction of features from ortophoto or stereo mapping from 3D aerial images.
+Building features in the SLU dataset have a position accuracy requirement of 2 meters.
+The SLU map contains building features and property limits, but no road data.
+The map is updated continously by Lantmäteriet, in conjoncture with the forming or reforming of property.
+Whereas OSM is a free service, the SLU map is available to paying customers or for free to students and researchers.
+\footnote{Lantmäteriet.
+Produktbeskrivning: GSD-Fastighetskartan vektor - Lantmäteriet.
+Acessed 16-04-2020.
+https://www.lantmateriet.se/globalassets/kartor-och-geografisk-information/kartor/fastshmi.pdf}
+
+\subsection{A word on coordinate systems}
+
+The SLU dataset is delivered in the SWEREF 99 TM coordinate system.
+SWEREF 99 TM is a projected coordinate system and there is no linear transformation to the WGS 48 system, which OSM uses.
+\footnote{Converting to WGS84 - OpenStreetMap Wiki.
+Accessed 13-03-2020.
+https://wiki.openstreetmap.org/wiki/Converting\_to\_WGS84}
+The coordinate conversions for this paper were obtained using proj: a Linux commandline application for geospatial coordinate conversion.
+
+\subsection{Specific geodata preprocessing}
+
+Upon delivery of the OSM dataset, all features that intersect with the user-specified domain are included, with their full geometry.
+The SLU map however, is delivered with building footprints cropped to exactly match the query coordinates, meaning that buildings at the edge of the user-specified domain can have cropped geometries.
+Any building area outside of the query domain will be excluded.
+Since the principles behind what features are delivered and how differ between the datasets, it is neccessary to crop both datasets to ensure that all buildings are complete and have matching candidates in the other set.
+Any features that intersect the edge of the rectagular boundary, in both datasets, were therefore excluded from the study.
+
 \chapter{Results and metrics}

### 2020-05-27 report/report.tex
Humorous swedish abstract placeholder

@@
 \begin{abstract}
-    Träutensilierna i ett tryckeri äro ingalunda en oviktig faktor,
-    för trevnadens, ordningens och ekonomiens upprätthållande, och
-    dock är det icke sällan som sorgliga erfarenheter göras på grund
-    af det oförstånd med hvilket kaster, formbräden och regaler
-    tillverkas och försäljas Kaster som äro dåligt hopkomna och af
-    otillräckligt.
+Haha, He-Man!
+There is no possible way that you feeble-brained barbarian and your furball mount can foil my plan this time.
+With the ancient power at my command I shall doom this kingdom to crumble in meager ashes and..
+Oh shit, he has run off with the crystal skull hasn't he?
+Damn you He-Man!!!
+
 \end{abstract}

### 2020-05-27 report/report.tex
Separate final chapter into analysis and conclusions
Update todo
Conclusion headlines

@@
 
-\chapter{Conclusions and future work}
+\chapter{Analysis}
 
 \section{Quality of the OSM dataset}
-
-One could reasonably assume that Stockholm, being one of the more developed cities in one of the most technologically connected country in the world, would be fairly well mapped in OSM.
-Let us take the positional accuracy of the Stockholm area, as obtained by this study, as reprentative of that of major cities in developed countries in the OSM dataset.
-Then the positional accuracy has improved by roughly 3.5 meters since the study by \textbf{(Haklay, 2010)}, and by roughly 1.5 meters since the study by \textbf{(Fan et al, 2014)}.
-This of course assumes that the SLU reference dataset is taken as absolute.
 
@@
 % Report, restructuring
-% TODO: Improve connectivity between two first paragraphs in the abstract
-% TODO: Merge section 3 with the background chapter
-% TODO: Move talk about specific geodata to Implementation section
-% TODO: Move 4.1 -> Evaluation/Methodology (new section)
+% TODO: Summarize Analysis in Conclusions
 % TODO: Overview at the beginning of each chapter
-% TODO: Move conclusions to a new Part 2 of Results and Metrics
-% TODO: New conclusions and future work section
-% TODO: Make research questions even clearer
 
@@
 
-\section{Scientific field study}
+\chapter{Conclusions and future work}
+
+\section{Data conclusions}
+
+One could reasonably assume that Stockholm, being one of the more developed cities in one of the most technologically connected country in the world, would be fairly well mapped in OSM.
+Let us take the positional accuracy of the Stockholm area, as obtained by this study, as reprentative of that of major cities in developed countries in the OSM dataset.
+Then the positional accuracy has improved by roughly 3.5 meters since the study by \textbf{(Haklay, 2010)}, and by roughly 1.5 meters since the study by \textbf{(Fan et al, 2014)}.
+This of course assumes that the SLU reference dataset is taken as absolute.
+
+\section{Remaining challenges in city reproduction}
 
@@
 
-Such procedural techniques have long been available.
+\section{Using procedural methods}
+
+Procedural techniques are, as mentioned, methods for reconstructing or creating 3D automatically, with minimal manual work required.
+Adding such techniques for content creation into the workflow of city reconstruction will increase the level of detail that is possible, while sacrificing accuracy to the real-world.
 \textbf{(Mũller et al, 2006)} present how to accomplish this using Shape Grammar: a method for automatically and iteratively constructing 3D models based in a hierarchical set of rules.
@@
 
-It even turns out, by looking at such studies, that a lot of research in urban development require detailed morpological information that goes beyond simply knowing the geometry of the city.
+\section{Feasibility in the urban planning field}
+By looking at such studies in urban development, one will discover that a lot of research in this field requires detailed morpological information that goes beyond simply knowing the geometry of the city.
 Urban architects often work with morpological structures where a single feature, such as a city neighbourhood, can be examined from many different levels of perspective.
@@
 
+\section{Final conclusions}
+
 The final conclusion is that generating virtual cities purely from OpenStreetMap data is a perfectly feasibile procedure in the media industry, with the inclusion of procedural methods such as shape grammars to generate neccessary detail.

### 2020-05-27 report/report.tex
Summary of analysis chapter

@@
 
-\subsection{Assesment of geodata precision}
+Programmatic collection of statistics has been neccessary for both the primary and secondary research questions, and this section presents a broad overview of the technical implementation behind the map analysis and data collection processes for both the geodata precision study and the collision study.
+
+\subsection{Geodata precision study}
 
@@
 
-The estimate of the positional error has was used to assess the integrity of the OSM Stockholm road network, by locating critical points where assigning standard widths to the roads would cause collision with other features, and evaluating how many of those points could be eliminated by making ajustments within the found margin of error.
+Once the estimate of the positional error was obtained from the precision study, this estimate was used to assess the integrity of the OSM Stockholm road network, by locating critical points where assigning standard widths to the roads would cause collision with other features, and evaluating how many of those points could be eliminated by making ajustments within the found margin of error.
 For the purpose of the collision study, each road has been assigned a standard width that is in line with Swedish regulations, since in the OSM dataset, roads are represented as simple polylines, lacking width.
@@
 
-For the secondary study, the average positional accuracy per building was calculated to roughly 2.06 meters, and this will be taken as an estaimate of the positional accuracy of the OSM dataset.
+For the secondary study, the average positional accuracy per building was calculated to roughly 2.06 meters, and this will be taken as an estimate of the positional accuracy of the OSM dataset.
 Table \ref{table:positional-accuracy} shows the estimate in positional accuracy that was obtained from the study.
@@
 
+Now that the colliison metrics have been obtained, we are in a good position to draw conclusions about the technical feasibility of 3D city reproduction from OSM data.
+We will begin by looking at the geodata precision study, and use the outcome of that study to assess the quality of the positional accuracy estimate, by looking at the dataset completeness, the low semantic accuracy and the unique characteristics of the Stockholm area.
+Secondly we will look at the collision study.
+We will make the claim that footpaths will not be more problematic than other road types in city reproduction from OSM data, and motivate this with the data obtained in section 6.1.
+We will look at the character of the feature collision cases and identify the most common cases, as well as present geometrical algorithms for how to correct each case automatically.
+For conclusions about the usage of OSM city reproduction in engineering practise and remaining challenges, see section 7.
+
 \section{Quality of the OSM dataset}
@@
 \chapter{Conclusions and future work}
-
-\section{Data conclusions}
-
-One could reasonably assume that Stockholm, being one of the more developed cities in one of the most technologically connected country in the world, would be fairly well mapped in OSM.
-Let us take the positional accuracy of the Stockholm area, as obtained by this study, as reprentative of that of major cities in developed countries in the OSM dataset.
-Then the positional accuracy has improved by roughly 3.5 meters since the study by \textbf{(Haklay, 2010)}, and by roughly 1.5 meters since the study by \textbf{(Fan et al, 2014)}.
-This of course assumes that the SLU reference dataset is taken as absolute.
 

### 2020-05-27 report/report.tex
Conclusions summary

@@
 
+We now have a firm idea of the technical feasibility of generating virtual cities from OpenStreetMap-data, including different types of road networks.
+We have however yet to touch upon the possibilities of road network modelling in the broader context of 3D city reproduction, and what are the remaining challenges for accurate 3D city reproduction?
+What techniques can we use to complement our basic system of importing street and building data and create a rich and living virtual city?
+In this section we will place everything we have learnt so far in the broader context of GIS and city modeling, focusing on the research fields of 3D media and city planning.
+
 \section{Remaining challenges in city reproduction}
-
-We now have a firm idea of the technical feasibility of generating virtual cities from OpenStreetMap-data, including different types of road networks.
-We have however yet to touch upon the importance for road network modelling in the broader context of 3D city reproduction.
-Where exactly in the pipeline of reconstructing a virtual city does this happen, and what are the remaining challenges?
-This will be the goal of this final scientific study: to place everything we have learnt so far in the broader context of GIS and city modeling, and the respective research fields.
 

### 2020-05-27 report/report.tex
Results section summary

@@
 
+Here we present the results of the statistics collections that were made as part of this project.
+First the results of the collision study is presented as how many of the road features of each type display collision with some other feature.
+Then the results of the geodata precision is presented.
+We will present our estimate of the positional accuracy of the OSM dataset, as well as the dataset completeness and statistics about building correspondence and shape accuracy.
+
 \section{Road collision study}
@@
 
-\subsection{Building matching statistics}
+\subsection{Building correspondence statistics}
 

### 2020-05-27 report/report.tex
Methodology and evaluation, summary

@@
 
+In this chapter we will look in details at the evaluation metrics that this project served to obtain, and how they relate to the eventual collision analysis.
+For the geodata precision study specifically, we will look in detail at how geodata is qualifyied in general, by examining van Oorts quality criteria.
+We will also present how van Oorts critera will be used in the context of this project.
+See section 4 for a detailed technical implementation overview that presents how the relevant metrics were obtained during the statistics collection process.
+For the resulting metrics that were obtained, see section 5.
+
 \section{Evaluation metrics}
 
-The primary study of the rate of collision between roads and other features, will serve to obtain the following three metrics:
+The primary study of the rate of collision between roads and other features will serve to obtain the following three metrics:
 
@@
 See section 5.3 for more details about the shape accuracy measure.
-
-\section{Problem constraints}
-
-In this study, when working with OSM geodata, the following assumptions have been made about the dataset to reduce the size of the problem domain:
-
-\begin{enumerate}
-    \item The program will not handle terrain features and altitude. It will be assumed that the road mesh can be projected on a flat 2D surface without feature intersection.
-    \item It will be assumed that the city map consists of only arterial (primary) roads, secondary roads and footpaths.
-    \item It will be assumed that the OSM dataset represents all roads as simple polylines, and building footprints as simple connected polygons.
-Any other features (such as squares, which are roads represented as polygons) will be omitted from the dataset.
-    \item The road network is assumed to be a 2D simple graph.
-This means that any overlapping roads, such as tunnels and bridges that cross over street-level roads, will be eliminated from the dataset.
-Any self-connected nodes will also be eliminated from the dataset.
-\end{enumerate}
 
@@
 The temporal accuracy will not be assessed as this projects focuses on comparing a single snapshot of the two datasets at the same point in time, and finally the logical accuracy will not be assessed because the underlying logical relations between features in both maps serve no purpose to this project.
+
+\section{Problem constraints}
+
+In this study, when working with OSM geodata, the following assumptions have been made about the dataset to reduce the size of the problem domain:
+
+\begin{enumerate}
+    \item The program will not handle terrain features and altitude. It will be assumed that the road mesh can be projected on a flat 2D surface without feature intersection.
+    \item It will be assumed that the city map consists of only arterial (primary) roads, secondary roads and footpaths.
+    \item It will be assumed that the OSM dataset represents all roads as simple polylines, and building footprints as simple connected polygons.
+Any other features (such as squares, which are roads represented as polygons) will be omitted from the dataset.
+    \item The road network is assumed to be a 2D simple graph.
+This means that any overlapping roads, such as tunnels and bridges that cross over street-level roads, will be eliminated from the dataset.
+Any self-connected nodes will also be eliminated from the dataset.
+\end{enumerate}
 

### 2020-05-27 report/report.tex
Implementation summary

@@
 \chapter{Implementation}
+
+In section 4.1 we will look at the data collection process for the geodata precision study.
+We will look at how to qualify dataset completeness and building correspondence (how many of the buildings in one dataset that can be found in the other) by collecting the cumulative building areas and looking at what buildings have overlapping areas between the two datasets.
+We will also look at the difinitions and algorithms neccessary for calculating posiitonal accuracy and a measure for shape accuracy.
+In section 4.2 we will look at the data collection process for the collision study.
+We will look at the how standard road withs were obtained, as well as the algorithm for detemining feature collision that was used.
+In section 4.3 we will discuss briefly the data sources used in this project, why they were chosen and specific issues with the data collection process.
 

### 2020-05-27 report/report.tex
Background section summary

@@
 \chapter{Background}
+
+This section will dive into the state-of-the-art use of city reconstruction methods within the research field of urban design.
+The first section will present the motivations for the digitization of urban planning, as well as the most common programs and tools in this area, which are after all not very commonly used in the practise of urban design due to the failure in being able to reflect the unique workspace of urban designers.
+The Related work section will present a brief history of the field of OpenStreetMap data qualification, and present the research papers whose methods are at the foundation of this project.
 

### 2020-06-03 report/report.tex
Update todo
Make research questions even clearer

@@
 This project has dealt with the feasibility of including footpath data when generating 3D representations of OSM data, and preserving the integrity of the 3D road network, avoiding overlap and collision between road meshes.
-The main contributions are as follows:
+The main research question is as follows:
+\begin{center}
+    \textbf{Is it technically feasibile to generate a 3D virtual city with footpath road data, with minimal (and preferably no) overlap between 3D features?}
+\end{center}
+
+To investigate this, the problem will specifically be divided into the following two subquestion:
 
 \begin{enumerate}
-    \item The primary contribution is an \textbf{investigation into the rate of collision between map features in OSM}.
-How many roads of each type will collide with some other map feature once they have been given a geometrical width?
-    \item The secondary contribution will be an \textbf{assesment of the geometrical precision of OSM data}. What is the positional accuracy of OSM map features compared to a reference map?
+    \item The primary subquestion is: \textbf{How many roads of each type will collide with some other map feature, once they have been given a geometrical width?}.
+This question aims to find the rate of collision between road features in OSM.
+    \item The secondary subquestion is: \textbf{What is the positional accuracy of OSM map features compared to a reference map?}.
+The positional accuracy will be neccessary to carry out the investigation of the first subquestion: the rate of collisions between roads, since it will determine the maximum tolerance in distance between road points before a collision is certain to have occured.
 \end{enumerate}
 
-For the primary contribution: the collision study, a program has taken an OSM dataset of the Stockholm area and assigned each road with standard withds according to swedish regulation.
+For the primary research question: the collision study, a program has taken an OSM dataset of the Stockholm area and assigned each road with standard withds according to swedish regulation.
 It has then identified critical areas in the OSM dataset, where the paths are so wide that they collide with existing map features.
@@
 
-For the secondary contribution: the geodata precision study, the positional accuracy will be estimated by comparing feature points between the OSM dataset and a reference dataset.
+For the secondary research question: the geodata precision study, the positional accuracy will be estimated by comparing feature points between the OSM dataset and a reference dataset.
 The positional accuracy will be needed to assess where road collisions occur and whether they can easily be corrected.
@@
 
-% Report, restructuring
-% TODO: Summarize Analysis in Conclusions
-% TODO: Overview at the beginning of each chapter
-
-% After next meeting with Chris
+% Onsdag:
+% TODO: Split up introduction, more subsections
+% TODO: Delimitations in Introduction, after RQ
+% TODO: Go over all images, add a line about their origins
+% TODO: Merge: Results and metrics + Analysis -> Evaluation, with result/analysis subsections
+% TODO: Rename: Implementation overview -> Methodology
+% TODO: Background should be the second chapter
+
+% Torsdag:
+% TODO: Add source of image on page 1
+% TODO: Mention that positional accuracy will be the focus early on
+% TODO: Move: 2.1.1: Accuracy criteria that will not be focused on -> Delimitations
+% TODO: Formal tone, never use us, we, me
+% TODO: Move content from chapters 1, 3 and 4 to Background
+% TODO: Switch to numerical citations
+
+% After Chris approves
+% TODO: Update section references
 % TODO: Spell check

### 2020-06-03 report/report.tex
Update on RQ2
Merge results and analysis sections, new section: evaluation
Background is now the second chapter
Update todo
Formal tone, remove ocurrences of we, our, my, etc

@@
     \item The primary subquestion is: \textbf{How many roads of each type will collide with some other map feature, once they have been given a geometrical width?}.
-This question aims to find the rate of collision between road features in OSM.
+This question aims to find the rate of collision between road features in OSM, and investigate the nature of the most common types of collisions and how they can be corrected.
     \item The secondary subquestion is: \textbf{What is the positional accuracy of OSM map features compared to a reference map?}.
@@
 
+\chapter{Background}
+
+This section will dive into the state-of-the-art use of city reconstruction methods within the research field of urban design.
+The first section will present the motivations for the digitization of urban planning, as well as the most common programs and tools in this area, which are after all not very commonly used in the practise of urban design due to the failure in being able to reflect the unique workspace of urban designers.
+The Related work section will present a brief history of the field of OpenStreetMap data qualification, and present the research papers whose methods are at the foundation of this project.
+
+\section{Background of 3D city reproduction}
+
+The project falls within the broader topic of geodata extraction and modelling, primarily of roads and building parcels.
+Geodata modeling is at the foundation of procedural modelling of real-world cities, since any application which models cities based on geodata will need a robust method of aqquiring and laying out road and building data without compromising the data's integrity.
+\textbf{(Müller et al, 2006)}
+
+Procedural modeling of road networks is of interest to the fields of urban planning and urban architecture.
+In recent years there has been an advent of interactive tools that aid urban planners in placing or generating features of an urban plan, particularly roads and road networks.
+Procedural and AI solutions that aid the arcitect in a form of human-machine symbiosis.
+
+Procedural city-generation models combine architectural elemets, such as streets, lots, buildings, building facades.
+Urban design practises with conventional two-dimenstional deliverables range from delivering neighbourhood designs, urban zoning and building plans, to writing design codes and building ordinances.
+The last decade has seen multiple startups in the area of software that aid urban arcitecture.
+The software uses procedural models to render a city grid based on real-world locations, or generate new user-specified parameters, and allow for varying degrees of manual editing on grid, road, neighbourhood, city block or individual feature level.
+Some of the most prominent examples are ArcGIS Pro (made by Esri), CityEngine (made by the startup Procedural, bought by ESRI) and Urban Canvas (made by the startup Synthicity, bought by Autodesk, now defunct).
+
+But such tools are as of today not commonly applied in urbanist practises, due to failures in reflecting the unique workspace of arcitects and their design needs.
+Urban arcitecture is a social factors, stakeholders.
+Each arcitect is an actor in a continously developing environment, and affecting that environment in a complex manner through their action \textbf{(Stojanovski et al, 2020)}.
+
+In recent years, there have been an advent of tools that extract geodata from OSM and generate 3D representations of real world locations by placing terrain, buildings, roads and other features.
+However, freely available such tools seldom include smaller roads and pedestrian footpaths such as walkways, sidewalks or trails.
+
+Additionally, the topic is of interest for 3D content creators in for example the fields of video game design or 3D animation.
+A trend in these fields is using methods of procedural modeling to create large amounts of 3D content quickly and efficiently, while requiring as little hand modeling as possible.
+Studios who create large worlds are often interested in solutions that generate large road networks without any hand modeling.
+\textbf{(Parish and Müller, 2001)}
+
+\section{Related work}
+When assessing the completeness of geodata, a very common scale is Van Oorts criteria for evaluating the quality of geographical information, which at the time of publication received attention from surveyors and cartographers.
+\textbf{(Van Oort, 2006)} Since then a number of publications have examined slices of geodata from van Oort's criteria.
+In 2008, \textbf{(Haklay, 2010)} conducted such a study to estimate of the quality of OSM positional geodata in London and England.
+Since the OSM project started in London it was thought that OSM data in the London metropolitan area would be representative of the highest quality data available, and therefore a good indicator for the whole global OSM dataset.
+The study was conducted by comparing OSM data to Ordnance Survey datasets.
+The study showed that the OSM dataset had a rough positional accuracy of 6 meters from the reference dataset.
+However, at the time of the study, the OSM project had captured roughly 29\% of the area of England, meaning that the data completeness was fairly low.
+
+The next notable study to assess the quality of OSM data was published by \textbf{(Kunze, 2012)}.
+The study applied different methods to assess the completeness of OSM data in two federal states in Germany, mainly by analysing the area difference between the OSM dataset and an administrative dataset.
+Finally, a notable study in 2013-2014, conducted by researchers from various universities in Norway and Germany \textbf{(Fan et al, 2014)}, examined the quality and accuracy in building footprint data of the Munich area, at a time and place where the completeness of the OSM dataset was significantly higher (the results of their study showed that the OSM had captured close to 100\% of all building footprints).
+At the time Munich was one of the most developed cities in OSM.
+Their study included an insight in the geometrical calculations neccessary to match features and points between two datasets, and how to reliably calculate the metrics needed for the most important and assessible of van Oorts criteria.
+Their study revealed that although the feature completeness of OSM was high, some architectural details were missing.
+The results reveal that the positional accuracy of OSM data at the time was about 4 meters.
+
 \chapter{Methodology and evaluation}
 
-In this chapter we will look in details at the evaluation metrics that this project served to obtain, and how they relate to the eventual collision analysis.
-For the geodata precision study specifically, we will look in detail at how geodata is qualifyied in general, by examining van Oorts quality criteria.
-We will also present how van Oorts critera will be used in the context of this project.
+This chapter will present the evaluation metrics that this project served to obtain, and how they relate to the eventual collision analysis.
+For the geodata precision study specifically, it will examine how geodata is qualifyied in general, by examining van Oorts quality criteria, and how these will be used in this project.
 See section 4 for a detailed technical implementation overview that presents how the relevant metrics were obtained during the statistics collection process.
@@
 
-\chapter{Background}
-
-This section will dive into the state-of-the-art use of city reconstruction methods within the research field of urban design.
-The first section will present the motivations for the digitization of urban planning, as well as the most common programs and tools in this area, which are after all not very commonly used in the practise of urban design due to the failure in being able to reflect the unique workspace of urban designers.
-The Related work section will present a brief history of the field of OpenStreetMap data qualification, and present the research papers whose methods are at the foundation of this project.
-
-\section{Background of 3D city reproduction}
-
-The project falls within the broader topic of geodata extraction and modelling, primarily of roads and building parcels.
-Geodata modeling is at the foundation of procedural modelling of real-world cities, since any application which models cities based on geodata will need a robust method of aqquiring and laying out road and building data without compromising the data's integrity.
-\textbf{(Müller et al, 2006)}
-
-Procedural modeling of road networks is of interest to the fields of urban planning and urban architecture.
-In recent years there has been an advent of interactive tools that aid urban planners in placing or generating features of an urban plan, particularly roads and road networks.
-Procedural and AI solutions that aid the arcitect in a form of human-machine symbiosis.
-
-Procedural city-generation models combine architectural elemets, such as streets, lots, buildings, building facades.
-Urban design practises with conventional two-dimenstional deliverables range from delivering neighbourhood designs, urban zoning and building plans, to writing design codes and building ordinances.
-The last decade has seen multiple startups in the area of software that aid urban arcitecture.
-The software uses procedural models to render a city grid based on real-world locations, or generate new user-specified parameters, and allow for varying degrees of manual editing on grid, road, neighbourhood, city block or individual feature level.
-Some of the most prominent examples are ArcGIS Pro (made by Esri), CityEngine (made by the startup Procedural, bought by ESRI) and Urban Canvas (made by the startup Synthicity, bought by Autodesk, now defunct).
-
-But such tools are as of today not commonly applied in urbanist practises, due to failures in reflecting the unique workspace of arcitects and their design needs.
-Urban arcitecture is a social factors, stakeholders.
-Each arcitect is an actor in a continously developing environment, and affecting that environment in a complex manner through their action \textbf{(Stojanovski et al, 2020)}.
-
-In recent years, there have been an advent of tools that extract geodata from OSM and generate 3D representations of real world locations by placing terrain, buildings, roads and other features.
-However, freely available such tools seldom include smaller roads and pedestrian footpaths such as walkways, sidewalks or trails.
-
-Additionally, the topic is of interest for 3D content creators in for example the fields of video game design or 3D animation.
-A trend in these fields is using methods of procedural modeling to create large amounts of 3D content quickly and efficiently, while requiring as little hand modeling as possible.
-Studios who create large worlds are often interested in solutions that generate large road networks without any hand modeling.
-\textbf{(Parish and Müller, 2001)}
-
-\section{Related work}
-When assessing the completeness of geodata, a very common scale is Van Oorts criteria for evaluating the quality of geographical information, which at the time of publication received attention from surveyors and cartographers.
-\textbf{(Van Oort, 2006)} Since then a number of publications have examined slices of geodata from van Oort's criteria.
-In 2008, \textbf{(Haklay, 2010)} conducted such a study to estimate of the quality of OSM positional geodata in London and England.
-Since the OSM project started in London it was thought that OSM data in the London metropolitan area would be representative of the highest quality data available, and therefore a good indicator for the whole global OSM dataset.
-The study was conducted by comparing OSM data to Ordnance Survey datasets.
-The study showed that the OSM dataset had a rough positional accuracy of 6 meters from the reference dataset.
-However, at the time of the study, the OSM project had captured roughly 29\% of the area of England, meaning that the data completeness was fairly low.
-
-The next notable study to assess the quality of OSM data was published by \textbf{(Kunze, 2012)}.
-The study applied different methods to assess the completeness of OSM data in two federal states in Germany, mainly by analysing the area difference between the OSM dataset and an administrative dataset.
-Finally, a notable study in 2013-2014, conducted by researchers from various universities in Norway and Germany \textbf{(Fan et al, 2014)}, examined the quality and accuracy in building footprint data of the Munich area, at a time and place where the completeness of the OSM dataset was significantly higher (the results of their study showed that the OSM had captured close to 100\% of all building footprints).
-At the time Munich was one of the most developed cities in OSM.
-Their study included an insight in the geometrical calculations neccessary to match features and points between two datasets, and how to reliably calculate the metrics needed for the most important and assessible of van Oorts criteria.
-Their study revealed that although the feature completeness of OSM was high, some architectural details were missing.
-The results reveal that the positional accuracy of OSM data at the time was about 4 meters.
-
 \chapter{Implementation}
 
-In section 4.1 we will look at the data collection process for the geodata precision study.
-We will look at how to qualify dataset completeness and building correspondence (how many of the buildings in one dataset that can be found in the other) by collecting the cumulative building areas and looking at what buildings have overlapping areas between the two datasets.
-We will also look at the difinitions and algorithms neccessary for calculating posiitonal accuracy and a measure for shape accuracy.
-In section 4.2 we will look at the data collection process for the collision study.
-We will look at the how standard road withs were obtained, as well as the algorithm for detemining feature collision that was used.
-In section 4.3 we will discuss briefly the data sources used in this project, why they were chosen and specific issues with the data collection process.
+Section 4.1 will present the data collection process for the geodata precision study.
+It will present how to qualify dataset completeness and building correspondence (how many of the buildings in one dataset that can be found in the other) by collecting the cumulative building areas and looking at what buildings have overlapping areas between the two datasets.
+The definitions and algorithms neccessary for calculating posiitonal accuracy and a measure for shape accuracy will also be presented.
+Section 4.2 will present at the data collection process for the collision study.
+It will present how standard road withs were obtained, as well as the algorithm for detemining feature collision that was used.
+Section 4.3 will discuss briefly the data sources used in this project, why they were chosen and specific issues with the data collection process.
 
@@
 
-Finally, when two building footprints and their vertices have been mapped, we will use the offset of matching vertices as a measure of positional error.
+Finally, when two building footprints and their vertices have been mapped, the offset of matching vertices will be used as a measure of positional error.
 The average and variance in vertex spread will be calculated and used to make an assessment of the geometric precision of the OSM dataset, and this precision will further be used to form an upper boundary on geometrical corrections that can be made to the map once it has been imported into the application.
@@
 
-The maximum distance between two features was set to 1 times the positional accuracy, since then there will always be ample room to make changes to any features that would collide with the newly moved nodes, and thus we do not need to worry about feature collisions propagating.
+The maximum distance between two features was set to 1 times the positional accuracy, since then there will always be ample room to make changes to any features that would collide with the newly moved nodes, and thus collisions propagating will not be an issue.
 In this implementation, roads of different types were allowed to merge into one another to form common intersections.
@@
 
-\chapter{Results and metrics}
-
-Here we present the results of the statistics collections that were made as part of this project.
+\chapter{Evaluation}
+
+\section{Results}
+
+Here the results of the statistics collections will be presented.
 First the results of the collision study is presented as how many of the road features of each type display collision with some other feature.
 Then the results of the geodata precision is presented.
-We will present our estimate of the positional accuracy of the OSM dataset, as well as the dataset completeness and statistics about building correspondence and shape accuracy.
-
-\section{Road collision study}
+The final estimate of the positional accuracy of the OSM dataset will be presented, as well as the dataset completeness and statistics about building correspondence and shape accuracy.
+
+\subsection{Road collision study}
 
@@
 
-\section{Positional accuracy study}
+\subsection{Positional accuracy study}
 
@@
 
-\section{Secondary metrics}
-
-\subsection{Completeness by building area}
+\subsection{Secondary metrics}
+
+\subsubsection{Completeness by building area}
 
@@
 
-\subsection{Building correspondence statistics}
+\subsubsection{Building correspondence statistics}
 
@@
 
-\subsection{Shape accuracy statistics}
+\subsubsection{Shape accuracy statistics}
 
@@
 
-\chapter{Analysis}
-
-Now that the colliison metrics have been obtained, we are in a good position to draw conclusions about the technical feasibility of 3D city reproduction from OSM data.
-We will begin by looking at the geodata precision study, and use the outcome of that study to assess the quality of the positional accuracy estimate, by looking at the dataset completeness, the low semantic accuracy and the unique characteristics of the Stockholm area.
-Secondly we will look at the collision study.
-We will make the claim that footpaths will not be more problematic than other road types in city reproduction from OSM data, and motivate this with the data obtained in section 6.1.
-We will look at the character of the feature collision cases and identify the most common cases, as well as present geometrical algorithms for how to correct each case automatically.
+\section{Analysis}
+
+Now that the colliison metrics have been obtained, conclusions may be drawn about the technical feasibility of 3D city reproduction from OSM data.
+First the geodata precision study will be discussed.
+The outcome of that study will be used to assess the quality of the positional accuracy estimate by looking at the dataset completeness, the low semantic accuracy and the unique characteristics of the Stockholm area.
+Secondly the collision study will be discussed.
+The main outcome is that footpaths will not be more problematic than other road types in city reproduction from OSM data.
+This claim will be motivated with the data obtained in section 6.1.
+Finally, the character of the feature collision cases will be discussed and the most common cases will be identified.
+The chapter will conclude with a presentation of geometrical algorithms for how to correct each case automatically.
 For conclusions about the usage of OSM city reproduction in engineering practise and remaining challenges, see section 7.
 
-\section{Quality of the OSM dataset}
+\subsection{Quality of the OSM dataset}
 
@@
 % Onsdag:
+% TODO: Delimitations in Introduction, after RQ
+% TODO: Move: 2.1.1: Accuracy criteria that will not be focused on -> Delimitations
+
+% Torsdag:
+% TODO: Mention that positional accuracy will be the focus early on
+% TODO: Move content from chapters 1, 3 and 4 to Background
 % TODO: Split up introduction, more subsections
-% TODO: Delimitations in Introduction, after RQ
 % TODO: Go over all images, add a line about their origins
-% TODO: Merge: Results and metrics + Analysis -> Evaluation, with result/analysis subsections
 % TODO: Rename: Implementation overview -> Methodology
-% TODO: Background should be the second chapter
-
-% Torsdag:
+
+% Internet:
+% TODO: Switch to numerical citations
 % TODO: Add source of image on page 1
-% TODO: Mention that positional accuracy will be the focus early on
-% TODO: Move: 2.1.1: Accuracy criteria that will not be focused on -> Delimitations
-% TODO: Formal tone, never use us, we, me
-% TODO: Move content from chapters 1, 3 and 4 to Background
-% TODO: Switch to numerical citations
 
@@
 
-\section{Collision study}
+\subsection{Collision study}
 
@@
 Looking closer at the individual collision cases, the next conclusion is that most often it is actually a small part of a road that is colliding, and will need correction.
-We make an exemplifying calculation to support this claim.
-Looking in tables \ref{table:road-collision} and \ref{table:road-collision-corrected}, we note that before road extrusion, 3189 features and 7330 of their edges are intersecting some other feature.
+Now will follow an exemplifying calculation to support this claim.
+Looking in tables \ref{table:road-collision} and \ref{table:road-collision-corrected}, note that before road extrusion, 3189 features and 7330 of their edges are intersecting some other feature.
 This means that for each colliding road, on average 2.3 of its edges will be intersecting some other feature, thus bearing responsibility for the collision.
@@
 
-\section{Suggested algorithms for collision correction}
+\subsection{Suggested algorithms for collision correction}
 
@@
 Cases 1 and 3-5 will be handled, while 2 will be left since case 2 collisions do not inherently cause feature overlap when extruding a 3D road mesh from the way edges.
-Thus it will me up to the individual developer to decide what to do with Case 2 collisions.
+Thus it will be up to the individual developer to decide what to do with Case 2 collisions.
 
@@
 
-We now have a firm idea of the technical feasibility of generating virtual cities from OpenStreetMap-data, including different types of road networks.
-We have however yet to touch upon the possibilities of road network modelling in the broader context of 3D city reproduction, and what are the remaining challenges for accurate 3D city reproduction?
-What techniques can we use to complement our basic system of importing street and building data and create a rich and living virtual city?
-In this section we will place everything we have learnt so far in the broader context of GIS and city modeling, focusing on the research fields of 3D media and city planning.
+This report has now established a firm idea of the technical feasibility of generating virtual cities from OpenStreetMap-data, including different types of road networks.
+It remains to touch upon the possibilities of road network modelling in the broader context of 3D city reproduction. What are the remaining challenges for accurate 3D city reproduction?
+What techniques can be used to complement the basic system of importing street and building data that is already in place, and create a rich and living virtual city?
+In this section everything we have learnt so far will be placed in the broader context of GIS and city modeling, focusing on the research fields of 3D media and city planning.
 
@@
 As has been mentiond in the introduction to this report, procedural generaion of road networks is a basal requirement for quick and low-effort modeling virtual cities.
-Since the road network ties together the city, and forms the grid that divides the lots, quarters and neighbourhoods, it make sense that we can only continue with the remaining challenges once we can a robust procedure for extracting road geodata and generating the 3D represention of it without compromising the data integrity.
+Since the road network ties together the city, and forms the grid that divides the lots, quarters and neighbourhoods, it make sense that one can only continue with the remaining challenges once there exists a robust procedure for extracting road geodata and generating the 3D represention of it without compromising the data integrity.
 The same is true for building modelling.
 It turns out that the same methods that have been used in this project to obtain and qualify road network data can be applied to building footprint data from OSM. \textbf{(Fan et al, 2014)}
-And only when we have such a fundamental procedure for accuratly importing building footprints without compromising the data integritey, we can proceed with the remaining challenges in building modeling.
+And only when there exists such a fundamental procedure for accuratly importing building footprints without compromising the data integritey, one can proceed with the remaining challenges in building modeling.
 \textbf{(Müller et al, 2006)}
@@
 \textbf{(Biljecki et al, 2016)}
-We may chose what LOD class(es) to use depending on data availablity and application.
-It turns out that, with the current means of data aquisition from OSM being limited to building footprints, using OSM data we can only produce LOB-B models at best.
-If we wish to aqquire building assets at a LOD-C level, we would have to complete the data with other sources, or rely on procedural methods to populate our crude city model with additional details and building features that give the model more aestethic depth, but do not neccessarily reflect reality.
+A modeler may chose what LOD class(es) to use depending on data availablity and application.
+It turns out that, with the current means of data aquisition from OSM being limited to building footprints, using OSM data a model can only produce LOB-B models at best.
+If a modeler wishes to aqquire building assets at a LOD-C level, they would have to complete the data with other sources, or rely on procedural methods to populate crude city models with additional details and building features that give the models more aestethic depth, but do not neccessarily reflect reality.
 

### 2020-06-03 report/report.tex
Rename section: Implementation overview -> Methodology
Add notes about figure origins

@@
     \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_map_residential_footpath_compare}
-    \caption{Example segments of the Stockholm metropolitan area with highlighted residential streets (left) and footpaths (right).
+    \caption{Example segments of the Stockholm metropolitan area in OpenStreetMap, with highlighted residential streets (left) and footpaths (right).
 Here can be seen the unique character of footpaths.
-Walkways in parks appear to grow much more organically, and pavements form rough grid systems with organic branches that bridge over city blocks.}
+Walkways in parks appear to grow much more organically, and pavements form rough grid systems with organic branches that bridge over city blocks.
+The image was generated using GitHub's geojson viewer.}
     \label{fig:residential-footpath-map}
@@
 
-\section{Implementation overview}
+\section{Methodology}
 
@@
     \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_building_match}
-    \caption{Examples of a 1:1 building match (left), where the sought-after building was found both in the OSM (red) and SLU (blue) datasets, and of a 1:n building match (right), where the sought-after building was represented as one polygon in the OSM (red) dataset, but divided into many smaller lots in the SLU (multicolored) dataset.}
+    \caption{Examples of a 1:1 building match (left), where the sought-after building was found both in the OSM (red) and SLU (blue) datasets, and of a 1:n building match (right), where the sought-after building was represented as one polygon in the OSM (red) dataset, but divided into many smaller lots in the SLU (multicolored) dataset.
+The buildings are examples from the OSM and SLU datasets that were prcessed in this projects, and were taken from the Stockholm area.}
     \label{fig:building-match-types}
@@
 
-The building footprint correspondence will be evaluated using a similarity function that depends on the difference between the turning funcitons of matching building footprints in both datasets.
+The building footprint correspondence will be evaluated using a similarity function that depends on the difference between the turning functions of matching building footprints in both datasets.
 The turning function was first defined by \textbf{(Arkin et al, 1991)}, as a method for measuring the similarity of two polygons.
@@
     \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_turn_function}
-    \caption{An example of the turning function of a polygon. The edges, vertices and angle of the initial step of the turning funciton are marked.}
+    \caption{An example of the turning function of a polygon.
+The edges, vertices and angle of the initial step of the turning function are marked.
+The building in the figure was taken from the OSM dataset in the Stockholm area.}
     \label{fig:turning-function}
@@
 % TODO: Go over all images, add a line about their origins
-% TODO: Rename: Implementation overview -> Methodology
 

### 2020-06-04 report/report.tex
Update todo
Delimitations section

@@
 For the positional accuracy study, see sections 1.3.2 for an implementation overview, 4.1 for a walkthrough of the program implementation, and 5.2 for the results.
+
+\section{Delimitations}
+
+in this study, when working with osm geodata, the following assumptions have been made about the dataset to reduce the size of the problem domain:
+
+\begin{enumerate}
+    \item the program will not handle terrain features and altitude. it will be assumed that the road mesh can be projected on a flat 2d surface without feature intersection.
+    \item it will be assumed that the city map consists of only arterial (primary) roads, secondary roads and footpaths.
+    \item it will be assumed that the osm dataset represents all roads as simple polylines, and building footprints as simple connected polygons.
+any other features (such as squares, which are roads represented as polygons) will be omitted from the dataset.
+    \item the road network is assumed to be a 2d simple graph.
+this means that any overlapping roads, such as tunnels and bridges that cross over street-level roads, will be eliminated from the dataset.
+any self-connected nodes will also be eliminated from the dataset.
+\end{enumerate}
 
@@
 The temporal accuracy will not be assessed as this projects focuses on comparing a single snapshot of the two datasets at the same point in time, and finally the logical accuracy will not be assessed because the underlying logical relations between features in both maps serve no purpose to this project.
-
-\section{Problem constraints}
-
-In this study, when working with OSM geodata, the following assumptions have been made about the dataset to reduce the size of the problem domain:
-
-\begin{enumerate}
-    \item The program will not handle terrain features and altitude. It will be assumed that the road mesh can be projected on a flat 2D surface without feature intersection.
-    \item It will be assumed that the city map consists of only arterial (primary) roads, secondary roads and footpaths.
-    \item It will be assumed that the OSM dataset represents all roads as simple polylines, and building footprints as simple connected polygons.
-Any other features (such as squares, which are roads represented as polygons) will be omitted from the dataset.
-    \item The road network is assumed to be a 2D simple graph.
-This means that any overlapping roads, such as tunnels and bridges that cross over street-level roads, will be eliminated from the dataset.
-Any self-connected nodes will also be eliminated from the dataset.
-\end{enumerate}
 
@@
 
-% Onsdag:
-% TODO: Delimitations in Introduction, after RQ
-% TODO: Move: 2.1.1: Accuracy criteria that will not be focused on -> Delimitations
-
 % Torsdag:
@@
 % TODO: Split up introduction, more subsections
-% TODO: Go over all images, add a line about their origins
+% TODO: Delimitations in Introduction, after RQ
+% TODO: Move: 2.1.1: Accuracy criteria that will not be focused on -> Delimitations
 

### 2020-06-04 report/report.tex
Split up Intructions in subsections
TOC depth 1
Update todo
Organize delimitations and implementation overview into Methodology
Organize background section

@@
 \setlength{\parskip}{1em}
-\setcounter{tocdepth}{2}
+\setcounter{tocdepth}{1}
 
@@
 
+\section{Introduction to virtual cities}
+
 Virtual cities has in recent years been a popular research topic in the field of computer visualization.
@@
 
-Cities consist of complex collections of buildings, blocks, parcels and neighbourhoods which are all interconnected via a street network, where individual streets can serve different types of traffic.
-These features are commonly represented as geodata in map databases, and processed and organized into visualization-ready data by a Geographic Information System (GIS).
-In practise, cities are often made up out of many square kilometers worth of geodata, and it would be highly labour intensive to recreate all the features in 3D.
-Instead, procedural models for city reproduction instead rely on programmatic solutions for converting map data to 3D road and building models, and composing them to a 3D world that the designer may continue to fine-tweak through interfaces fo interactive editing.
-At the very base of this process is the procedural layout of road networks in 3D,
-\textbf{(Müller et al, 2006)}
-and so far there have been a lot of research in modelling of 3D road networks, but these projects have often focused on the road networks designated for cars and public transit, as a part of the sustainable mobility paradigm and the topic of sustainable cities, particularily due to the interest in self-propelled cars.
-The study of energy efficient mobility such as walking and cycling is only in its emerging phase. \textbf{(Stojanovski, 2019)}
-
 \begin{figure}[H]
@@
 \end{figure}
+
+\section{Accuracy limitations in 3D city reproduction}
 
@@
 
-\section{Delimitations}
-
-in this study, when working with osm geodata, the following assumptions have been made about the dataset to reduce the size of the problem domain:
-
-\begin{enumerate}
-    \item the program will not handle terrain features and altitude. it will be assumed that the road mesh can be projected on a flat 2d surface without feature intersection.
-    \item it will be assumed that the city map consists of only arterial (primary) roads, secondary roads and footpaths.
-    \item it will be assumed that the osm dataset represents all roads as simple polylines, and building footprints as simple connected polygons.
-any other features (such as squares, which are roads represented as polygons) will be omitted from the dataset.
-    \item the road network is assumed to be a 2d simple graph.
-this means that any overlapping roads, such as tunnels and bridges that cross over street-level roads, will be eliminated from the dataset.
-any self-connected nodes will also be eliminated from the dataset.
-\end{enumerate}
-
-\section{Methodology}
-
-Programmatic collection of statistics has been neccessary for both the primary and secondary research questions, and this section presents a broad overview of the technical implementation behind the map analysis and data collection processes for both the geodata precision study and the collision study.
-
-\subsection{Geodata precision study}
-
-The geometrical precision study has been carried out in a number of steps that are all detailed below.
-First a feature map was found, which mapped individual features between the OSM dataset and a reference dataset.
-A program then took matching building pairs and extracted matching geometrical points after first simplifying the polygons using the Douglas-Peucker algorithm \textbf{(Douglas and Peucker, 1973)}.
-Once a set of matching points has been found across the whole dataset, the average distance between them was taken as a measure of the positional accuracy.
-
-Additionally, a number of auxillary calculations were neccessary for this study.
-The estimated completeness of the OSM dataset (a measure of how many features are represented in both the OSM dataset and the reference dataset) was calculated as the relative difference in area of all building footprints, between the two datasets.
-The completeness is needed to estimate the error boundary of the positional accuracy.
-Additionally, the shape accuracy of the OSM dataset has also been calculated and graphed as an extra verification of the similarity between the datasets.
-
-\subsection{Collision study}
-
-Once the estimate of the positional error was obtained from the precision study, this estimate was used to assess the integrity of the OSM Stockholm road network, by locating critical points where assigning standard widths to the roads would cause collision with other features, and evaluating how many of those points could be eliminated by making ajustments within the found margin of error.
-For the purpose of the collision study, each road has been assigned a standard width that is in line with Swedish regulations, since in the OSM dataset, roads are represented as simple polylines, lacking width.
-In a 3D representation however, it is crucial that each road object has a spacial width in order to be rendered and seen.
-After the width assignment, a program has iterated over all features and calculated how many of them that overlap with any other feature, as well as the total length of the road polylines that is intersecting some other feature.
-The program has then calculated how many of these colliding features that can be corrected by making adjustments that are within the positional accuracy of the OSM data (As obtained by the previous study in this report: Assessment of Geodata Precision).
-
 \chapter{Background}
@@
 This section will dive into the state-of-the-art use of city reconstruction methods within the research field of urban design.
-The first section will present the motivations for the digitization of urban planning, as well as the most common programs and tools in this area, which are after all not very commonly used in the practise of urban design due to the failure in being able to reflect the unique workspace of urban designers.
-The Related work section will present a brief history of the field of OpenStreetMap data qualification, and present the research papers whose methods are at the foundation of this project.
-
-\section{Background of 3D city reproduction}
+This chapter will present the motivations for the digitization of urban planning, as well as the most common programs and tools in this area, which are after all not very commonly used in the practise of urban design due to the failure in being able to reflect the unique workspace of urban designers.
+The final section will present a brief history of the field of OpenStreetMap data qualification, and present the research papers whose methods are at the foundation of this project.
+
+\section{Overview of virtual cities}
+
+Cities consist of complex collections of \textbf{buildings}, \textbf{blocks}, \textbf{parcels} and \textbf{neighbourhoods} which are all interconnected via a \textbf{street} network, where individual streets can serve different types of traffic.
+These features are commonly represented as \textbf{geodata} in map databases, and processed and organized into visualization-ready data by a \textbf{Geographic Information System (GIS)}.
+In practise, cities are often made up out of many square kilometers worth of geodata, and it would be highly labour intensive to recreate all the features in 3D.
+Instead, \textbf{procedural models} for city reproduction instead rely on programmatic solutions for converting map data to 3D road and building models, and composing them to a 3D world that the designer may continue to fine-tweak through interfaces fo interactive editing.
+At the very base of this process is the procedural layout of road networks in 3D,
+\textbf{(Müller et al, 2006)}
+and so far there have been a lot of research in modelling of 3D road networks, but these projects have often focused on the road networks designated for cars and public transit, as a part of the sustainable mobility paradigm and the topic of sustainable cities, particularily due to the interest in self-propelled cars.
+The study of energy efficient mobility such as walking and cycling is only in its emerging phase. \textbf{(Stojanovski, 2019)}
+
+\section{3D city reproduction as a research field}
 
@@
 
-\section{Related work}
+\section{Previous work in qualifying OpenStreetMap data}
+
 When assessing the completeness of geodata, a very common scale is Van Oorts criteria for evaluating the quality of geographical information, which at the time of publication received attention from surveyors and cartographers.
@@
 
-\subsection{Other accuracy criteria}
+\section{Delimitations}
+
+in this study, when working with osm geodata, the following assumptions have been made about the dataset to reduce the size of the problem domain:
+
+\begin{enumerate}
+    \item the program will not handle terrain features and altitude. it will be assumed that the road mesh can be projected on a flat 2d surface without feature intersection.
+    \item it will be assumed that the city map consists of only arterial (primary) roads, secondary roads and footpaths.
+    \item it will be assumed that the osm dataset represents all roads as simple polylines, and building footprints as simple connected polygons.
+any other features (such as squares, which are roads represented as polygons) will be omitted from the dataset.
+    \item the road network is assumed to be a 2d simple graph.
+this means that any overlapping roads, such as tunnels and bridges that cross over street-level roads, will be eliminated from the dataset.
+any self-connected nodes will also be eliminated from the dataset.
+\end{enumerate}
 
@@
 The temporal accuracy will not be assessed as this projects focuses on comparing a single snapshot of the two datasets at the same point in time, and finally the logical accuracy will not be assessed because the underlying logical relations between features in both maps serve no purpose to this project.
+
+\section{Implementation overview}
+
+Programmatic collection of statistics has been neccessary for both the primary and secondary research questions, and this section presents a broad overview of the technical implementation behind the map analysis and data collection processes for both the geodata precision study and the collision study.
+
+\subsection{Geodata precision study}
+
+The geometrical precision study has been carried out in a number of steps that are all detailed below.
+First a feature map was found, which mapped individual features between the OSM dataset and a reference dataset.
+A program then took matching building pairs and extracted matching geometrical points after first simplifying the polygons using the Douglas-Peucker algorithm \textbf{(Douglas and Peucker, 1973)}.
+Once a set of matching points has been found across the whole dataset, the average distance between them was taken as a measure of the positional accuracy.
+
+Additionally, a number of auxillary calculations were neccessary for this study.
+The estimated completeness of the OSM dataset (a measure of how many features are represented in both the OSM dataset and the reference dataset) was calculated as the relative difference in area of all building footprints, between the two datasets.
+The completeness is needed to estimate the error boundary of the positional accuracy.
+Additionally, the shape accuracy of the OSM dataset has also been calculated and graphed as an extra verification of the similarity between the datasets.
+
+\subsection{Collision study}
+
+Once the estimate of the positional error was obtained from the precision study, this estimate was used to assess the integrity of the OSM Stockholm road network, by locating critical points where assigning standard widths to the roads would cause collision with other features, and evaluating how many of those points could be eliminated by making ajustments within the found margin of error.
+For the purpose of the collision study, each road has been assigned a standard width that is in line with Swedish regulations, since in the OSM dataset, roads are represented as simple polylines, lacking width.
+In a 3D representation however, it is crucial that each road object has a spacial width in order to be rendered and seen.
+After the width assignment, a program has iterated over all features and calculated how many of them that overlap with any other feature, as well as the total length of the road polylines that is intersecting some other feature.
+The program has then calculated how many of these colliding features that can be corrected by making adjustments that are within the positional accuracy of the OSM data (As obtained by the previous study in this report: Assessment of Geodata Precision).
 
@@
 % Torsdag:
+% TODO: Move content from chapters 1, 3 and 4 to Background
 % TODO: Mention that positional accuracy will be the focus early on
-% TODO: Move content from chapters 1, 3 and 4 to Background
-% TODO: Split up introduction, more subsections
-% TODO: Delimitations in Introduction, after RQ
-% TODO: Move: 2.1.1: Accuracy criteria that will not be focused on -> Delimitations
 

### 2020-06-04 report/report.tex
Specify todo about problem specification in the introduction, do this lates when the layout is done

@@
 
-When it comes to the quality of available data, it turns out that in practise, 3D GIS datasets without errors are rare.
+% TODO: Mention that positional accuracy will be focused on, and the implications of not fixing errors when converting to a 3D model
+When it comes to the quality of available data, it turns out that in practise, geodata databases without errors are rare.
 Among the most common errors are missing or invalid geometry, and invalid semantics.
 However, \textbf{(Biljecki et al, 2016)} note that there are coming more and more sophisticated methods for automatically correcting the most common errors.
-GIS validation and correction is only an emerging field of research.
+Geodata validation and correction is only an emerging field of research.
 and many such datasets were usually obtained years ago, and creating error-free datasets was neither possible due to the lack of validation techniques, nor a high priority.
@@
 
-% Torsdag:
-% TODO: Move content from chapters 1, 3 and 4 to Background
-% TODO: Mention that positional accuracy will be the focus early on
-
 % Internet:

### 2020-06-04 report/report.tex
Add source for virtual city image

@@
 Creating 3D virtual cities using methods of procedual generation is an emerging research field that comes with its own set of challenges.
-This project deals with generating virtual cities from OpenStreetMap data, and the technical challenges therein.}
+This project deals with generating virtual cities from OpenStreetMap data, and the technical challenges therein.$^0$}
     \label{fig:virtual-city}
 \end{figure}
+
+\footnotetext{Image source: Downtown Boston 3D City Model in Infraworks 360, available via: https://gallery.autodesk.com/civilinfrastructure/projects/490/downtown-boston-3d-city-model-in-infraworks-360}
 

### 2020-06-04 report/report.tex
Update todo
Format citations to provide references

@@
 
-Virtual cities has in recent years been a popular research topic in the field of computer visualization.
+Virtual cities has in recent years been a popular research topic in the field of computer visualization, according to \textbf{(Vanegas et al, 2010)}.
 The idea of recreating real world cities in 3D applications has a growing number of applications, including visualization and modelling, urban planning and architecture, as well as entertainment in the form of video games and 3D animation.
@@
 Urban environments are typically dense and large, spanning a few to hundreds of square kilometers, and are influenced by variables that are hard to quantify, including land policies, government plans, population changes and transportation infrastructure.
-\textbf{(Vanegas et al, 2010)}.
 
@@
 
-Cities consist of complex collections of \textbf{buildings}, \textbf{blocks}, \textbf{parcels} and \textbf{neighbourhoods} which are all interconnected via a \textbf{street} network, where individual streets can serve different types of traffic.
+In their analysis of prodedural city generation models, \textbf{(Müller et al, 2006)} provide an overview of the process of recreating virtual 3D citites from their real life conterparts.
+They present cities as complex collections of \textbf{buildings}, \textbf{blocks}, \textbf{parcels} and \textbf{neighbourhoods} which are all interconnected via a \textbf{street} network, where individual streets can serve different types of traffic.
 These features are commonly represented as \textbf{geodata} in map databases, and processed and organized into visualization-ready data by a \textbf{Geographic Information System (GIS)}.
@@
 Instead, \textbf{procedural models} for city reproduction instead rely on programmatic solutions for converting map data to 3D road and building models, and composing them to a 3D world that the designer may continue to fine-tweak through interfaces fo interactive editing.
-At the very base of this process is the procedural layout of road networks in 3D,
-\textbf{(Müller et al, 2006)}
-and so far there have been a lot of research in modelling of 3D road networks, but these projects have often focused on the road networks designated for cars and public transit, as a part of the sustainable mobility paradigm and the topic of sustainable cities, particularily due to the interest in self-propelled cars.
-The study of energy efficient mobility such as walking and cycling is only in its emerging phase. \textbf{(Stojanovski, 2019)}
+At the very base of this process is the procedural layout of road networks in 3D, and so far there have been a lot of research in modelling of 3D road networks.
+\textbf{(Stojanovski, 2019)} however claims that these projects have often focused on the road networks designated for cars and public transit, as a part of the sustainable mobility paradigm and the topic of sustainable cities, particularily due to the interest in self-propelled cars.
+The study of energy efficient mobility such as walking and cycling is only in its emerging phase.
 
@@
 The project falls within the broader topic of geodata extraction and modelling, primarily of roads and building parcels.
-Geodata modeling is at the foundation of procedural modelling of real-world cities, since any application which models cities based on geodata will need a robust method of aqquiring and laying out road and building data without compromising the data's integrity.
-\textbf{(Müller et al, 2006)}
-
+Geodata modeling is at the foundation of procedural modelling of real-world cities (according to \textbf{(Müller et al, 2006)}), since any application which models cities based on geodata will need a robust method of aqquiring and laying out road and building data without compromising the data's integrity.
 Procedural modeling of road networks is of interest to the fields of urban planning and urban architecture.
@@
 Some of the most prominent examples are ArcGIS Pro (made by Esri), CityEngine (made by the startup Procedural, bought by ESRI) and Urban Canvas (made by the startup Synthicity, bought by Autodesk, now defunct).
-
-But such tools are as of today not commonly applied in urbanist practises, due to failures in reflecting the unique workspace of arcitects and their design needs.
+However, freely available such tools seldom include smaller roads and pedestrian footpaths such as walkways, sidewalks or trails.
+
+Even as such tools are well available, \textbf{(Stojanovski et al, 2020)} notes that they are as of today not commonly applied in urbanist practises, due to failures in reflecting the unique workspace of arcitects and their design needs.
 Urban arcitecture is a social factors, stakeholders.
-Each arcitect is an actor in a continously developing environment, and affecting that environment in a complex manner through their action \textbf{(Stojanovski et al, 2020)}.
-
-In recent years, there have been an advent of tools that extract geodata from OSM and generate 3D representations of real world locations by placing terrain, buildings, roads and other features.
-However, freely available such tools seldom include smaller roads and pedestrian footpaths such as walkways, sidewalks or trails.
-
-Additionally, the topic is of interest for 3D content creators in for example the fields of video game design or 3D animation.
-A trend in these fields is using methods of procedural modeling to create large amounts of 3D content quickly and efficiently, while requiring as little hand modeling as possible.
+Each arcitect is an actor in a continously developing environment, and affecting that environment in a complex manner through their action.
+
+Finally, wen it comes to the fields of video game design or 3D animation, tools for city reproduction is undoubtedly of use to 3D content creators.
+In their study of procedurally generated buildings, \textbf{(Parish and Müller, 2001)} claim that there is a widespread trend in these fields to use methods of procedural modeling to create large amounts of 3D content quickly and efficiently, while requiring as little hand modeling as possible.
 Studios who create large worlds are often interested in solutions that generate large road networks without any hand modeling.
-\textbf{(Parish and Müller, 2001)}
 
@@
 
-When assessing the completeness of geodata, a very common scale is Van Oorts criteria for evaluating the quality of geographical information, which at the time of publication received attention from surveyors and cartographers.
-\textbf{(Van Oort, 2006)} Since then a number of publications have examined slices of geodata from van Oort's criteria.
+When assessing the completeness of geodata, a very common scale is Van Oorts criteria for evaluating the quality of geographical information, which at the time of publication received attention from surveyors and cartographers (\textbf{(Van Oort, 2006)}).
+Since then a number of publications have examined slices of geodata from van Oort's criteria.
 In 2008, \textbf{(Haklay, 2010)} conducted such a study to estimate of the quality of OSM positional geodata in London and England.
@@
 The study applied different methods to assess the completeness of OSM data in two federal states in Germany, mainly by analysing the area difference between the OSM dataset and an administrative dataset.
-Finally, a notable study in 2013-2014, conducted by researchers from various universities in Norway and Germany \textbf{(Fan et al, 2014)}, examined the quality and accuracy in building footprint data of the Munich area, at a time and place where the completeness of the OSM dataset was significantly higher (the results of their study showed that the OSM had captured close to 100\% of all building footprints).
+Finally, a notable study in 2013-2014, conducted by researchers from various universities in Norway and Germany (\textbf{(Fan et al, 2014)}), examined the quality and accuracy in building footprint data of the Munich area, at a time and place where the completeness of the OSM dataset was significantly higher (the results of their study showed that the OSM had captured close to 100\% of all building footprints).
 At the time Munich was one of the most developed cities in OSM.
@@
 
-\textbf{(Haklay, 2010)} presents a comprehesive list of 8 accuracy classes when it comes to evaluating the quality of geographic information (van Oorts criteria)
+\textbf{(Haklay, 2010)} presents in a comprehesive manner the list of van Oort's 8 accuracy classes for evaluating the quality of geographic information:
 
@@
 First a feature map was found, which mapped individual features between the OSM dataset and a reference dataset.
-A program then took matching building pairs and extracted matching geometrical points after first simplifying the polygons using the Douglas-Peucker algorithm \textbf{(Douglas and Peucker, 1973)}.
+A program then took matching building pairs and extracted matching geometrical points after first simplifying the polygons using the Douglas-Peucker algorithm by \textbf{(Douglas and Peucker, 1973)}.
 Once a set of matching points has been found across the whole dataset, the average distance between them was taken as a measure of the positional accuracy.
@@
 Problems that arise are e.g. that the vertex counts are dissimilar, or that vertex clusters may be found at different parts of the polygon in the two datasets.
-To avoid this effect, key points are extracted using the Douglas-Peucker algorithm \textbf{(Douglas and Peucker, 1973)}, to create a simplifed footprint with less points that still retain information about the rough features of the detailed footprint.
+To avoid this effect, key points are extracted using the Douglas-Peucker algorithm (\textbf{(Douglas and Peucker, 1973)}), to create a simplifed footprint with less points that still retain information about the rough features of the detailed footprint.
 The idea behind Douglas-Peucker is to recursively divide a polyline.
@@
 OSM road data and building footprints are commonly obtained by manually tracing features in commercially available satellite images.
-Such images have a limit on their resolution which puts a theoretical limit on the accuracy of map features compared to their real-world equivalents (Haklay 2010).
-Particularily the high resolution imagery from Bing in 2010 led to an increase in building information in OSM, and the positional error is largely due to the limited resolution in such satellite imagery.
-\textbf{(Fan et al, 2014)}.
+Such images have a limit on their resolution which puts a theoretical limit on the accuracy of map features compared to their real-world equivalents
+The driving force behind the availablity of VGI is, according to \textbf{(Fan et al, 2014)}, the availablity of high resolution satellite imagery to private individuals.
+They claim that particularily the imagery from Bing Maps, which launched in 2010, led to an increase in building information in OSM, and the positional error is largely due to the limited resolution in such satellite imagery.
 
@@
 
-% Internet:
-% TODO: Switch to numerical citations
-% TODO: Add source of image on page 1
-
 % After Chris approves
@@
 The same is true for building modelling.
-It turns out that the same methods that have been used in this project to obtain and qualify road network data can be applied to building footprint data from OSM. \textbf{(Fan et al, 2014)}
+As \textbf{(Fan et al, 2014)} notes, it turns out that the same methods that have been used in this project to obtain and qualify road network data can be applied to building footprint data from OSM. 
 And only when there exists such a fundamental procedure for accuratly importing building footprints without compromising the data integritey, one can proceed with the remaining challenges in building modeling.
-\textbf{(Müller et al, 2006)}
 
@@
 Another crucial concept in city modelling is the Level Of Detail (LOD) aspect.
-LOD is a definition of a series of different 3D representation of the same data, at different detail levels, and the LOD chosen for a particulary application suggests how the data has been aqquired and modelled.
+\textbf{(Biljecki et al, 2016)} provide an extensive overview of the concept of LOD in GIS mapping.
+They provide the GityGML definition of LOD as a series of different 3D representation of the same data, at different detail levels, and the LOD chosen for a particulary application suggests how the data has been aqquired and modelled.
 The CityGML standard divides the LOD representation of 3D models into three classes: A B and C.
@@
 \end{enumerate}
-\textbf{(Biljecki et al, 2016)}
 A modeler may chose what LOD class(es) to use depending on data availablity and application.
@@
 Many studies urban development have been done before that requred detailed knowledge about a city's geometry, e.g. for the purposes of modeling heat development
-\textbf{(Nakata-Osaki, Souza and Rodrigues, 2018)}
-and reducing noise.
-\textbf{(Joon, Seo and Byung, 2011)}
+(\textbf{(Nakata-Osaki, Souza and Rodrigues, 2018)})
+and reducing noise
+(\textbf{(Joon, Seo and Byung, 2011)}).
 
@@
 A single inner-city street for example, often contains sidewalks and lanes reserved for taxis or public transit, all of a certain placement and dimension that determines what type of traffic it can serve.
-This type of representation is known as a multimodal structures, due to the possibility of incorporating mulitple modes of transportation, but it may even describe subtilities like building facades and distances between curbline and building lot.
+According to \textbf{(Stojanovski, 2019)}, this type of representation is known as a multimodal structures, due to the possibility of incorporating mulitple modes of transportation, but it may even describe subtilities like building facades and distances between curbline and building lot.
 OSM, while being an excellent source of city geometry, is simply not a sophisticated enough dataset to cater to all these possible areas of research.
-\textbf{(Stojanovski, 2019)}
 For media applications, such morppological and multimodal information could in theory be generated procedurally, by using shape grammars and some sort of wayfinding to create logical routes for different types of traffic.

### 2020-06-08 report/report.tex
Update todo
Move delimitations to introduction

@@
 For the positional accuracy study, see sections 1.3.2 for an implementation overview, 4.1 for a walkthrough of the program implementation, and 5.2 for the results.
+
+\section{Delimitations}
+
+In this study, when working with OSM geodata, the following assumptions have been made about the OSM dataset to reduce the size of the problem domain:
+
+\begin{enumerate}
+    \item the program will not handle terrain features and altitude. it will be assumed that the road mesh can be projected on a flat 2d surface without feature intersection.
+    \item it will be assumed that the city map consists of only arterial (primary) roads, secondary roads and footpaths.
+    \item it will be assumed that the osm dataset represents all roads as simple polylines, and building footprints as simple connected polygons.
+any other features (such as squares, which are roads represented as polygons) will be omitted from the dataset.
+    \item the road network is assumed to be a 2d simple graph.
+this means that any overlapping roads, such as tunnels and bridges that cross over street-level roads, will be eliminated from the dataset.
+any self-connected nodes will also be eliminated from the dataset.
+\end{enumerate}
+
+This study will not consider the lineage aspect of the geodata.
+Lineage is defined by (\textbf{Van Oort, 2006}) as the historical aspect of the dataset: how it has been obtained, with what instruments and how it has evolved between the moment it was obtained and the moment it was queried.
+The entire OSM dataset is under constant modification by volonteers, so it is expected that the accuracy will improve over time.
+Previous studies highlight how the posiitonal accuracy in similar-sized cities have improved over several years (\textbf{Fan et al, 2014}, \textbf{Haklay, 2010}).
+The historical aspect of the OSM dataset is available through the OSM History Viewer
+\footnote{OSM History Viewer - OpenStreetMap Wiki. Accessed 20-03-2020. https://wiki.openstreetmap.org/wiki/OSM\_History\_Viewer}
+, an online debugging tool that lets anyone freely view the change history of individual fetures in a commit history-like fashion.
+There is an option for editors to include a personal note with their changesets to include additional details or motivate why a change was made, but there is no guaratee that the data includes any information on the aquisition method used.
+
+Another aspect of geodata precision is attribute accuracy, which will neither be considered in this study.
+Attributes are small flags of metadata that are included with geodata features, letting the user know for example the type of a road, how many files it has, if the base is asphalt or gravel, or how many stories a building has.
+Attribute accurcay is the meausere of how complete attribute information is in a dataset.
+How much attribute data is missing and how much of it does not comform with reality?
+In general, attribute accuracy is hard to quantify, but might be interesting, as both the OSM map and the referece map used in this project use encoded metadata on a per feature basis.
+The attribute accuracy of the two datasets could in theory be assessed by doing a simple feature comparison and creating a translation table between the two attribute maps.
+To the knowledge of the author, this has never been done with specifically these two datasets, at the time of writing.
+
+The temporal accuracy will not be assessed as this projects focuses on comparing a single snapshot of the two datasets at the same point in time, and finally the logical accuracy will not be assessed because the underlying logical relations between features in both maps serve no purpose to this project.
 
@@
 \end{figure}
-
-\section{Delimitations}
-
-in this study, when working with osm geodata, the following assumptions have been made about the dataset to reduce the size of the problem domain:
-
-\begin{enumerate}
-    \item the program will not handle terrain features and altitude. it will be assumed that the road mesh can be projected on a flat 2d surface without feature intersection.
-    \item it will be assumed that the city map consists of only arterial (primary) roads, secondary roads and footpaths.
-    \item it will be assumed that the osm dataset represents all roads as simple polylines, and building footprints as simple connected polygons.
-any other features (such as squares, which are roads represented as polygons) will be omitted from the dataset.
-    \item the road network is assumed to be a 2d simple graph.
-this means that any overlapping roads, such as tunnels and bridges that cross over street-level roads, will be eliminated from the dataset.
-any self-connected nodes will also be eliminated from the dataset.
-\end{enumerate}
-
-Let us briefly go over all of van Oorts accuracy criteria that will not be assessed through this study.
-Let us see how they could be obtained and assessed for both the OSM and SLU maps, staring with the data lineage.
-The lineage aspect of the OSM data is available through the OSM History Viewer, an online debugging tool that lets anyone freely view the change history of individual fetures in a commit history-like fashion.
-There is an option for editors to include a personal note with their changesets to include additional details or motivate why a change was made, but there is no guaratee that the data includes any information on the aquisition method used \footnote{OSM History Viewer - OpenStreetMap Wiki.
-Accessed 20-03-2020.
-https://wiki.openstreetmap.org/wiki/OSM\_History\_Viewer}
-The history aspect of the SLU dataset is not freely available online, but the acquistion method is included in the file metadata on a per-object basis.
-Lantmäteriet uses internal codes to specifify the acquistion method, and these codes can be referred to in the product description the accompanies the map files.
-As the scope of this project is limited to making a broad comparison of the positional accuracy of the two datasets, only the required positional accuracy as described by the SLU product description will be used.
-
-The attribute accuracy measure is interesting, as both maps use encoded metadata on a per feature basis, the attribute accuracy of the two datasets could be assessed by doing a simple feature comparison and creating a translation table between the two attribute maps.
-To the knowledge of the author, this has never been done with specifically these two datasets, at the time of writing.
-
-The temporal accuracy will not be assessed as this projects focuses on comparing a single snapshot of the two datasets at the same point in time, and finally the logical accuracy will not be assessed because the underlying logical relations between features in both maps serve no purpose to this project.
 
@@
 
-The SLU map is maintained by The Swedish National Land Survey (Swedish: Lantmäteriet).
+The SLU map is maintained by The Swedish National Land Survey (Swedish: Lantmäteriet)
+\footnote{Lantmäteriet.
+Produktbeskrivning: GSD-Fastighetskartan vektor - Lantmäteriet.
+Acessed 16-04-2020.
+https://www.lantmateriet.se/globalassets/kartor-och-geografisk-information/kartor/fastshmi.pdf}.
 The data is collected by geodetic professionals, by land surveying methods such as GPS or DGNSS positioning, or by reproduction of features from ortophoto or stereo mapping from 3D aerial images.
@@
 Whereas OSM is a free service, the SLU map is available to paying customers or for free to students and researchers.
-\footnote{Lantmäteriet.
-Produktbeskrivning: GSD-Fastighetskartan vektor - Lantmäteriet.
-Acessed 16-04-2020.
-https://www.lantmateriet.se/globalassets/kartor-och-geografisk-information/kartor/fastshmi.pdf}
+
+The history aspect of the SLU dataset is not freely available online, but the acquistion method is included in the file metadata on a per-object basis.
+Lantmäteriet uses internal codes to specifify the acquistion method, and these codes can be referred to in the product description the accompanies the map files.
+As the scope of this project is limited to making a broad comparison of the positional accuracy of the two datasets, only the required positional accuracy as described by the SLU product description will be used.
 
@@
 \end{figure}
+
+% TODO: Move some of the text from method and evaluation to background
+% TODO: Read over the sample report to see what they have in the background section
+% TODO: Move the rest of the "method and evaluation" text  into the implementation and evaluations section, perhaps creating a new subsection in the implementation section called "Method"
 

### 2020-06-08 report/report.tex
Split up methodology chapter

@@
 
-This section will dive into the state-of-the-art use of city reconstruction methods within the research field of urban design.
-This chapter will present the motivations for the digitization of urban planning, as well as the most common programs and tools in this area, which are after all not very commonly used in the practise of urban design due to the failure in being able to reflect the unique workspace of urban designers.
-The final section will present a brief history of the field of OpenStreetMap data qualification, and present the research papers whose methods are at the foundation of this project.
+This chapter will dive into the state-of-the-art use of city reconstruction methods within the research field of urban design.
+This chapter will present the motivations for the digitization of urban planning, as well as a brief history of the field of OpenStreetMap data qualification, and present the research papers whose methods are at the foundation of this project.
+Furthermore, it will present the evaluation metrics that this project served to obtain, and how they relate to the eventual collision analysis.
+For the geodata precision study specifically, it will examine how geodata is qualifyied in general, by examining van Oorts quality criteria, and how these will be used in this project.
 
@@
 
-\chapter{Methodology and evaluation}
-
-This chapter will present the evaluation metrics that this project served to obtain, and how they relate to the eventual collision analysis.
-For the geodata precision study specifically, it will examine how geodata is qualifyied in general, by examining van Oorts quality criteria, and how these will be used in this project.
-See section 4 for a detailed technical implementation overview that presents how the relevant metrics were obtained during the statistics collection process.
-For the resulting metrics that were obtained, see section 5.
-
 \section{Evaluation metrics}
@@
 
-\section{Quality criteria}
+\subsection{Quality criteria}
 
@@
 
+\chapter{Implementation}
+
+This chapter will provide a technical overview of the data collection process and the algorithms used for the purpose of this study, starting with a brief overview in section 4.1.
+Section 4.2 will present the data collection process for the geodata precision study.
+It will present how to qualify dataset completeness and building correspondence (how many of the buildings in one dataset that can be found in the other) by collecting the cumulative building areas and looking at what buildings have overlapping areas between the two datasets.
+The definitions and algorithms neccessary for calculating posiitonal accuracy and a measure for shape accuracy will also be presented.
+Section 4.3 will present at the data collection process for the collision study.
+It will present how standard road withs were obtained, as well as the algorithm for detemining feature collision that was used.
+Section 4.4 will discuss briefly the data sources used in this project, why they were chosen and specific issues with the data collection process.
+
 \section{Implementation overview}
@@
 The program has then calculated how many of these colliding features that can be corrected by making adjustments that are within the positional accuracy of the OSM data (As obtained by the previous study in this report: Assessment of Geodata Precision).
-
-\chapter{Implementation}
-
-Section 4.1 will present the data collection process for the geodata precision study.
-It will present how to qualify dataset completeness and building correspondence (how many of the buildings in one dataset that can be found in the other) by collecting the cumulative building areas and looking at what buildings have overlapping areas between the two datasets.
-The definitions and algorithms neccessary for calculating posiitonal accuracy and a measure for shape accuracy will also be presented.
-Section 4.2 will present at the data collection process for the collision study.
-It will present how standard road withs were obtained, as well as the algorithm for detemining feature collision that was used.
-Section 4.3 will discuss briefly the data sources used in this project, why they were chosen and specific issues with the data collection process.
 
@@
 
-% TODO: Move some of the text from method and evaluation to background
 % TODO: Read over the sample report to see what they have in the background section
-% TODO: Move the rest of the "method and evaluation" text  into the implementation and evaluations section, perhaps creating a new subsection in the implementation section called "Method"
 

### 2020-06-09 report/report.tex
Update todo
Increase TOC depth and add subsections to Delimitations section

@@
 \setlength{\parskip}{1em}
-\setcounter{tocdepth}{1}
+\setcounter{tocdepth}{2}
 
@@
 \begin{enumerate}
-    \item The primary subquestion is: \textbf{How many roads of each type will collide with some other map feature, once they have been given a geometrical width?}.
-This question aims to find the rate of collision between road features in OSM, and investigate the nature of the most common types of collisions and how they can be corrected.
+    \item The primary subquestion is: \textbf{How many roads of each type will collide with some other map feature, once they have been given a geometrical width?} This question aims to find the rate of collision between road features in OSM, and investigate the nature of the most common types of collisions and how they can be corrected.
     \item The secondary subquestion is: \textbf{What is the positional accuracy of OSM map features compared to a reference map?}.
@@
 \section{Delimitations}
+
+\subsection{Delimitations of the OSM map data}
 
@@
 \end{enumerate}
+
+\subsection{Delimitations of accuracy classes}
 
@@
 
-% TODO: Read over the sample report to see what they have in the background section
+% TODO: Extend 2.3 with more content and references
 

### 2020-06-09 report/report.tex
Background, new subsections + Move LOD paragraph to Background

@@
 Such services are however usually limited in the accuracy and type of data that they can provide.
-While they can provide both builing footprints and road networks, they usually provide road data with no respect to overlap between features.
+While they can provide both building footprints and road networks, they usually provide road data with no respect to overlap between features.
 Collision between road meshes in 3D will lead to strange-looking mesh intersections and inaccuracies that modelers often want to avoid.
@@
 The results reveal that the positional accuracy of OSM data at the time was about 4 meters.
+
+\section{Procedural models for city reproduction}
+
+\section{Methods for geodata qualification}
+
+\section{Level of detail}
+
+Another crucial concept in city modelling is the Level Of Detail (LOD) aspect.
+\textbf{(Biljecki et al, 2016)} provide an extensive overview of the concept of LOD in GIS mapping.
+They provide the GityGML definition of LOD as a series of different 3D representation of the same data, at different detail levels, and the LOD chosen for a particulary application suggests how the data has been aqquired and modelled.
+The CityGML standard divides the LOD representation of 3D models into three classes: A B and C.
+\begin{enumerate}
+    \item LOD-A models are renderings of the building footprint itself, or rectangular building blocks that have been extruded from the footprint to give the building height.
+    \item LOD-B models are building blocks with simplifed roof shapes.
+    \item LOD-C models are LOD-B models which include features such as doors, windows, alcoves, building decor, indoor features or vegetation.
+\end{enumerate}
+A modeler may chose what LOD class(es) to use depending on data availablity and application.
 
@@
 
-Another crucial concept in city modelling is the Level Of Detail (LOD) aspect.
-\textbf{(Biljecki et al, 2016)} provide an extensive overview of the concept of LOD in GIS mapping.
-They provide the GityGML definition of LOD as a series of different 3D representation of the same data, at different detail levels, and the LOD chosen for a particulary application suggests how the data has been aqquired and modelled.
-The CityGML standard divides the LOD representation of 3D models into three classes: A B and C.
-\begin{enumerate}
-    \item LOD-A models are renderings of the building footprint itself, or rectangular building blocks that have been extruded from the footprint to give the building height.
-    \item LOD-B models are building blocks with simplifed roof shapes.
-    \item LOD-C models are LOD-B models which include features such as doors, windows, alcoves, building decor, indoor features or vegetation.
-\end{enumerate}
-A modeler may chose what LOD class(es) to use depending on data availablity and application.
+Yet another challenge lies in producing highly detailed 3D models of buildings at a consistent rate.
+Recall the concept of Level of Detail, as presented by \textbf{(Biljecki et al, 2016)} in section *** of this report.
 It turns out that, with the current means of data aquisition from OSM being limited to building footprints, using OSM data a model can only produce LOB-B models at best.

### 2020-06-09 report/report.tex
Convert related works sections to subswctions
Move procedural methods section to Background, add todos
Move geodata qualification content from Implementation to Background

@@
 
-\section{Previous work in qualifying OpenStreetMap data}
+\section{Related work}
+
+\subsection{OpenStreetMap data qualification}
 
@@
 
-\section{Procedural models for city reproduction}
-
-\section{Methods for geodata qualification}
-
-\section{Level of detail}
+\subsection{Procedural models for city reproduction}
+
+Procedural techniques are, as mentioned, methods for reconstructing or creating 3D automatically, with minimal manual work required.
+Adding such techniques for content creation into the workflow of city reconstruction will increase the level of detail that is possible, while sacrificing accuracy to the real-world.
+\textbf{(Mũller et al, 2006)} present how to accomplish this using Shape Grammar: a method for automatically and iteratively constructing 3D models based in a hierarchical set of rules.
+
+By starting with a crude volumetric model of a building and iteratively remodeling it from a set of shape grammars, a procedural model can structure the façade, add texture and add details for windows, doors and ornaments, and so on.
+Shape grammars often work from a databank of 3D models and a set of hierarchical rules for how these can be pieced together.
+This creates a possible generative space of many different models that can be created from the same data and rule set, and with the induction of random fluctuations, many different models can be generated from the same dataset and rules.
+
+%TODO: Mention a variety of procedural methods from different sources, start by reading Müller et al and look for more references
+
+\subsection{Methods for geodata qualification}
+
+\textbf{(Fan et al, 2014)} used the fraction of the total building area in their reference dataset and the OSM dataset to assess the completeness of OSM data.
+The motivation for this is to eliminate semantic differences between both datasets.
+A very common example of a semantic error when comparing geographic data is that a large building in one dataset may be segmented into several smaller ones in the other, that is to say a building in one set is represented as an aggreation of mutliple buildings in the other set.
+This makes a complete one-to-one object mapping impossible.
+Since it is impossible to find a one-to-one feature map for building footprints, using building area as an estimator for completeness is much better than i.e.
+using the number of buildings or other objects in each dataset.
+
+\textbf{(Fan et al, 2014)} further discussed how the relative overlap in building area can be used as an estimate of the building correspondence between datasets.
+This can only be done in cases where there is not much displacement between OSM building footprints data and the reference data set.
+The relative building overlap between the OSM and SLU datasets will be defined as folows, given the footprints of any building in the OSM set ($foot_{OSM}$) and the SLU set ($foot_{SLU}$):
+\begin{center}
+    $S_{RO}(foot_{OSM}, foot_{SLU}) = \frac{A_{Overlap}}{min(A_{foot_{OSM}}, A_{foot_{SLU}}))}$
+\end{center}
+
+The relative overlap may be used to determine building matching relations even when the semantic accuracy is low.
+If a large building is represented by a single footprint in one dataset but by several smaller footprints in the other,
+\textbf{(Rutzinger, Rottensteiner and Pfeifer, 2009)}
+found that if $S_{RO}(foot_{A}, foot_{B}) < 30\%$ for two buildings $A$ and $B$ from different sets, then $A$ and $B$ are highly likely to be separate, neighbouring buildings and not in fact identical.
+
+%TODO: DP, describe the algorithm with reference to DP and Fan
+Douglas-Peucker algorithm for polygon reduction. The full procedure for matching points will be presented in section ***.
+
+%TODO: Turning function for shape accuracy, find source.
+The building footprint correspondence will be evaluated using a similarity function that depends on the difference between the turning functions of matching building footprints in both datasets.
+The turning function was first defined by \textbf{(Arkin et al, 1991)}, as a method for measuring the similarity of two polygons.
+
+%TODO: Read Fan et al and look for references
+
+\subsection{Level of detail}
 
@@
 A modeler may chose what LOD class(es) to use depending on data availablity and application.
+
+%TODO: Summarize both sources by Biljecki
 
@@
 
-\subsection{Assesment of dataset completeness}
-
-\textbf{(Fan et al, 2014)} used the fraction of the total building area in their reference dataset and the OSM dataset to assess the completeness of OSM data.
+\subsection{Dataset completeness and correspondence}
+
+In the same way as the study by \textbf{(Fan et al, 2014)}, this project will use the building area in the OSM dataset versus that of the SLU dataset to determine the completeness of OSM building data in the Stockholm area, which will then be used as an argument for how accurate the building correspondence and point proximity measures can be.
 The motivation for this is to eliminate semantic differences between both datasets.
-A very common example of a semantic error when comparing geographic data is that a large building in one dataset may be segmented into several smaller ones in the other, that is to say a building in one set is represented as an aggreation of mutliple buildings in the other set.
-This makes a complete one-to-one object mapping impossible.
-Rather, buildings have to be identified as 1:1 (one to one), 1:n (one to many) or 1:0 (one to none) cases, depending on if a building is fully represented in the other dataset, represented as an aggregation of smaller plots, or not found at all.
+Buildings were identified as 1:1 (one to one), 1:n (one to many) or 1:0 (one to none) cases, depending on if a building is fully represented in the other dataset, represented as an aggregation of smaller plots, or not found at all.
 Figure \ref{fig:building-match-types} shows examples of two buildings that were found in the OSM and SLU datasets as 1:1 and 1:n matches respectively.
-
-Since it is impossible to find a one-to-one feature map for building footprints, using building area as an estimator for completeness is much better than i.e.
-using the number of buildings or other objects in each dataset.
-This project will use the building area in the OSM dataset versus that of the SLU dataset to determine the completeness of OSM building data in the Stockholm area, which will then be used as an argument for how accurate the building correspondence and point proximity measures can be.
 
@@
 
-\subsection{Correspondence by building area}
-
-\textbf{(Fan et al, 2014)} further discussed how the relative overlap in building area can be used as an estimate of the building correspondence between datasets.
+I accordance with \textbf{Fan et al, 2014}, 
 In this project, relative area overlap was used in order to pair buildings from the two datasets together and create a feature map.
-This can only be done in cases where there is not much displacement between OSM building footprints data and the reference data set.
+
 Figure \ref{fig:osm-slu-map} under Section 5 shows a typical comparison between the OSM and SLU datasets.
-By eyesight it was determined that the displacement of features is small.
-
-The relative building overlap between the OSM and SLU datasets will be defined as folows, given the footprints of any building in the OSM set ($foot_{OSM}$) and the SLU set ($foot_{SLU}$):
-\begin{center}
-    $S_{RO}(foot_{OSM}, foot_{SLU}) = \frac{A_{Overlap}}{min(A_{foot_{OSM}}, A_{foot_{SLU}}))}$
-\end{center}
-
-The relative overlap may be used to determine building matching relations even when the semantic accuracy is low.
-If a large building is represented by a single footprint in one dataset but by several smaller footprints in the other, \textbf{(Rutzinger, Rottensteiner and Pfeifer, 2009)} found that if $S_{RO}(foot_{A}, foot_{B}) < 30\%$ for two buildings $A$ and $B$ from different sets, then $A$ and $B$ are highly likely to be separate, neighbouring buildings and not in fact identical.
+
+\textbf{(Rutzinger, Rottensteiner and Pfeifer, 2009)}, 30% area overlap
 Therefore it will here be assumed that two buildings are matching candidates if their relative area overlap is greater than 30\%.
@@
 
-The building footprint correspondence will be evaluated using a similarity function that depends on the difference between the turning functions of matching building footprints in both datasets.
-The turning function was first defined by \textbf{(Arkin et al, 1991)}, as a method for measuring the similarity of two polygons.
 The Turning function $T_c(l)$ measures the cumulative angle of the polygon's counter-clockwise tangent, as a function of the cumulative normalized length l.
@@
 
-Procedural techniques are, as mentioned, methods for reconstructing or creating 3D automatically, with minimal manual work required.
-Adding such techniques for content creation into the workflow of city reconstruction will increase the level of detail that is possible, while sacrificing accuracy to the real-world.
-\textbf{(Mũller et al, 2006)} present how to accomplish this using Shape Grammar: a method for automatically and iteratively constructing 3D models based in a hierarchical set of rules.
-By starting with a crude volumetric model of a building and iteratively remodeling it from a set of shape grammars, a procedural model can structure the façade, add texture and add details for windows, doors and ornaments, and so on.
-Shape grammars often work from a databank of 3D models and a set of hierarchical rules for how these can be pieced together.
-This creates a possible generative space of many different models that can be created from the same data and rule set, and with the induction of random fluctuations, many different models can be generated from the same dataset and rules.
 For simple presentation in e.g. animation and videogames or even urban planning software, such procedurally generated cities could very well be enough.

### 2020-06-10 report/report.tex
Add more scientific sources for procedural models
Add Goetz and Ziph source

@@
 
-%TODO: Mention a variety of procedural methods from different sources, start by reading Müller et al and look for more references
+Shape grammer, developed by \textbf{Stiny, 1975}
+Space syntax, developed by \textbf{Hillier, 1997}. Analytical method that relies on the existence of
+
+Wayfinding
+\textbf{(Peponis, Zimring and Choi, 1990)}.
 
@@
 Such images have a limit on their resolution which puts a theoretical limit on the accuracy of map features compared to their real-world equivalents
-The driving force behind the availablity of VGI is, according to \textbf{(Fan et al, 2014)}, the availablity of high resolution satellite imagery to private individuals.
+The driving force behind the availablity of VGI is, according to
+\textbf{Goetz and Zipf, 2012},
+the availablity of high resolution satellite imagery to private individuals.
 They claim that particularily the imagery from Bing Maps, which launched in 2010, led to an increase in building information in OSM, and the positional error is largely due to the limited resolution in such satellite imagery.
@@
 
+\bibitem{hillier-97}
+Hillier, B. (1997) Space is the Machine: A Configurational Theory of Architecture. Cambridge University Press, Cambridge, UK.
+
 \bibitem{kunze-12}
@@
 
+\bibitem{peponis-90}
+Peponis, J., Zimring, C. and Choi, Y.K. (1990). Finding the Building in Wayfinding. Environment and Behavior, Vol. 22, 555-590.
+
 \bibitem{rutzinger-09}
@@
 IEEE Journal of Selected Topics in Applied Earth Observations and Remote Sensing 2(1): 11-20.
+
+\bibitem{stiny-75}
+Stiny, G. (1975) Pictorial and Formal Aspects of Shapes and Shape Grammars. Birkhauser, Basel, Switzerland.
 
@@
 \bibitem{van-oort-06}
-Van Oort P.A.J. (2006). Spatial Data Quality: From Description to Application. PhD Thesis, Wageningen University.
+Van Oort, P.A.J. (2006). Spatial Data Quality: From Description to Application. PhD Thesis, Wageningen University.
+
+\bibitem{goetz-12}
+Goetz, M. and Zipf, A. (2012). Towards defining a framework for the automatic derivation of 3D CityGML models from volunteered geographic information. Int. J. 3-D Inf. Model. 2012, 1, 496–507.
 

### 2020-06-10 report/report.tex
Organize Biljecki references and todos

@@
 Among the most common errors are missing or invalid geometry, and invalid semantics.
-However, \textbf{(Biljecki et al, 2016)} note that there are coming more and more sophisticated methods for automatically correcting the most common errors.
+However, \textbf{(Biljecki, Ledoux and Stoter, 2016, 1)} note that there are coming more and more sophisticated methods for automatically correcting the most common errors.
 Geodata validation and correction is only an emerging field of research.
@@
 
+%TODO: Sumarize these
 Shape grammer, developed by \textbf{Stiny, 1975}
@@
 
-%TODO: DP, describe the algorithm with reference to DP and Fan
+%TODO: summarize these (DP with reference to DP and Fan)
 Douglas-Peucker algorithm for polygon reduction. The full procedure for matching points will be presented in section ***.
-
-%TODO: Turning function for shape accuracy, find source.
 The building footprint correspondence will be evaluated using a similarity function that depends on the difference between the turning functions of matching building footprints in both datasets.
 The turning function was first defined by \textbf{(Arkin et al, 1991)}, as a method for measuring the similarity of two polygons.
-
-%TODO: Read Fan et al and look for references
+and used by \textbf{Fan et al, 2014} to obtain a measure of the shape accuracy of building footprints.
 
@@
 Another crucial concept in city modelling is the Level Of Detail (LOD) aspect.
-\textbf{(Biljecki et al, 2016)} provide an extensive overview of the concept of LOD in GIS mapping.
+\textbf{(Biljecki, Ledoux and Stoter, 2016, 1)} provide an extensive overview of the concept of LOD in GIS mapping.
 They provide the GityGML definition of LOD as a series of different 3D representation of the same data, at different detail levels, and the LOD chosen for a particulary application suggests how the data has been aqquired and modelled.
@@
 %TODO: Summarize both sources by Biljecki
+\textbf{Biljecki, Ledoux and Stoter, 2016, 1}
+\textbf{Biljecki et al, 2016, 2}
 
@@
 
-% TODO: Extend 2.3 with more content and references
-
 % After Chris approves
@@
 Yet another challenge lies in producing highly detailed 3D models of buildings at a consistent rate.
-Recall the concept of Level of Detail, as presented by \textbf{(Biljecki et al, 2016)} in section *** of this report.
+Recall the concept of Level of Detail, as presented by \textbf{(Biljecki, Ledoux and Stoter, 2016, 1)} in section *** of this report.
 It turns out that, with the current means of data aquisition from OSM being limited to building footprints, using OSM data a model can only produce LOB-B models at best.
@@
 
-\bibitem{biljecki-16}
-Biljecki, F., Ledoux, H., Du, X., Stoter, J., Soon, K., and Khoo, V. (2016). The most common geometric and semantic errors in CityGML datasets. 10.5194/isprs-annals-IV-2-W1-13-2016.
-
+\bibitem{biljecki-16-1}
+Biljecki, F., Ledoux, H., Stoter, J. (2016, 1): An improved LOD specification for 3D building models. Computers, Environment, and Urban Systems, vol. 59, 25-37.
+
+\bibitem{biljecki-16-2}
+Biljecki, F., Ledoux, H., Du, X., Stoter, J., Soon, K., and Khoo, V. (2016, 2). The most common geometric and semantic errors in CityGML datasets. 10.5194/isprs-annals-IV-2-W1-13-2016. 
 \bibitem{biljecki-17}

### 2020-06-10 report/report.tex
References in alphabetical order
Add more noteable OSM quality studies

@@
 The study applied different methods to assess the completeness of OSM data in two federal states in Germany, mainly by analysing the area difference between the OSM dataset and an administrative dataset.
+
+Other notable studies about OSM data accuracy were made around the same time by in Germany by \textbf{Zielstra and Zipf, 2010} and \textbf{Neis et al, 2012}, as well as in France by \textbf{Girres and Touya, 2010}.
+
 Finally, a notable study in 2013-2014, conducted by researchers from various universities in Norway and Germany (\textbf{(Fan et al, 2014)}), examined the quality and accuracy in building footprint data of the Munich area, at a time and place where the completeness of the OSM dataset was significantly higher (the results of their study showed that the OSM had captured close to 100\% of all building footprints).
@@
 
+\bibitem{girres-10}
+Girres, J.F. and Touya, G. (2010). Quality assessment of the French OpenStreetMap dataset. Trans. GIS 2010, 14, 435–459.
+
+\bibitem{goetz-12}
+Goetz, M. and Zipf, A. (2012). Towards defining a framework for the automatic derivation of 3D CityGML models from volunteered geographic information. Int. J. 3-D Inf. Model. 2012, 1, 496–507.
+
 \bibitem{joon-11}
@@
 
+\bibitem{neis-12}
+Neis, P., Zielstra, D. and Zipf, A. (2012). The Street Network Evolution of Crowdsourced Maps: OpenStreetMap in Germany 2007–2011. Future Internet. 2012; 4(1):1-21.
+
 \bibitem{parish-01}
@@
 
-\bibitem{goetz-12}
-Goetz, M. and Zipf, A. (2012). Towards defining a framework for the automatic derivation of 3D CityGML models from volunteered geographic information. Int. J. 3-D Inf. Model. 2012, 1, 496–507.
+\bibitem{zielstra-10}
+Zielstra, D.
+and Zipf, A.
+(2010).
+A Comparative Study of Proprietary Geodata and Volunteered Geographic Information for Germany.
+In Proceedings of 13th AGILE International Conference on Geographic Information Science, Guimarães, Portugal, 10–14 May 2010.
 

### 2020-06-10 report/report.tex
Include papers by Peponis and Zimring

@@
 %TODO: Sumarize these
-Shape grammer, developed by \textbf{Stiny, 1975}
-Space syntax, developed by \textbf{Hillier, 1997}. Analytical method that relies on the existence of
+Shape grammer, developed by \textbf{Stiny, 1975} (p_Par01.pdf)
+Space syntax, developed by \textbf{Hillier, 1997}. Analytical method that relies on the existence of (p_Par01.pdf, SITM_Introduction.pdf)
 
 Wayfinding
-\textbf{(Peponis, Zimring and Choi, 1990)}.
+\textbf{(Peponis, Zimring and Choi, 1990)}. (p_Par01.pdf, Peponis...pdf)
 

### 2020-06-10 report/report.tex
Fix duplicate citations

@@
 Among the most common errors are missing or invalid geometry, and invalid semantics.
-However, \textbf{(Biljecki, Ledoux and Stoter, 2016, 1)} note that there are coming more and more sophisticated methods for automatically correcting the most common errors.
+However, \textbf{(Biljecki, Ledoux and Stoter, 2016 (1))} note that there are coming more and more sophisticated methods for automatically correcting the most common errors.
 Geodata validation and correction is only an emerging field of research.
@@
 Another crucial concept in city modelling is the Level Of Detail (LOD) aspect.
-\textbf{(Biljecki, Ledoux and Stoter, 2016, 1)} provide an extensive overview of the concept of LOD in GIS mapping.
+\textbf{(Biljecki, Ledoux and Stoter, 2016 (1))} provide an extensive overview of the concept of LOD in GIS mapping.
 They provide the GityGML definition of LOD as a series of different 3D representation of the same data, at different detail levels, and the LOD chosen for a particulary application suggests how the data has been aqquired and modelled.
@@
 %TODO: Summarize both sources by Biljecki
-\textbf{Biljecki, Ledoux and Stoter, 2016, 1}
+\textbf{Biljecki, Ledoux and Stoter, 2016 (1)}
 \textbf{Biljecki et al, 2016, 2}
@@
 Yet another challenge lies in producing highly detailed 3D models of buildings at a consistent rate.
-Recall the concept of Level of Detail, as presented by \textbf{(Biljecki, Ledoux and Stoter, 2016, 1)} in section *** of this report.
+Recall the concept of Level of Detail, as presented by \textbf{(Biljecki, Ledoux and Stoter, 2016 (1))} in section *** of this report.
 It turns out that, with the current means of data aquisition from OSM being limited to building footprints, using OSM data a model can only produce LOB-B models at best.
@@
 \bibitem{biljecki-16-1}
-Biljecki, F., Ledoux, H., Stoter, J. (2016, 1): An improved LOD specification for 3D building models. Computers, Environment, and Urban Systems, vol. 59, 25-37.
+Biljecki, F., Ledoux, H., Stoter, J. (2016 (1)): An improved LOD specification for 3D building models. Computers, Environment, and Urban Systems, vol. 59, 25-37.
 
 \bibitem{biljecki-16-2}
-Biljecki, F., Ledoux, H., Du, X., Stoter, J., Soon, K., and Khoo, V. (2016, 2). The most common geometric and semantic errors in CityGML datasets. 10.5194/isprs-annals-IV-2-W1-13-2016. 
+    Biljecki, F., Ledoux, H., Du, X., Stoter, J., Soon, K., and Khoo, V. (2016 (2)). The most common geometric and semantic errors in CityGML datasets. 10.5194/isprs-annals-IV-2-W1-13-2016. 
 \bibitem{biljecki-17}

### 2020-06-10 report/report.tex
Incorporate Stiny into the Background text

@@
 Adding such techniques for content creation into the workflow of city reconstruction will increase the level of detail that is possible, while sacrificing accuracy to the real-world.
-\textbf{(Mũller et al, 2006)} present how to accomplish this using Shape Grammar: a method for automatically and iteratively constructing 3D models based in a hierarchical set of rules.
-
+\textbf{(Müller et al, 2006)} present how to accomplish this using Shape Grammar: a method developed by \textbf{Stiny, 1975}, for automatically and iteratively constructing 3D models based in a hierarchical set of rules.
 By starting with a crude volumetric model of a building and iteratively remodeling it from a set of shape grammars, a procedural model can structure the façade, add texture and add details for windows, doors and ornaments, and so on.
@@
 %TODO: Sumarize these
-Shape grammer, developed by \textbf{Stiny, 1975} (p_Par01.pdf)
 Space syntax, developed by \textbf{Hillier, 1997}. Analytical method that relies on the existence of (p_Par01.pdf, SITM_Introduction.pdf)

### 2020-06-10 report/report.tex
Space syntax subsection

@@
 
-Finally, wen it comes to the fields of video game design or 3D animation, tools for city reproduction is undoubtedly of use to 3D content creators.
+Finally, when it comes to the fields of video game design or 3D animation, tools for city reproduction is undoubtedly of use to 3D content creators.
 In their study of procedurally generated buildings, \textbf{(Parish and Müller, 2001)} claim that there is a widespread trend in these fields to use methods of procedural modeling to create large amounts of 3D content quickly and efficiently, while requiring as little hand modeling as possible.
@@
 \textbf{(Müller et al, 2006)} present how to accomplish this using Shape Grammar: a method developed by \textbf{Stiny, 1975}, for automatically and iteratively constructing 3D models based in a hierarchical set of rules.
-By starting with a crude volumetric model of a building and iteratively remodeling it from a set of shape grammars, a procedural model can structure the façade, add texture and add details for windows, doors and ornaments, and so on.
+Shape grammar was originally developed from a method known as L-systems: a system for model description which was originally used to generate plants for 3D animation \textbf{Parish and Müller, 2001}.
+Shape grammar works by starting with a crude volumetric model of a building and iteratively remodeling it from a set of shape grammars, a procedural model can structure the façade, add texture and add details for windows, doors and ornaments, and so on.
 Shape grammars often work from a databank of 3D models and a set of hierarchical rules for how these can be pieced together.
@@
 
-%TODO: Sumarize these
-Space syntax, developed by \textbf{Hillier, 1997}. Analytical method that relies on the existence of (p_Par01.pdf, SITM_Introduction.pdf)
-
-Wayfinding
-\textbf{(Peponis, Zimring and Choi, 1990)}. (p_Par01.pdf, Peponis...pdf)
+\subsection{Space syntax}
+
+Space syntax, developed by \textbf{Hillier, 1997}, is a set of theories that analyze complex, large-scale spaces such as cities, and tries to explain human behavior from a spatial point of view.
+The central idea is that spatial structures sucha as building patterns both reflect and create the patterns for human interaction with their surroundings.
+Dividing spacial structures into different levels of observation, and identifying different units of space and the pattern in which they connect to each other.
+Space syntax is an analytical method that relies on the availablity of city maps. Has since been developed into an extensive research fields on its own.
+Space syntax is commonly used in the analysis of procedural generation models of urban environments
+\textbf{Parish and Müller, 2001},
+and has been widelry used in the analysis of pedestrian flows and wayfinding
+\textbf{(Peponis, Zimring and Choi, 1990)}.
+The idea of using space syntax to divide spatial structures into different levels of observation is frequently used in the research field of urban design and planning, to gain an understanding how space is utilized and connected at different levels of usage (one can for example analyze a city on an interior-by-interior basis, or use city blocks and streets as the least unit of space)
+\textbf{(Stojanovski et al, 2020)}.
 

### 2020-06-10 report/report.tex
Fix Biljecki citations

@@
 Among the most common errors are missing or invalid geometry, and invalid semantics.
-However, \textbf{(Biljecki, Ledoux and Stoter, 2016 (1))} note that there are coming more and more sophisticated methods for automatically correcting the most common errors.
+However, \textbf{(Biljecki et al, 2016 (2))} note that there are coming more and more sophisticated methods for automatically correcting the most common errors.
 Geodata validation and correction is only an emerging field of research.
@@
 \textbf{Biljecki, Ledoux and Stoter, 2016 (1)}
-\textbf{Biljecki et al, 2016, 2}
+\textbf{Biljecki et al, 2014}
 

### 2020-06-10 report/report.tex
Updated LOD summary

@@
 
-Another crucial concept in city modelling is the Level Of Detail (LOD) aspect.
-\textbf{(Biljecki, Ledoux and Stoter, 2016 (1))} provide an extensive overview of the concept of LOD in GIS mapping.
-They provide the GityGML definition of LOD as a series of different 3D representation of the same data, at different detail levels, and the LOD chosen for a particulary application suggests how the data has been aqquired and modelled.
+Another crucial concept in city modelling and GIS mapping is the Level Of Detail (LOD) aspect.
+The LOD defines a number of possible different representation of the same GIS data, that vary in the amount of detail in its presentation.
+A modeler can choose to present the same data (e.g. building, roads, or decorative models) at different detail levels, depending on the use and application.
+Furthermore, the LOD chosen for a particulary application also suggests how the data has been aqquired and modelled.
+It turns out however, in practise, that the terminology and use of LOD is still ambigous and the definitions vary greatly between practitioners, standards and institutions, despite its usefulness.
+\textbf{(Biljecki et al, 2014)}
+
+\textbf{(Biljecki, Ledoux and Stoter, 2016 (1))} provide the GityGML definition of LOD.
 The CityGML standard divides the LOD representation of 3D models into three classes: A B and C.
@@
 A modeler may chose what LOD class(es) to use depending on data availablity and application.
-
-%TODO: Summarize both sources by Biljecki
-\textbf{Biljecki, Ledoux and Stoter, 2016 (1)}
-\textbf{Biljecki et al, 2014}
 

### 2020-06-10 report/report.tex
Fix citation parenthesises

@@
 This study will not consider the lineage aspect of the geodata.
-Lineage is defined by (\textbf{Van Oort, 2006}) as the historical aspect of the dataset: how it has been obtained, with what instruments and how it has evolved between the moment it was obtained and the moment it was queried.
+Lineage is defined by \textbf{(Van Oort, 2006)} as the historical aspect of the dataset: how it has been obtained, with what instruments and how it has evolved between the moment it was obtained and the moment it was queried.
 The entire OSM dataset is under constant modification by volonteers, so it is expected that the accuracy will improve over time.
@@
 
-When assessing the completeness of geodata, a very common scale is Van Oorts criteria for evaluating the quality of geographical information, which at the time of publication received attention from surveyors and cartographers (\textbf{(Van Oort, 2006)}).
+When assessing the completeness of geodata, a very common scale is Van Oorts criteria for evaluating the quality of geographical information, which at the time of publication received attention from surveyors and cartographers \textbf{(Van Oort, 2006)}.
 Since then a number of publications have examined slices of geodata from van Oort's criteria.
@@
 
-Other notable studies about OSM data accuracy were made around the same time by in Germany by \textbf{Zielstra and Zipf, 2010} and \textbf{Neis et al, 2012}, as well as in France by \textbf{Girres and Touya, 2010}.
-
-Finally, a notable study in 2013-2014, conducted by researchers from various universities in Norway and Germany (\textbf{(Fan et al, 2014)}), examined the quality and accuracy in building footprint data of the Munich area, at a time and place where the completeness of the OSM dataset was significantly higher (the results of their study showed that the OSM had captured close to 100\% of all building footprints).
+Other notable studies about OSM data accuracy were made around the same time by in Germany by \textbf{(Zielstra and Zipf, 2010)} and \textbf{(Neis et al, 2012)}, as well as in France by \textbf{(Girres and Touya, 2010)}.
+
+Finally, a notable study in 2013-2014, conducted by researchers from various universities in Norway and Germany \textbf{(Fan et al, 2014)}, examined the quality and accuracy in building footprint data of the Munich area, at a time and place where the completeness of the OSM dataset was significantly higher (the results of their study showed that the OSM had captured close to 100\% of all building footprints).
 At the time Munich was one of the most developed cities in OSM.
@@
 Adding such techniques for content creation into the workflow of city reconstruction will increase the level of detail that is possible, while sacrificing accuracy to the real-world.
-\textbf{(Müller et al, 2006)} present how to accomplish this using Shape Grammar: a method developed by \textbf{Stiny, 1975}, for automatically and iteratively constructing 3D models based in a hierarchical set of rules.
-Shape grammar was originally developed from a method known as L-systems: a system for model description which was originally used to generate plants for 3D animation \textbf{Parish and Müller, 2001}.
+\textbf{(Müller et al, 2006)} present how to accomplish this using Shape Grammar: a method developed by \textbf{(Stiny, 1975)}, for automatically and iteratively constructing 3D models based in a hierarchical set of rules.
+Shape grammar was originally developed from a method known as L-systems: a system for model description which was originally used to generate plants for 3D animation \textbf{(Parish and Müller, 2001)}.
 Shape grammar works by starting with a crude volumetric model of a building and iteratively remodeling it from a set of shape grammars, a procedural model can structure the façade, add texture and add details for windows, doors and ornaments, and so on.
@@
 
-Space syntax, developed by \textbf{Hillier, 1997}, is a set of theories that analyze complex, large-scale spaces such as cities, and tries to explain human behavior from a spatial point of view.
+Space syntax, developed by \textbf{(Hillier, 1997)}, is a set of theories that analyze complex, large-scale spaces such as cities, and tries to explain human behavior from a spatial point of view.
 The central idea is that spatial structures sucha as building patterns both reflect and create the patterns for human interaction with their surroundings.
@@
 Space syntax is commonly used in the analysis of procedural generation models of urban environments
-\textbf{Parish and Müller, 2001},
+\textbf{(Parish and Müller, 2001)},
 and has been widelry used in the analysis of pedestrian flows and wayfinding
@@
 The turning function was first defined by \textbf{(Arkin et al, 1991)}, as a method for measuring the similarity of two polygons.
-and used by \textbf{Fan et al, 2014} to obtain a measure of the shape accuracy of building footprints.
+and used by \textbf{(Fan et al, 2014)} to obtain a measure of the shape accuracy of building footprints.
 
@@
 
-I accordance with \textbf{Fan et al, 2014}, 
+I accordance with \textbf{(Fan et al, 2014)}, 
 In this project, relative area overlap was used in order to pair buildings from the two datasets together and create a feature map.
@@
 
-\textbf{(Rutzinger, Rottensteiner and Pfeifer, 2009)}, 30% area overlap
+\textbf{(Rutzinger, Rottensteiner and Pfeifer, 2009)} determined in their study of geodata comparison that if two building footprints share an area overlap of at least 30\%, then they can be safely assumed to point to each other in a matching set of buildings.
 Therefore it will here be assumed that two buildings are matching candidates if their relative area overlap is greater than 30\%.
@@
 Problems that arise are e.g. that the vertex counts are dissimilar, or that vertex clusters may be found at different parts of the polygon in the two datasets.
-To avoid this effect, key points are extracted using the Douglas-Peucker algorithm (\textbf{(Douglas and Peucker, 1973)}), to create a simplifed footprint with less points that still retain information about the rough features of the detailed footprint.
+To avoid this effect, key points are extracted using the Douglas-Peucker algorithm \textbf{(Douglas and Peucker, 1973)}, to create a simplifed footprint with less points that still retain information about the rough features of the detailed footprint.
 The idea behind Douglas-Peucker is to recursively divide a polyline.
@@
 Such images have a limit on their resolution which puts a theoretical limit on the accuracy of map features compared to their real-world equivalents
-The driving force behind the availablity of VGI is, according to
-\textbf{Goetz and Zipf, 2012},
-the availablity of high resolution satellite imagery to private individuals.
+The driving force behind the availablity of VGI is, according to \textbf{(Goetz and Zipf, 2012)}, the availablity of high resolution satellite imagery to private individuals.
 They claim that particularily the imagery from Bing Maps, which launched in 2010, led to an increase in building information in OSM, and the positional error is largely due to the limited resolution in such satellite imagery.
@@
 Many studies urban development have been done before that requred detailed knowledge about a city's geometry, e.g. for the purposes of modeling heat development
-(\textbf{(Nakata-Osaki, Souza and Rodrigues, 2018)})
+\textbf{(Nakata-Osaki, Souza and Rodrigues, 2018)}
 and reducing noise
-(\textbf{(Joon, Seo and Byung, 2011)}).
+\textbf{(Joon, Seo and Byung, 2011)}.
 

### 2020-06-10 report/report.tex
Update mathods for geodata qualification section

@@
 
-The relative overlap may be used to determine building matching relations even when the semantic accuracy is low.
+The relative overlap may also be used to determine building matching relations even when the semantic accuracy is low.
 If a large building is represented by a single footprint in one dataset but by several smaller footprints in the other,
@@
 
-%TODO: summarize these (DP with reference to DP and Fan)
-Douglas-Peucker algorithm for polygon reduction. The full procedure for matching points will be presented in section ***.
-The building footprint correspondence will be evaluated using a similarity function that depends on the difference between the turning functions of matching building footprints in both datasets.
-The turning function was first defined by \textbf{(Arkin et al, 1991)}, as a method for measuring the similarity of two polygons.
-and used by \textbf{(Fan et al, 2014)} to obtain a measure of the shape accuracy of building footprints.
+Once the building correspondence has been found and a matching building set has been established, where any building in one dataset points to one or more (or no) buildings in the other, the challenge becomes evaluating the similarity of mathcing buildings.
+When assessing the positional accuracy between the points of two building footprint polygons, first one must find a matching set of vertices between any pair of matching buildings, which is not always trivial since the vertex counts of the two buildings may not be the same.
+The two buildings might have been modeled at different levels of detail, or vertex clusters may be found at different parts of the polygon in the two datasets.
+
+The OSM data precision study by \textbf{(Fan et al, 2014)} used the Douglas-Peucker algorithm, first defined by \textbf{(Douglas and Peucker, 1973)}, to reduce both building polygons to a simliar vertex count, and then proceed with locating the most likely matching vertices in the polygons.
+The full procedure for matching points will be presented in section 3.2.3 of this report.
+The study by \textbf{(Fan et al, 2014)} also obtained a measure of the shape accuracy of building footprints.
+The shape accuracy between two matching buildings was quantified by using a similarity function that depends on the difference between the turning functions of matching building footprints in both datasets.
+The turning function was first defined by \textbf{(Arkin et al, 1991)}, as a method for measuring the similarity of two polygons, and measures the cumulative angle of a single polygon's counter-clockwise tangent, as a function of the cumulative normalized length l.
+The turning function effectively steps over the tangent of a polygon continously for a full cycle, and logs the total turning angle by adding together the angle of each turn.
+It includes consideration of wether the individual turns are clockwise or counter-clockwise, and is invariant to rotation and scaling of the polygon.
+Once the turning functions of two similar polygons have been obtained, the integral between them is taken as a measure of shape accuracy.
 

### 2020-06-10 report/report.tex
Add definition of shape accuracy

@@
 The full procedure for matching points will be presented in section 3.2.3 of this report.
-The study by \textbf{(Fan et al, 2014)} also obtained a measure of the shape accuracy of building footprints.
+The study by \textbf{(Fan et al, 2014)} also obtained a measure of the \textbf{shape accuracy} of building footprints.
+While the positional accuracy accounts for the error between individual points in a polygon, the shape accuracy is a measure of how similar two polygons at a higher level of observation, regardless of relative scaling and rotation of the two polygons.
 The shape accuracy between two matching buildings was quantified by using a similarity function that depends on the difference between the turning functions of matching building footprints in both datasets.

### 2020-06-11 report/report.tex
Update todo
Update shape accuracy section references
Shorten title

@@
 
-\title{Quality Assesment of OpenStreetMap Footpath Data for 3D City Recreation}
-\alttitle{Kvalitetiv undersökning av OpenStreetMaps gångvägsdata för stadsmodellering i 3D}
+\title{Quality Assesment of OpenStreetMap Footpath Data}
+\alttitle{Kvalitetiv undersökning av OpenStreetMaps gångvägsdata}
 \author{Alexander Hjelm}
@@
 
-% TODO: Mention that positional accuracy will be focused on, and the implications of not fixing errors when converting to a 3D model
 When it comes to the quality of available data, it turns out that in practise, geodata databases without errors are rare.
@@
 Once the turning functions of two similar polygons have been obtained, the integral between them is taken as a measure of shape accuracy.
+The full procedure for calculating shape accuracy will be presented in section 3.2.3 of this report.
 

### 2020-06-11 report/report.tex
Update instroduction

@@
 
+This project assesses the feasibility of reconstructing a city from OpenStreetMap (OSM) road and building data, and will focus on the feasibility of including pedestrian footpaths, by focusing on OSM road and building data in the Stockholm metropolitan area.
+
+\section{Introduction to virtual cities}
+
+Virtual cities has since long been a popular research topic in the field of computer visualization.
+The idea of recreating real world cities in 3D applications has a growing number of applications, including visualization and modelling, urban planning and architecture, as well as media and entertainment in the form of video games and 3D animation.
+Traditionally, modelling virtual cities is a labour-intensive and time-consuming task due to the need for manual modeling, texturing and design and placement of individual model in the large 3D scene.
+This type of tasks is commonly done by professional 3D graphical artists and designers.
+However, in recent years there has been major research work in the field of procedural content generation, meaning automatic or semi automatic generation of 3D models without the direct involvment of a 3D artist.
+
+Procedural methods for 3D city generation commonly rely on the availablity of spatial \textbf{geodata}, which serves as the initial input to the procedure of model generation and placement.
+Geodata can contain any form of map features, from terrain to road networks and buildings, but should ideally be easily retrievable in a format that the procedural model can easily work with, e.g.
+as a list of polygons with world coordinates.
+Such methods are indeed used in the field of virtual city reproduction to reduce the labour required to create accurate 3D representations of cities, but the field of procedual city reproduction comes with its own host of challenges.
+Urban environments are typically dense and large, spanning a few to hundreds of square kilometers, and are influenced by variables that are hard to quantify, including land policies, government plans, population changes and transportation infrastructure, making it non-trivial to produce 3D city representations with both high accuracy and high detail, while simultaneuosly having high utility in the intended field of study.
+\textbf{(Vanegas et al, 2010)}.
+
 \begin{figure}[H]
@@
 \footnotetext{Image source: Downtown Boston 3D City Model in Infraworks 360, available via: https://gallery.autodesk.com/civilinfrastructure/projects/490/downtown-boston-3d-city-model-in-infraworks-360}
-
-This project assesses the feasibility of reconstructing a city from OpenStreetMap (OSM) road and building data, and will focus on the feasibility of including pedestrian footpaths, by focusing on OSM road and building data in the Stockholm metropolitan area.
-
-\section{Introduction to virtual cities}
-
-Virtual cities has in recent years been a popular research topic in the field of computer visualization, according to \textbf{(Vanegas et al, 2010)}.
-The idea of recreating real world cities in 3D applications has a growing number of applications, including visualization and modelling, urban planning and architecture, as well as entertainment in the form of video games and 3D animation.
-Traditionally, modelling virtual cities is a labour-intensive and time-consuming task.
-However, in recent years there has been major research work in the field of procedural content generation, meaning automatic or semi automatic generation of 3D models without the direct involvment of a 3D artist.
-Such methods can indeed be used in the field of virtual city reproduction to reduce the labour required to reproduce accurate 3D representations of cities, but the field of procedual city reproduction comes with its own host of challenges.
-Urban environments are typically dense and large, spanning a few to hundreds of square kilometers, and are influenced by variables that are hard to quantify, including land policies, government plans, population changes and transportation infrastructure.
 
@@
 When it comes to the quality of available data, it turns out that in practise, geodata databases without errors are rare.
-Among the most common errors are missing or invalid geometry, and invalid semantics.
-However, \textbf{(Biljecki et al, 2016 (2))} note that there are coming more and more sophisticated methods for automatically correcting the most common errors.
-Geodata validation and correction is only an emerging field of research.
-and many such datasets were usually obtained years ago, and creating error-free datasets was neither possible due to the lack of validation techniques, nor a high priority.
+Among the most common errors are missing data or invalid geometry.
+In the case of rendering road networks and buildings, intersection between features can lead to rendering artifacts.
+This is why ideally one would like to import road data with minimal feature overlap.
+There are however coming more and more sophisticated methods for automatically correcting the most common errors.
+Geodata validation and correction is only an emerging field of research, and many such datasets were usually obtained years ago, and creating error-free datasets was neither possible due to the lack of validation techniques, nor a high priority
+\textbf{(Biljecki et al, 2016 (2))}.
 

### 2020-06-11 report/report.tex
Update Introduction: accuracy limitations chapter

@@
 Among the most common errors are missing data or invalid geometry.
-In the case of rendering road networks and buildings, intersection between features can lead to rendering artifacts.
-This is why ideally one would like to import road data with minimal feature overlap.
-There are however coming more and more sophisticated methods for automatically correcting the most common errors.
-Geodata validation and correction is only an emerging field of research, and many such datasets were usually obtained years ago, and creating error-free datasets was neither possible due to the lack of validation techniques, nor a high priority
+Geodata validation and correction is only an emerging field of research, and many such datasets were usually obtained years ago, and creating error-free datasets was neither possible due to the lack of validation techniques, nor a high priority.
+Luckily there are coming more and more sophisticated methods for automatically correcting the most common errors
 \textbf{(Biljecki et al, 2016 (2))}.
 
-A possible way of mitigating this problem is looking towards gathering geodata from volunteer-based geoservices such as OSM, where the temporal aspect of data accuracy is perhaps more assured by the crowd contribution nature.
-There are already a few such services that use OSM data to generate 3D geodata visualizations for data visualization and video game applications.
+Specifically in the case of rendering road networks and buildings, a common issue is overlap between the 3D models of the features, which can lead to strange-looking mesh intersections and inaccuracies that modelers often want to avoid.
+In particular, the risk of feature collision may be higher between streets and roads designated for energy efficient mobility, such as footpaths, since they are much more organic and dense in their nature than main streets.
+Thus, it is often in the designer's best interest to have a procedure for importing road data that results in minimal feature overlap, while still retaining accuracy to the real world.
+A possible way of mitigating this problem is looking towards gathering geodata from volunteer-based geoservices such as OSM.
+The OSM project is a freely available wiki-like geodatabase based on Volunteered Geographical Information (VGI), meaning anyone can contribute by adding or editing features.
+The OSM world dataset is being consistently updated by volonteers from all over the world, and the main idea is that by using OSM geodata for city reproduction, the temporal aspect of data accuracy would in theory be more assured by the crowd contribution nature.
+There are already a few such services that use OSM data to generate 3D geodata visualizations for media and video game applications.
 A few prominent examples are Mantle
@@
 While they can provide both building footprints and road networks, they usually provide road data with no respect to overlap between features.
-Collision between road meshes in 3D will lead to strange-looking mesh intersections and inaccuracies that modelers often want to avoid.
-In particular, streets and roads designated for energy efficient mobility, such as footpaths, are much more organic and dense in their nature than main streets, and thus the risk of feature collision may be higher.
 
@@
 
-The OSM project is a wiki-like geodatabase based on Volunteered Geographical Information (VGI), meaning anyone can contribute by adding or editing features.
+The OSM project is a wiki-like geodatabase based on Volunteered Geographical Information (VGI).
 OSM road data and building footprints are commonly obtained by manually tracing features in commercially available satellite images.

### 2020-06-11 report/report.tex
Update section references in Research Question section

@@
 
-For the primary research question: the collision study, a program has taken an OSM dataset of the Stockholm area and assigned each road with standard withds according to swedish regulation.
+For the primary research question: the collision study, a program has taken a slice of the OSM dataset in the Stockholm area and assigned each road with standard withds according to swedish regulation.
 It has then identified critical areas in the OSM dataset, where the paths are so wide that they collide with existing map features.
@@
 For the purpose of this report, features will refer mainly to roads and building footprints.
-For the collision study, see sections 1.3.1 for an implementation overview, 4.2 for a detailed walkthrough of the program implementation, and 5.1 for the results.
+For the collision study, see sections 3.1.2 for an implementation overview, 3.3 for a detailed walkthrough of the program implementation, and 4.1.1 for the results.
 
@@
 The positional accuracy will be needed to assess where road collisions occur and whether they can easily be corrected.
-For the positional accuracy study, see sections 1.3.2 for an implementation overview, 4.1 for a walkthrough of the program implementation, and 5.2 for the results.
+For the positional accuracy study, see sections 3.1.1 for an implementation overview, 3.2 for a walkthrough of the program implementation, and 4.1.2 for the results.
 

### 2020-06-11 report/report.tex
Update delimitations section

@@
 \begin{enumerate}
-    \item the program will not handle terrain features and altitude. it will be assumed that the road mesh can be projected on a flat 2d surface without feature intersection.
-    \item it will be assumed that the city map consists of only arterial (primary) roads, secondary roads and footpaths.
-    \item it will be assumed that the osm dataset represents all roads as simple polylines, and building footprints as simple connected polygons.
-any other features (such as squares, which are roads represented as polygons) will be omitted from the dataset.
-    \item the road network is assumed to be a 2d simple graph.
-this means that any overlapping roads, such as tunnels and bridges that cross over street-level roads, will be eliminated from the dataset.
-any self-connected nodes will also be eliminated from the dataset.
+    \item The program will not handle terrain features and altitude. It will be assumed that the road mesh can be projected on a flat 2d surface without feature intersection.
+    \item It will be assumed that the city map consists of only arterial (primary) roads, secondary roads, residential roads and footpaths.
+    \item It will be assumed that the OSM dataset represents all roads as simple polylines, and building footprints as simple connected polygons.
+Any other features (such as squares, which are roads represented as polygons) will be omitted from the dataset.
+    \item The road network is assumed to be a 2D simple graph.
+This means that any overlapping roads, such as tunnels and bridges that cross over street-level roads, will be eliminated from the dataset.
+Any self-connected nodes will also be eliminated from the dataset.
 \end{enumerate}
@@
 This study will not consider the lineage aspect of the geodata.
-Lineage is defined by \textbf{(Van Oort, 2006)} as the historical aspect of the dataset: how it has been obtained, with what instruments and how it has evolved between the moment it was obtained and the moment it was queried.
+Lineage is defined by \textbf{(Van Oort, 2006)} as the historical aspect of the dataset: how it has been obtained, with what instruments and how it has evolved between the moment it was obtained and the moment it was retrieved.
 The entire OSM dataset is under constant modification by volonteers, so it is expected that the accuracy will improve over time.
@@
 Another aspect of geodata precision is attribute accuracy, which will neither be considered in this study.
-Attributes are small flags of metadata that are included with geodata features, letting the user know for example the type of a road, how many files it has, if the base is asphalt or gravel, or how many stories a building has.
-Attribute accurcay is the meausere of how complete attribute information is in a dataset.
+Attributes are small flags of metadata that are included with geodata features, letting the user know for example the type of a road, how many lanes it has, if the base is asphalt or gravel, or how many stories a building has.
+Attribute accuracy is the meausere of how complete attribute information is in a dataset.
 How much attribute data is missing and how much of it does not comform with reality?
-In general, attribute accuracy is hard to quantify, but might be interesting, as both the OSM map and the referece map used in this project use encoded metadata on a per feature basis.
+In general, attribute accuracy is hard to quantify.
+In this case howere, it might be interesting to compare attributes between the OSM map and the referece map that was used in this project, as both use encoded metadata on a per feature basis, and matching features can be easily identified.
+See section 3.4.2 for an overview of the reference dataset that was used in this study.
 The attribute accuracy of the two datasets could in theory be assessed by doing a simple feature comparison and creating a translation table between the two attribute maps.
 To the knowledge of the author, this has never been done with specifically these two datasets, at the time of writing.
-
-The temporal accuracy will not be assessed as this projects focuses on comparing a single snapshot of the two datasets at the same point in time, and finally the logical accuracy will not be assessed because the underlying logical relations between features in both maps serve no purpose to this project.
 

### 2020-06-11 report/report.tex
Update overview and research fields sections in the Background chapter

@@
 
-This chapter will dive into the state-of-the-art use of city reconstruction methods within the research field of urban design.
-This chapter will present the motivations for the digitization of urban planning, as well as a brief history of the field of OpenStreetMap data qualification, and present the research papers whose methods are at the foundation of this project.
-Furthermore, it will present the evaluation metrics that this project served to obtain, and how they relate to the eventual collision analysis.
+This chapter will dive into the state-of-the-art use of city reconstruction methods within the research fields of visualization and urban design.
+This chapter will present the motivations for the digitization of urban planning, as well as a brief history of the field of OpenStreetMap data qualification, and furthermore present the research papers whose methods are at the foundation of this project.
+Lastly, it will present the evaluation metrics that this project served to obtain, and how they relate to the eventual collision analysis.
 For the geodata precision study specifically, it will examine how geodata is qualifyied in general, by examining van Oorts quality criteria, and how these will be used in this project.
@@
 These features are commonly represented as \textbf{geodata} in map databases, and processed and organized into visualization-ready data by a \textbf{Geographic Information System (GIS)}.
-In practise, cities are often made up out of many square kilometers worth of geodata, and it would be highly labour intensive to recreate all the features in 3D.
+In practise, cities are often made up out of many square kilometers worth of geodata, and it would be highly labour intensive to manually recreate all these features in 3D.
 Instead, \textbf{procedural models} for city reproduction instead rely on programmatic solutions for converting map data to 3D road and building models, and composing them to a 3D world that the designer may continue to fine-tweak through interfaces fo interactive editing.
-At the very base of this process is the procedural layout of road networks in 3D, and so far there have been a lot of research in modelling of 3D road networks.
+At the very base of this process is the procedural layout of road networks in 3D, and so far there has been a lot of research in modelling of 3D road networks.
 \textbf{(Stojanovski, 2019)} however claims that these projects have often focused on the road networks designated for cars and public transit, as a part of the sustainable mobility paradigm and the topic of sustainable cities, particularily due to the interest in self-propelled cars.
@@
 
-The project falls within the broader topic of geodata extraction and modelling, primarily of roads and building parcels.
+This project falls within the broader topic of geodata extraction and modelling, primarily of roads and building parcels.
 Geodata modeling is at the foundation of procedural modelling of real-world cities (according to \textbf{(Müller et al, 2006)}), since any application which models cities based on geodata will need a robust method of aqquiring and laying out road and building data without compromising the data's integrity.
-Procedural modeling of road networks is of interest to the fields of urban planning and urban architecture.
 In recent years there has been an advent of interactive tools that aid urban planners in placing or generating features of an urban plan, particularly roads and road networks.
-Procedural and AI solutions that aid the arcitect in a form of human-machine symbiosis.
-
-Procedural city-generation models combine architectural elemets, such as streets, lots, buildings, building facades.
-Urban design practises with conventional two-dimenstional deliverables range from delivering neighbourhood designs, urban zoning and building plans, to writing design codes and building ordinances.
-The last decade has seen multiple startups in the area of software that aid urban arcitecture.
-The software uses procedural models to render a city grid based on real-world locations, or generate new user-specified parameters, and allow for varying degrees of manual editing on grid, road, neighbourhood, city block or individual feature level.
+Such tools use procedural and AI solutions to move the burden of labour from the human designer to the software that the designer uses.
+
+The last decade has seen multiple startups in the area of procedural software for urban arcitecture.
+Such software uses procedural models to render a city grid based on real-world locations, or generate new grids and features based on user-specified parameters, and allow for varying degrees of manual editing on grid, road, neighbourhood, city block or individual feature level.
 Some of the most prominent examples are ArcGIS Pro (made by Esri), CityEngine (made by the startup Procedural, bought by ESRI) and Urban Canvas (made by the startup Synthicity, bought by Autodesk, now defunct).
-However, freely available such tools seldom include smaller roads and pedestrian footpaths such as walkways, sidewalks or trails.
 
 Even as such tools are well available, \textbf{(Stojanovski et al, 2020)} notes that they are as of today not commonly applied in urbanist practises, due to failures in reflecting the unique workspace of arcitects and their design needs.
-Urban arcitecture is a social factors, stakeholders.
-Each arcitect is an actor in a continously developing environment, and affecting that environment in a complex manner through their action.
+Urban arcitecture is a field that not only accounts for spatial data, but also captures many social factors and the individual wills of the many stakeholders.
+Each arcitect is an actor in a continously developing environment, and even the archtiect themselves is affecting that environment in a complex manner through their actions.
 

### 2020-06-11 report/report.tex
Related work, update first 3 sections

@@
 
-When assessing the completeness of geodata, a very common scale is Van Oorts criteria for evaluating the quality of geographical information, which at the time of publication received attention from surveyors and cartographers \textbf{(Van Oort, 2006)}.
-Since then a number of publications have examined slices of geodata from van Oort's criteria.
+When assessing the completeness of geodata, a very common scale is Van Oorts criteria for evaluating the quality of geographical information.
+At the time of publication this method of evaluation received much attention from surveyors and cartographers \textbf{(Van Oort, 2006)}, and since then a number of publications have examined slices of geodata by using van Oort's criteria.
 In 2008, \textbf{(Haklay, 2010)} conducted such a study to estimate of the quality of OSM positional geodata in London and England.
 Since the OSM project started in London it was thought that OSM data in the London metropolitan area would be representative of the highest quality data available, and therefore a good indicator for the whole global OSM dataset.
-The study was conducted by comparing OSM data to Ordnance Survey datasets.
-The study showed that the OSM dataset had a rough positional accuracy of 6 meters from the reference dataset.
+The study was conducted by comparing OSM data to Ordnance Survey datasets, and it showed that the OSM dataset had a rough positional accuracy of 6 meters from the reference dataset.
 However, at the time of the study, the OSM project had captured roughly 29\% of the area of England, meaning that the data completeness was fairly low.
 
-The next notable study to assess the quality of OSM data was published by \textbf{(Kunze, 2012)}.
+The next major study to assess the quality of OSM data was published by \textbf{(Kunze, 2012)}.
 The study applied different methods to assess the completeness of OSM data in two federal states in Germany, mainly by analysing the area difference between the OSM dataset and an administrative dataset.
-
-Other notable studies about OSM data accuracy were made around the same time by in Germany by \textbf{(Zielstra and Zipf, 2010)} and \textbf{(Neis et al, 2012)}, as well as in France by \textbf{(Girres and Touya, 2010)}.
-
-Finally, a notable study in 2013-2014, conducted by researchers from various universities in Norway and Germany \textbf{(Fan et al, 2014)}, examined the quality and accuracy in building footprint data of the Munich area, at a time and place where the completeness of the OSM dataset was significantly higher (the results of their study showed that the OSM had captured close to 100\% of all building footprints).
-At the time Munich was one of the most developed cities in OSM.
-Their study included an insight in the geometrical calculations neccessary to match features and points between two datasets, and how to reliably calculate the metrics needed for the most important and assessible of van Oorts criteria.
-Their study revealed that although the feature completeness of OSM was high, some architectural details were missing.
-The results reveal that the positional accuracy of OSM data at the time was about 4 meters.
+Other notable studies about OSM data accuracy were made around the same time in Germany by \textbf{(Zielstra and Zipf, 2010)} and \textbf{(Neis et al, 2012)}, as well as in France by \textbf{(Girres and Touya, 2010)}.
+
+Finally, a major study in 2013-2014, conducted by researchers from various universities in Norway and Germany \textbf{(Fan et al, 2014)}, examined the quality and accuracy in building footprint data of the Munich area, at a time and place where the completeness of the OSM dataset was significantly higher than the study by \textbf{(Haklay, 2010)} (the results of their study showed that the OSM had captured close to 100\% of all building footprints).
+At the time Munich was one of the most developed cities in OSM, but although feature completeness of the OSM dataset was high, some architectural details were missing.
+This new study included an insight in the geometrical calculations neccessary to match features and points between two datasets, and how to reliably calculate the metrics needed for the most important and assessible of van Oorts criteria.
+The results reveal that the positional accuracy of OSM data at the time was about 4 meters
 
@@
 Procedural techniques are, as mentioned, methods for reconstructing or creating 3D automatically, with minimal manual work required.
-Adding such techniques for content creation into the workflow of city reconstruction will increase the level of detail that is possible, while sacrificing accuracy to the real-world.
-\textbf{(Müller et al, 2006)} present how to accomplish this using Shape Grammar: a method developed by \textbf{(Stiny, 1975)}, for automatically and iteratively constructing 3D models based in a hierarchical set of rules.
-Shape grammar was originally developed from a method known as L-systems: a system for model description which was originally used to generate plants for 3D animation \textbf{(Parish and Müller, 2001)}.
-Shape grammar works by starting with a crude volumetric model of a building and iteratively remodeling it from a set of shape grammars, a procedural model can structure the façade, add texture and add details for windows, doors and ornaments, and so on.
+Adding such techniques for content creation into the workflow of city reconstruction will increase the level of detail that is possible to generate quickly and efficiently, while sacrificing accuracy to the real-world.
+This is because most commercially available procedural city reproduction models uses methods to guess what the finer detail might look like when map data of the desired level of detail is not available.
+
+\textbf{(Müller et al, 2006)} present how this is accomplished using Shape Grammar: a method developed by \textbf{(Stiny, 1975)}, for automatically and iteratively constructing 3D models based in a hierarchical set of rules.
+Shape grammar was originally developed from a method known as L-systems: a system for hierarchical model description which was originally used to generate plants for 3D animation \textbf{(Parish and Müller, 2001)}.
+Shape grammar works by starting with a crude volumetric model of a building and iteratively remodeling it from a set of shape grammars.
 Shape grammars often work from a databank of 3D models and a set of hierarchical rules for how these can be pieced together.
 This creates a possible generative space of many different models that can be created from the same data and rule set, and with the induction of random fluctuations, many different models can be generated from the same dataset and rules.
+Using this procedure, a procedural model can structure the façade of a building, add texture and add details for windows, doors and ornaments, and so on.
+The maximum accuracy that can be obtained in these finer details is largely dependent on the availablity of such data, and as most commercially available GIS data only focuses on crude building footprints and road placement, the model will often have to make guesses about the finer detail, which sacrifices accuracy to the real world.
 
@@
 Space syntax, developed by \textbf{(Hillier, 1997)}, is a set of theories that analyze complex, large-scale spaces such as cities, and tries to explain human behavior from a spatial point of view.
-The central idea is that spatial structures sucha as building patterns both reflect and create the patterns for human interaction with their surroundings.
-Dividing spacial structures into different levels of observation, and identifying different units of space and the pattern in which they connect to each other.
-Space syntax is an analytical method that relies on the availablity of city maps. Has since been developed into an extensive research fields on its own.
-Space syntax is commonly used in the analysis of procedural generation models of urban environments
-\textbf{(Parish and Müller, 2001)},
-and has been widelry used in the analysis of pedestrian flows and wayfinding
+It is commonly used in the analysis of procedural generation models of urban environments.
+The central idea is that spatial structures such as building patterns both reflect and create the patterns for human interaction with their surroundings.
+A common theme in space syntax is the division of spacial structures into different levels of observation, and identifying different units of space and the pattern in which they connect to each other.
+Space syntax is an analytical method that relies on the availablity of city maps, which has since its invention been developed into an extensive research fields on its own
+\textbf{(Parish and Müller, 2001)}.
+It has also been widely used in the analysis of pedestrian flows and wayfinding
 \textbf{(Peponis, Zimring and Choi, 1990)}.
-The idea of using space syntax to divide spatial structures into different levels of observation is frequently used in the research field of urban design and planning, to gain an understanding how space is utilized and connected at different levels of usage (one can for example analyze a city on an interior-by-interior basis, or use city blocks and streets as the least unit of space)
+Furthermore, the idea of using space syntax to divide spatial structures into different levels of observation is frequently used in the research field of urban design and urban planning, to gain an understanding how space is utilized and connected at different levels of usage (one can for example analyze a city on an interior-by-interior basis, or use city blocks and streets as the least unit of space)
 \textbf{(Stojanovski et al, 2020)}.

### 2020-06-11 report/report.tex
Update geodata qualification methods section

@@
 
-\textbf{(Fan et al, 2014)} used the fraction of the total building area in their reference dataset and the OSM dataset to assess the completeness of OSM data.
-The motivation for this is to eliminate semantic differences between both datasets.
-A very common example of a semantic error when comparing geographic data is that a large building in one dataset may be segmented into several smaller ones in the other, that is to say a building in one set is represented as an aggreation of mutliple buildings in the other set.
-This makes a complete one-to-one object mapping impossible.
-Since it is impossible to find a one-to-one feature map for building footprints, using building area as an estimator for completeness is much better than i.e.
+To assess the completeness of their OSM data, the study by \textbf{(Fan et al, 2014)} used the fraction of the total building area in the reference dataset and the OSM dataset.
+The motivation for this is to eliminate semantic differences between both datasets, such as the fact that a large building in one dataset may be segmented into several smaller ones in the other.
+That is to say a building in one set is represented as an aggregation of multiple buildings in the other set.
+This makes a complete one-to-one object mapping impossible, and since it is impossible to find a one-to-one feature map for building footprints, the idea is that using building area as an estimator for completeness is much better than i.e.
 using the number of buildings or other objects in each dataset.
@@
 This can only be done in cases where there is not much displacement between OSM building footprints data and the reference data set.
-The relative building overlap between the OSM and SLU datasets will be defined as folows, given the footprints of any building in the OSM set ($foot_{OSM}$) and the SLU set ($foot_{SLU}$):
+The relative building overlap between the OSM and the reference datasets is defined as follows, given the footprints of any building in the OSM set ($foot_{OSM}$) and the reference set ($foot_{REF}$):
 \begin{center}
-    $S_{RO}(foot_{OSM}, foot_{SLU}) = \frac{A_{Overlap}}{min(A_{foot_{OSM}}, A_{foot_{SLU}}))}$
+    $S_{RO}(foot_{OSM}, foot_{REF}) = \frac{A_{Overlap}}{min(A_{foot_{OSM}}, A_{foot_{REF}}))}$
 \end{center}
@@
 
-Once the building correspondence has been found and a matching building set has been established, where any building in one dataset points to one or more (or no) buildings in the other, the challenge becomes evaluating the similarity of mathcing buildings.
+Once the building correspondence has been found and a matching building set has been established, where any building in one dataset points to one or more (or no) buildings in the other, the challenge becomes evaluating the similarity of individual matching pairs of buildings.
 When assessing the positional accuracy between the points of two building footprint polygons, first one must find a matching set of vertices between any pair of matching buildings, which is not always trivial since the vertex counts of the two buildings may not be the same.
 The two buildings might have been modeled at different levels of detail, or vertex clusters may be found at different parts of the polygon in the two datasets.
-
 The OSM data precision study by \textbf{(Fan et al, 2014)} used the Douglas-Peucker algorithm, first defined by \textbf{(Douglas and Peucker, 1973)}, to reduce both building polygons to a simliar vertex count, and then proceed with locating the most likely matching vertices in the polygons.
 The full procedure for matching points will be presented in section 3.2.3 of this report.
+
 The study by \textbf{(Fan et al, 2014)} also obtained a measure of the \textbf{shape accuracy} of building footprints.
@@
 Once the turning functions of two similar polygons have been obtained, the integral between them is taken as a measure of shape accuracy.
-The full procedure for calculating shape accuracy will be presented in section 3.2.3 of this report.
+The full procedure for calculating shape accuracy will be presented in section 3.2.2 of this report.
 

### 2020-06-11 report/report.tex
Update LOD section

@@
 Another crucial concept in city modelling and GIS mapping is the Level Of Detail (LOD) aspect.
-The LOD defines a number of possible different representation of the same GIS data, that vary in the amount of detail in its presentation.
+The LOD defines a number of possible different representation of the same GIS data, which vary in the amount of detail in its presentation.
 A modeler can choose to present the same data (e.g. building, roads, or decorative models) at different detail levels, depending on the use and application.
@@
     \item LOD-B models are building blocks with simplifed roof shapes.
-    \item LOD-C models are LOD-B models which include features such as doors, windows, alcoves, building decor, indoor features or vegetation.
+    \item LOD-C models are LOD-B models which include finer detail features such as doors, windows, alcoves, building decor, indoor features or vegetation.
 \end{enumerate}
-A modeler may chose what LOD class(es) to use depending on data availablity and application.
+A modeler may chose which of these LOD class(es) to use depending on data availablity and application.
 

### 2020-06-11 report/report.tex
Update the remaining sections under Background

@@
 Shape accuracy is a measure of polygon similarity that was obtained by comparing polygon turning function between matching features, feature by feature.
-See section 5.3 for more details about the shape accuracy measure.
 
@@
 
-\textbf{(Haklay, 2010)} presents in a comprehesive manner the list of van Oort's 8 accuracy classes for evaluating the quality of geographic information:
+\textbf{(Haklay, 2010)} presents the list of van Oort's 8 accuracy classes for evaluating the quality of geographic information:
 
@@
 
-Other than for the purpose of mentioning various sources or error, this project will focus exclusively on the Positional accuracy, Completeness and Semantic accuracy classes.
-The comparison of the OSM and SLU maps will focus on these accuracy classes to determine the error tolerance when manipulating features in the 3D map presentation and the road network generating phase.
-Figure \ref{fig:osm-slu-map} shows two segments of the Stockholm map where the OSM and SLU datasets show lower semantic and positional accuracy respectively.
+Other than for the purpose of assessing various sources or error, this project will focus exclusively on the positional accuracy, completeness and semantic accuracy classes.
+The comparison of the OSM and reference maps will focus on these accuracy classes to determine the error tolerance when manipulating features in the 3D map presentation and the road network generating phase.
+Figure \ref{fig:osm-slu-map} shows two segments of the Stockholm map where the OSM and reference datasets show lower semantic and positional accuracy respectively.
 
@@
     \caption{Example segments of the Stockholm metropolitan area with unsystematic errors in precision.
-The SLU building dataset (blue) in superimposed on the OSM dataset (dark gray).
+The reference building dataset (SLU, blue) in superimposed on the OSM dataset (dark gray).
 The top picture is centered on a city block which shows a semantic mismatching between the two datasets.

### 2020-06-11 report/report.tex
Update section references in Implementation

@@
 
-This chapter will provide a technical overview of the data collection process and the algorithms used for the purpose of this study, starting with a brief overview in section 4.1.
-Section 4.2 will present the data collection process for the geodata precision study.
+This chapter will provide a technical overview of the data collection process and the algorithms used for the purpose of this study, starting with a brief overview in section 3.1.
+Section 3.2 will present the data collection process for the geodata precision study.
 It will present how to qualify dataset completeness and building correspondence (how many of the buildings in one dataset that can be found in the other) by collecting the cumulative building areas and looking at what buildings have overlapping areas between the two datasets.
 The definitions and algorithms neccessary for calculating posiitonal accuracy and a measure for shape accuracy will also be presented.
-Section 4.3 will present at the data collection process for the collision study.
+Section 3.3 will present at the data collection process for the collision study.
 It will present how standard road withs were obtained, as well as the algorithm for detemining feature collision that was used.
-Section 4.4 will discuss briefly the data sources used in this project, why they were chosen and specific issues with the data collection process.
+Section 3.4 will discuss briefly the data sources used in this project, why they were chosen and specific issues with the data collection process.
 

### 2020-06-11 report/report.tex
Update implementation overview

@@
 
-The geometrical precision study has been carried out in a number of steps that are all detailed below.
+The geodata precision study has been carried out in a number of steps that are all detailed below.
 First a feature map was found, which mapped individual features between the OSM dataset and a reference dataset.
@@
 
-Additionally, a number of auxillary calculations were neccessary for this study.
-The estimated completeness of the OSM dataset (a measure of how many features are represented in both the OSM dataset and the reference dataset) was calculated as the relative difference in area of all building footprints, between the two datasets.
-The completeness is needed to estimate the error boundary of the positional accuracy.
+Additionally, a number of auxillary calculations were needed to estimate the error boundary of the positional accuracy.
+The completeness of the OSM dataset (a measure of how many features are represented in both the OSM dataset and the reference dataset) was calculated as the relative difference in area of all building footprints, between the two datasets.
 Additionally, the shape accuracy of the OSM dataset has also been calculated and graphed as an extra verification of the similarity between the datasets.
@@
 
-Once the estimate of the positional error was obtained from the precision study, this estimate was used to assess the integrity of the OSM Stockholm road network, by locating critical points where assigning standard widths to the roads would cause collision with other features, and evaluating how many of those points could be eliminated by making ajustments within the found margin of error.
+Once the positional error estimate was obtained from the precision study, it was used to assess the integrity of the OSM Stockholm road network.
+This was done by firstly locating critical points where assigning standard widths to the roads would cause collision with other features, and then evaluating how many of those points could be eliminated by making ajustments within the found margin of error.
 For the purpose of the collision study, each road has been assigned a standard width that is in line with Swedish regulations, since in the OSM dataset, roads are represented as simple polylines, lacking width.
-In a 3D representation however, it is crucial that each road object has a spacial width in order to be rendered and seen.
-After the width assignment, a program has iterated over all features and calculated how many of them that overlap with any other feature, as well as the total length of the road polylines that is intersecting some other feature.
-The program has then calculated how many of these colliding features that can be corrected by making adjustments that are within the positional accuracy of the OSM data (As obtained by the previous study in this report: Assessment of Geodata Precision).
+After the width assignment, a program iterated over all features and calculated how many of them that show overlap with any other feature, as well as the total length of the road polylines that is intersecting some other feature.
+Finally, the program calculated how many of these colliding features that can be corrected by making adjustments that are within the positional accuracy of the OSM data, as obtained by the geodata precision study.
 

### 2020-06-11 report/report.tex
Update geodata precision study implementation section

@@
 
-In the same way as the study by \textbf{(Fan et al, 2014)}, this project will use the building area in the OSM dataset versus that of the SLU dataset to determine the completeness of OSM building data in the Stockholm area, which will then be used as an argument for how accurate the building correspondence and point proximity measures can be.
-The motivation for this is to eliminate semantic differences between both datasets.
+In the same way as the study by \textbf{(Fan et al, 2014)}, this project will use the building area in the OSM dataset versus that of the SLU dataset to determine the completeness of the OSM building data in the Stockholm area, which will then be used as an argument for how accurate the building correspondence and point proximity measures can be.
 Buildings were identified as 1:1 (one to one), 1:n (one to many) or 1:0 (one to none) cases, depending on if a building is fully represented in the other dataset, represented as an aggregation of smaller plots, or not found at all.
@@
 
-I accordance with \textbf{(Fan et al, 2014)}, 
-In this project, relative area overlap was used in order to pair buildings from the two datasets together and create a feature map.
-
-Figure \ref{fig:osm-slu-map} under Section 5 shows a typical comparison between the OSM and SLU datasets.
-
+To evaluate feature correspondence, relative area overlap was used as a measure of which buildings from either dataset corresponded with each other.
+In this was a feature map could be generated where any building in either dataset points to one, many or no buildings in the other dataset.
 \textbf{(Rutzinger, Rottensteiner and Pfeifer, 2009)} determined in their study of geodata comparison that if two building footprints share an area overlap of at least 30\%, then they can be safely assumed to point to each other in a matching set of buildings.
 Therefore it will here be assumed that two buildings are matching candidates if their relative area overlap is greater than 30\%.
-If a one-to-many object matching is found, the compound perimeter of the footprints in the many-set will be used when calculating the shape accuracy and finding closest vertices between the polygons.
-
-When calculating the area overlap, it was utilized the fact that multipolygons in geojson are represented with the main, bounding polygon first.
-Any subsequent polygons always represent cutouts or crop-outs in the footprint.
-When calculating the area of a multipolygon building, the area of the first polygon was therefore taken and subtracted with those of the subsequent polygons.
+Finally, if a one-to-many object matching is found, the compound perimeter of the footprints in the many-set will be used when calculating the shape accuracy and finding closest vertices between the polygons.
 
@@
 
-The Turning function $T_c(l)$ measures the cumulative angle of the polygon's counter-clockwise tangent, as a function of the cumulative normalized length l.
+The Turning function $T_c(l)$ measures the cumulative angle of the polygon's counter-clockwise tangent, as a function of the cumulative normalized length $l$.
 This project uses the turning function as it is defined by \textbf{(Fan et al, 2014)}.
@@
 The turning function has some nice geometric properties, in that it is invariant to both rotation and scaling of the polygon.
-The function contains no information of the orientation of the polygon, only of the relative angle between successive line segments, thus it does not change under roation.
+The function contains no information of the orientation of the polygon, only of the relative angle between successive line segments, thus it does not change under rotation.
 It also measures only the normalized cumulative length, which does not change under scaling.
@@
 
-The final problem is that even when building polygons have been matched between the two datasets, they may not have one-to-one vertex relationship.
-Footprints from different datasets may be formed at a different level of detail.
-Problems that arise are e.g. that the vertex counts are dissimilar, or that vertex clusters may be found at different parts of the polygon in the two datasets.
+The final problem is that even when building polygons have been matched between the two datasets, they may not have one-to-one vertex relationship, since footprints from different datasets may be formed at a different level of detail.
+Possible problems are that the vertex counts could be dissimilar, or that vertex clusters may be found at different parts of the two polygons.
 To avoid this effect, key points are extracted using the Douglas-Peucker algorithm \textbf{(Douglas and Peucker, 1973)}, to create a simplifed footprint with less points that still retain information about the rough features of the detailed footprint.
@@
 See Algorithm 1 for a detailed view.
-In either case The Oriented Minimum Bounding Rectangle (OMBR) is calculated for the two polygons.
+Following this, the Oriented Minimum Bounding Rectangles (OMBR) are calculated for both the simplifed polygons.
 Finally the OMBR for the OSM building footprint is shifted so that its centroid aligns with the centroid of the OMBR of the SLU building footprint.
@@
 
-Finally, when two building footprints and their vertices have been mapped, the offset of matching vertices will be used as a measure of positional error.
-The average and variance in vertex spread will be calculated and used to make an assessment of the geometric precision of the OSM dataset, and this precision will further be used to form an upper boundary on geometrical corrections that can be made to the map once it has been imported into the application.
+Finally, when two building footprints and their vertices have been mapped, the average offset of matching vertices will be used as a measure of positional error.
+The average and variance in vertex spread will be calculated and used to make an assessment of the geometric precision of the OSM dataset, and this precision will further be used to form an upper boundary on geometrical corrections that can be made in the collision study.
 

### 2020-06-11 report/report.tex
Update collision study, implementation

@@
 
-The maximum distance between two features was set to 1 times the positional accuracy, since then there will always be ample room to make changes to any features that would collide with the newly moved nodes, and thus collisions propagating will not be an issue.
-In this implementation, roads of different types were allowed to merge into one another to form common intersections.
-To facilitate this, it was neccessary to work with a path-by-path representation of the point data.
-When two path share a node with common coordinates, they form an intersection which can be of multiple different types.
-
-See the footnote for the publication that the default minimum widths were obtained from.
+The standard road widths were obtained from a publication by the Swedish Transport Administration
 \footnote{Vägars och gators utformning - Trafikverket.
@@
 https://trafikverket.ineko.se/Files/sv-SE/71830/\\Ineko.Product.RelatedFiles/2020\_029\_vagar\_och\_gators\_utformning\_krav.pdf}
-The standard recommended minimum width for a two-lane road in Sweden is 6.5 meters.
-
-Subsequent visual measurements of a few key locations in the Stockholm area revealed the following minimum way widths:
-Table \ref{table:road-widths} shows the minimum and maximum road widths that were found during these measurements, and that will be used in the collision study.
+, and were presented as requirements on the minimum lane width for certain types of roads and applications.
+These regulated minimum widths were used as a lower bound road with when examining the OSM dataset for potential collisions.
+The upper bound was found by visual measurements of a few key locations in the Stockholm area using the Bing maps service.
+Table \ref{table:road-widths} shows the minimum and maximum road widths that will be used in the collision study.
 
@@
 
-Here a few special cases are noted:
+Finally, a few special cases where features or feature collisions should not be registered were neccessary to implement:
 \begin{enumerate}

### 2020-06-11 report/report.tex
Update remaning Implementation chapter

@@
 
-For the purpose of this project, the OSM road dataset has been sampled around Stockholm metropolitan area.
-To assess and correct path widths and path node locations, the OSM building footprint dataset areound the same area has also been used.
+This project used a slice of the OSM road and building datasets around the Stockholm metropolitan area.
 As a reference dataset, the property map from the Swedish National Land Survey was used in comparison with the OSM map.
@@
 https://www.lantmateriet.se/globalassets/kartor-och-geografisk-information/kartor/fastshmi.pdf}.
-The data is collected by geodetic professionals, by land surveying methods such as GPS or DGNSS positioning, or by reproduction of features from ortophoto or stereo mapping from 3D aerial images.
+The data is collected by geodetic professionals, by using land surveying methods such as GPS or DGNSS positioning, or by reproduction of features from ortophoto or stereo mapping from 3D aerial images.
 Building features in the SLU dataset have a position accuracy requirement of 2 meters.

### 2020-06-11 report/report.tex
Update results sections

@@
 For the primary collision study.
-Table \ref{table:road-collision} shows the statistics of the collision study, and the number of roads, number of edges and total edge length that showed intersections with other features before road extrusion, by road category.
+Table \ref{table:road-collision} shows the statistics of the collision study, and the number of roads, number of edges and total edge length that showed intersections with other features, by road category.
 For footpaths in particular, it was shown that 12.05\% of all roads collide with some other feature after extrusion to the minimum width.
@@
 Table \ref{table:road-collision-corrected} shows the remaining roads, edges and total edges length whose intersections with other features could not be resolved by geometrical translation of a distance smaller than the positional accuracy.
-For footpaths this was shown to hold of 0.51\% of all roads, 
+For footpaths this was shown to hold for 0.51\% of all roads, 
 The number of road edges and percentage of the total road length that could not be corrected by geometric translation was significantly small.

### 2020-06-11 report/report.tex
Update intro to analysis section

@@
 The main outcome is that footpaths will not be more problematic than other road types in city reproduction from OSM data.
-This claim will be motivated with the data obtained in section 6.1.
+This claim will be motivated with the data obtained in section 4.1.
 Finally, the character of the feature collision cases will be discussed and the most common cases will be identified.
 The chapter will conclude with a presentation of geometrical algorithms for how to correct each case automatically.
-For conclusions about the usage of OSM city reproduction in engineering practise and remaining challenges, see section 7.
+For conclusions about the usage of OSM city reproduction in engineering practise and remaining challenges, see section 5.
 

### 2020-06-11 report/report.tex
Update analysis: precision study

@@
 
-The completeness analysis by building area shows that the OSM Stockholm dataset has high completeness in terms of both building area and counted features, in the same level as the study of Munich by \textbf{(Fan et al, 2014)}.
-The difference in area cover can partially be explained by small utility buildings that are often not represented in the OSM dataset.
+The building completeness analysis shows that the OSM Stockholm dataset has high completeness in terms of both building area and counted features.
+It even shows that the building completeness is around the same level as the study of Munich by \textbf{(Fan et al, 2014)}.
+In the particular case of Stockholm, the small difference in area cover that can be seen could partially be explained by the large number of small utility buildings that are often not represented in the OSM dataset.
 Particularily the residential areas in downtown, Stockholm, are characteristic in that each city block consists of a street front and an open, communal compound in the middle, inside of which there are usually a number of small utility buildings such as bicycle garages, toolsheds, refuse rooms and such.
 See figure \ref{fig:osm-slu-map-utility-buildings} for an example of an area in stockholm with such characteristics.
-These buildings are likely hard to make out and trace on publicly available satellite image data.
-They are, according to \textbf{(Fan et al, 2014)} occluded by their surroundings due to shadows or forestry.
-This appears to be a large consensus in the field.
-
-The low semantic accuracy has made analysis hard.
-The available SLU dataset from Lantmäteriet uses a plot-based subdivision of buildings, whereas OSM uses a subdivision that more closely resembles the actual building footprint as seen from above.
-In many cases whole city blocks are represented as single buildings, which means that many less meaningful feature matches can be found.
+These buildings are likely hard to make out and trace on publicly available satellite image data, most likely due to occlusion by their surroundings due to shadows or forestry.
+
+When it comes to evaluating semantic accuracy, this study found that the available SLU dataset from Lantmäteriet uses a plot-based subdivision of buildings, whereas OSM uses a subdivision that more closely resembles the actual building footprint as seen from above.
+In many cases even whole city blocks are represented as single buildings in OSM, which means that there are less meaning to the 1:n feature matches can be found.
 As seen in table \ref{table:building-matching}, about 80\% of all OSM features were found to have a 1:1 mapping to the SLU dataset, but upon further visual comparison it can be seen that many of the 1:n matches are found in a very specified area, namely the residential areas in downtown, Stockholm.
 This is problematic since this typical area with the same level of accessibility was likely measured with the same measuring technique, and the unsystematic positional error will likely be similar in these areas.
-The question is wether it comforms well to the positional error in the rest of the dataset or not.
-This analysis will not be carried out in this project, but could be a good subject for a future master's thesis project.
+The question is wether it comforms well to the positional error in the rest of the dataset or not, though this analysis will not be carried out in this project, but could be a good subject for a future master's thesis project.
 

### 2020-06-11 report/report.tex
Update remaining analysis section

@@
 As a result, it can promptly be assumed that city modelling services which already utilize OSM car road data could also use OSM footpath data without a drastic increase in feature collision.
-The biggest reason for this is thought to be that footpaths in their nature are narrow.
-So narrow in fact that they can be displaced by more than half of their full width (minimum 1 meter) before the limit of half the positional accuracy (1.025 meters) is reached.
-
-Looking closer at the individual collision cases, the next conclusion is that most often it is actually a small part of a road that is colliding, and will need correction.
+The biggest reason for this is thought to be that footpaths in their nature are narrow, and so narrow in fact that they can be displaced by more than half of their full width (minimum 1 meter) before the limit of half the positional accuracy (1.025 meters) is reached.
+
+Looking closer at the individual collision cases, the next conclusion is that most often it is actually a small part of a road that is responsible for the collision, and will need correction.
 Now will follow an exemplifying calculation to support this claim.
-Looking in tables \ref{table:road-collision} and \ref{table:road-collision-corrected}, note that before road extrusion, 3189 features and 7330 of their edges are intersecting some other feature.
+Looking in tables \ref{table:road-collision} and \ref{table:road-collision-corrected}, note that before road extrusion, 3189 features and 7330 of their edges were intersecting some other feature.
 This means that for each colliding road, on average 2.3 of its edges will be intersecting some other feature, thus bearing responsibility for the collision.
@@
 
-After visual inspection of the cases where features collide, it was determined that all collision occurences can be almost exclusively classified into one of five categories.
+After visual inspection, it was determined that all collision occurences can be almost exclusively classified into one of five categories.
 These categories are explained in the list below, and each case is exemplified in figure \ref{fig:collision-cases}.
@@
 \begin{enumerate}
-\item Case 1: A building lies in paralel with a road, and the distance between the two is smaller than the minimum width of the road.
+\item Case 1: A building lies in parallel with a road, and the distance between the two is smaller than the minimum width of the road.
 \item Case 2: A road shares a node with a building.

### 2020-06-11 report/report.tex
Update conclusions

@@
 
-As has been mentiond in the introduction to this report, procedural generaion of road networks is a basal requirement for quick and low-effort modeling virtual cities.
+As has been mentioned in the introduction to this report, procedural generaion of road networks is a basal requirement for quick and low-effort modeling virtual cities.
 Since the road network ties together the city, and forms the grid that divides the lots, quarters and neighbourhoods, it make sense that one can only continue with the remaining challenges once there exists a robust procedure for extracting road geodata and generating the 3D represention of it without compromising the data integrity.
@@
 As \textbf{(Fan et al, 2014)} notes, it turns out that the same methods that have been used in this project to obtain and qualify road network data can be applied to building footprint data from OSM. 
-And only when there exists such a fundamental procedure for accuratly importing building footprints without compromising the data integritey, one can proceed with the remaining challenges in building modeling.
-
-Such challenges, which have not been mentioned up until this point, include infering or guessing building height, since OSM does not provide data about building elevation.
+Only when there exists such a fundamental procedure for accuratly importing building footprints without compromising the data integritey, one can proceed with the remaining challenges in building modeling.
+
+Such challenges, which have not been mentioned, include for example infering or guessing building height, since OSM does not provide data about building elevation.
 Occasinally buildings in OSM may be tagged, but the norm is that no elevation data is available.
@@
 Yet another challenge lies in producing highly detailed 3D models of buildings at a consistent rate.
-Recall the concept of Level of Detail, as presented by \textbf{(Biljecki, Ledoux and Stoter, 2016 (1))} in section *** of this report.
+Recall the concept of Level of Detail, as presented by \textbf{(Biljecki, Ledoux and Stoter, 2016 (1))} in section 2.3.5 of this report.
 It turns out that, with the current means of data aquisition from OSM being limited to building footprints, using OSM data a model can only produce LOB-B models at best.
@@
 
-\section{Using procedural methods}
+\section{Feasibility in the urban planning field}
 
 For simple presentation in e.g. animation and videogames or even urban planning software, such procedurally generated cities could very well be enough.
-As said however, the big problem is that these features will rarely match real life, such models will be sufficient as long as one can be satisified with looking at rough representations of buildings and does not require additional knowledge about minor features.
-Many studies urban development have been done before that requred detailed knowledge about a city's geometry, e.g. for the purposes of modeling heat development
+As said however, the big problem is that these features will rarely match real life.
+Such models will be sufficient as long as one can be satisified with looking at rough representations of buildings and does not require additional knowledge about minor features.
+However, many studies in urban development have been done before that requred detailed knowledge about a city's geometry, e.g. for the purposes of modeling heat development
 \textbf{(Nakata-Osaki, Souza and Rodrigues, 2018)}
-and reducing noise
+or reducing noise
 \textbf{(Joon, Seo and Byung, 2011)}.
 
-\section{Feasibility in the urban planning field}
 By looking at such studies in urban development, one will discover that a lot of research in this field requires detailed morpological information that goes beyond simply knowing the geometry of the city.
@@
 A neighbourhood can be analysed from the level of its city blocks, individual lots, properties or building limits, building facades or even indoor spaces.
-Simlarily a road network can be analysed by looking at it either at its connectivity by viewing it a a simple 2D graph
-But it might as well be seen as a hierarchical structure where primary roads branch off into sub-streets. 
+Simlarily a road network can be analysed by looking at it either at its connectivity by viewing it a a simple 2D graph, but it might as well be seen as a hierarchical structure where primary roads branch off into sub-streets.
 The very unfortunate fact is that much of that morpological information is not present in the OSM dataset.
@@
 OSM, while being an excellent source of city geometry, is simply not a sophisticated enough dataset to cater to all these possible areas of research.
-For media applications, such morppological and multimodal information could in theory be generated procedurally, by using shape grammars and some sort of wayfinding to create logical routes for different types of traffic.
+For media applications, such morphological and multimodal information could in theory be generated procedurally, by using shape grammars and some sort of wayfinding to create logical routes for different types of traffic.
 But for urban planners requiring an exact morphological representation of some specification, this complementary data will have to be obtained from another dataset.

### 2020-06-12 report/report.tex
Update todo
Swedish abstract

@@
 Modeling cities in 3D has long been a topic of interest in field of computer graphics.
-It is often desirable to recreate real-world cities in 3D applications such as urban design software and video games, and the topic of 3D city reproduction comes with its own set of unique challenges.
-So far there have been a lot of research in modelling of 3D road networks, but these projects have often focused on the road networks designated for cars and public transit, and seldom on pedestrian footpaths.
+It is often desirable to recreate real-world cities in 3D applications such as urban design software and video games, and the topic of 3D city reproduction comes with its own set of unique challenges, at the base of which lies procedual rendering of 3D road networks.
+So far there have been a lot of research in modelling of road networks, but these projects have often focused on the road networks designated for cars and public transit, and seldom on pedestrian footpaths.
 
@@
 It was found that footpaths were less problematic than other road types (primary, secondary, residential), in terms of feature collisions, and thus should pose no more problems in 3D city reproduction than other road types.
-The report concludes with a scientific study of the field of 3D city reproduction.
-It serves to give an overview of the challenges in the field and place the study and findings of this project in a broader scientific context.
+The report concludes with a study of the remaining challenges in 3D city reproduction.
+
 \end{abstract}
@@
 \begin{abstract}
-Haha, He-Man!
-There is no possible way that you feeble-brained barbarian and your furball mount can foil my plan this time.
-With the ancient power at my command I shall doom this kingdom to crumble in meager ashes and..
-Oh shit, he has run off with the crystal skull hasn't he?
-Damn you He-Man!!!
+Modellering av städer i 3D har länge varit ett ämne av intresse inom datorgrafik.
+Ofta vill man återskapa verkliga städer i 3D-miljöer som till exempel mjukvara för stadsplanering eller datorspel.
+Avbildning av städer i 3D är ett ämne som kommer med många unika utmaningar, bland vilka en central utmaning är procedurell rendrering av vägnät.
+Tidigare har det bedrivits mycket forskning inom detta område, men många projekt har valt att fokusera på de delar av vägnätet som används av bilar och kollektivtraffik, och sällan gångvägar.
+
+Det här projektet har undersökt hur genomförbart det är att återskapa en stad utifrån data från OpenStreetMap (OSM), och kommer särskilt fokusera på hur genomförbart det är att inkludera gångvägar i en sådan modell.
+Studien har undersökt hur vanligt det är att gångvägar kolliderar kolliderar med andra kartdetaljer såsom byggnader och andra vägar.
+I denna process har felmarginalen för OSM-datamängden tagits fram genom att jämföra datamängen med en referenskarta med känd felmarginal.
+OSM-datamängdens felmarginal har använts som felmarginal i kollisionsstudien.
+
+Studien fann att gångvägar är mindre problematiska än andra vägtyper beträffande hur många vägar som kolliderar, och således borde inte gångvägar vara svårare att inkludera i en 3D-miljö än andra vägtyper.
+Rapporten avslutas med en studie över kvarvarande utmaningar inom området återskapande av städer i 3D.
 
@@
 % After Chris approves
-% TODO: Update section references
 % TODO: Spell check
-% TODO: Go over report one final time
 % TODO: Swedish abstract

### 2020-06-15 report/report.tex
Update todo
Mention SLU dataset in abstract and introduction

@@
 This project assesses the feasibility of reconstructing a city from OpenStreetMap (OSM) road and building data, and will focus on the feasibility of including pedestrian footpaths, by examining how common it is that these footpaths collide with other map features.
-In the collision detection process, the margin of error has been taken as the positional error in the OSM dataset, which has been obtained by comparing OSM data to a dataset with a known accuracy.
+In the collision detection process, the margin of error has been taken as the positional error in the OSM dataset, which has been obtained by comparing OSM data to the Swedish University of Agricultural Sciences (SLU) building datset with known accuracy.
 
@@
 
-For the secondary research question: the geodata precision study, the positional accuracy will be estimated by comparing feature points between the OSM dataset and a reference dataset.
+For the secondary research question: the geodata precision study, the positional accuracy will be estimated by comparing feature points between the OSM building dataset and the Swedish University of Agricultural Sciences (SLU) building datset with known accuracy.
 The positional accuracy will be needed to assess where road collisions occur and whether they can easily be corrected.
@@
 
-The SLU map is maintained by The Swedish National Land Survey (Swedish: Lantmäteriet)
+The SLU map is maintained and published by The Swedish National Land Survey (Swedish: Lantmäteriet)
 \footnote{Lantmäteriet.
@@
 
+% NOW-NOW:
+% TODO: Make clear that a footpath is a type of road. Even sidewalks that run along main roads are classified as roads in OSM.
+% TODO: Abstract: Reduce first paragraph (background)
+% TODO: Use overlap instead of collision outside of technical contexts
+
+% NOW:
+% TODO: Introduce concepts: Collision, overlap, error
+% TODO: Come up with a sentence that explains why errors could cause overlap. Repeat it in Introduction, research question and the abstract
+% TODO: The word footpath does not appear anywhere in the TOC
+% TODO: Abstract: Provide more detail about what was done and how it was evaluated - rephrase "In the collision detection process"
+
+% LATER:
+% TODO: Image with road collision (Accurate data vs inaccurate data)
+% TODO: Swedish abstract
+
 % After Chris approves
+% TODO: New thesis name
 % TODO: Spell check
-% TODO: Swedish abstract
 

### 2020-06-15 report/report.tex
Reduced first paragraph in the abstract
More technically oriented abstract

@@
 It is often desirable to recreate real-world cities in 3D applications such as urban design software and video games, and the topic of 3D city reproduction comes with its own set of unique challenges, at the base of which lies procedual rendering of 3D road networks.
-So far there have been a lot of research in modelling of road networks, but these projects have often focused on the road networks designated for cars and public transit, and seldom on pedestrian footpaths.
 
 This project assesses the feasibility of reconstructing a city from OpenStreetMap (OSM) road and building data, and will focus on the feasibility of including pedestrian footpaths, by examining how common it is that these footpaths collide with other map features.
-In the collision detection process, the margin of error has been taken as the positional error in the OSM dataset, which has been obtained by comparing OSM data to the Swedish University of Agricultural Sciences (SLU) building datset with known accuracy.
-
+To examine this, a slice of the OSM dataset in the Stockholm area has been taken, and each road has been assigned with standard withds according to swedish regulation, after which a geometrical algorithm was applied to all road pairs to obtain the average amount of road overlap.
 It was found that footpaths were less problematic than other road types (primary, secondary, residential), in terms of feature collisions, and thus should pose no more problems in 3D city reproduction than other road types.
-The report concludes with a study of the remaining challenges in 3D city reproduction.
+
+As a part of this study, the positional accuracy of the OSM dataset was calculated by comparing points of OSM building footprints to those in the Swedish University of Agricultural Sciences (SLU) building datset, whose accuracy is known.
+The positional accuracy determines the maximum tolerance in distance between points before it is certain that a road overlaps with another road or building.
 
@@
 % TODO: Make clear that a footpath is a type of road. Even sidewalks that run along main roads are classified as roads in OSM.
-% TODO: Abstract: Reduce first paragraph (background)
 % TODO: Use overlap instead of collision outside of technical contexts
@@
 % TODO: The word footpath does not appear anywhere in the TOC
-% TODO: Abstract: Provide more detail about what was done and how it was evaluated - rephrase "In the collision detection process"
 

### 2020-06-15 report/report.tex
Add definiton of roads and footpaths

@@
 \textbf{(Vanegas et al, 2010)}.
+
+This project will strive to determine the feasibility of procuderally reconstructing a city road network from OSM data by using such models as have been mentioned in this section, focusing on pedestrian footpaths.
+\textbf{Footpaths} include any form of path that is designated for pedestrians, including sidewalks, walkways and trails.
+OSM commonly divides city road networks into any of the following 4 categories: arterial (primary) roads, secondary roads, residential roads and footpaths.
+For the purpose of this report, a \textbf{road} will be any road whose primary use is human travel, either via a vehicle or by foot.
+Note that by this definition even sidewalks are classifed as roads, even though they commonly run in connection to larger roads (which are usually reserved for motor vehicles), and thus may not always be thought of as roads in plain language.
 
@@
 For the secondary research question: the geodata precision study, the positional accuracy will be estimated by comparing feature points between the OSM building dataset and the Swedish University of Agricultural Sciences (SLU) building datset with known accuracy.
+See section 3.4.2 for an overview of the SLU dataset and how it is obtained.
 The positional accuracy will be needed to assess where road collisions occur and whether they can easily be corrected.
@@
 
-% NOW-NOW:
-% TODO: Make clear that a footpath is a type of road. Even sidewalks that run along main roads are classified as roads in OSM.
+% NOW:
 % TODO: Use overlap instead of collision outside of technical contexts
-
-% NOW:
 % TODO: Introduce concepts: Collision, overlap, error

### 2020-06-16 report/report.tex
Change section headlines to be more representative of the footpath study
Organize todo
Swedish abstract, add clarity to english abstract

@@
 
-As a part of this study, the positional accuracy of the OSM dataset was calculated by comparing points of OSM building footprints to those in the Swedish University of Agricultural Sciences (SLU) building datset, whose accuracy is known.
+As a part of this study, the positional accuracy of the OSM dataset was calculated.
+Any geodataset has a limit on its positional accuracy, that is determined by the theoretical precision of the measuring methods.
+For the case of the OSM data used in this project, the accuracy was obtained by comparing coordinate points of OSM building footprints to those in the Swedish University of Agricultural Sciences (SLU) building datset, whose accuracy is known.
 The positional accuracy determines the maximum tolerance in distance between points before it is certain that a road overlaps with another road or building.
@@
 Avbildning av städer i 3D är ett ämne som kommer med många unika utmaningar, bland vilka en central utmaning är procedurell rendrering av vägnät.
-Tidigare har det bedrivits mycket forskning inom detta område, men många projekt har valt att fokusera på de delar av vägnätet som används av bilar och kollektivtraffik, och sällan gångvägar.
 
 Det här projektet har undersökt hur genomförbart det är att återskapa en stad utifrån data från OpenStreetMap (OSM), och kommer särskilt fokusera på hur genomförbart det är att inkludera gångvägar i en sådan modell.
-Studien har undersökt hur vanligt det är att gångvägar kolliderar kolliderar med andra kartdetaljer såsom byggnader och andra vägar.
-I denna process har felmarginalen för OSM-datamängden tagits fram genom att jämföra datamängen med en referenskarta med känd felmarginal.
-OSM-datamängdens felmarginal har använts som felmarginal i kollisionsstudien.
-
+Studien har undersökt hur vanligt det är att gångvägar överlappar med andra kartdetaljer såsom byggnader och andra vägar.
+I denna undersökning har en delmängd av OSM-datamängden över Stockholm tagits, och varje väg har givits en standardbredd enligt svenska bestämmelser.
+Sedan har en geometrisk algoritm tagit alla möjliga vägpar och beräknat hur stor mängd av vägarna som skulle överlappa med andra kartdetaljer.
 Studien fann att gångvägar är mindre problematiska än andra vägtyper beträffande hur många vägar som kolliderar, och således borde inte gångvägar vara svårare att inkludera i en 3D-miljö än andra vägtyper.
-Rapporten avslutas med en studie över kvarvarande utmaningar inom området återskapande av städer i 3D.
+
+I studien har felmarginalen för OSM-datamängden också tagits fram.
+Varje geodatamängd har en gräns på hur väl den stämmer överrens med verkligheten, som bestäms av vilka mätmedoter som har använts i datans insamling.
+I detta projekt uppskattades felmarginalen genom att jämföra byggnaders koordinatpunkter i datamängen med de från en referenskarta från Svenska Lantbruksuniversitetet (SLU) med känd felmarginal.
+Felmarginalen bestämmer toleransen mellan två punkter som kan tillåtas innan man är säker på att en väg börjar överlappa med en annan väg eller en byggnad.
 
@@
 
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_virtual_city}
+    \caption{Real-world cities are complex collections of buildings, blocks and road networks that interconnect them all.
+Creating 3D virtual cities using methods of procedual generation is an emerging research field that comes with its own set of challenges.
+This project deals with generating virtual cities from OpenStreetMap data, and the technical challenges therein.$^0$}
+    \label{fig:virtual-city}
+\end{figure}
+
+\footnotetext{Image source: Downtown Boston 3D City Model in Infraworks 360, available via: https://gallery.autodesk.com/civilinfrastructure/projects/490/downtown-boston-3d-city-model-in-infraworks-360}
+
+\section{Footpath definition}
+
 This project will strive to determine the feasibility of procuderally reconstructing a city road network from OSM data by using such models as have been mentioned in this section, focusing on pedestrian footpaths.
@@
 Note that by this definition even sidewalks are classifed as roads, even though they commonly run in connection to larger roads (which are usually reserved for motor vehicles), and thus may not always be thought of as roads in plain language.
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_virtual_city}
-    \caption{Real-world cities are complex collections of buildings, blocks and road networks that interconnect them all.
-Creating 3D virtual cities using methods of procedual generation is an emerging research field that comes with its own set of challenges.
-This project deals with generating virtual cities from OpenStreetMap data, and the technical challenges therein.$^0$}
-    \label{fig:virtual-city}
-\end{figure}
-
-\footnotetext{Image source: Downtown Boston 3D City Model in Infraworks 360, available via: https://gallery.autodesk.com/civilinfrastructure/projects/490/downtown-boston-3d-city-model-in-infraworks-360}
 
@@
 \textbf{(Biljecki et al, 2016 (2))}.
+
+% TODO: Use overlap instead of collision outside of technical contexts
+% TODO: Introduce concepts: Collision, overlap, error
+% TODO: Come up with a sentence that explains why errors could cause overlap. Repeat it in Introduction, research question and the abstract
+% TODO: Image with road collision (Accurate data vs inaccurate data)
 
@@
 
-\subsection{Road collision study}
+\subsection{Road overlap study}
 
@@
 
-% NOW:
-% TODO: Use overlap instead of collision outside of technical contexts
-% TODO: Introduce concepts: Collision, overlap, error
-% TODO: Come up with a sentence that explains why errors could cause overlap. Repeat it in Introduction, research question and the abstract
-% TODO: The word footpath does not appear anywhere in the TOC
-
-% LATER:
-% TODO: Image with road collision (Accurate data vs inaccurate data)
-% TODO: Swedish abstract
-
 % After Chris approves
@@
 
-\subsection{Collision study}
+\subsection{Footpath overlap study}
 
@@
 
-\subsection{Suggested algorithms for collision correction}
+\subsection{Suggested algorithms for overlap correction}
 
@@
 In this section everything we have learnt so far will be placed in the broader context of GIS and city modeling, focusing on the research fields of 3D media and city planning.
+
+\section{Feasibility of exporting footpaths}
+
+% TODO: Write this section
 

### 2020-06-17 report/report.tex
Add road collision case figure to report

@@
 % TODO: Image with road collision (Accurate data vs inaccurate data)
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_road_error_cases}
+    \caption{Example segments of a 3D reconstruciton of the Stockholm road map, with clearly visible geometric errors.
+The road overlap is marked with red boundaries.
+The segments show firstly a healthy intersection with main car roads, sidewalks and pedestrian crossings (top left), secondly a city segment where roads have been placed so they overlap with existing buildings due to insufficient data precision (top right), thirdly a similar case where roads overlap with each other (bottom left) and lastly an intersection where the road geomerty twists and self-intersects in strange ways due to invalid source data.
+This last case will not be examined in this project.}
+    \label{fig:residential-footpath-map}
+\end{figure}
 

### 2020-06-17 report/report.tex
Introduce overlap and collision

@@
 % TODO: Use overlap instead of collision outside of technical contexts
-% TODO: Introduce concepts: Collision, overlap, error
 % TODO: Come up with a sentence that explains why errors could cause overlap. Repeat it in Introduction, research question and the abstract
-% TODO: Image with road collision (Accurate data vs inaccurate data)
+
+Specifically in the case of rendering road networks and buildings, a common issue is that the 3D models being generated from road and building data migh overlap with each other, which can lead to strange-looking mesh intersections and inaccuracies that modelers often want to avoid.
+Such feature overlap is a type of geomerty error that arises due to inaccuracies in the source data, often due to limited precision in the measuring methods used to obtain the data.
+For the purpose of this project, the terms \textbf{feature overlap} and \textbf{feature collision} will be used synonymously, referring to the amount of features that show overlap with each other in the dataset, or the total feature area that show overlap in the same way.
+See figure \ref{fig:img_road_error_cases} for a few visual examples of what feature overlap could look like in the final 3D city model.
 
@@
 This last case will not be examined in this project.}
-    \label{fig:residential-footpath-map}
-\end{figure}
-
-Specifically in the case of rendering road networks and buildings, a common issue is overlap between the 3D models of the features, which can lead to strange-looking mesh intersections and inaccuracies that modelers often want to avoid.
-In particular, the risk of feature collision may be higher between streets and roads designated for energy efficient mobility, such as footpaths, since they are much more organic and dense in their nature than main streets.
+    \label{fig:img_road_error_cases}
+\end{figure}
+
+In this project it was hypothised that the risk of feature collision may be higher between streets and roads designated for energy efficient mobility, such as footpaths, since they are much more organic and dense in their nature than main streets.
 Thus, it is often in the designer's best interest to have a procedure for importing road data that results in minimal feature overlap, while still retaining accuracy to the real world.

### 2020-06-17 report/report.tex
Introduce feature overlap in abstract and RQ

@@
 \begin{abstract}
-Modeling cities in 3D has long been a topic of interest in field of computer graphics.
-It is often desirable to recreate real-world cities in 3D applications such as urban design software and video games, and the topic of 3D city reproduction comes with its own set of unique challenges, at the base of which lies procedual rendering of 3D road networks.
-
-This project assesses the feasibility of reconstructing a city from OpenStreetMap (OSM) road and building data, and will focus on the feasibility of including pedestrian footpaths, by examining how common it is that these footpaths collide with other map features.
+This project assesses the feasibility of reconstructing virtual cities from OpenStreetMap (OSM) road and building data, and will focus on the feasibility of including pedestrian footpaths, by examining how common it is that these footpaths overlap with other map features.
+Feature overlap is a type of geomerty error that arises to inaccuracies in the source data, which results in 3D models overlapping, which can in turn lead to strange-looking mesh intersections and artifacts in the final city model
+
 To examine this, a slice of the OSM dataset in the Stockholm area has been taken, and each road has been assigned with standard withds according to swedish regulation, after which a geometrical algorithm was applied to all road pairs to obtain the average amount of road overlap.
@@
 \begin{abstract}
-Modellering av städer i 3D har länge varit ett ämne av intresse inom datorgrafik.
-Ofta vill man återskapa verkliga städer i 3D-miljöer som till exempel mjukvara för stadsplanering eller datorspel.
-Avbildning av städer i 3D är ett ämne som kommer med många unika utmaningar, bland vilka en central utmaning är procedurell rendrering av vägnät.
-
-Det här projektet har undersökt hur genomförbart det är att återskapa en stad utifrån data från OpenStreetMap (OSM), och kommer särskilt fokusera på hur genomförbart det är att inkludera gångvägar i en sådan modell.
+Det här projektet har undersökt hur genomförbart det är att återskapa virtuella städer utifrån data från OpenStreetMap (OSM), och kommer särskilt fokusera på hur genomförbart det är att inkludera gångvägar i en sådan modell.
 Studien har undersökt hur vanligt det är att gångvägar överlappar med andra kartdetaljer såsom byggnader och andra vägar.
+Överlappning är en typ av fel i 3D-geometrin som uppstår på grund av osäkerhet i källdatan, vilket resulterar i att 3D-modeller överlappar med varandra.
+Detta kan leda till besvärliga skärningar och artifakter i den färdiga stadsmodellen.
+
 I denna undersökning har en delmängd av OSM-datamängden över Stockholm tagits, och varje väg har givits en standardbredd enligt svenska bestämmelser.
@@
 % TODO: Use overlap instead of collision outside of technical contexts
-% TODO: Come up with a sentence that explains why errors could cause overlap. Repeat it in Introduction, research question and the abstract
-
-Specifically in the case of rendering road networks and buildings, a common issue is that the 3D models being generated from road and building data migh overlap with each other, which can lead to strange-looking mesh intersections and inaccuracies that modelers often want to avoid.
+
+Specifically in the case of rendering road networks and buildings, a common issue is that the 3D models being generated from road and building data might overlap with each other, which can lead to strange-looking mesh intersections and inaccuracies that modelers often want to avoid.
 Such feature overlap is a type of geomerty error that arises due to inaccuracies in the source data, often due to limited precision in the measuring methods used to obtain the data.
@@
 
-To investigate this, the problem will specifically be divided into the following two subquestion:
+Again, note that feature overlap is defined here as geomerty errors due to inaccuracies in the source data, which results in 3D models overlapping, which can in turn lead to strange-looking mesh intersections.
+To investigate the feasibility of generating a 3D virtual city with minimal footpath overlap, the main research question will specifically be divided into the following two subquestion:
 
@@
 % After Chris approves
-% TODO: New thesis name
 % TODO: Spell check

### 2020-06-17 report/report.tex
Use overlap instead of collision ouside of techincal contexts

@@
 To examine this, a slice of the OSM dataset in the Stockholm area has been taken, and each road has been assigned with standard withds according to swedish regulation, after which a geometrical algorithm was applied to all road pairs to obtain the average amount of road overlap.
-It was found that footpaths were less problematic than other road types (primary, secondary, residential), in terms of feature collisions, and thus should pose no more problems in 3D city reproduction than other road types.
+It was found that footpaths were less problematic than other road types (primary, secondary, residential), in terms of feature overlap, and thus should pose no more problems in 3D city reproduction than other road types.
 
@@
 
-% TODO: Use overlap instead of collision outside of technical contexts
-
 Specifically in the case of rendering road networks and buildings, a common issue is that the 3D models being generated from road and building data might overlap with each other, which can lead to strange-looking mesh intersections and inaccuracies that modelers often want to avoid.
@@
 
-In this project it was hypothised that the risk of feature collision may be higher between streets and roads designated for energy efficient mobility, such as footpaths, since they are much more organic and dense in their nature than main streets.
+In this project it was hypothised that the risk of feature overlap may be higher between streets and roads designated for energy efficient mobility, such as footpaths, since they are much more organic and dense in their nature than main streets.
 Thus, it is often in the designer's best interest to have a procedure for importing road data that results in minimal feature overlap, while still retaining accuracy to the real world.
@@
 
-For the primary research question: the collision study, a program has taken a slice of the OSM dataset in the Stockholm area and assigned each road with standard withds according to swedish regulation.
+For the primary research question: the road overlap study, a program has taken a slice of the OSM dataset in the Stockholm area and assigned each road with standard withds according to swedish regulation.
 It has then identified critical areas in the OSM dataset, where the paths are so wide that they collide with existing map features.
@@
 For the purpose of this report, features will refer mainly to roads and building footprints.
-For the collision study, see sections 3.1.2 for an implementation overview, 3.3 for a detailed walkthrough of the program implementation, and 4.1.1 for the results.
+For the overlap study, see sections 3.1.2 for an implementation overview, 3.3 for a detailed walkthrough of the program implementation, and 4.1.1 for the results.
 
@@
 See section 3.4.2 for an overview of the SLU dataset and how it is obtained.
-The positional accuracy will be needed to assess where road collisions occur and whether they can easily be corrected.
+The positional accuracy will be needed to assess where road overlap occurs and whether the collisions can easily be corrected.
 For the positional accuracy study, see sections 3.1.1 for an implementation overview, 3.2 for a walkthrough of the program implementation, and 4.1.2 for the results.
@@
 This chapter will present the motivations for the digitization of urban planning, as well as a brief history of the field of OpenStreetMap data qualification, and furthermore present the research papers whose methods are at the foundation of this project.
-Lastly, it will present the evaluation metrics that this project served to obtain, and how they relate to the eventual collision analysis.
+Lastly, it will present the evaluation metrics that this project served to obtain, and how they relate to the overlap analysis.
 For the geodata precision study specifically, it will examine how geodata is qualifyied in general, by examining van Oorts quality criteria, and how these will be used in this project.
@@
 
-The primary study of the rate of collision between roads and other features will serve to obtain the following three metrics:
+The primary study of the rate of overlap between roads and other features will serve to obtain the following three metrics:
 
@@
 
-The road features will be categorized according to type, so that the rate of collisions between footpaths and other features can be compared to that of other road types.
+The road features will be categorized according to type, so that the rate of overlap between footpaths and other features can be compared to that of other road types.
 OSM mainly makes the distinction between primary, secondary, residential and footpath roads, so this classification will also be used in this report.
@@
 The definitions and algorithms neccessary for calculating posiitonal accuracy and a measure for shape accuracy will also be presented.
-Section 3.3 will present at the data collection process for the collision study.
+Section 3.3 will present at the data collection process for the overlap study.
 It will present how standard road withs were obtained, as well as the algorithm for detemining feature collision that was used.
@@
 
-Programmatic collection of statistics has been neccessary for both the primary and secondary research questions, and this section presents a broad overview of the technical implementation behind the map analysis and data collection processes for both the geodata precision study and the collision study.
+Programmatic collection of statistics has been neccessary for both the primary and secondary research questions, and this section presents a broad overview of the technical implementation behind the map analysis and data collection processes for both the geodata precision study and the overlap study.
 
@@
 
-\subsection{Collision study}
+\subsection{Overlap study}
 
@@
 This was done by firstly locating critical points where assigning standard widths to the roads would cause collision with other features, and then evaluating how many of those points could be eliminated by making ajustments within the found margin of error.
-For the purpose of the collision study, each road has been assigned a standard width that is in line with Swedish regulations, since in the OSM dataset, roads are represented as simple polylines, lacking width.
+For the purpose of the overlap study, each road has been assigned a standard width that is in line with Swedish regulations, since in the OSM dataset, roads are represented as simple polylines, lacking width.
 After the width assignment, a program iterated over all features and calculated how many of them that show overlap with any other feature, as well as the total length of the road polylines that is intersecting some other feature.
@@
 Finally, when two building footprints and their vertices have been mapped, the average offset of matching vertices will be used as a measure of positional error.
-The average and variance in vertex spread will be calculated and used to make an assessment of the geometric precision of the OSM dataset, and this precision will further be used to form an upper boundary on geometrical corrections that can be made in the collision study.
-
-\section{Collision study}
+The average and variance in vertex spread will be calculated and used to make an assessment of the geometric precision of the OSM dataset, and this precision will further be used to form an upper boundary on geometrical corrections that can be made in the overlap study.
+
+\section{Overlap study}
 
@@
 , and were presented as requirements on the minimum lane width for certain types of roads and applications.
-These regulated minimum widths were used as a lower bound road with when examining the OSM dataset for potential collisions.
+These regulated minimum widths were used as a lower bound road with when examining the OSM dataset for potential overlap.
 The upper bound was found by visual measurements of a few key locations in the Stockholm area using the Bing maps service.
-Table \ref{table:road-widths} shows the minimum and maximum road widths that will be used in the collision study.
+Table \ref{table:road-widths} shows the minimum and maximum road widths that will be used in the overlap study.
 
@@
     \end{tabular}
-    \captionof{table}{Standard road widths used for the collision study}
+    \captionof{table}{Standard road widths used for the overlap study}
     \label{table:road-widths}
@@
 Let $A$ be the current road feature.
-Since only road collisions are of interests, all roads were added to a queue and gone over one by one.
+Since this study only looks at the overlap of road features, all roads were added to a queue and gone over one by one.
 The polyline of the current road feature is extracted.
-Then, all other candidate features for collision are iterated over, and their boundary polygons (for buildings) or polylines (for roads) are extracted.
+Then, all other candidate features for overlap are iterated over, and their boundary polygons (for buildings) or polylines (for roads) are extracted.
 Let $B$ be the current feature, road or building, that is being compared to $A$.
@@
 
-To speed up the collision checking process, a geometric hashing solution was developed.
-In this way, feature collision lookup could be made without comparing a feature to all other available features, but instead from a selection of the closest candidates.
+To speed up the overlap checking process, a geometric hashing solution was developed.
+In this way, feature lookup could be made without comparing a feature to all other available features, but instead from a selection of the closest candidates.
 The hash bucket size was 230 by 230 meters and a feature could be in multiple hash buckets.
@@
 Here the results of the statistics collections will be presented.
-First the results of the collision study is presented as how many of the road features of each type display collision with some other feature.
+First the results of the overlap study is presented as how many of the road features of each type that overlap with some other feature.
 Then the results of the geodata precision is presented.
@@
 
-For the primary collision study.
-Table \ref{table:road-collision} shows the statistics of the collision study, and the number of roads, number of edges and total edge length that showed intersections with other features, by road category.
+For the primary overlap study.
+Table \ref{table:road-collision} shows the statistics of the overlap study, and the number of roads, number of edges and total edge length that showed intersections with other features, by road category.
 For footpaths in particular, it was shown that 12.05\% of all roads collide with some other feature after extrusion to the minimum width.
@@
     \end{tabular}
-    \captionof{table}{The first result table of the collision study.
+    \captionof{table}{The first result table of the overlap study.
 This table shows the amount of features and edges, as well as the cumulative edge length, in total and which are intersecting with any other feature.}
@@
     \end{tabular}
-    \captionof{table}{The second result table of the collision study.
+    \captionof{table}{The second result table of the overlap study.
 This table shows the amount of features and edges, as well as the cumulative edge length, which can be corrected by simple geometric translation.}
@@
 The outcome of that study will be used to assess the quality of the positional accuracy estimate by looking at the dataset completeness, the low semantic accuracy and the unique characteristics of the Stockholm area.
-Secondly the collision study will be discussed.
+Secondly the road overlap study will be discussed.
 The main outcome is that footpaths will not be more problematic than other road types in city reproduction from OSM data.
@@
 
-The immediate conclusion of this study is that, in terms of feature collision, footpaths were not significantly worse than other road categories.
-As a result, it can promptly be assumed that city modelling services which already utilize OSM car road data could also use OSM footpath data without a drastic increase in feature collision.
+The immediate conclusion of this study is that, in terms of feature overlap, footpaths were not significantly worse than other road categories.
+As a result, it can promptly be assumed that city modelling services which already utilize OSM car road data could also use OSM footpath data without a drastic increase in feature overlap.
 The biggest reason for this is thought to be that footpaths in their nature are narrow, and so narrow in fact that they can be displaced by more than half of their full width (minimum 1 meter) before the limit of half the positional accuracy (1.025 meters) is reached.

### 2020-06-17 report/report.tex
Introduce the concept of path width

@@
 Note that by this definition even sidewalks are classifed as roads, even though they commonly run in connection to larger roads (which are usually reserved for motor vehicles), and thus may not always be thought of as roads in plain language.
+
+In the 3D city model, each road will be represented by a polyline and a fixed \textbf{road width}.
+The width, or thickness, of a road at a certain point will be defined as the width of the perpendicular road cross section at that point, which is in its entirety usable by the intended types of traffic.
+It is important to note that OSM does commonly not include a defition of the width of any road.
+An optional tag exists that lets users specify the average path width of a road, but this tag is rarely used in practise.
 

### 2020-06-17 report/report.tex
Update introduction to improve clarity

@@
 This project assesses the feasibility of reconstructing virtual cities from OpenStreetMap (OSM) road and building data, and will focus on the feasibility of including pedestrian footpaths, by examining how common it is that these footpaths overlap with other map features.
-Feature overlap is a type of geomerty error that arises to inaccuracies in the source data, which results in 3D models overlapping, which can in turn lead to strange-looking mesh intersections and artifacts in the final city model
-
-To examine this, a slice of the OSM dataset in the Stockholm area has been taken, and each road has been assigned with standard withds according to swedish regulation, after which a geometrical algorithm was applied to all road pairs to obtain the average amount of road overlap.
+Feature overlap is a type of geomerty error that arises to inaccuracies in the source data, which results in 3D models overlapping, which can in turn lead to strange-looking mesh intersections and artifacts in the final city model.
+
+In this study, a slice of the OSM dataset in the Stockholm area has been taken, and each road has been assigned with standard withds according to swedish regulation, after which a geometrical algorithm was applied to all road pairs to obtain the average amount of road overlap.
 It was found that footpaths were less problematic than other road types (primary, secondary, residential), in terms of feature overlap, and thus should pose no more problems in 3D city reproduction than other road types.
@@
 As a part of this study, the positional accuracy of the OSM dataset was calculated.
-Any geodataset has a limit on its positional accuracy, that is determined by the theoretical precision of the measuring methods.
+Any geodataset has a limit on its positional accuracy, which is determined by the theoretical precision of the measuring methods.
 For the case of the OSM data used in this project, the accuracy was obtained by comparing coordinate points of OSM building footprints to those in the Swedish University of Agricultural Sciences (SLU) building datset, whose accuracy is known.
@@
 The idea of recreating real world cities in 3D applications has a growing number of applications, including visualization and modelling, urban planning and architecture, as well as media and entertainment in the form of video games and 3D animation.
-Traditionally, modelling virtual cities is a labour-intensive and time-consuming task due to the need for manual modeling, texturing and design and placement of individual model in the large 3D scene.
+Traditionally, modelling virtual cities is a labour-intensive and time-consuming task due to the need for manual modeling, texturing and design and placement of individual models in the large 3D scene.
 This type of tasks is commonly done by professional 3D graphical artists and designers.
 However, in recent years there has been major research work in the field of procedural content generation, meaning automatic or semi automatic generation of 3D models without the direct involvment of a 3D artist.
-
-Procedural methods for 3D city generation commonly rely on the availablity of spatial \textbf{geodata}, which serves as the initial input to the procedure of model generation and placement.
+\textbf{Procedural models} typically involve a machine which algorithmically composes together or alters human-generated 3D assets.
+Sometimes a degree of computer-generated randomness is induced into the generative algoritm in order to create a larger and less predictable generative space of 3D models that can be created from the same algorithm.
+
+Various procedural methods are indeed used in the field of virtual city reproduction to reduce the labour required to create accurate 3D representations of cities.
+These models rely on the availablity of spatial \textbf{geodata}, which serves as the initial input to the procedure of model generation and placement.
 Geodata can contain any form of map features, from terrain to road networks and buildings, but should ideally be easily retrievable in a format that the procedural model can easily work with, e.g.
-as a list of polygons with world coordinates.
-Such methods are indeed used in the field of virtual city reproduction to reduce the labour required to create accurate 3D representations of cities, but the field of procedual city reproduction comes with its own host of challenges.
-Urban environments are typically dense and large, spanning a few to hundreds of square kilometers, and are influenced by variables that are hard to quantify, including land policies, government plans, population changes and transportation infrastructure, making it non-trivial to produce 3D city representations with both high accuracy and high detail, while simultaneuosly having high utility in the intended field of study.
+as a list of polygons with points as world coordinates.
+
+The field of procedual city reproduction comes with its own host of challenges however.
+In practise, urban environments are typically dense and large, spanning a few to hundreds of square kilometers, and are influenced by variables that are hard to quantify, including land policies, government plans, population changes and transportation infrastructure, making it non-trivial to produce 3D city representations with both high accuracy and high detail, while simultaneuosly having high utility in the intended field of study.
 \textbf{(Vanegas et al, 2010)}.
@@
 
-This project will strive to determine the feasibility of procuderally reconstructing a city road network from OSM data by using such models as have been mentioned in this section, focusing on pedestrian footpaths.
+This project will strive to determine the feasibility of procedurally reconstructing a city road network from OSM data by using such models as have been mentioned in this section, focusing on the feasibility of including pedestrian footpaths.
 \textbf{Footpaths} include any form of path that is designated for pedestrians, including sidewalks, walkways and trails.

### 2020-06-17 report/report.tex
Update todo
Done updating Introduction for clarity, add todos

@@
 When it comes to the quality of available data, it turns out that in practise, geodata databases without errors are rare.
-Among the most common errors are missing data or invalid geometry.
 Geodata validation and correction is only an emerging field of research, and many such datasets were usually obtained years ago, and creating error-free datasets was neither possible due to the lack of validation techniques, nor a high priority.
@@
 
+% TODO: Introduce the idea of positional error due to low accuracy in this section
+\textbf{(Fan et al, 2014)}.
+
 Specifically in the case of rendering road networks and buildings, a common issue is that the 3D models being generated from road and building data might overlap with each other, which can lead to strange-looking mesh intersections and inaccuracies that modelers often want to avoid.
-Such feature overlap is a type of geomerty error that arises due to inaccuracies in the source data, often due to limited precision in the measuring methods used to obtain the data.
+In this project it was hypothised that the risk of feature overlap may be higher between streets and roads designated for energy efficient mobility, such as footpaths, since they are much more organic and dense in their nature than main streets.
+\textbf{Feature overlap} is a type of geomerty error that arises due to inaccuracies in the source data, often due to limited precision in the measuring methods used to obtain the data.
 For the purpose of this project, the terms \textbf{feature overlap} and \textbf{feature collision} will be used synonymously, referring to the amount of features that show overlap with each other in the dataset, or the total feature area that show overlap in the same way.
 See figure \ref{fig:img_road_error_cases} for a few visual examples of what feature overlap could look like in the final 3D city model.
+
+% TODO: The idea is to use road width and point coordinates as variables when making adjustments against errors.
 
@@
 
-In this project it was hypothised that the risk of feature overlap may be higher between streets and roads designated for energy efficient mobility, such as footpaths, since they are much more organic and dense in their nature than main streets.
-Thus, it is often in the designer's best interest to have a procedure for importing road data that results in minimal feature overlap, while still retaining accuracy to the real world.
+To limit visual quirks such as mesh self intersection, it is often in the designer's best interest to have a stable procedure for importing road data that results in minimal feature overlap, while still retaining accuracy to the real world.
 A possible way of mitigating this problem is looking towards gathering geodata from volunteer-based geoservices such as OSM.
@@
 
+%TODO: Include a note here about the challenges of identifying what buildings belong together and then finding what vertices belong together.
+
 For the secondary research question: the geodata precision study, the positional accuracy will be estimated by comparing feature points between the OSM building dataset and the Swedish University of Agricultural Sciences (SLU) building datset with known accuracy.
@@
 The entire OSM dataset is under constant modification by volonteers, so it is expected that the accuracy will improve over time.
-Previous studies highlight how the posiitonal accuracy in similar-sized cities have improved over several years (\textbf{Fan et al, 2014}, \textbf{Haklay, 2010}).
+Previous studies highlight how the posiitonal accuracy in similar-sized cities have improved over the course of several years (\textbf{Fan et al, 2014}, \textbf{Haklay, 2010}).
 The historical aspect of the OSM dataset is available through the OSM History Viewer
@@
 Another aspect of geodata precision is attribute accuracy, which will neither be considered in this study.
-Attributes are small flags of metadata that are included with geodata features, letting the user know for example the type of a road, how many lanes it has, if the base is asphalt or gravel, or how many stories a building has.
+Attributes are small flags of metadata that are included with geodata features, letting the user know additional details about a feature that is not reflected in the map, e.g.
+how many lanes a road has, if the base is asphalt or gravel, or how many stories a building has.
 Attribute accuracy is the meausere of how complete attribute information is in a dataset.
@@
 In general, attribute accuracy is hard to quantify.
-In this case howere, it might be interesting to compare attributes between the OSM map and the referece map that was used in this project, as both use encoded metadata on a per feature basis, and matching features can be easily identified.
+In this case however, it might be interesting to compare attributes between the OSM map and the SLU referece map that was used in this project, as both use encoded metadata on a per feature basis, and matching features can be easily identified.
 See section 3.4.2 for an overview of the reference dataset that was used in this study.

### 2020-06-18 report/report.tex
Background clarity update

@@
 This chapter will dive into the state-of-the-art use of city reconstruction methods within the research fields of visualization and urban design.
-This chapter will present the motivations for the digitization of urban planning, as well as a brief history of the field of OpenStreetMap data qualification, and furthermore present the research papers whose methods are at the foundation of this project.
+This chapter will present the motivations behind digitization of urban planning, as well as a brief history of the research field of OpenStreetMap data qualification, and furthermore present the research papers whose methods are at the foundation of this project.
 Lastly, it will present the evaluation metrics that this project served to obtain, and how they relate to the overlap analysis.
-For the geodata precision study specifically, it will examine how geodata is qualifyied in general, by examining van Oorts quality criteria, and how these will be used in this project.
+For the geodata precision study specifically, this chapter will examine how geodata is qualifyed in general, by examining van Oorts quality criteria, and how these will be used in this project.
 
@@
 These features are commonly represented as \textbf{geodata} in map databases, and processed and organized into visualization-ready data by a \textbf{Geographic Information System (GIS)}.
+In commercially available geodata, roads are most often defined as polylines: arrays of global coordinate points that come in the same order as they would normally be visited by a person travelling along the road.
+Buildings in turn are most often defined by their \textbf{building footprints}, a single polygon with global coordinates as vertices, which defines the extents of the building as seen on the map from above.
 In practise, cities are often made up out of many square kilometers worth of geodata, and it would be highly labour intensive to manually recreate all these features in 3D.
 Instead, \textbf{procedural models} for city reproduction instead rely on programmatic solutions for converting map data to 3D road and building models, and composing them to a 3D world that the designer may continue to fine-tweak through interfaces fo interactive editing.
-At the very base of this process is the procedural layout of road networks in 3D, and so far there has been a lot of research in modelling of 3D road networks.
-\textbf{(Stojanovski, 2019)} however claims that these projects have often focused on the road networks designated for cars and public transit, as a part of the sustainable mobility paradigm and the topic of sustainable cities, particularily due to the interest in self-propelled cars.
+
+The \textbf{road network} is what ties the city together as a whole, and forms the grid that divides the lots, quarters and neighbourhoods, giving the city grid its unqigue aestethic character and defining possible building arrangements.
+Thus, for any procedural city reproduction model it will be neccessary to have a robust procedure for extracting road geodata and generating the 3D represention of it without compromising the data integrity.
+So far there has been a lot of research into this area, but \textbf{(Stojanovski, 2019)} claims that most of these projects have often focused on the road networks designated for cars and public transit, as a part of the sustainable mobility paradigm and the topic of sustainable cities, particularily due to the interest in self-propelled cars.
 The study of energy efficient mobility such as walking and cycling is only in its emerging phase.
@@
 This project falls within the broader topic of geodata extraction and modelling, primarily of roads and building parcels.
-Geodata modeling is at the foundation of procedural modelling of real-world cities (according to \textbf{(Müller et al, 2006)}), since any application which models cities based on geodata will need a robust method of aqquiring and laying out road and building data without compromising the data's integrity.
+Geodata modeling is at the foundation of procedural modelling of real-world cities, since any software which models cities based on geodata will need a robust method of aqquiring and laying out road and building data without compromising the data integrity.
 In recent years there has been an advent of interactive tools that aid urban planners in placing or generating features of an urban plan, particularly roads and road networks.
-Such tools use procedural and AI solutions to move the burden of labour from the human designer to the software that the designer uses.
+Such tools use procedural methods and AI solutions to move the burden of labour from the human designer to the software that the designer uses
+\textbf{(Müller et al, 2006)}.
 
@@
 Such software uses procedural models to render a city grid based on real-world locations, or generate new grids and features based on user-specified parameters, and allow for varying degrees of manual editing on grid, road, neighbourhood, city block or individual feature level.
-Some of the most prominent examples are ArcGIS Pro (made by Esri), CityEngine (made by the startup Procedural, bought by ESRI) and Urban Canvas (made by the startup Synthicity, bought by Autodesk, now defunct).
-
-Even as such tools are well available, \textbf{(Stojanovski et al, 2020)} notes that they are as of today not commonly applied in urbanist practises, due to failures in reflecting the unique workspace of arcitects and their design needs.
+Some of the most prominent examples are ArcGIS Pro (made by Esri), CityEngine (made by the startup Procedural, bought by ESRI) and Urban Canvas (made by the startup Synthicity, bought by Autodesk, now defunct) \textbf{(Stojanovski et al, 2019)}.
+
+Even as such tools are well available, \textbf{(Stojanovski et al, 2020)} note that they are as of today not commonly applied in urbanist practises, due to failures in reflecting the unique workspace of arcitects and their design needs.
 Urban arcitecture is a field that not only accounts for spatial data, but also captures many social factors and the individual wills of the many stakeholders.
@@
 In their study of procedurally generated buildings, \textbf{(Parish and Müller, 2001)} claim that there is a widespread trend in these fields to use methods of procedural modeling to create large amounts of 3D content quickly and efficiently, while requiring as little hand modeling as possible.
-Studios who create large worlds are often interested in solutions that generate large road networks without any hand modeling.
+Studios who create large worlds are often interested in solutions that generate large road networks with minimal hand modeling and hand editing.
 
@@
 When assessing the completeness of geodata, a very common scale is Van Oorts criteria for evaluating the quality of geographical information.
-At the time of publication this method of evaluation received much attention from surveyors and cartographers \textbf{(Van Oort, 2006)}, and since then a number of publications have examined slices of geodata by using van Oort's criteria.
+At the time of publication, this method of evaluation received much attention from surveyors and cartographers \textbf{(Van Oort, 2006)}, and since then a number of publications have examined slices of geodata by using van Oort's criteria.
 In 2008, \textbf{(Haklay, 2010)} conducted such a study to estimate of the quality of OSM positional geodata in London and England.
@@
 The study was conducted by comparing OSM data to Ordnance Survey datasets, and it showed that the OSM dataset had a rough positional accuracy of 6 meters from the reference dataset.
-However, at the time of the study, the OSM project had captured roughly 29\% of the area of England, meaning that the data completeness was fairly low.
+However, at the time of the study, the OSM project had only captured roughly 29\% of the area of England, meaning that the data completeness was fairly low, and thus the uncertanity was relatively high.
 
@@
 Finally, a major study in 2013-2014, conducted by researchers from various universities in Norway and Germany \textbf{(Fan et al, 2014)}, examined the quality and accuracy in building footprint data of the Munich area, at a time and place where the completeness of the OSM dataset was significantly higher than the study by \textbf{(Haklay, 2010)} (the results of their study showed that the OSM had captured close to 100\% of all building footprints).
-At the time Munich was one of the most developed cities in OSM, but although feature completeness of the OSM dataset was high, some architectural details were missing.
 This new study included an insight in the geometrical calculations neccessary to match features and points between two datasets, and how to reliably calculate the metrics needed for the most important and assessible of van Oorts criteria.
-The results reveal that the positional accuracy of OSM data at the time was about 4 meters
+The results of this study revealed that the positional accuracy of the OSM Munich data at the time was about 4 meters
 
@@
 
-Procedural techniques are, as mentioned, methods for reconstructing or creating 3D automatically, with minimal manual work required.
+Procedural techniques are, as previously mentioned, methods for reconstructing or creating 3D automatically, with minimal manual work required.
 Adding such techniques for content creation into the workflow of city reconstruction will increase the level of detail that is possible to generate quickly and efficiently, while sacrificing accuracy to the real-world.
-This is because most commercially available procedural city reproduction models uses methods to guess what the finer detail might look like when map data of the desired level of detail is not available.
-
-\textbf{(Müller et al, 2006)} present how this is accomplished using Shape Grammar: a method developed by \textbf{(Stiny, 1975)}, for automatically and iteratively constructing 3D models based in a hierarchical set of rules.
+This is because most commercially available procedural city reproduction models use methods to guess what the finer detail might look like when map data of the desired level of detail is not available.
+
+\textbf{(Müller et al, 2006)} present how procedural model generaiton is accomplished using Shape Grammar: a method developed by \textbf{(Stiny, 1975)} for automatically and iteratively constructing 3D models based in a hierarchical set of rules.
 Shape grammar was originally developed from a method known as L-systems: a system for hierarchical model description which was originally used to generate plants for 3D animation \textbf{(Parish and Müller, 2001)}.
@@
 This creates a possible generative space of many different models that can be created from the same data and rule set, and with the induction of random fluctuations, many different models can be generated from the same dataset and rules.
-Using this procedure, a procedural model can structure the façade of a building, add texture and add details for windows, doors and ornaments, and so on.
-The maximum accuracy that can be obtained in these finer details is largely dependent on the availablity of such data, and as most commercially available GIS data only focuses on crude building footprints and road placement, the model will often have to make guesses about the finer detail, which sacrifices accuracy to the real world.
+Using this procedure, a procedural model can structure the façade of a building, add texture and add minor detail models such as windows, doors and ornaments, and so on.
+The maximum accuracy that can be obtained in these finer details is largely dependent on the availablity of such data, and as most commercially available GIS data only serve crude building footprints and road placement, the model will often have to make guesses about the finer detail, which sacrifices accuracy to the real world.
 
@@
 The central idea is that spatial structures such as building patterns both reflect and create the patterns for human interaction with their surroundings.
-A common theme in space syntax is the division of spacial structures into different levels of observation, and identifying different units of space and the pattern in which they connect to each other.
+A common theme in space syntax is the division of spacial structures into different levels of observation, and then identifying different units of space and the pattern in which they connect to each other.
 Space syntax is an analytical method that relies on the availablity of city maps, which has since its invention been developed into an extensive research fields on its own
@@
 \textbf{(Peponis, Zimring and Choi, 1990)}.
-Furthermore, the idea of using space syntax to divide spatial structures into different levels of observation is frequently used in the research field of urban design and urban planning, to gain an understanding how space is utilized and connected at different levels of usage (one can for example analyze a city on an interior-by-interior basis, or use city blocks and streets as the least unit of space)
+Furthermore, the idea of using space syntax to divide spatial structures into different levels of observation is frequently used in the research field of urban design and urban planning, to gain an understanding how space is utilized and connected at different levels of usage.
+One can for example analyze a city on an interior-by-interior basis, or use city blocks and streets as the least unit of space, depending on the goal of the study.
 \textbf{(Stojanovski et al, 2020)}.
@@
 To assess the completeness of their OSM data, the study by \textbf{(Fan et al, 2014)} used the fraction of the total building area in the reference dataset and the OSM dataset.
-The motivation for this is to eliminate semantic differences between both datasets, such as the fact that a large building in one dataset may be segmented into several smaller ones in the other.
-That is to say a building in one set is represented as an aggregation of multiple buildings in the other set.
-This makes a complete one-to-one object mapping impossible, and since it is impossible to find a one-to-one feature map for building footprints, the idea is that using building area as an estimator for completeness is much better than i.e.
-using the number of buildings or other objects in each dataset.
-
-\textbf{(Fan et al, 2014)} further discussed how the relative overlap in building area can be used as an estimate of the building correspondence between datasets.
-This can only be done in cases where there is not much displacement between OSM building footprints data and the reference data set.
+The motivation for this was to eliminate semantic error, which is a type of systematic error due to differences in how data is presented in both datasets.
+The most common such difference is the fact that a large building in one dataset may be segmented into several smaller ones in the other.
+Because of this, a building in one set might be represented as an aggregation of multiple buildings in the other set (see figure \ref{fig:building-match-types} for a visual example).
+This makes it impossible to find a complete one-to-one building footprint mapping between the two datasets, and thus the idea is that using building area as an estimator for completeness will be less sensitive to semantic errors than using the number of buildings or other objects in each dataset.
+
+\textbf{(Fan et al, 2014)} further discussed how the relative overlap in building area can be used as an estimate of the \textbf{building correspondence} between datasets.
+Building correspondence is a term that refers to how well it can be judged which building footprints between two datasets that correspond to the same building in real life.
+Using the relative area overlap between buildings to find which footprints represent the same building can only be done in cases where there is not much displacement between OSM building footprints data and the reference data set.
 The relative building overlap between the OSM and the reference datasets is defined as follows, given the footprints of any building in the OSM set ($foot_{OSM}$) and the reference set ($foot_{REF}$):
@@
 
-The relative overlap may also be used to determine building matching relations even when the semantic accuracy is low.
+The relative overlap may also be used to determine what building footprints match with each other even when the semantic accuracy is low.
 If a large building is represented by a single footprint in one dataset but by several smaller footprints in the other,
@@
 
-Once the building correspondence has been found and a matching building set has been established, where any building in one dataset points to one or more (or no) buildings in the other, the challenge becomes evaluating the similarity of individual matching pairs of buildings.
-When assessing the positional accuracy between the points of two building footprint polygons, first one must find a matching set of vertices between any pair of matching buildings, which is not always trivial since the vertex counts of the two buildings may not be the same.
+Once the building correspondence has been found and a building footprint mapping has been established (where any building in one dataset points to one, more or no buildings in the other dataset) the next challenge is evaluating how similar two matching building footprints are.
+When assessing the positional accuracy between the points of two building footprint polygons, first one must find a matching set of vertices between a pair of matching buildings, which is not always trivial since the vertex counts of the two buildings may not be the same.
 The two buildings might have been modeled at different levels of detail, or vertex clusters may be found at different parts of the polygon in the two datasets.
-The OSM data precision study by \textbf{(Fan et al, 2014)} used the Douglas-Peucker algorithm, first defined by \textbf{(Douglas and Peucker, 1973)}, to reduce both building polygons to a simliar vertex count, and then proceed with locating the most likely matching vertices in the polygons.
+To adress this problwm, the OSM data precision study by \textbf{(Fan et al, 2014)} used the Douglas-Peucker algorithm, first defined by \textbf{(Douglas and Peucker, 1973)}, to reduce both building polygons to a simliar vertex count, and then proceed with locating the most likely matching vertices in the polygons.
 The full procedure for matching points will be presented in section 3.2.3 of this report.
@@
 The shape accuracy between two matching buildings was quantified by using a similarity function that depends on the difference between the turning functions of matching building footprints in both datasets.
-The turning function was first defined by \textbf{(Arkin et al, 1991)}, as a method for measuring the similarity of two polygons, and measures the cumulative angle of a single polygon's counter-clockwise tangent, as a function of the cumulative normalized length l.
+The turning function was first defined by \textbf{(Arkin et al, 1991)}, as a method for measuring the similarity of two polygons, and measures the cumulative angle of a single polygon's counter-clockwise tangent, as a function of the cumulative normalized length when travelling along the edge of the polygon.
 The turning function effectively steps over the tangent of a polygon continously for a full cycle, and logs the total turning angle by adding together the angle of each turn.
@@
 
-The primary study of the rate of overlap between roads and other features will serve to obtain the following three metrics:
+The feature overlap study of this project will serve to obtain the following three metrics:
 
@@
 A number of auxillary metrics will be needed to estimate the domain of error on the positional accuracy, namely the feature completeness and shape accuracy of the OSM dataset.
-The completeness of the OSM dataset will be taken as a the number of features are represented in both datasets, as a perentage of the total numer of features in the reference dataset.
-Shape accuracy is a measure of polygon similarity that was obtained by comparing polygon turning function between matching features, feature by feature.
+The completeness of the OSM dataset will be taken as a the number of features that are represented in both datasets, as a perentage of the total numer of features in the reference dataset.
+Shape accuracy (the measure of polygon similarity) was obtained by comparing polygon turning function between matching features, feature by feature.
 

### 2020-06-18 report/report.tex
Clarify that the analysis section deals with footpaths

@@
 
-Looking closer at the individual collision cases, the next conclusion is that most often it is actually a small part of a road that is responsible for the collision, and will need correction.
+Looking closer at the individual collision cases for footpaths, the next conclusion is that most often it is actually a small part of a path that is responsible for the collision, and will need correction.
 Now will follow an exemplifying calculation to support this claim.
@@
 
-After visual inspection, it was determined that all collision occurences can be almost exclusively classified into one of five categories.
+After visual inspection, it was determined that footpath all collision occurences can be almost exclusively classified into one of five categories.
 These categories are explained in the list below, and each case is exemplified in figure \ref{fig:collision-cases}.
@@
 
-\subsection{Suggested algorithms for overlap correction}
-
-All collision cases can be solved with relatively simply geometric algorithms.
+\subsection{Suggested algorithms for footpath overlap correction}
+
+All footpath collision cases can be solved with relatively simply geometric algorithms.
 This section will present suggestions on algorithms that will resolve collisions by identifying individual colliding nodes and translating them, after which some degree of smoothing (such as linear interpolation) can be applied to preserve visual features.

### 2020-06-18 report/report.tex
Clarify the hypothesis of generating an overlap-free map

@@
 % TODO: Introduce the idea of positional error due to low accuracy in this section
+
 \textbf{(Fan et al, 2014)}.
@@
 Specifically in the case of rendering road networks and buildings, a common issue is that the 3D models being generated from road and building data might overlap with each other, which can lead to strange-looking mesh intersections and inaccuracies that modelers often want to avoid.
+\textbf{Feature overlap} is a type of geomerty error that arises due to inaccuracies in the source data, often due to limited precision in the measuring methods used to obtain the data.
 In this project it was hypothised that the risk of feature overlap may be higher between streets and roads designated for energy efficient mobility, such as footpaths, since they are much more organic and dense in their nature than main streets.
-\textbf{Feature overlap} is a type of geomerty error that arises due to inaccuracies in the source data, often due to limited precision in the measuring methods used to obtain the data.
+It was also hypothised that one can use the individual road widths and road vertex coordinates as variables when making adjustments against overlap errors.
+By first detecting critical locations where feature overlap occurs and subsequently correcting this overlap by adjusting the road width and individual vertex positions within the margin of the positional accuracy, this project examined wether or not it was possible to obtain a slightly altered map with little to no overlap between footpaths and other features.
+
 For the purpose of this project, the terms \textbf{feature overlap} and \textbf{feature collision} will be used synonymously, referring to the amount of features that show overlap with each other in the dataset, or the total feature area that show overlap in the same way.
 See figure \ref{fig:img_road_error_cases} for a few visual examples of what feature overlap could look like in the final 3D city model.
-
-% TODO: The idea is to use road width and point coordinates as variables when making adjustments against errors.
 
@@
 This project has dealt with the feasibility of including footpath data when generating 3D representations of OSM data, and preserving the integrity of the 3D road network, avoiding overlap and collision between road meshes.
+It was hypothised that most road overlap can be corrected by first detecting critical locations where feature overlap occurs, and subsequently adjust the footpath road widths and individual vertex positions within the margin of the OSM positional accuracy, in order to obtain a slightly altered map with little to no overlap between footpaths and other features.
+
 The main research question is as follows:

### 2020-06-18 report/report.tex
Introduce positional error

@@
 
-% TODO: Introduce the idea of positional error due to low accuracy in this section
-
+Early on, at the stage of importing geodata into a procedural city reproduction model, the most critical source of error is the \textbf{positional accuracy} that the data provider can ensure.
+Every geodata service has an upper limit on the positional accuracy of their data, which depends primarily on the method of data acquistion that was used.
+Different instruments, such as GPS measurements or satellite image stereoscoping, have theoretical upper limits on their precision, which propagate through the data collection and preservation process as a source of positional error
 \textbf{(Fan et al, 2014)}.
-
-Specifically in the case of rendering road networks and buildings, a common issue is that the 3D models being generated from road and building data might overlap with each other, which can lead to strange-looking mesh intersections and inaccuracies that modelers often want to avoid.
+Specifically in the case of rendering road networks and buildings, a common issue that stems from low positional accuracy is that the 3D models being generated from road and building data might overlap with each other, which can lead to strange-looking mesh intersections and inaccuracies that modelers often want to avoid.
 \textbf{Feature overlap} is a type of geomerty error that arises due to inaccuracies in the source data, often due to limited precision in the measuring methods used to obtain the data.
+
 In this project it was hypothised that the risk of feature overlap may be higher between streets and roads designated for energy efficient mobility, such as footpaths, since they are much more organic and dense in their nature than main streets.

### 2020-06-18 report/report.tex
Update todo
Footpath importing conclusions section

@@
 For the overlap study, see sections 3.1.2 for an implementation overview, 3.3 for a detailed walkthrough of the program implementation, and 4.1.1 for the results.
-
-%TODO: Include a note here about the challenges of identifying what buildings belong together and then finding what vertices belong together.
 
@@
 
-\subsection{Feature Overlap algorithm}
-
-The algorithm for determining feature overlap will calculate how many features and how many of their edges are intersecting with some other feature or edge.
+\subsection{Feature overlap detection algorithm}
+
+The algorithm for detecting feature overlap will calculate how many features and how many of their edges are intersecting with some other feature or edge.
 It will also calculate the total length of all edge which are overlapping with some other edge.
@@
 
-\section{Feasibility of exporting footpaths}
-
-% TODO: Write this section
+\section{Feasibility of importing footpaths}
+
+As has already been established in the analysis of the footpath overlap study, primary conclusion of the footpath overlap study is that, in terms of feature overlap, footpaths were not significantly worse than other road categories.
+Thus it can very well be assumed that city modelling software could use the same routine for importing OSM car roads as well as OSM footpaths, without a drastic increase in feature overlap.
+If it is desired to obtain a map will signifcantly less overlap than what OSM can provide due to limitations on positional accuracy, the designer could build upon their existing procedure and add algorithms for overlap detection and correction.
+Such a procedure could for example utilize the method for overlap detection that was presented in section 3.3.2 of this report, as well as some of the methods for overlap correction that were suggested in section 4.2.3.
+Such a procedure would work for all types of roads, including footpaths, without any special treatment of footpath data in particular.
+
+It was found in the feature overlap study that the individual cases of overlap could be classified into types, where all collisions of the same type could in theory be corrected by the same algoritm.
+Note however that since this study focused exclusively on the Stockholm area, and the feature collisions detected in Stockholm could be of types that are characteristic to Stockholm in particular.
+There could very well be additional types of collisions that are exclusive to cities with other unique city grid aestethics, which would not have been considered by this particular project.
 

### 2020-06-24 report/report.tex
Update todo
Shorten abstract

@@
 \begin{abstract}
-This project assesses the feasibility of reconstructing virtual cities from OpenStreetMap (OSM) road and building data, and will focus on the feasibility of including pedestrian footpaths, by examining how common it is that these footpaths overlap with other map features.
-Feature overlap is a type of geomerty error that arises to inaccuracies in the source data, which results in 3D models overlapping, which can in turn lead to strange-looking mesh intersections and artifacts in the final city model.
-
-In this study, a slice of the OSM dataset in the Stockholm area has been taken, and each road has been assigned with standard withds according to swedish regulation, after which a geometrical algorithm was applied to all road pairs to obtain the average amount of road overlap.
-It was found that footpaths were less problematic than other road types (primary, secondary, residential), in terms of feature overlap, and thus should pose no more problems in 3D city reproduction than other road types.
-
-As a part of this study, the positional accuracy of the OSM dataset was calculated.
+This project assesses the feasibility of reconstructing virtual cities from OpenStreetMap (OSM) road and building data, and will focus on the feasibility of including pedestrian footpaths.
+It will be examined how common it is that footpaths overlap with other map features.
+Feature overlap is a type of geomerty error that arises to inaccuracies in the source data, which can lead to strange-looking mesh intersections and artifacts in the final city model.
+In this study, a slice of the OSM dataset in the Stockholm area has been taken, and a geometrical algorithm was applied to all road pairs to obtain the average amount of road overlap.
+It was found that footpaths were less problematic than car roads in terms of feature overlap, and thus footpaths should pose no more problems in 3D city reproduction than car roads.
+
+As a secondary study, the positional accuracy of the OSM dataset was calculated, since the positional accuracy determines the maximum tolerance in distance between points before it is certain that a road overlaps with another road or building.
 Any geodataset has a limit on its positional accuracy, which is determined by the theoretical precision of the measuring methods.
-For the case of the OSM data used in this project, the accuracy was obtained by comparing coordinate points of OSM building footprints to those in the Swedish University of Agricultural Sciences (SLU) building datset, whose accuracy is known.
-The positional accuracy determines the maximum tolerance in distance between points before it is certain that a road overlaps with another road or building.
+In this project, the accuracy was obtained by comparing coordinate points of OSM building footprints to those in the Swedish University of Agricultural Sciences (SLU) building datset, whose accuracy is known.
 
@@
 
-% After Chris approves
+% TODO: Even shorter abstract
+% TODO: New structure: Background, problem/question, implementation/method, evaluation, results, implications
+% TODO: Ensure in the abstract that it is visible that there are two research questions
+% TODO: Restructure research subquestions according to Chris's suggestions (see meeting log)
+
+% With internet
+% TODO: Fix result table margins
 % TODO: Spell check

### 2020-06-24 report/report.tex
Swedish abstract

@@
 This project assesses the feasibility of reconstructing virtual cities from OpenStreetMap (OSM) road and building data, and will focus on the feasibility of including pedestrian footpaths.
-It will be examined how common it is that footpaths overlap with other map features.
+It will be examined how common it is that footpaths overlap with other map features such as roads and buildings.
 Feature overlap is a type of geomerty error that arises to inaccuracies in the source data, which can lead to strange-looking mesh intersections and artifacts in the final city model.
@@
 Studien har undersökt hur vanligt det är att gångvägar överlappar med andra kartdetaljer såsom byggnader och andra vägar.
-Överlappning är en typ av fel i 3D-geometrin som uppstår på grund av osäkerhet i källdatan, vilket resulterar i att 3D-modeller överlappar med varandra.
-Detta kan leda till besvärliga skärningar och artifakter i den färdiga stadsmodellen.
-
-I denna undersökning har en delmängd av OSM-datamängden över Stockholm tagits, och varje väg har givits en standardbredd enligt svenska bestämmelser.
-Sedan har en geometrisk algoritm tagit alla möjliga vägpar och beräknat hur stor mängd av vägarna som skulle överlappa med andra kartdetaljer.
-Studien fann att gångvägar är mindre problematiska än andra vägtyper beträffande hur många vägar som kolliderar, och således borde inte gångvägar vara svårare att inkludera i en 3D-miljö än andra vägtyper.
-
-I studien har felmarginalen för OSM-datamängden också tagits fram.
+Överlappning är en typ av fel i 3D-geometrin som uppstår på grund av osäkerhet i källdatan, vilket kan leda till besvärliga skärningar och artifakter i den färdiga stadsmodellen.
+I denna undersökning har en delmängd av OSM-datamängden över Stockholm tagits, och en geometrisk algoritm tagit alla möjliga vägpar och beräknat hur stor mängd av vägarna som skulle överlappa med andra kartdetaljer.
+Studien fann att gångvägar är mindre problematiska än bilvägar beträffande hur många vägar som kolliderar, och således borde inte gångvägar vara svårare att inkludera i en 3D-miljö än bilvägar.
+
+Som en andra del av studien har felmarginalen för OSM-datamängden också tagits fram, eftersom felmarginalen bestämmer toleransen mellan två punkter som kan tillåtas innan man är säker på att en väg börjar överlappa med en annan väg eller en byggnad.
 Varje geodatamängd har en gräns på hur väl den stämmer överrens med verkligheten, som bestäms av vilka mätmedoter som har använts i datans insamling.
 I detta projekt uppskattades felmarginalen genom att jämföra byggnaders koordinatpunkter i datamängen med de från en referenskarta från Svenska Lantbruksuniversitetet (SLU) med känd felmarginal.
-Felmarginalen bestämmer toleransen mellan två punkter som kan tillåtas innan man är säker på att en väg börjar överlappa med en annan väg eller en byggnad.
 

### 2020-06-24 report/report.tex
Update todo
Rename implementation -> method

@@
 For the purpose of this report, features will refer mainly to roads and building footprints.
-For the overlap study, see sections 3.1.2 for an implementation overview, 3.3 for a detailed walkthrough of the program implementation, and 4.1.1 for the results.
+For the overlap study, see sections 3.1.2 for a method overview, 3.3 for a detailed walkthrough of the method, and 4.1.1 for the results.
 
@@
 The positional accuracy will be needed to assess where road overlap occurs and whether the collisions can easily be corrected.
-For the positional accuracy study, see sections 3.1.1 for an implementation overview, 3.2 for a walkthrough of the program implementation, and 4.1.2 for the results.
+For the positional accuracy study, see sections 3.1.1 for a method overview, 3.2 for a walkthrough of the method, and 4.1.2 for the results.
 
@@
 
-\chapter{Implementation}
+\chapter{Method}
 
@@
 
-\section{Implementation overview}
-
-Programmatic collection of statistics has been neccessary for both the primary and secondary research questions, and this section presents a broad overview of the technical implementation behind the map analysis and data collection processes for both the geodata precision study and the overlap study.
+\section{Method overview}
+
+Programmatic collection of statistics has been neccessary for both the primary and secondary research questions, and this section presents a broad overview of the technical details behind the map analysis and data collection processes for both the geodata precision study and the overlap study.
 
@@
 
-% TODO: Even shorter abstract
 % TODO: New structure: Background, problem/question, implementation/method, evaluation, results, implications
-% TODO: Ensure in the abstract that it is visible that there are two research questions
+
+% With internet
 % TODO: Restructure research subquestions according to Chris's suggestions (see meeting log)
-
-% With internet
 % TODO: Fix result table margins

### 2020-06-24 report/report.tex
Move evaluation metrics section to the end of the method chapter
Restructure research questions to make the division clearer

@@
 Again, note that feature overlap is defined here as geomerty errors due to inaccuracies in the source data, which results in 3D models overlapping, which can in turn lead to strange-looking mesh intersections.
-To investigate the feasibility of generating a 3D virtual city with minimal footpath overlap, the main research question will specifically be divided into the following two subquestion:
-
-\begin{enumerate}
-    \item The primary subquestion is: \textbf{How many roads of each type will collide with some other map feature, once they have been given a geometrical width?} This question aims to find the rate of collision between road features in OSM, and investigate the nature of the most common types of collisions and how they can be corrected.
-    \item The secondary subquestion is: \textbf{What is the positional accuracy of OSM map features compared to a reference map?}.
-The positional accuracy will be neccessary to carry out the investigation of the first subquestion: the rate of collisions between roads, since it will determine the maximum tolerance in distance between road points before a collision is certain to have occured.
-\end{enumerate}
-
-For the primary research question: the road overlap study, a program has taken a slice of the OSM dataset in the Stockholm area and assigned each road with standard withds according to swedish regulation.
+To investigate the feasibility of generating a 3D virtual city with minimal footpath overlap, the main research question will specifically be divided into two subquestions. The primary subquestion is:
+
+\begin{center}
+    \textbf{How many roads of each type will collide with some other map feature, once they have been given a geometrical width?}
+\end{center}
+
+This question aims to find the rate of collision between road features in OSM, and investigate the nature of the most common types of collisions and how they can be corrected.
+For the road overlap study, a program has taken a slice of the OSM dataset in the Stockholm area and assigned each road with standard withds according to swedish regulation.
 It has then identified critical areas in the OSM dataset, where the paths are so wide that they collide with existing map features.
@@
 
+The secondary subquestion is:
+
+\begin{center}
+    \textbf{What is the positional accuracy of OSM map features compared to a reference map?}.
+\end{center}
+
+The positional accuracy will be neccessary to carry out the investigation of the first subquestion: the rate of collisions between roads, since it will determine the maximum tolerance in distance between road points before a collision is certain to have occured.
 For the secondary research question: the geodata precision study, the positional accuracy will be estimated by comparing feature points between the OSM building dataset and the Swedish University of Agricultural Sciences (SLU) building datset with known accuracy.
@@
 A modeler may chose which of these LOD class(es) to use depending on data availablity and application.
-
-\section{Evaluation metrics}
-
-The feature overlap study of this project will serve to obtain the following three metrics:
-
-\begin{enumerate}
-    \item The number of road features that collide with any other feature (road or building).
-    \item The number of road edges that intersect with any other building polygon or road polyline.
-    \item The total length in meters of these edges will also be calculated as a measure of the total length of road.
-\end{enumerate}
-
-The road features will be categorized according to type, so that the rate of overlap between footpaths and other features can be compared to that of other road types.
-OSM mainly makes the distinction between primary, secondary, residential and footpath roads, so this classification will also be used in this report.
-The above features will be calculated for each road type separately by summing over each road feature of a certain type.
-Following this, an algorithm will find the exact same metrics with regards to how many road features that can in theory be corrected by simple geometrical translation of their vertices.
-This will also be preseneted in terms of the number of road features, the number of edges and the total length in meters of remaining colliding road.
-
-The secondary study about geodata precision will primarily serve to obtain an estimate of the positional accuracy of the OSM dataset, in meters.
-A number of auxillary metrics will be needed to estimate the domain of error on the positional accuracy, namely the feature completeness and shape accuracy of the OSM dataset.
-The completeness of the OSM dataset will be taken as a the number of features that are represented in both datasets, as a perentage of the total numer of features in the reference dataset.
-Shape accuracy (the measure of polygon similarity) was obtained by comparing polygon turning function between matching features, feature by feature.
-
-\subsection{Quality criteria}
-
-\textbf{(Haklay, 2010)} presents the list of van Oort's 8 accuracy classes for evaluating the quality of geographic information:
-
-\begin{itemize}
-  \item Lineage. This is the historical aspect of the dataset, which concerns the collection process and evolution.
-  \item Positional accuracy. This relates the coordinate value of an object in the database to the actual location of the ground in the real world.
-  \item Attibute accuracy. In a geographical database, objects are commonly tagged with meta-information. This class assesses how correct those values are.
-  \item Logical consistency - This assesses the internal consistency of the dataset.
-For every dataset there may be internal rules and relationships that objects and features must follow, and this class assesses the degree to which these are adhered to.
-  \item Completeness - This assesses the lack of data in a dataset, and the coverage of real-world objects. Objects or features may be missing from a dataset, which reduces its quality.
-  \item Semantic accuracy - This links the way in which an object or feature is recorded and represented to how it should be interpreted.
-  \item Usage, purpose and constraints - This concerns the validity of the dataset in relation to its purpose and how it is used.
-  \item Temporal quality - This assesses the validity of the dataset in relation to real-world changes over time.
-\end{itemize}
-
-Other than for the purpose of assessing various sources or error, this project will focus exclusively on the positional accuracy, completeness and semantic accuracy classes.
-The comparison of the OSM and reference maps will focus on these accuracy classes to determine the error tolerance when manipulating features in the 3D map presentation and the road network generating phase.
-Figure \ref{fig:osm-slu-map} shows two segments of the Stockholm map where the OSM and reference datasets show lower semantic and positional accuracy respectively.
-
-\begin{figure}[H]
-    \centering
-    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_map_compare}
-    \caption{Example segments of the Stockholm metropolitan area with unsystematic errors in precision.
-The reference building dataset (SLU, blue) in superimposed on the OSM dataset (dark gray).
-The top picture is centered on a city block which shows a semantic mismatching between the two datasets.
-The whole block consists of only two buildings in the OSM dataset, but is divided into 15 smaller building lots in the SLU dataset.
-The bottom picture shows 4 smaller buildings whose the positional accuracy is low, due to the buildings being incorrectly scaled, skewed and rotated.}
-    \label{fig:osm-slu-map}
-\end{figure}
 
@@
 
+\section{Evaluation metrics}
+
+The feature overlap study of this project will serve to obtain the following three metrics:
+
+\begin{enumerate}
+    \item The number of road features that collide with any other feature (road or building).
+    \item The number of road edges that intersect with any other building polygon or road polyline.
+    \item The total length in meters of these edges will also be calculated as a measure of the total length of road.
+\end{enumerate}
+
+The road features will be categorized according to type, so that the rate of overlap between footpaths and other features can be compared to that of other road types.
+OSM mainly makes the distinction between primary, secondary, residential and footpath roads, so this classification will also be used in this report.
+The above features will be calculated for each road type separately by summing over each road feature of a certain type.
+Following this, an algorithm will find the exact same metrics with regards to how many road features that can in theory be corrected by simple geometrical translation of their vertices.
+This will also be preseneted in terms of the number of road features, the number of edges and the total length in meters of remaining colliding road.
+
+The secondary study about geodata precision will primarily serve to obtain an estimate of the positional accuracy of the OSM dataset, in meters.
+A number of auxillary metrics will be needed to estimate the domain of error on the positional accuracy, namely the feature completeness and shape accuracy of the OSM dataset.
+The completeness of the OSM dataset will be taken as a the number of features that are represented in both datasets, as a perentage of the total numer of features in the reference dataset.
+Shape accuracy (the measure of polygon similarity) was obtained by comparing polygon turning function between matching features, feature by feature.
+
+\subsection{Quality criteria}
+
+\textbf{(Haklay, 2010)} presents the list of van Oort's 8 accuracy classes for evaluating the quality of geographic information:
+
+\begin{itemize}
+  \item Lineage. This is the historical aspect of the dataset, which concerns the collection process and evolution.
+  \item Positional accuracy. This relates the coordinate value of an object in the database to the actual location of the ground in the real world.
+  \item Attibute accuracy. In a geographical database, objects are commonly tagged with meta-information. This class assesses how correct those values are.
+  \item Logical consistency - This assesses the internal consistency of the dataset.
+For every dataset there may be internal rules and relationships that objects and features must follow, and this class assesses the degree to which these are adhered to.
+  \item Completeness - This assesses the lack of data in a dataset, and the coverage of real-world objects. Objects or features may be missing from a dataset, which reduces its quality.
+  \item Semantic accuracy - This links the way in which an object or feature is recorded and represented to how it should be interpreted.
+  \item Usage, purpose and constraints - This concerns the validity of the dataset in relation to its purpose and how it is used.
+  \item Temporal quality - This assesses the validity of the dataset in relation to real-world changes over time.
+\end{itemize}
+
+Other than for the purpose of assessing various sources or error, this project will focus exclusively on the positional accuracy, completeness and semantic accuracy classes.
+The comparison of the OSM and reference maps will focus on these accuracy classes to determine the error tolerance when manipulating features in the 3D map presentation and the road network generating phase.
+Figure \ref{fig:osm-slu-map} shows two segments of the Stockholm map where the OSM and reference datasets show lower semantic and positional accuracy respectively.
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_map_compare}
+    \caption{Example segments of the Stockholm metropolitan area with unsystematic errors in precision.
+The reference building dataset (SLU, blue) in superimposed on the OSM dataset (dark gray).
+The top picture is centered on a city block which shows a semantic mismatching between the two datasets.
+The whole block consists of only two buildings in the OSM dataset, but is divided into 15 smaller building lots in the SLU dataset.
+The bottom picture shows 4 smaller buildings whose the positional accuracy is low, due to the buildings being incorrectly scaled, skewed and rotated.}
+    \label{fig:osm-slu-map}
+\end{figure}
+
 \chapter{Evaluation}
@@
 
-% TODO: New structure: Background, problem/question, implementation/method, evaluation, results, implications
-
 % With internet

### 2020-06-24 report/report.tex
Spell check

@@
 
-\title{Quality Assesment of OpenStreetMap Footpath Data}
-\alttitle{Kvalitetiv undersökning av OpenStreetMaps gångvägsdata}
+\title{Quality Assessment of OpenStreetMap Footpath Data}
+\alttitle{Kvalitativ undersökning av OpenStreetMaps gångvägsdata}
 \author{Alexander Hjelm}
@@
 It will be examined how common it is that footpaths overlap with other map features such as roads and buildings.
-Feature overlap is a type of geomerty error that arises to inaccuracies in the source data, which can lead to strange-looking mesh intersections and artifacts in the final city model.
+Feature overlap is a type of geometry error that arises to inaccuracies in the source data, which can lead to strange-looking mesh intersections and artifacts in the final city model.
 In this study, a slice of the OSM dataset in the Stockholm area has been taken, and a geometrical algorithm was applied to all road pairs to obtain the average amount of road overlap.
@@
 Any geodataset has a limit on its positional accuracy, which is determined by the theoretical precision of the measuring methods.
-In this project, the accuracy was obtained by comparing coordinate points of OSM building footprints to those in the Swedish University of Agricultural Sciences (SLU) building datset, whose accuracy is known.
+In this project, the accuracy was obtained by comparing coordinate points of OSM building footprints to those in the Swedish University of Agricultural Sciences (SLU) building dataset, whose accuracy is known.
 
@@
 Virtual cities has since long been a popular research topic in the field of computer visualization.
-The idea of recreating real world cities in 3D applications has a growing number of applications, including visualization and modelling, urban planning and architecture, as well as media and entertainment in the form of video games and 3D animation.
-Traditionally, modelling virtual cities is a labour-intensive and time-consuming task due to the need for manual modeling, texturing and design and placement of individual models in the large 3D scene.
+The idea of recreating real world cities in 3D applications has a growing number of applications, including visualization and modeling, urban planning and architecture, as well as media and entertainment in the form of video games and 3D animation.
+Traditionally, modeling virtual cities is a labor-intensive and time-consuming task due to the need for manual modeling, texturing and design and placement of individual models in the large 3D scene.
 This type of tasks is commonly done by professional 3D graphical artists and designers.
-However, in recent years there has been major research work in the field of procedural content generation, meaning automatic or semi automatic generation of 3D models without the direct involvment of a 3D artist.
+However, in recent years there has been major research work in the field of procedural content generation, meaning automatic or semi automatic generation of 3D models without the direct involvement of a 3D artist.
 \textbf{Procedural models} typically involve a machine which algorithmically composes together or alters human-generated 3D assets.
@@
 
-Various procedural methods are indeed used in the field of virtual city reproduction to reduce the labour required to create accurate 3D representations of cities.
-These models rely on the availablity of spatial \textbf{geodata}, which serves as the initial input to the procedure of model generation and placement.
+Various procedural methods are indeed used in the field of virtual city reproduction to reduce the labor required to create accurate 3D representations of cities.
+These models rely on the availability of spatial \textbf{geodata}, which serves as the initial input to the procedure of model generation and placement.
 Geodata can contain any form of map features, from terrain to road networks and buildings, but should ideally be easily retrievable in a format that the procedural model can easily work with, e.g.
@@
 
-The field of procedual city reproduction comes with its own host of challenges however.
-In practise, urban environments are typically dense and large, spanning a few to hundreds of square kilometers, and are influenced by variables that are hard to quantify, including land policies, government plans, population changes and transportation infrastructure, making it non-trivial to produce 3D city representations with both high accuracy and high detail, while simultaneuosly having high utility in the intended field of study.
+The field of procedural city reproduction comes with its own host of challenges however.
+In practice, urban environments are typically dense and large, spanning a few to hundreds of square kilometers, and are influenced by variables that are hard to quantify, including land policies, government plans, population changes and transportation infrastructure, making it non-trivial to produce 3D city representations with both high accuracy and high detail, while simultaneously having high utility in the intended field of study.
 \textbf{(Vanegas et al, 2010)}.
@@
     \caption{Real-world cities are complex collections of buildings, blocks and road networks that interconnect them all.
-Creating 3D virtual cities using methods of procedual generation is an emerging research field that comes with its own set of challenges.
+Creating 3D virtual cities using methods of procedural generation is an emerging research field that comes with its own set of challenges.
 This project deals with generating virtual cities from OpenStreetMap data, and the technical challenges therein.$^0$}
@@
 For the purpose of this report, a \textbf{road} will be any road whose primary use is human travel, either via a vehicle or by foot.
-Note that by this definition even sidewalks are classifed as roads, even though they commonly run in connection to larger roads (which are usually reserved for motor vehicles), and thus may not always be thought of as roads in plain language.
+Note that by this definition even sidewalks are classified as roads, even though they commonly run in connection to larger roads (which are usually reserved for motor vehicles), and thus may not always be thought of as roads in plain language.
 
@@
 The width, or thickness, of a road at a certain point will be defined as the width of the perpendicular road cross section at that point, which is in its entirety usable by the intended types of traffic.
-It is important to note that OSM does commonly not include a defition of the width of any road.
-An optional tag exists that lets users specify the average path width of a road, but this tag is rarely used in practise.
+It is important to note that OSM does commonly not include a definition of the width of any road.
+An optional tag exists that lets users specify the average path width of a road, but this tag is rarely used in practice.
 
@@
 
-When it comes to the quality of available data, it turns out that in practise, geodata databases without errors are rare.
+When it comes to the quality of available data, it turns out that in practice, geodata databases without errors are rare.
 Geodata validation and correction is only an emerging field of research, and many such datasets were usually obtained years ago, and creating error-free datasets was neither possible due to the lack of validation techniques, nor a high priority.
@@
 Early on, at the stage of importing geodata into a procedural city reproduction model, the most critical source of error is the \textbf{positional accuracy} that the data provider can ensure.
-Every geodata service has an upper limit on the positional accuracy of their data, which depends primarily on the method of data acquistion that was used.
+Every geodata service has an upper limit on the positional accuracy of their data, which depends primarily on the method of data acquisition that was used.
 Different instruments, such as GPS measurements or satellite image stereoscoping, have theoretical upper limits on their precision, which propagate through the data collection and preservation process as a source of positional error
@@
 Specifically in the case of rendering road networks and buildings, a common issue that stems from low positional accuracy is that the 3D models being generated from road and building data might overlap with each other, which can lead to strange-looking mesh intersections and inaccuracies that modelers often want to avoid.
-\textbf{Feature overlap} is a type of geomerty error that arises due to inaccuracies in the source data, often due to limited precision in the measuring methods used to obtain the data.
-
-In this project it was hypothised that the risk of feature overlap may be higher between streets and roads designated for energy efficient mobility, such as footpaths, since they are much more organic and dense in their nature than main streets.
-It was also hypothised that one can use the individual road widths and road vertex coordinates as variables when making adjustments against overlap errors.
-By first detecting critical locations where feature overlap occurs and subsequently correcting this overlap by adjusting the road width and individual vertex positions within the margin of the positional accuracy, this project examined wether or not it was possible to obtain a slightly altered map with little to no overlap between footpaths and other features.
+\textbf{Feature overlap} is a type of geometry error that arises due to inaccuracies in the source data, often due to limited precision in the measuring methods used to obtain the data.
+
+In this project it was hypothesized that the risk of feature overlap may be higher between streets and roads designated for energy efficient mobility, such as footpaths, since they are much more organic and dense in their nature than main streets.
+It was also hypothesized that one can use the individual road widths and road vertex coordinates as variables when making adjustments against overlap errors.
+By first detecting critical locations where feature overlap occurs and subsequently correcting this overlap by adjusting the road width and individual vertex positions within the margin of the positional accuracy, this project examined whether or not it was possible to obtain a slightly altered map with little to no overlap between footpaths and other features.
 
@@
     \includegraphics[width=\textwidth,height=0.5\textheight,keepaspectratio]{img_road_error_cases}
-    \caption{Example segments of a 3D reconstruciton of the Stockholm road map, with clearly visible geometric errors.
+    \caption{Example segments of a 3D reconstruction of the Stockholm road map, with clearly visible geometric errors.
 The road overlap is marked with red boundaries.
-The segments show firstly a healthy intersection with main car roads, sidewalks and pedestrian crossings (top left), secondly a city segment where roads have been placed so they overlap with existing buildings due to insufficient data precision (top right), thirdly a similar case where roads overlap with each other (bottom left) and lastly an intersection where the road geomerty twists and self-intersects in strange ways due to invalid source data.
+The segments show firstly a healthy intersection with main car roads, sidewalks and pedestrian crossings (top left), secondly a city segment where roads have been placed so they overlap with existing buildings due to insufficient data precision (top right), thirdly a similar case where roads overlap with each other (bottom left) and lastly an intersection where the road geometry twists and self-intersects in strange ways due to invalid source data.
 This last case will not be examined in this project.}
@@
 The OSM project is a freely available wiki-like geodatabase based on Volunteered Geographical Information (VGI), meaning anyone can contribute by adding or editing features.
-The OSM world dataset is being consistently updated by volonteers from all over the world, and the main idea is that by using OSM geodata for city reproduction, the temporal aspect of data accuracy would in theory be more assured by the crowd contribution nature.
+The OSM world dataset is being consistently updated by volunteers from all over the world, and the main idea is that by using OSM geodata for city reproduction, the temporal aspect of data accuracy would in theory be more assured by the crowd contribution nature.
 There are already a few such services that use OSM data to generate 3D geodata visualizations for media and video game applications.
@@
 This project has dealt with the feasibility of including footpath data when generating 3D representations of OSM data, and preserving the integrity of the 3D road network, avoiding overlap and collision between road meshes.
-It was hypothised that most road overlap can be corrected by first detecting critical locations where feature overlap occurs, and subsequently adjust the footpath road widths and individual vertex positions within the margin of the OSM positional accuracy, in order to obtain a slightly altered map with little to no overlap between footpaths and other features.
+It was  that most road overlap can be corrected by first detecting critical locations where feature overlap occurs, and subsequently adjust the footpath road widths and individual vertex positions within the margin of the OSM positional accuracy, in order to obtain a slightly altered map with little to no overlap between footpaths and other features.
 
@@
 \begin{center}
-    \textbf{Is it technically feasibile to generate a 3D virtual city with footpath road data, with minimal (and preferably no) overlap between 3D features?}
+    \textbf{Is it technically feasible to generate a 3D virtual city with footpath road data, with minimal (and preferably no) overlap between 3D features?}
 \end{center}
 
-Again, note that feature overlap is defined here as geomerty errors due to inaccuracies in the source data, which results in 3D models overlapping, which can in turn lead to strange-looking mesh intersections.
+Again, note that feature overlap is defined here as geometry errors due to inaccuracies in the source data, which results in 3D models overlapping, which can in turn lead to strange-looking mesh intersections.
 To investigate the feasibility of generating a 3D virtual city with minimal footpath overlap, the main research question will specifically be divided into two subquestions. The primary subquestion is:
@@
 This question aims to find the rate of collision between road features in OSM, and investigate the nature of the most common types of collisions and how they can be corrected.
-For the road overlap study, a program has taken a slice of the OSM dataset in the Stockholm area and assigned each road with standard withds according to swedish regulation.
+For the road overlap study, a program has taken a slice of the OSM dataset in the Stockholm area and assigned each road with standard widths according to swedish regulation.
 It has then identified critical areas in the OSM dataset, where the paths are so wide that they collide with existing map features.
@@
 For the purpose of this report, features will refer mainly to roads and building footprints.
-For the overlap study, see sections 3.1.2 for a method overview, 3.3 for a detailed walkthrough of the method, and 4.1.1 for the results.
+For the overlap study, see sections 3.1.2 for a method overview, 3.3 for a detailed walk through of the method, and 4.1.1 for the results.
 
@@
 
-The positional accuracy will be neccessary to carry out the investigation of the first subquestion: the rate of collisions between roads, since it will determine the maximum tolerance in distance between road points before a collision is certain to have occured.
-For the secondary research question: the geodata precision study, the positional accuracy will be estimated by comparing feature points between the OSM building dataset and the Swedish University of Agricultural Sciences (SLU) building datset with known accuracy.
+The positional accuracy will be necessary to carry out the investigation of the first subquestion: the rate of collisions between roads, since it will determine the maximum tolerance in distance between road points before a collision is certain to have occurred.
+For the secondary research question: the geodata precision study, the positional accuracy will be estimated by comparing feature points between the OSM building dataset and the Swedish University of Agricultural Sciences (SLU) building dataset with known accuracy.
 See section 3.4.2 for an overview of the SLU dataset and how it is obtained.
 The positional accuracy will be needed to assess where road overlap occurs and whether the collisions can easily be corrected.
-For the positional accuracy study, see sections 3.1.1 for a method overview, 3.2 for a walkthrough of the method, and 4.1.2 for the results.
+For the positional accuracy study, see sections 3.1.1 for a method overview, 3.2 for a walk through of the method, and 4.1.2 for the results.
 
@@
 Lineage is defined by \textbf{(Van Oort, 2006)} as the historical aspect of the dataset: how it has been obtained, with what instruments and how it has evolved between the moment it was obtained and the moment it was retrieved.
-The entire OSM dataset is under constant modification by volonteers, so it is expected that the accuracy will improve over time.
-Previous studies highlight how the posiitonal accuracy in similar-sized cities have improved over the course of several years (\textbf{Fan et al, 2014}, \textbf{Haklay, 2010}).
+The entire OSM dataset is under constant modification by volunteers, so it is expected that the accuracy will improve over time.
+Previous studies highlight how the positional accuracy in similar-sized cities have improved over the course of several years (\textbf{Fan et al, 2014}, \textbf{Haklay, 2010}).
 The historical aspect of the OSM dataset is available through the OSM History Viewer
 \footnote{OSM History Viewer - OpenStreetMap Wiki. Accessed 20-03-2020. https://wiki.openstreetmap.org/wiki/OSM\_History\_Viewer}
-, an online debugging tool that lets anyone freely view the change history of individual fetures in a commit history-like fashion.
-There is an option for editors to include a personal note with their changesets to include additional details or motivate why a change was made, but there is no guaratee that the data includes any information on the aquisition method used.
+, an online debugging tool that lets anyone freely view the change history of individual features in a commit history-like fashion.
+There is an option for editors to include a personal note with their changesets to include additional details or motivate why a change was made, but there is no guarantee that the data includes any information on the acquisition method used.
 
@@
 how many lanes a road has, if the base is asphalt or gravel, or how many stories a building has.
-Attribute accuracy is the meausere of how complete attribute information is in a dataset.
-How much attribute data is missing and how much of it does not comform with reality?
+Attribute accuracy is the measure of how complete attribute information is in a dataset.
+How much attribute data is missing and how much of it does not conform with reality?
 In general, attribute accuracy is hard to quantify.
-In this case however, it might be interesting to compare attributes between the OSM map and the SLU referece map that was used in this project, as both use encoded metadata on a per feature basis, and matching features can be easily identified.
+In this case however, it might be interesting to compare attributes between the OSM map and the SLU reference map that was used in this project, as both use encoded metadata on a per feature basis, and matching features can be easily identified.
 See section 3.4.2 for an overview of the reference dataset that was used in this study.
@@
 Lastly, it will present the evaluation metrics that this project served to obtain, and how they relate to the overlap analysis.
-For the geodata precision study specifically, this chapter will examine how geodata is qualifyed in general, by examining van Oorts quality criteria, and how these will be used in this project.
+For the geodata precision study specifically, this chapter will examine how geodata is qualified in general, by examining van Oort's quality criteria, and how these will be used in this project.
 
@@
 
-In their analysis of prodedural city generation models, \textbf{(Müller et al, 2006)} provide an overview of the process of recreating virtual 3D citites from their real life conterparts.
-They present cities as complex collections of \textbf{buildings}, \textbf{blocks}, \textbf{parcels} and \textbf{neighbourhoods} which are all interconnected via a \textbf{street} network, where individual streets can serve different types of traffic.
+In their analysis of procedural city generation models, \textbf{(Müller et al, 2006)} provide an overview of the process of recreating virtual 3D cities from their real life counterparts.
+They present cities as complex collections of \textbf{buildings}, \textbf{blocks}, \textbf{parcels} and \textbf{neighborhoods} which are all interconnected via a \textbf{street} network, where individual streets can serve different types of traffic.
 These features are commonly represented as \textbf{geodata} in map databases, and processed and organized into visualization-ready data by a \textbf{Geographic Information System (GIS)}.
-In commercially available geodata, roads are most often defined as polylines: arrays of global coordinate points that come in the same order as they would normally be visited by a person travelling along the road.
+In commercially available geodata, roads are most often defined as polylines: arrays of global coordinate points that come in the same order as they would normally be visited by a person traveling along the road.
 Buildings in turn are most often defined by their \textbf{building footprints}, a single polygon with global coordinates as vertices, which defines the extents of the building as seen on the map from above.
-In practise, cities are often made up out of many square kilometers worth of geodata, and it would be highly labour intensive to manually recreate all these features in 3D.
-Instead, \textbf{procedural models} for city reproduction instead rely on programmatic solutions for converting map data to 3D road and building models, and composing them to a 3D world that the designer may continue to fine-tweak through interfaces fo interactive editing.
-
-The \textbf{road network} is what ties the city together as a whole, and forms the grid that divides the lots, quarters and neighbourhoods, giving the city grid its unqigue aestethic character and defining possible building arrangements.
-Thus, for any procedural city reproduction model it will be neccessary to have a robust procedure for extracting road geodata and generating the 3D represention of it without compromising the data integrity.
-So far there has been a lot of research into this area, but \textbf{(Stojanovski, 2019)} claims that most of these projects have often focused on the road networks designated for cars and public transit, as a part of the sustainable mobility paradigm and the topic of sustainable cities, particularily due to the interest in self-propelled cars.
+In practice, cities are often made up out of many square kilometers worth of geodata, and it would be highly labor intensive to manually recreate all these features in 3D.
+Instead, \textbf{procedural models} for city reproduction instead rely on programmatic solutions for converting map data to 3D road and building models, and composing them to a 3D world that the designer may continue to fine-tweak through interfaces for interactive editing.
+
+The \textbf{road network} is what ties the city together as a whole, and forms the grid that divides the lots, quarters and neighborhoods, giving the city grid its unique aestethic character and defining possible building arrangements.
+Thus, for any procedural city reproduction model it will be necessary to have a robust procedure for extracting road geodata and generating the 3D representation of it without compromising the data integrity.
+So far there has been a lot of research into this area, but \textbf{(Stojanovski, 2019)} claims that most of these projects have often focused on the road networks designated for cars and public transit, as a part of the sustainable mobility paradigm and the topic of sustainable cities, particularly due to the interest in self-propelled cars.
 The study of energy efficient mobility such as walking and cycling is only in its emerging phase.
@@
 
-This project falls within the broader topic of geodata extraction and modelling, primarily of roads and building parcels.
-Geodata modeling is at the foundation of procedural modelling of real-world cities, since any software which models cities based on geodata will need a robust method of aqquiring and laying out road and building data without compromising the data integrity.
+This project falls within the broader topic of geodata extraction and modeling, primarily of roads and building parcels.
+Geodata modeling is at the foundation of procedural modeling of real-world cities, since any software which models cities based on geodata will need a robust method of acquiring and laying out road and building data without compromising the data integrity.
 In recent years there has been an advent of interactive tools that aid urban planners in placing or generating features of an urban plan, particularly roads and road networks.
-Such tools use procedural methods and AI solutions to move the burden of labour from the human designer to the software that the designer uses
+Such tools use procedural methods and AI solutions to move the burden of labor from the human designer to the software that the designer uses
 \textbf{(Müller et al, 2006)}.
 
-The last decade has seen multiple startups in the area of procedural software for urban arcitecture.
-Such software uses procedural models to render a city grid based on real-world locations, or generate new grids and features based on user-specified parameters, and allow for varying degrees of manual editing on grid, road, neighbourhood, city block or individual feature level.
+The last decade has seen multiple startups in the area of procedural software for urban architecture.
+Such software uses procedural models to render a city grid based on real-world locations, or generate new grids and features based on user-specified parameters, and allow for varying degrees of manual editing on grid, road, neighborhood, city block or individual feature level.
 Some of the most prominent examples are ArcGIS Pro (made by Esri), CityEngine (made by the startup Procedural, bought by ESRI) and Urban Canvas (made by the startup Synthicity, bought by Autodesk, now defunct) \textbf{(Stojanovski et al, 2019)}.
 
-Even as such tools are well available, \textbf{(Stojanovski et al, 2020)} note that they are as of today not commonly applied in urbanist practises, due to failures in reflecting the unique workspace of arcitects and their design needs.
-Urban arcitecture is a field that not only accounts for spatial data, but also captures many social factors and the individual wills of the many stakeholders.
-Each arcitect is an actor in a continously developing environment, and even the archtiect themselves is affecting that environment in a complex manner through their actions.
+Even as such tools are well available, \textbf{(Stojanovski et al, 2020)} note that they are as of today not commonly applied in urbanist practices, due to failures in reflecting the unique workspace of architects and their design needs.
+Urban architecture is a field that not only accounts for spatial data, but also captures many social factors and the individual wills of the many stake holders.
+Each architect is an actor in a continuously developing environment, and even the architect themselves is affecting that environment in a complex manner through their actions.
 
@@
 
-When assessing the completeness of geodata, a very common scale is Van Oorts criteria for evaluating the quality of geographical information.
+When assessing the completeness of geodata, a very common scale is Van Oort's criteria for evaluating the quality of geographical information.
 At the time of publication, this method of evaluation received much attention from surveyors and cartographers \textbf{(Van Oort, 2006)}, and since then a number of publications have examined slices of geodata by using van Oort's criteria.
@@
 The study was conducted by comparing OSM data to Ordnance Survey datasets, and it showed that the OSM dataset had a rough positional accuracy of 6 meters from the reference dataset.
-However, at the time of the study, the OSM project had only captured roughly 29\% of the area of England, meaning that the data completeness was fairly low, and thus the uncertanity was relatively high.
+However, at the time of the study, the OSM project had only captured roughly 29\% of the area of England, meaning that the data completeness was fairly low, and thus the uncertainty was relatively high.
 
 The next major study to assess the quality of OSM data was published by \textbf{(Kunze, 2012)}.
-The study applied different methods to assess the completeness of OSM data in two federal states in Germany, mainly by analysing the area difference between the OSM dataset and an administrative dataset.
+The study applied different methods to assess the completeness of OSM data in two federal states in Germany, mainly by analyzing the area difference between the OSM dataset and an administrative dataset.
 Other notable studies about OSM data accuracy were made around the same time in Germany by \textbf{(Zielstra and Zipf, 2010)} and \textbf{(Neis et al, 2012)}, as well as in France by \textbf{(Girres and Touya, 2010)}.
@@
 Finally, a major study in 2013-2014, conducted by researchers from various universities in Norway and Germany \textbf{(Fan et al, 2014)}, examined the quality and accuracy in building footprint data of the Munich area, at a time and place where the completeness of the OSM dataset was significantly higher than the study by \textbf{(Haklay, 2010)} (the results of their study showed that the OSM had captured close to 100\% of all building footprints).
-This new study included an insight in the geometrical calculations neccessary to match features and points between two datasets, and how to reliably calculate the metrics needed for the most important and assessible of van Oorts criteria.
+This new study included an insight in the geometrical calculations necessary to match features and points between two datasets, and how to reliably calculate the metrics needed for the most important and assessable of van Oort's criteria.
 The results of this study revealed that the positional accuracy of the OSM Munich data at the time was about 4 meters
@@
 
-\textbf{(Müller et al, 2006)} present how procedural model generaiton is accomplished using Shape Grammar: a method developed by \textbf{(Stiny, 1975)} for automatically and iteratively constructing 3D models based in a hierarchical set of rules.
+\textbf{(Müller et al, 2006)} present how procedural model generation is accomplished using Shape Grammar: a method developed by \textbf{(Stiny, 1975)} for automatically and iteratively constructing 3D models based in a hierarchical set of rules.
 Shape grammar was originally developed from a method known as L-systems: a system for hierarchical model description which was originally used to generate plants for 3D animation \textbf{(Parish and Müller, 2001)}.
 Shape grammar works by starting with a crude volumetric model of a building and iteratively remodeling it from a set of shape grammars.
-Shape grammars often work from a databank of 3D models and a set of hierarchical rules for how these can be pieced together.
+Shape grammars often work from a data bank of 3D models and a set of hierarchical rules for how these can be pieced together.
 This creates a possible generative space of many different models that can be created from the same data and rule set, and with the induction of random fluctuations, many different models can be generated from the same dataset and rules.
 Using this procedure, a procedural model can structure the façade of a building, add texture and add minor detail models such as windows, doors and ornaments, and so on.
-The maximum accuracy that can be obtained in these finer details is largely dependent on the availablity of such data, and as most commercially available GIS data only serve crude building footprints and road placement, the model will often have to make guesses about the finer detail, which sacrifices accuracy to the real world.
+The maximum accuracy that can be obtained in these finer details is largely dependent on the availability of such data, and as most commercially available GIS data only serve crude building footprints and road placement, the model will often have to make guesses about the finer detail, which sacrifices accuracy to the real world.
 
@@
 A common theme in space syntax is the division of spacial structures into different levels of observation, and then identifying different units of space and the pattern in which they connect to each other.
-Space syntax is an analytical method that relies on the availablity of city maps, which has since its invention been developed into an extensive research fields on its own
+Space syntax is an analytical method that relies on the availability of city maps, which has since its invention been developed into an extensive research fields on its own
 \textbf{(Parish and Müller, 2001)}.
-It has also been widely used in the analysis of pedestrian flows and wayfinding
+It has also been widely used in the analysis of pedestrian flows and way finding
 \textbf{(Peponis, Zimring and Choi, 1990)}.
@@
 \textbf{(Rutzinger, Rottensteiner and Pfeifer, 2009)}
-found that if $S_{RO}(foot_{A}, foot_{B}) < 30\%$ for two buildings $A$ and $B$ from different sets, then $A$ and $B$ are highly likely to be separate, neighbouring buildings and not in fact identical.
+found that if $S_{RO}(foot_{A}, foot_{B}) < 30\%$ for two buildings $A$ and $B$ from different sets, then $A$ and $B$ are highly likely to be separate, neighboring buildings and not in fact identical.
 
@@
 The two buildings might have been modeled at different levels of detail, or vertex clusters may be found at different parts of the polygon in the two datasets.
-To adress this problwm, the OSM data precision study by \textbf{(Fan et al, 2014)} used the Douglas-Peucker algorithm, first defined by \textbf{(Douglas and Peucker, 1973)}, to reduce both building polygons to a simliar vertex count, and then proceed with locating the most likely matching vertices in the polygons.
+To address this problem, the OSM data precision study by \textbf{(Fan et al, 2014)} used the Douglas-Peucker algorithm, first defined by \textbf{(Douglas and Peucker, 1973)}, to reduce both building polygons to a similar vertex count, and then proceed with locating the most likely matching vertices in the polygons.
 The full procedure for matching points will be presented in section 3.2.3 of this report.
@@
 The shape accuracy between two matching buildings was quantified by using a similarity function that depends on the difference between the turning functions of matching building footprints in both datasets.
-The turning function was first defined by \textbf{(Arkin et al, 1991)}, as a method for measuring the similarity of two polygons, and measures the cumulative angle of a single polygon's counter-clockwise tangent, as a function of the cumulative normalized length when travelling along the edge of the polygon.
-The turning function effectively steps over the tangent of a polygon continously for a full cycle, and logs the total turning angle by adding together the angle of each turn.
-It includes consideration of wether the individual turns are clockwise or counter-clockwise, and is invariant to rotation and scaling of the polygon.
+The turning function was first defined by \textbf{(Arkin et al, 1991)}, as a method for measuring the similarity of two polygons, and measures the cumulative angle of a single polygon's counter-clockwise tangent, as a function of the cumulative normalized length when traveling along the edge of the polygon.
+The turning function effectively steps over the tangent of a polygon continuously for a full cycle, and logs the total turning angle by adding together the angle of each turn.
+It includes consideration of whether the individual turns are clockwise or counter-clockwise, and is invariant to rotation and scaling of the polygon.
 Once the turning functions of two similar polygons have been obtained, the integral between them is taken as a measure of shape accuracy.
@@
 
-Another crucial concept in city modelling and GIS mapping is the Level Of Detail (LOD) aspect.
+Another crucial concept in city modeling and GIS mapping is the Level Of Detail (LOD) aspect.
 The LOD defines a number of possible different representation of the same GIS data, which vary in the amount of detail in its presentation.
 A modeler can choose to present the same data (e.g. building, roads, or decorative models) at different detail levels, depending on the use and application.
-Furthermore, the LOD chosen for a particulary application also suggests how the data has been aqquired and modelled.
-It turns out however, in practise, that the terminology and use of LOD is still ambigous and the definitions vary greatly between practitioners, standards and institutions, despite its usefulness.
+Furthermore, the LOD chosen for a particularly application also suggests how the data has been acquired and modeled.
+It turns out however, in practice, that the terminology and use of LOD is still ambiguous and the definitions vary greatly between practitioners, standards and institutions, despite its usefulness.
 \textbf{(Biljecki et al, 2014)}
 
-\textbf{(Biljecki, Ledoux and Stoter, 2016 (1))} provide the GityGML definition of LOD.
+\textbf{(Biljecki, Ledoux and Stoter, 2016 (1))} provide the CityGML definition of LOD.
 The CityGML standard divides the LOD representation of 3D models into three classes: A B and C.
@@
     \item LOD-A models are renderings of the building footprint itself, or rectangular building blocks that have been extruded from the footprint to give the building height.
-    \item LOD-B models are building blocks with simplifed roof shapes.
+    \item LOD-B models are building blocks with simplified roof shapes.
     \item LOD-C models are LOD-B models which include finer detail features such as doors, windows, alcoves, building decor, indoor features or vegetation.
 \end{enumerate}
-A modeler may chose which of these LOD class(es) to use depending on data availablity and application.
+A modeler may chose which of these LOD class(es) to use depending on data availability and application.
 
@@
 It will present how to qualify dataset completeness and building correspondence (how many of the buildings in one dataset that can be found in the other) by collecting the cumulative building areas and looking at what buildings have overlapping areas between the two datasets.
-The definitions and algorithms neccessary for calculating posiitonal accuracy and a measure for shape accuracy will also be presented.
+The definitions and algorithms necessary for calculating positional accuracy and a measure for shape accuracy will also be presented.
 Section 3.3 will present at the data collection process for the overlap study.
-It will present how standard road withs were obtained, as well as the algorithm for detemining feature collision that was used.
+It will present how standard road widths were obtained, as well as the algorithm for determining feature collision that was used.
 Section 3.4 will discuss briefly the data sources used in this project, why they were chosen and specific issues with the data collection process.
@@
 
-Programmatic collection of statistics has been neccessary for both the primary and secondary research questions, and this section presents a broad overview of the technical details behind the map analysis and data collection processes for both the geodata precision study and the overlap study.
+Programmatic collection of statistics has been necessary for both the primary and secondary research questions, and this section presents a broad overview of the technical details behind the map analysis and data collection processes for both the geodata precision study and the overlap study.
 
@@
 
-Additionally, a number of auxillary calculations were needed to estimate the error boundary of the positional accuracy.
+Additionally, a number of auxiliary calculations were needed to estimate the error boundary of the positional accuracy.
 The completeness of the OSM dataset (a measure of how many features are represented in both the OSM dataset and the reference dataset) was calculated as the relative difference in area of all building footprints, between the two datasets.
@@
 Once the positional error estimate was obtained from the precision study, it was used to assess the integrity of the OSM Stockholm road network.
-This was done by firstly locating critical points where assigning standard widths to the roads would cause collision with other features, and then evaluating how many of those points could be eliminated by making ajustments within the found margin of error.
+This was done by firstly locating critical points where assigning standard widths to the roads would cause collision with other features, and then evaluating how many of those points could be eliminated by making adjustments within the found margin of error.
 For the purpose of the overlap study, each road has been assigned a standard width that is in line with Swedish regulations, since in the OSM dataset, roads are represented as simple polylines, lacking width.
@@
     \caption{Examples of a 1:1 building match (left), where the sought-after building was found both in the OSM (red) and SLU (blue) datasets, and of a 1:n building match (right), where the sought-after building was represented as one polygon in the OSM (red) dataset, but divided into many smaller lots in the SLU (multicolored) dataset.
-The buildings are examples from the OSM and SLU datasets that were prcessed in this projects, and were taken from the Stockholm area.}
+The buildings are examples from the OSM and SLU datasets that were processed in this projects, and were taken from the Stockholm area.}
     \label{fig:building-match-types}
@@
 Fix a starting vertex $v_1$.
-The tangent angle at $v_1$ is $\theta_{n,1}$. This is the angle between the neighbouring line segments $e_n$ and $e_1$.
+The tangent angle at $v_1$ is $\theta_{n,1}$. This is the angle between the neighboring line segments $e_n$ and $e_1$.
 For any $i$ such that $i>1$ and $n<i$, the tangent angle at $v_i$ is recursively defined as:
@@
 Possible problems are that the vertex counts could be dissimilar, or that vertex clusters may be found at different parts of the two polygons.
-To avoid this effect, key points are extracted using the Douglas-Peucker algorithm \textbf{(Douglas and Peucker, 1973)}, to create a simplifed footprint with less points that still retain information about the rough features of the detailed footprint.
+To avoid this effect, key points are extracted using the Douglas-Peucker algorithm \textbf{(Douglas and Peucker, 1973)}, to create a simplified footprint with less points that still retain information about the rough features of the detailed footprint.
 The idea behind Douglas-Peucker is to recursively divide a polyline.
@@
 See Algorithm 1 for a detailed view.
-Following this, the Oriented Minimum Bounding Rectangles (OMBR) are calculated for both the simplifed polygons.
+Following this, the Oriented Minimum Bounding Rectangles (OMBR) are calculated for both the simplified polygons.
 Finally the OMBR for the OSM building footprint is shifted so that its centroid aligns with the centroid of the OMBR of the SLU building footprint.
-Any edges in the simplified footprints that coinside with the OMBR from the same dataset are extracted, and the corresponding points in the original footprints will be matched with each other.
+Any edges in the simplified footprints that coincide with the OMBR from the same dataset are extracted, and the corresponding points in the original footprints will be matched with each other.
 See figure \ref{fig:vertex-match-algo} for an illustration of the algorithm that detects matching vertices between two similar polygons.
@@
 Starting from two highly similar polygons, the first step is to obtain both the Douglas-Peucker reduced polygons.
-The next step is to obtain the OMBRs of the reduced polygons, and for both polygons find the edges that coinside with the edges of the OMBRs.
+The next step is to obtain the OMBRs of the reduced polygons, and for both polygons find the edges that coincide with the edges of the OMBRs.
 These edges are corresponding, and thus their end vertices form a matching vertex map between the original polygons.}
@@
 
-Finally, a few special cases where features or feature collisions should not be registered were neccessary to implement:
+Finally, a few special cases where features or feature collisions should not be registered were necessary to implement:
 \begin{enumerate}
     \item If two roads intersect even though they share a node, it is a natural connection in the road network and should not be counted as a problematic case of feature overlap.
-    \item Collisions between tunnels or bridges and other roads were ignored since these are of different heights. Brdiges and tunnels were identified by looking at feature attributes in OSM.
-    \item Any roads that were represented as polygons or multipolygons instead of ploylines were also excluded from the study. These are typically town squares or areas designated for walking.
+    \item Collisions between tunnels or bridges and other roads were ignored since these are of different heights. Bridges and tunnels were identified by looking at feature attributes in OSM.
+    \item Any roads that were represented as polygons or multipolygons instead of polylines were also excluded from the study. These are typically town squares or areas designated for walking.
 \end{enumerate}
@@
 
-The Projected Line Segment Distance (PLSD) algorithm is an extented version of the Line Segment Distance (LSD) algorithm.
+The Projected Line Segment Distance (PLSD) algorithm is an extended version of the Line Segment Distance (LSD) algorithm.
 Whereas LSD simply measures the distance between the two absolute closest points on two line segments, PLSD will yield the distance of the shortest vector $v$ between line segments $e_a$ and $e_b$, with the condition that $v$ is perpendicular to the tangent of $e_a$.
@@
 \SetAlgoLined
-\KwResult{A float representing the shortests distance between $A$ and $B$}
+\KwResult{A float representing the shortest distance between $A$ and $B$}
     \textbf{Input:} Line segment $A$: [$\vec{a_1}$, $\vec{a_2}$], Line segment $B$: [$\vec{b_1}$, $\vec{b_2}$]\\
@@
 The reference map is provided by the Swedish University of Agricultural Sciences (SLU), and will be referred to in this report as the SLU map or the SLU dataset.
-The SLU map conatins only building and property footprints.
+The SLU map contains only building and property footprints.
 The exact map segments that were extracted were in both cases rectangular boundaries with the following coordinates in WGS84: (N: 59.42, E: 18.15, S: 59.23, W: 17.79).
-Both datasets were collcted on March 18th, 2020.
+Both datasets were collected on March 18th, 2020.
 
@@
 Such images have a limit on their resolution which puts a theoretical limit on the accuracy of map features compared to their real-world equivalents
-The driving force behind the availablity of VGI is, according to \textbf{(Goetz and Zipf, 2012)}, the availablity of high resolution satellite imagery to private individuals.
-They claim that particularily the imagery from Bing Maps, which launched in 2010, led to an increase in building information in OSM, and the positional error is largely due to the limited resolution in such satellite imagery.
+The driving force behind the availability of VGI is, according to \textbf{(Goetz and Zipf, 2012)}, the availability of high resolution satellite imagery to private individuals.
+They claim that particularly the imagery from Bing Maps, which launched in 2010, led to an increase in building information in OSM, and the positional error is largely due to the limited resolution in such satellite imagery.
 
@@
 The SLU map contains building features and property limits, but no road data.
-The map is updated continously by Lantmäteriet, in conjoncture with the forming or reforming of property.
+The map is updated continuously by Lantmäteriet, in conjuncture with the forming or reforming of property.
 Whereas OSM is a free service, the SLU map is available to paying customers or for free to students and researchers.
 
-The history aspect of the SLU dataset is not freely available online, but the acquistion method is included in the file metadata on a per-object basis.
-Lantmäteriet uses internal codes to specifify the acquistion method, and these codes can be referred to in the product description the accompanies the map files.
+The history aspect of the SLU dataset is not freely available online, but the acquisition method is included in the file metadata on a per-object basis.
+Lantmäteriet uses internal codes to specify the acquisition method, and these codes can be referred to in the product description the accompanies the map files.
 As the scope of this project is limited to making a broad comparison of the positional accuracy of the two datasets, only the required positional accuracy as described by the SLU product description will be used.
@@
 https://wiki.openstreetmap.org/wiki/Converting\_to\_WGS84}
-The coordinate conversions for this paper were obtained using proj: a Linux commandline application for geospatial coordinate conversion.
+The coordinate conversions for this paper were obtained using proj: a Linux command line application for geospatial coordinate conversion.
 
@@
 Any building area outside of the query domain will be excluded.
-Since the principles behind what features are delivered and how differ between the datasets, it is neccessary to crop both datasets to ensure that all buildings are complete and have matching candidates in the other set.
-Any features that intersect the edge of the rectagular boundary, in both datasets, were therefore excluded from the study.
+Since the principles behind what features are delivered and how differ between the datasets, it is necessary to crop both datasets to ensure that all buildings are complete and have matching candidates in the other set.
+Any features that intersect the edge of the rectangular boundary, in both datasets, were therefore excluded from the study.
 
@@
 Following this, an algorithm will find the exact same metrics with regards to how many road features that can in theory be corrected by simple geometrical translation of their vertices.
-This will also be preseneted in terms of the number of road features, the number of edges and the total length in meters of remaining colliding road.
+This will also be presented in terms of the number of road features, the number of edges and the total length in meters of remaining colliding road.
 
 The secondary study about geodata precision will primarily serve to obtain an estimate of the positional accuracy of the OSM dataset, in meters.
-A number of auxillary metrics will be needed to estimate the domain of error on the positional accuracy, namely the feature completeness and shape accuracy of the OSM dataset.
-The completeness of the OSM dataset will be taken as a the number of features that are represented in both datasets, as a perentage of the total numer of features in the reference dataset.
+A number of auxiliary metrics will be needed to estimate the domain of error on the positional accuracy, namely the feature completeness and shape accuracy of the OSM dataset.
+The completeness of the OSM dataset will be taken as a the number of features that are represented in both datasets, as a percentage of the total number of features in the reference dataset.
 Shape accuracy (the measure of polygon similarity) was obtained by comparing polygon turning function between matching features, feature by feature.
@@
   \item Positional accuracy. This relates the coordinate value of an object in the database to the actual location of the ground in the real world.
-  \item Attibute accuracy. In a geographical database, objects are commonly tagged with meta-information. This class assesses how correct those values are.
+  \item Attribute accuracy. In a geographical database, objects are commonly tagged with meta-information. This class assesses how correct those values are.
   \item Logical consistency - This assesses the internal consistency of the dataset.
@@
 Table \ref{table:num-of-buildings} shows the building coverage of the OSM and SLU datasets in terms of the number of buildings and their total area.
-Table \ref{table:num-of-buildings-percent} shows the building coverage of both datasets as perentages.
+Table \ref{table:num-of-buildings-percent} shows the building coverage of both datasets as percentages.
 The area cover of the OSM dataset in relation to the SLU dataset was estimated to 98.18\%. This will be taken as an estimate of the completeness of the OSM dataset.
@@
 
-Now that the colliison metrics have been obtained, conclusions may be drawn about the technical feasibility of 3D city reproduction from OSM data.
+Now that the collision metrics have been obtained, conclusions may be drawn about the technical feasibility of 3D city reproduction from OSM data.
 First the geodata precision study will be discussed.
@@
 The chapter will conclude with a presentation of geometrical algorithms for how to correct each case automatically.
-For conclusions about the usage of OSM city reproduction in engineering practise and remaining challenges, see section 5.
+For conclusions about the usage of OSM city reproduction in engineering practice and remaining challenges, see section 5.
 
@@
 In the particular case of Stockholm, the small difference in area cover that can be seen could partially be explained by the large number of small utility buildings that are often not represented in the OSM dataset.
-Particularily the residential areas in downtown, Stockholm, are characteristic in that each city block consists of a street front and an open, communal compound in the middle, inside of which there are usually a number of small utility buildings such as bicycle garages, toolsheds, refuse rooms and such.
-See figure \ref{fig:osm-slu-map-utility-buildings} for an example of an area in stockholm with such characteristics.
+Particularly the residential areas in downtown, Stockholm, are characteristic in that each city block consists of a street front and an open, communal compound in the middle, inside of which there are usually a number of small utility buildings such as bicycle garages, tool sheds, refuse rooms and such.
+See figure \ref{fig:osm-slu-map-utility-buildings} for an example of an area in Stockholm with such characteristics.
 These buildings are likely hard to make out and trace on publicly available satellite image data, most likely due to occlusion by their surroundings due to shadows or forestry.
@@
 This is problematic since this typical area with the same level of accessibility was likely measured with the same measuring technique, and the unsystematic positional error will likely be similar in these areas.
-The question is wether it comforms well to the positional error in the rest of the dataset or not, though this analysis will not be carried out in this project, but could be a good subject for a future master's thesis project.
+The question is whether it conforms well to the positional error in the rest of the dataset or not, though this analysis will not be carried out in this project, but could be a good subject for a future master's thesis project.
 
@@
 % With internet
-% TODO: Restructure research subquestions according to Chris's suggestions (see meeting log)
 % TODO: Fix result table margins
-% TODO: Spell check
 
@@
 The immediate conclusion of this study is that, in terms of feature overlap, footpaths were not significantly worse than other road categories.
-As a result, it can promptly be assumed that city modelling services which already utilize OSM car road data could also use OSM footpath data without a drastic increase in feature overlap.
+As a result, it can promptly be assumed that city modeling services which already utilize OSM car road data could also use OSM footpath data without a drastic increase in feature overlap.
 The biggest reason for this is thought to be that footpaths in their nature are narrow, and so narrow in fact that they can be displaced by more than half of their full width (minimum 1 meter) before the limit of half the positional accuracy (1.025 meters) is reached.
@@
 This means that for each colliding road, on average 2.3 of its edges will be intersecting some other feature, thus bearing responsibility for the collision.
-In total there are 26473 feautures and 6382579 edges, meaning that each road feature is on average contructed out of 241 edges.
+In total there are 26473 features and 6382579 edges, meaning that each road feature is on average constructed out of 241 edges.
 Thus the percentage of edges colliding per road is on average roughly 0.95\%.
@@
 
-After visual inspection, it was determined that footpath all collision occurences can be almost exclusively classified into one of five categories.
+After visual inspection, it was determined that footpath all collision occurrences can be almost exclusively classified into one of five categories.
 These categories are explained in the list below, and each case is exemplified in figure \ref{fig:collision-cases}.
@@
     \caption{Example of the solution algorithm for case 3 collision.
-The portruding building feature as seen in \ref{fig:collision-cases} has been eliminated.
+The protruding building feature as seen in \ref{fig:collision-cases} has been eliminated.
 The yellow line shows the original road placement.
@@
 This report has now established a firm idea of the technical feasibility of generating virtual cities from OpenStreetMap-data, including different types of road networks.
-It remains to touch upon the possibilities of road network modelling in the broader context of 3D city reproduction. What are the remaining challenges for accurate 3D city reproduction?
+It remains to touch upon the possibilities of road network modeling in the broader context of 3D city reproduction. What are the remaining challenges for accurate 3D city reproduction?
 What techniques can be used to complement the basic system of importing street and building data that is already in place, and create a rich and living virtual city?
-In this section everything we have learnt so far will be placed in the broader context of GIS and city modeling, focusing on the research fields of 3D media and city planning.
+In this section everything we have learned so far will be placed in the broader context of GIS and city modeling, focusing on the research fields of 3D media and city planning.
 
@@
 As has already been established in the analysis of the footpath overlap study, primary conclusion of the footpath overlap study is that, in terms of feature overlap, footpaths were not significantly worse than other road categories.
-Thus it can very well be assumed that city modelling software could use the same routine for importing OSM car roads as well as OSM footpaths, without a drastic increase in feature overlap.
-If it is desired to obtain a map will signifcantly less overlap than what OSM can provide due to limitations on positional accuracy, the designer could build upon their existing procedure and add algorithms for overlap detection and correction.
+Thus it can very well be assumed that city modeling software could use the same routine for importing OSM car roads as well as OSM footpaths, without a drastic increase in feature overlap.
+If it is desired to obtain a map will significantly less overlap than what OSM can provide due to limitations on positional accuracy, the designer could build upon their existing procedure and add algorithms for overlap detection and correction.
 Such a procedure could for example utilize the method for overlap detection that was presented in section 3.3.2 of this report, as well as some of the methods for overlap correction that were suggested in section 4.2.3.
@@
 
-As has been mentioned in the introduction to this report, procedural generaion of road networks is a basal requirement for quick and low-effort modeling virtual cities.
-Since the road network ties together the city, and forms the grid that divides the lots, quarters and neighbourhoods, it make sense that one can only continue with the remaining challenges once there exists a robust procedure for extracting road geodata and generating the 3D represention of it without compromising the data integrity.
-The same is true for building modelling.
+As has been mentioned in the introduction to this report, procedural generation of road networks is a basal requirement for quick and low-effort modeling virtual cities.
+Since the road network ties together the city, and forms the grid that divides the lots, quarters and neighborhoods, it make sense that one can only continue with the remaining challenges once there exists a robust procedure for extracting road geodata and generating the 3D representation of it without compromising the data integrity.
+The same is true for building modeling.
 As \textbf{(Fan et al, 2014)} notes, it turns out that the same methods that have been used in this project to obtain and qualify road network data can be applied to building footprint data from OSM. 
-Only when there exists such a fundamental procedure for accuratly importing building footprints without compromising the data integritey, one can proceed with the remaining challenges in building modeling.
-
-Such challenges, which have not been mentioned, include for example infering or guessing building height, since OSM does not provide data about building elevation.
-Occasinally buildings in OSM may be tagged, but the norm is that no elevation data is available.
+Only when there exists such a fundamental procedure for accurately importing building footprints without compromising the data integrity, one can proceed with the remaining challenges in building modeling.
+
+Such challenges, which have not been mentioned, include for example inferring or guessing building height, since OSM does not provide data about building elevation.
+Occasionally buildings in OSM may be tagged, but the norm is that no elevation data is available.
 There are however very sophisticated methods from guessing the building height after looking at reference datasets.
 \textbf{(Biljecki, Ledoux and Stoter, 2017)} presented a method for inferring building height exclusively from the building footprint and surrounding footprints.
-Surprisingly, their model was shown to be accurace within a meter on average.
+Surprisingly, their model was shown to be accuracy within a meter on average.
 
@@
 Even if high resolution terrain elevation data is available from other sources, it is not provided by OSM, and will have to be obtained from a complementary data source.
-Method exist for how to make a city model comform to an underlying terrain, however.
+Method exist for how to make a city model conform to an underlying terrain, however.
 \textbf{(Galin et al, 2010)} presented such a geometrical model for placing a 3D road asset onto a non-flat terrain surface, in the context of procedurally generated roads.
@@
 Recall the concept of Level of Detail, as presented by \textbf{(Biljecki, Ledoux and Stoter, 2016 (1))} in section 2.3.5 of this report.
-It turns out that, with the current means of data aquisition from OSM being limited to building footprints, using OSM data a model can only produce LOB-B models at best.
-If a modeler wishes to aqquire building assets at a LOD-C level, they would have to complete the data with other sources, or rely on procedural methods to populate crude city models with additional details and building features that give the models more aestethic depth, but do not neccessarily reflect reality.
+It turns out that, with the current means of data acquisition from OSM being limited to building footprints, using OSM data a model can only produce LOB-B models at best.
+If a modeler wishes to acquire building assets at a LOD-C level, they would have to complete the data with other sources, or rely on procedural methods to populate crude city models with additional details and building features that give the models more aestethic depth, but do not necessarily reflect reality.
 
@@
 
-For simple presentation in e.g. animation and videogames or even urban planning software, such procedurally generated cities could very well be enough.
+For simple presentation in e.g. animation and video games or even urban planning software, such procedurally generated cities could very well be enough.
 As said however, the big problem is that these features will rarely match real life.
-Such models will be sufficient as long as one can be satisified with looking at rough representations of buildings and does not require additional knowledge about minor features.
-However, many studies in urban development have been done before that requred detailed knowledge about a city's geometry, e.g. for the purposes of modeling heat development
+Such models will be sufficient as long as one can be satisfied with looking at rough representations of buildings and does not require additional knowledge about minor features.
+However, many studies in urban development have been done before that required detailed knowledge about a city's geometry, e.g. for the purposes of modeling heat development
 \textbf{(Nakata-Osaki, Souza and Rodrigues, 2018)}
@@
 
-By looking at such studies in urban development, one will discover that a lot of research in this field requires detailed morpological information that goes beyond simply knowing the geometry of the city.
-Urban architects often work with morpological structures where a single feature, such as a city neighbourhood, can be examined from many different levels of perspective.
-A neighbourhood can be analysed from the level of its city blocks, individual lots, properties or building limits, building facades or even indoor spaces.
-Simlarily a road network can be analysed by looking at it either at its connectivity by viewing it a a simple 2D graph, but it might as well be seen as a hierarchical structure where primary roads branch off into sub-streets.
-The very unfortunate fact is that much of that morpological information is not present in the OSM dataset.
+By looking at such studies in urban development, one will discover that a lot of research in this field requires detailed morphological information that goes beyond simply knowing the geometry of the city.
+Urban architects often work with morphological structures where a single feature, such as a city neighborhood, can be examined from many different levels of perspective.
+A neighborhood can be analyzed from the level of its city blocks, individual lots, properties or building limits, building facades or even indoor spaces.
+Similarly a road network can be analyzed by looking at it either at its connectivity by viewing it a a simple 2D graph, but it might as well be seen as a hierarchical structure where primary roads branch off into sub-streets.
+The very unfortunate fact is that much of that morphological information is not present in the OSM dataset.
 Furthermore, take the fact that even a single road can be comprised of many different sections with different purposes.
 A single inner-city street for example, often contains sidewalks and lanes reserved for taxis or public transit, all of a certain placement and dimension that determines what type of traffic it can serve.
-According to \textbf{(Stojanovski, 2019)}, this type of representation is known as a multimodal structures, due to the possibility of incorporating mulitple modes of transportation, but it may even describe subtilities like building facades and distances between curbline and building lot.
+According to \textbf{(Stojanovski, 2019)}, this type of representation is known as a multimodal structures, due to the possibility of incorporating multiple modes of transportation, but it may even describe subtilties like building facades and distances between curb line and building lot.
 OSM, while being an excellent source of city geometry, is simply not a sophisticated enough dataset to cater to all these possible areas of research.
-For media applications, such morphological and multimodal information could in theory be generated procedurally, by using shape grammars and some sort of wayfinding to create logical routes for different types of traffic.
+For media applications, such morphological and multimodal information could in theory be generated procedurally, by using shape grammars and some sort of way finding to create logical routes for different types of traffic.
 But for urban planners requiring an exact morphological representation of some specification, this complementary data will have to be obtained from another dataset.
@@
 
-The final conclusion is that generating virtual cities purely from OpenStreetMap data is a perfectly feasibile procedure in the media industry, with the inclusion of procedural methods such as shape grammars to generate neccessary detail.
-This will be a cheap method that reuires minimum effort in terms of source data collection, and the shape grammar method even allows for a great deal of scalability.
+The final conclusion is that generating virtual cities purely from OpenStreetMap data is a perfectly feasible procedure in the media industry, with the inclusion of procedural methods such as shape grammars to generate necessary detail.
+This will be a cheap method that requires minimum effort in terms of source data collection, and the shape grammar method even allows for a great deal of scalability.
 The level of detail of individual models and the composition as a whole can be fitted to the scope of the project and the team size.
 However, in the area of urban design, urban planners often require complementary data to build morphological representations of the data that they are working with.
-OSM does generally not serve this type of data, so the feasibility of creating virtual cities that are of use to urban planners will not depend upon the accuracy of the OSM dataset, but rather on the availablity of that complementary data elsewhere.
+OSM does generally not serve this type of data, so the feasibility of creating virtual cities that are of use to urban planners will not depend upon the accuracy of the OSM dataset, but rather on the availability of that complementary data elsewhere.
 

### 2020-06-24 report/report.tex
Import result table margins

@@
         Edges, total                & 6382579       & 6612383      & 968420      & 420810      \\
-        Edge length, total      & 129984.79 km  & 186089.55 km & 29304.57 km & 16162.95 km \\
+        Edge length, total (km)     & 129984.79     & 186089.55    & 29304.57    & 16162.95    \\
         Features, colliding         & 3189          & 1032         & 412         & 456         \\
         Edges, colliding            & 7330          & 2362         & 643         & 824         \\
-        Edge length, colliding  & 110.67 km     & 66.18 km     & 9389.30 m   & 20.41 km    \\
+        Edge length, colliding (km) & 110.67        & 66.18        & 0.94        & 20.41       \\
         Features, percent           & 12.05 \%      & 9.81 \%      & 13.53 \%    & 16.55 \%    \\
@@
 \end{figure}
-
-% With internet
-% TODO: Fix result table margins
 

### 2020-08-19 report/report.tex
Add spacing around the abstract section
Make it clear that the accuracy estimate methods are based on the Fan paper
Update information on DP

@@
 In this project, the accuracy was obtained by comparing coordinate points of OSM building footprints to those in the Swedish University of Agricultural Sciences (SLU) building dataset, whose accuracy is known.
-
 \end{abstract}
@@
 I detta projekt uppskattades felmarginalen genom att jämföra byggnaders koordinatpunkter i datamängen med de från en referenskarta från Svenska Lantbruksuniversitetet (SLU) med känd felmarginal.
-
 \end{abstract}
@@
 
-The geodata precision study has been carried out in a number of steps that are all detailed below.
+The methods used for the geodata precision study were largely based on those used by \textbf{(Fan et al, 2014)} in their study of the quality of the OSM Munich dataslice in 2014.
 First a feature map was found, which mapped individual features between the OSM dataset and a reference dataset.
@@
 
-The final problem is that even when building polygons have been matched between the two datasets, they may not have one-to-one vertex relationship, since footprints from different datasets may be formed at a different level of detail.
-Possible problems are that the vertex counts could be dissimilar, or that vertex clusters may be found at different parts of the two polygons.
-To avoid this effect, key points are extracted using the Douglas-Peucker algorithm \textbf{(Douglas and Peucker, 1973)}, to create a simplified footprint with less points that still retain information about the rough features of the detailed footprint.
+Once a feature map has been obtained, the next task will be to find a matching vertex set among the vertices of the corresponding building footprint polygons.
+Provided that the corresponding building footprints are true representations of the real life building, these pairs of matching verties must match the same part of the building.
+The most common case is that a vertex represents a corner in the footprint polygon, and in these cases it must hold that two matching vertices represent the same corner.
+The central problem is that even when building polygons have been matched between the two datasets, they may not have one-to-one vertex relationship, since footprints from different datasets may be formed at a different level of detail.
+Possible causes are that the vertex counts are dissimilar, or that vertex clusters may be found at different parts of the two polygons.
+Thus, simply comparing the closest vertices between the two building representations may not be valid, since those vertices may not represent the same part of the true building footprint.
+To avoid this effect, key points were extracted using the Douglas-Peucker algorithm \textbf{(Douglas and Peucker, 1973)}, to create a simplified footprint with less points that still retain information about the rough features of the detailed footprint.
+This will eliminate dense vertex clusters, as well as any vertices that lie in roughly straight lines, in favour of creating polygons with reduced vertex counts.
+A useful property of the Douglas-Peucker algorithm is that two highly similar polygons of different vertex counts will be reduced to resulting polygons with similar vertex counts.
+Thus, one can be much more certain that a closest vertex pair is indeed corresponding to the same feature of the building.
+
 The idea behind Douglas-Peucker is to recursively divide a polyline.
@@
 
-Finally, when two building footprints and their vertices have been mapped, the average offset of matching vertices will be used as a measure of positional error.
+Finally, when two building footprints and their vertices have been mapped, the average offset of the matching vertices will be used as a measure of positional error.
+The final estimate of the positional accuracy of the OSM dataset will be taken as the average of each building's positional error, over the entire dataset.
 The average and variance in vertex spread will be calculated and used to make an assessment of the geometric precision of the OSM dataset, and this precision will further be used to form an upper boundary on geometrical corrections that can be made in the overlap study.

### 2020-08-19 report/report.tex
Add info on shape accuracy

@@
 
-The Turning function $T_c(l)$ measures the cumulative angle of the polygon's counter-clockwise tangent, as a function of the cumulative normalized length $l$.
-This project uses the turning function as it is defined by \textbf{(Fan et al, 2014)}.
+This project will additionally calculation the shape accuracy measure, as it is defined by \textbf{(Fan et al, 2014)}, for building polygons over the entire OSM dataset.
+Their definition of shape accuracy uses the turning function $T_c(l)$, which measures the cumulative angle of the polygon's counter-clockwise tangent, as a function of the cumulative normalized length $l$.
 See figure \ref{fig:turning-function} for a side-by-side comparison of a polygon and its turning function.
@@
 \end{center}
-The turning function has some nice geometric properties, in that it is invariant to both rotation and scaling of the polygon.
-The function contains no information of the orientation of the polygon, only of the relative angle between successive line segments, thus it does not change under rotation.
-It also measures only the normalized cumulative length, which does not change under scaling.
+
 The similarity of two polygons A and B in terms of their turning function is defined as their distance of their cumulative turning functions:
@@
 See figure \ref{fig:turning-function-compare} for a side-by-side comparison of two similar polygons and their turning functions.
+
+The turning function has some nice geometric properties, in that it is invariant to any linear transformation of the polygon.
+The function contains no information of the orientation of the polygon, only of the relative angle between successive line segments, thus it does not change under rotation.
+It also measures only the normalized cumulative length, which does not change under scaling.
+Because of this, the shape accuracy measure will ignore any systematic, linear errors that are present in the entire dataset, and will only focus on the unsystematic errors in vertex displacement, that are due to limitations in measuring precision.
 

### 2020-08-19 report/report.tex
Add constraint about building taper
Add notes to urban planning feasibility discussion

@@
 Any self-connected nodes will also be eliminated from the dataset.
+    \item All building footprints are assumed to be coherent to their footprints along the z axis.
+Thus, it will be assumed that no buildings taper or extend beyond their footprints at a higher altitude.
 \end{enumerate}
@@
 
-For simple presentation in e.g. animation and video games or even urban planning software, such procedurally generated cities could very well be enough.
-As said however, the big problem is that these features will rarely match real life.
-Such models will be sufficient as long as one can be satisfied with looking at rough representations of buildings and does not require additional knowledge about minor features.
-However, many studies in urban development have been done before that required detailed knowledge about a city's geometry, e.g. for the purposes of modeling heat development
+Simple, procedurally generated cities, populated with generic 3D models, could very well be enough for simple presentation in e.g. animation and video games or even urban planning software, 
+In such a model however, features will only be rough representations, and rarely match their real life counterparts.
+Such models will be sufficient as long as one can be satisfied with looking at rough representations of major features, and do not require additional knowledge about minor features.
+It turns out however, that many studies in urban development have a highly case-by-case nature.
+Studies acroos many different areas have been done before that required very different knowledge about a city's geometry, e.g. for the purposes of modeling heat development
 \textbf{(Nakata-Osaki, Souza and Rodrigues, 2018)}
@@
 
-By looking at such studies in urban development, one will discover that a lot of research in this field requires detailed morphological information that goes beyond simply knowing the geometry of the city.
+According to \textbf{(Stojanovski, 2019)}, software that strives to aid in the urban development process is commonly not sophisticated enough to fully represent the needs and workspaces of urban planners in practise.
+Stojanovski goes on to claim that a lot of research in this field requires detailed information that goes beyond simply knowing the geometry of the city.
+Additionally, such research also typically requires additional logical data structures that a OSM cannot note serve.
 Urban architects often work with morphological structures where a single feature, such as a city neighborhood, can be examined from many different levels of perspective.
@@
 A single inner-city street for example, often contains sidewalks and lanes reserved for taxis or public transit, all of a certain placement and dimension that determines what type of traffic it can serve.
-According to \textbf{(Stojanovski, 2019)}, this type of representation is known as a multimodal structures, due to the possibility of incorporating multiple modes of transportation, but it may even describe subtilties like building facades and distances between curb line and building lot.
+This type of representation is known as a multimodal structures, due to the possibility of incorporating multiple modes of transportation, but it may even describe subtilties like building facades and distances between curb line and building lot.
 OSM, while being an excellent source of city geometry, is simply not a sophisticated enough dataset to cater to all these possible areas of research.
 For media applications, such morphological and multimodal information could in theory be generated procedurally, by using shape grammars and some sort of way finding to create logical routes for different types of traffic.
-But for urban planners requiring an exact morphological representation of some specification, this complementary data will have to be obtained from another dataset.
+For urban planners however, requiring an exact representation of some specification, this complementary data will have to be obtained from another dataset.
+It may also be necessary for a researcher to implement their own data structures on top of the existing city model, with will increase the complexity of the problem even further.
+The case-by-case nature of urban development projects also means that it is unlikely that a single piece of software can cater to the very differing needs of data and data structures for variuos projects.
 

### 2020-08-19 report/report.tex
Add note on why the positional accuracy was used as an upper bound for collision correction

@@
 After the width assignment, a program iterated over all features and calculated how many of them that show overlap with any other feature, as well as the total length of the road polylines that is intersecting some other feature.
-Finally, the program calculated how many of these colliding features that can be corrected by making adjustments that are within the positional accuracy of the OSM data, as obtained by the geodata precision study.
+Finally, the program calculated how many of these colliding features that can be corrected by making adjustments to the feature vertices, away from the collision point, that are within the positional accuracy of the OSM data, as obtained by the geodata precision study.
+The motivation behind using the positional accuracy as an upper bound for correction is to tread the positional accuracy as a maximum resolution of the map.
+Making translation withing the map resolution will preserve the logical consistency to OSM source data, and preserve aestetic characteristics of the map, both between features and in the map as a whole
 

### 2020-08-19 report/report.tex
Add note about where the road width measurements were obtained from

@@
 
-The standard road widths were obtained from a publication by the Swedish Transport Administration
+The standard road widths for the different types of car road were obtained from a publication by the Swedish Transport Administration
 \footnote{Vägars och gators utformning - Trafikverket.
@@
 , and were presented as requirements on the minimum lane width for certain types of roads and applications.
-These regulated minimum widths were used as a lower bound road with when examining the OSM dataset for potential overlap.
-The upper bound was found by visual measurements of a few key locations in the Stockholm area using the Bing maps service.
-Table \ref{table:road-widths} shows the minimum and maximum road widths that will be used in the overlap study.
+The minimum road widths were used as lower bounds when examining the OSM dataset for potential overlap.
+The estimate standard width for footpaths was obtained from a set of manual sidewalk and walkway width measurements, obtained from visual inspection of publicly available satellite imagery from Bing Maps.
+A rounded value was chosen that was weighted towards the maximum results of those measurements.
+Table \ref{table:road-widths} shows the road widths that were be used in the overlap study.
 
@@
                    & Footpath & Residential & Secondary & Primary \\
-        Minimum width & 2.0 m    & 6.0 m       & 6.5 m     & 8.0 m   \\
-        Maximum width & 5.0 m    & 9.0 m       & 10.0 m    & 16.0 m
+        Used width & 2.0 m    & 6.0 m       & 6.5 m     & 8.0 m   \\
     \end{tabular}

### 2020-08-24 report/report.tex
Add acknowledgement section to the report
Fix swedish spelling mistakes

@@
 Studien har undersökt hur vanligt det är att gångvägar överlappar med andra kartdetaljer såsom byggnader och andra vägar.
-Överlappning är en typ av fel i 3D-geometrin som uppstår på grund av osäkerhet i källdatan, vilket kan leda till besvärliga skärningar och artifakter i den färdiga stadsmodellen.
-I denna undersökning har en delmängd av OSM-datamängden över Stockholm tagits, och en geometrisk algoritm tagit alla möjliga vägpar och beräknat hur stor mängd av vägarna som skulle överlappa med andra kartdetaljer.
+Överlappning är en typ av fel i 3D-geometrin som uppstår på grund av osäkerhet i källdatan, vilket kan leda till besvärliga skärningar och artefakter i den färdiga stadsmodellen.
+I denna undersökning har en delmängd av OSM-datamängden över Stockholm använts, och en geometrisk algoritm har tagit alla möjliga vägpar och beräknat hur stor mängd av vägarna som skulle överlappa med andra kartdetaljer.
 Studien fann att gångvägar är mindre problematiska än bilvägar beträffande hur många vägar som kolliderar, och således borde inte gångvägar vara svårare att inkludera i en 3D-miljö än bilvägar.
@@
 
+\section*{Acknowledgements}
+The author would like to thank KTH associate professor Christopher Peters for supervising this master's thesis project, as well as KTH postdoctoral researcher Todor Stojanovski for expert consultation in the field of urban development.
+
 \begin{thebibliography}{9}

### 2020-08-24 report/build.sh
Add report build script

@@ -1,0 +1,1 @@
+pdflatex report.tex; pdftk cover_front.pdf report.pdf cover_rear.pdf output report_out.pdf; mv report_out.pdf report.pdf

### 2020-08-24 report/build.sh
Modify build script so there's an empty page between the front page and the title

@@ -1,1 +1,1 @@
-pdflatex report.tex; pdftk cover_front.pdf report.pdf cover_rear.pdf output report_out.pdf; mv report_out.pdf report.pdf
+pdflatex report.tex; pdftk cover_front.pdf empty_page.pdf report.pdf cover_rear.pdf output report_out.pdf; mv report_out.pdf report.pdf

### 2020-08-24 report/report.tex
Fix opponent's suggestions

@@
 \setcounter{tocdepth}{2}
+\interfootnotelinepenalty=10000
 
@@
 Traditionally, modeling virtual cities is a labor-intensive and time-consuming task due to the need for manual modeling, texturing and design and placement of individual models in the large 3D scene.
-This type of tasks is commonly done by professional 3D graphical artists and designers.
+This type of task is commonly done by professional 3D graphical artists and designers.
 However, in recent years there has been major research work in the field of procedural content generation, meaning automatic or semi automatic generation of 3D models without the direct involvement of a 3D artist.
@@
 In practice, urban environments are typically dense and large, spanning a few to hundreds of square kilometers, and are influenced by variables that are hard to quantify, including land policies, government plans, population changes and transportation infrastructure, making it non-trivial to produce 3D city representations with both high accuracy and high detail, while simultaneously having high utility in the intended field of study.
-\textbf{(Vanegas et al, 2010)}.
+\textbf{(Vanegas et al., 2010)}.
 
@@
 
-This project will strive to determine the feasibility of procedurally reconstructing a city road network from OSM data by using such models as have been mentioned in this section, focusing on the feasibility of including pedestrian footpaths.
+This project will strive to determine the feasibility of procedurally reconstructing a city road network from OSM data, focusing on the feasibility of including pedestrian footpaths.
 \textbf{Footpaths} include any form of path that is designated for pedestrians, including sidewalks, walkways and trails.
@@
 When it comes to the quality of available data, it turns out that in practice, geodata databases without errors are rare.
-Geodata validation and correction is only an emerging field of research, and many such datasets were usually obtained years ago, and creating error-free datasets was neither possible due to the lack of validation techniques, nor a high priority.
+Geodata validation and correction is only an emerging field of research, and many such datasets were obtained years ago, when creating error-free datasets was neither possible due to the lack of validation techniques, nor a high priority.
 Luckily there are coming more and more sophisticated methods for automatically correcting the most common errors
-\textbf{(Biljecki et al, 2016 (2))}.
+\textbf{(Biljecki et al., 2016 (2))}.
 
@@
 Different instruments, such as GPS measurements or satellite image stereoscoping, have theoretical upper limits on their precision, which propagate through the data collection and preservation process as a source of positional error
-\textbf{(Fan et al, 2014)}.
+\textbf{(Fan et al., 2014)}.
 Specifically in the case of rendering road networks and buildings, a common issue that stems from low positional accuracy is that the 3D models being generated from road and building data might overlap with each other, which can lead to strange-looking mesh intersections and inaccuracies that modelers often want to avoid.
@@
 Such services are however usually limited in the accuracy and type of data that they can provide.
-While they can provide both building footprints and road networks, they usually provide road data with no respect to overlap between features.
+While they can provide both building footprints (the polygon that defines the building's extents) and road networks, they usually provide road data with no respect to overlap between features.
 
@@
 This project has dealt with the feasibility of including footpath data when generating 3D representations of OSM data, and preserving the integrity of the 3D road network, avoiding overlap and collision between road meshes.
-It was  that most road overlap can be corrected by first detecting critical locations where feature overlap occurs, and subsequently adjust the footpath road widths and individual vertex positions within the margin of the OSM positional accuracy, in order to obtain a slightly altered map with little to no overlap between footpaths and other features.
 
@@
 
-The positional accuracy will be necessary to carry out the investigation of the first subquestion: the rate of collisions between roads, since it will determine the maximum tolerance in distance between road points before a collision is certain to have occurred.
+The positional accuracy will be necessary to carry out the investigation of the second subquestion: the rate of collisions between roads, since it will determine the maximum tolerance in distance between road points before a collision is certain to have occurred.
 For the secondary research question: the geodata precision study, the positional accuracy will be estimated by comparing feature points between the OSM building dataset and the Swedish University of Agricultural Sciences (SLU) building dataset with known accuracy.
@@
 \begin{enumerate}
+    \item The OSM data will be limited to a slice of the Stockholm. See section 3.4 for an exact description of the data that was used used.
     \item The program will not handle terrain features and altitude. It will be assumed that the road mesh can be projected on a flat 2d surface without feature intersection.
@@
 The entire OSM dataset is under constant modification by volunteers, so it is expected that the accuracy will improve over time.
-Previous studies highlight how the positional accuracy in similar-sized cities have improved over the course of several years (\textbf{Fan et al, 2014}, \textbf{Haklay, 2010}).
+Previous studies highlight how the positional accuracy in similar-sized cities have improved over the course of several years (\textbf{Fan et al., 2014}, \textbf{Haklay, 2010}).
 The historical aspect of the OSM dataset is available through the OSM History Viewer
@@
 
-Another aspect of geodata precision is attribute accuracy, which will neither be considered in this study.
-Attributes are small flags of metadata that are included with geodata features, letting the user know additional details about a feature that is not reflected in the map, e.g.
-how many lanes a road has, if the base is asphalt or gravel, or how many stories a building has.
-Attribute accuracy is the measure of how complete attribute information is in a dataset.
-How much attribute data is missing and how much of it does not conform with reality?
-In general, attribute accuracy is hard to quantify.
-In this case however, it might be interesting to compare attributes between the OSM map and the SLU reference map that was used in this project, as both use encoded metadata on a per feature basis, and matching features can be easily identified.
-See section 3.4.2 for an overview of the reference dataset that was used in this study.
-The attribute accuracy of the two datasets could in theory be assessed by doing a simple feature comparison and creating a translation table between the two attribute maps.
-To the knowledge of the author, this has never been done with specifically these two datasets, at the time of writing.
-
 \chapter{Background}
@@
 
-In their analysis of procedural city generation models, \textbf{(Müller et al, 2006)} provide an overview of the process of recreating virtual 3D cities from their real life counterparts.
+In their analysis of procedural city generation models, \textbf{(Müller et al., 2006)} provide an overview of the process of recreating virtual 3D cities from their real life counterparts.
 They present cities as complex collections of \textbf{buildings}, \textbf{blocks}, \textbf{parcels} and \textbf{neighborhoods} which are all interconnected via a \textbf{street} network, where individual streets can serve different types of traffic.
@@
 Such tools use procedural methods and AI solutions to move the burden of labor from the human designer to the software that the designer uses
-\textbf{(Müller et al, 2006)}.
+\textbf{(Müller et al., 2006)}.
 
@@
 Such software uses procedural models to render a city grid based on real-world locations, or generate new grids and features based on user-specified parameters, and allow for varying degrees of manual editing on grid, road, neighborhood, city block or individual feature level.
-Some of the most prominent examples are ArcGIS Pro (made by Esri), CityEngine (made by the startup Procedural, bought by ESRI) and Urban Canvas (made by the startup Synthicity, bought by Autodesk, now defunct) \textbf{(Stojanovski et al, 2019)}.
-
-Even as such tools are well available, \textbf{(Stojanovski et al, 2020)} note that they are as of today not commonly applied in urbanist practices, due to failures in reflecting the unique workspace of architects and their design needs.
+Some of the most prominent examples are ArcGIS Pro (made by Esri), CityEngine (made by the startup Procedural, bought by ESRI) and Urban Canvas (made by the startup Synthicity, bought by Autodesk, now defunct) \textbf{(Stojanovski et al., 2019)}.
+
+Even as such tools are well available, \textbf{(Stojanovski et al., 2020)} note that they are as of today not commonly applied in urbanist practices, due to failures in reflecting the unique workspace of architects and their design needs.
 Urban architecture is a field that not only accounts for spatial data, but also captures many social factors and the individual wills of the many stake holders.
@@
 The study applied different methods to assess the completeness of OSM data in two federal states in Germany, mainly by analyzing the area difference between the OSM dataset and an administrative dataset.
-Other notable studies about OSM data accuracy were made around the same time in Germany by \textbf{(Zielstra and Zipf, 2010)} and \textbf{(Neis et al, 2012)}, as well as in France by \textbf{(Girres and Touya, 2010)}.
-
-Finally, a major study in 2013-2014, conducted by researchers from various universities in Norway and Germany \textbf{(Fan et al, 2014)}, examined the quality and accuracy in building footprint data of the Munich area, at a time and place where the completeness of the OSM dataset was significantly higher than the study by \textbf{(Haklay, 2010)} (the results of their study showed that the OSM had captured close to 100\% of all building footprints).
+Other notable studies about OSM data accuracy were made around the same time in Germany by \textbf{(Zielstra and Zipf, 2010)} and \textbf{(Neis et al., 2012)}, as well as in France by \textbf{(Girres and Touya, 2010)}.
+
+Finally, a major study in 2013-2014, conducted by researchers from various universities in Norway and Germany \textbf{(Fan et al., 2014)}, examined the quality and accuracy in building footprint data of the Munich area, at a time and place where the completeness of the OSM dataset was significantly higher than the study by \textbf{(Haklay, 2010)} (the results of their study showed that the OSM had captured close to 100\% of all building footprints).
 This new study included an insight in the geometrical calculations necessary to match features and points between two datasets, and how to reliably calculate the metrics needed for the most important and assessable of van Oort's criteria.
@@
 
-\textbf{(Müller et al, 2006)} present how procedural model generation is accomplished using Shape Grammar: a method developed by \textbf{(Stiny, 1975)} for automatically and iteratively constructing 3D models based in a hierarchical set of rules.
+\textbf{(Müller et al., 2006)} present how procedural model generation is accomplished using Shape Grammar: a method developed by \textbf{(Stiny, 1975)} for automatically and iteratively constructing 3D models based in a hierarchical set of rules.
 Shape grammar was originally developed from a method known as L-systems: a system for hierarchical model description which was originally used to generate plants for 3D animation \textbf{(Parish and Müller, 2001)}.
@@
 One can for example analyze a city on an interior-by-interior basis, or use city blocks and streets as the least unit of space, depending on the goal of the study.
-\textbf{(Stojanovski et al, 2020)}.
+\textbf{(Stojanovski et al., 2020)}.
 
@@
 
-To assess the completeness of their OSM data, the study by \textbf{(Fan et al, 2014)} used the fraction of the total building area in the reference dataset and the OSM dataset.
+To assess the completeness of their OSM data, the study by \textbf{(Fan et al., 2014)} used the fraction of the total building area in the reference dataset and the OSM dataset.
 The motivation for this was to eliminate semantic error, which is a type of systematic error due to differences in how data is presented in both datasets.
@@
 
-\textbf{(Fan et al, 2014)} further discussed how the relative overlap in building area can be used as an estimate of the \textbf{building correspondence} between datasets.
+\textbf{(Fan et al., 2014)} further discussed how the relative overlap in building area can be used as an estimate of the \textbf{building correspondence} between datasets.
 Building correspondence is a term that refers to how well it can be judged which building footprints between two datasets that correspond to the same building in real life.
@@
 The two buildings might have been modeled at different levels of detail, or vertex clusters may be found at different parts of the polygon in the two datasets.
-To address this problem, the OSM data precision study by \textbf{(Fan et al, 2014)} used the Douglas-Peucker algorithm, first defined by \textbf{(Douglas and Peucker, 1973)}, to reduce both building polygons to a similar vertex count, and then proceed with locating the most likely matching vertices in the polygons.
+To address this problem, the OSM data precision study by \textbf{(Fan et al., 2014)} used the Douglas-Peucker algorithm, first defined by \textbf{(Douglas and Peucker, 1973)}, to reduce both building polygons to a similar vertex count, and then proceed with locating the most likely matching vertices in the polygons.
 The full procedure for matching points will be presented in section 3.2.3 of this report.
 
-The study by \textbf{(Fan et al, 2014)} also obtained a measure of the \textbf{shape accuracy} of building footprints.
+The study by \textbf{(Fan et al., 2014)} also obtained a measure of the \textbf{shape accuracy} of building footprints.
 While the positional accuracy accounts for the error between individual points in a polygon, the shape accuracy is a measure of how similar two polygons at a higher level of observation, regardless of relative scaling and rotation of the two polygons.
 The shape accuracy between two matching buildings was quantified by using a similarity function that depends on the difference between the turning functions of matching building footprints in both datasets.
-The turning function was first defined by \textbf{(Arkin et al, 1991)}, as a method for measuring the similarity of two polygons, and measures the cumulative angle of a single polygon's counter-clockwise tangent, as a function of the cumulative normalized length when traveling along the edge of the polygon.
+The turning function was first defined by \textbf{(Arkin et al., 1991)}, as a method for measuring the similarity of two polygons, and measures the cumulative angle of a single polygon's counter-clockwise tangent, as a function of the cumulative normalized length when traveling along the edge of the polygon.
 The turning function effectively steps over the tangent of a polygon continuously for a full cycle, and logs the total turning angle by adding together the angle of each turn.
@@
 It turns out however, in practice, that the terminology and use of LOD is still ambiguous and the definitions vary greatly between practitioners, standards and institutions, despite its usefulness.
-\textbf{(Biljecki et al, 2014)}
+\textbf{(Biljecki et al., 2014)}
 
@@
 
-The methods used for the geodata precision study were largely based on those used by \textbf{(Fan et al, 2014)} in their study of the quality of the OSM Munich dataslice in 2014.
-First a feature map was found, which mapped individual features between the OSM dataset and a reference dataset.
+The methods used for the geodata precision study were largely based on those used by \textbf{(Fan et al., 2014)} in their study of the quality of the OSM Munich dataslice in 2014.
+First a feature map was found, which mapped individual buildings between the OSM dataset and a reference dataset.
 A program then took matching building pairs and extracted matching geometrical points after first simplifying the polygons using the Douglas-Peucker algorithm by \textbf{(Douglas and Peucker, 1973)}.
@@
 
-Additionally, a number of auxiliary calculations were needed to estimate the error boundary of the positional accuracy.
-The completeness of the OSM dataset (a measure of how many features are represented in both the OSM dataset and the reference dataset) was calculated as the relative difference in area of all building footprints, between the two datasets.
-Additionally, the shape accuracy of the OSM dataset has also been calculated and graphed as an extra verification of the similarity between the datasets.
+Additionally, the completeness of the OSM dataset (a measure of how many features are represented in both the OSM dataset and the reference dataset) was calculated as the relative difference in area of all building footprints, between the two datasets.
+Finally, the shape accuracy of the OSM dataset has also been calculated and graphed as an extra verification of the similarity between the datasets.
 
@@
 
-In the same way as the study by \textbf{(Fan et al, 2014)}, this project will use the building area in the OSM dataset versus that of the SLU dataset to determine the completeness of the OSM building data in the Stockholm area, which will then be used as an argument for how accurate the building correspondence and point proximity measures can be.
+In the same way as the OSM quality assessment study by \textbf{(Fan et al., 2014)}, this project will use the building area in the OSM dataset versus that of the SLU dataset to determine the completeness of the OSM building data in the Stockholm area, which will then be used as an argument for how accurate the building correspondence and point proximity measures can be.
 Buildings were identified as 1:1 (one to one), 1:n (one to many) or 1:0 (one to none) cases, depending on if a building is fully represented in the other dataset, represented as an aggregation of smaller plots, or not found at all.
@@
 
-To evaluate feature correspondence, relative area overlap was used as a measure of which buildings from either dataset corresponded with each other.
-In this was a feature map could be generated where any building in either dataset points to one, many or no buildings in the other dataset.
+To evaluate feature correspondence, the relative area overlap was used as a measure of which buildings from either dataset corresponded with each other.
+In this way a feature map could be generated where any building in either dataset points to one, many or no buildings in the other dataset.
 \textbf{(Rutzinger, Rottensteiner and Pfeifer, 2009)} determined in their study of geodata comparison that if two building footprints share an area overlap of at least 30\%, then they can be safely assumed to point to each other in a matching set of buildings.
@@
 
-This project will additionally calculation the shape accuracy measure, as it is defined by \textbf{(Fan et al, 2014)}, for building polygons over the entire OSM dataset.
+This project will additionally calculation the shape accuracy measure, as it is defined by \textbf{(Fan et al., 2014)}, for building polygons over the entire OSM dataset.
 Their definition of shape accuracy uses the turning function $T_c(l)$, which measures the cumulative angle of the polygon's counter-clockwise tangent, as a function of the cumulative normalized length $l$.
@@
 Thus, simply comparing the closest vertices between the two building representations may not be valid, since those vertices may not represent the same part of the true building footprint.
-To avoid this effect, key points were extracted using the Douglas-Peucker algorithm \textbf{(Douglas and Peucker, 1973)}, to create a simplified footprint with less points that still retain information about the rough features of the detailed footprint.
+To avoid this effect, key points were extracted using the Douglas-Peucker algorithm, in order to create a simplified footprint with less points that still retain information about the rough features of the detailed footprint.
 This will eliminate dense vertex clusters, as well as any vertices that lie in roughly straight lines, in favour of creating polygons with reduced vertex counts.
@@
 
-For the primary overlap study.
 Table \ref{table:road-collision} shows the statistics of the overlap study, and the number of roads, number of edges and total edge length that showed intersections with other features, by road category.
-For footpaths in particular, it was shown that 12.05\% of all roads collide with some other feature after extrusion to the minimum width.
+For footpaths in particular, it was shown that 12.05\% of all roads collide with some other feature after being assigned with widths.
 The same was shown to be true for 0.11\% of the road polyline edges and 0.09\% of the total road length.
@@
 For footpaths this was shown to hold for 0.51\% of all roads, 
-The number of road edges and percentage of the total road length that could not be corrected by geometric translation was significantly small.
+The number of road edges and percentage of the total road length that could not be corrected by geometric translation was determined to be insignificantly small for any practical application.
 
@@
         & Total number of buildings & Area cover \\
-        OSM & 102755 & 33105023.32... $m^2$ \\
-        SLU & 170783 & 33718628.97... $m^2$
+        OSM & 102755 & 33.11 $km^2$ \\
+        SLU & 170783 & 33.72 $m^2$
     \end{tabular}
@@
 The building completeness analysis shows that the OSM Stockholm dataset has high completeness in terms of both building area and counted features.
-It even shows that the building completeness is around the same level as the study of Munich by \textbf{(Fan et al, 2014)}.
+It even shows that the building completeness is around the same level as the study of Munich by \textbf{(Fan et al., 2014)}.
 In the particular case of Stockholm, the small difference in area cover that can be seen could partially be explained by the large number of small utility buildings that are often not represented in the OSM dataset.
@@
 As a result, it can promptly be assumed that city modeling services which already utilize OSM car road data could also use OSM footpath data without a drastic increase in feature overlap.
-The biggest reason for this is thought to be that footpaths in their nature are narrow, and so narrow in fact that they can be displaced by more than half of their full width (minimum 1 meter) before the limit of half the positional accuracy (1.025 meters) is reached.
+The biggest reason for this was thought to be that footpaths in their nature are narrow, and so narrow in fact that they can be displaced by more than half of their full width (minimum 1 meter) before the limit of half the positional accuracy (1.025 meters) is reached.
 
@@
 
-After visual inspection, it was determined that footpath all collision occurrences can be almost exclusively classified into one of five categories.
+After visual inspection, it was determined that for footpaths all collision occurrences can be almost exclusively classified into one of five categories.
 These categories are explained in the list below, and each case is exemplified in figure \ref{fig:collision-cases}.
@@
 
-As has already been established in the analysis of the footpath overlap study, primary conclusion of the footpath overlap study is that, in terms of feature overlap, footpaths were not significantly worse than other road categories.
+As has already been established in the analysis of the footpath overlap study, the primary conclusion of the footpath overlap study is that, in terms of feature overlap, footpaths were not significantly worse than other road categories.
 Thus it can very well be assumed that city modeling software could use the same routine for importing OSM car roads as well as OSM footpaths, without a drastic increase in feature overlap.
-If it is desired to obtain a map will significantly less overlap than what OSM can provide due to limitations on positional accuracy, the designer could build upon their existing procedure and add algorithms for overlap detection and correction.
+If it is desired to obtain a map with significantly less overlap than what OSM can provide due to limitations on positional accuracy, the designer could build upon their existing procedure and add algorithms for overlap detection and correction.
 Such a procedure could for example utilize the method for overlap detection that was presented in section 3.3.2 of this report, as well as some of the methods for overlap correction that were suggested in section 4.2.3.
@@
 The same is true for building modeling.
-As \textbf{(Fan et al, 2014)} notes, it turns out that the same methods that have been used in this project to obtain and qualify road network data can be applied to building footprint data from OSM. 
+As \textbf{(Fan et al., 2014)} notes, it turns out that the same methods that have been used in this project to obtain and qualify road network data can be applied to building footprint data from OSM.
 Only when there exists such a fundamental procedure for accurately importing building footprints without compromising the data integrity, one can proceed with the remaining challenges in building modeling.
@@
 Even if high resolution terrain elevation data is available from other sources, it is not provided by OSM, and will have to be obtained from a complementary data source.
-Method exist for how to make a city model conform to an underlying terrain, however.
-\textbf{(Galin et al, 2010)} presented such a geometrical model for placing a 3D road asset onto a non-flat terrain surface, in the context of procedurally generated roads.
+Method for making a city model conform to an underlying terrain does exist, however.
+\textbf{(Galin et al., 2010)} presented such a geometrical model for placing a 3D road asset onto a non-flat terrain surface, in the context of procedurally generated roads.
 

### 2020-08-25 report/report.tex
English spell check

@@
 As a secondary study, the positional accuracy of the OSM dataset was calculated, since the positional accuracy determines the maximum tolerance in distance between points before it is certain that a road overlaps with another road or building.
-Any geodataset has a limit on its positional accuracy, which is determined by the theoretical precision of the measuring methods.
+Any geographic dataset has a limit on its positional accuracy, which is determined by the theoretical precision of the measuring methods.
 In this project, the accuracy was obtained by comparing coordinate points of OSM building footprints to those in the Swedish University of Agricultural Sciences (SLU) building dataset, whose accuracy is known.
@@
 Som en andra del av studien har felmarginalen för OSM-datamängden också tagits fram, eftersom felmarginalen bestämmer toleransen mellan två punkter som kan tillåtas innan man är säker på att en väg börjar överlappa med en annan väg eller en byggnad.
-Varje geodatamängd har en gräns på hur väl den stämmer överrens med verkligheten, som bestäms av vilka mätmedoter som har använts i datans insamling.
+Varje geografisk datamängd har en gräns på hur väl den stämmer överrens med verkligheten, som bestäms av vilka mätmetoder som har använts i datans insamling.
 I detta projekt uppskattades felmarginalen genom att jämföra byggnaders koordinatpunkter i datamängen med de från en referenskarta från Svenska Lantbruksuniversitetet (SLU) med känd felmarginal.
@@
 \textbf{Procedural models} typically involve a machine which algorithmically composes together or alters human-generated 3D assets.
-Sometimes a degree of computer-generated randomness is induced into the generative algoritm in order to create a larger and less predictable generative space of 3D models that can be created from the same algorithm.
+Sometimes a degree of computer-generated randomness is induced into the generative algorithm in order to create a larger and less predictable generative space of 3D models that can be created from the same algorithm.
 
@@
 To limit visual quirks such as mesh self intersection, it is often in the designer's best interest to have a stable procedure for importing road data that results in minimal feature overlap, while still retaining accuracy to the real world.
-A possible way of mitigating this problem is looking towards gathering geodata from volunteer-based geoservices such as OSM.
-The OSM project is a freely available wiki-like geodatabase based on Volunteered Geographical Information (VGI), meaning anyone can contribute by adding or editing features.
+A possible way of mitigating this problem is looking towards gathering geodata from volunteer-based map services such as OSM.
+The OSM project is a freely available wiki-like geographic database based on Volunteered Geographical Information (VGI), meaning anyone can contribute by adding or editing features.
 The OSM world dataset is being consistently updated by volunteers from all over the world, and the main idea is that by using OSM geodata for city reproduction, the temporal aspect of data accuracy would in theory be more assured by the crowd contribution nature.
@@
 Again, note that feature overlap is defined here as geometry errors due to inaccuracies in the source data, which results in 3D models overlapping, which can in turn lead to strange-looking mesh intersections.
-To investigate the feasibility of generating a 3D virtual city with minimal footpath overlap, the main research question will specifically be divided into two subquestions. The primary subquestion is:
+To investigate the feasibility of generating a 3D virtual city with minimal footpath overlap, the main research question will specifically be divided into two sub-questions.
+The primary sub-question is:
 
@@
 This question aims to find the rate of collision between road features in OSM, and investigate the nature of the most common types of collisions and how they can be corrected.
-For the road overlap study, a program has taken a slice of the OSM dataset in the Stockholm area and assigned each road with standard widths according to swedish regulation.
+For the road overlap study, a program has taken a slice of the OSM dataset in the Stockholm area and assigned each road with standard widths according to Swedish regulation.
 It has then identified critical areas in the OSM dataset, where the paths are so wide that they collide with existing map features.
@@
 
-The secondary subquestion is:
+The secondary sub-question is:
 
@@
 
-The positional accuracy will be necessary to carry out the investigation of the second subquestion: the rate of collisions between roads, since it will determine the maximum tolerance in distance between road points before a collision is certain to have occurred.
+The positional accuracy will be necessary to carry out the investigation of the second sub-question: the rate of collisions between roads, since it will determine the maximum tolerance in distance between road points before a collision is certain to have occurred.
 For the secondary research question: the geodata precision study, the positional accuracy will be estimated by comparing feature points between the OSM building dataset and the Swedish University of Agricultural Sciences (SLU) building dataset with known accuracy.
@@
 
-The \textbf{road network} is what ties the city together as a whole, and forms the grid that divides the lots, quarters and neighborhoods, giving the city grid its unique aestethic character and defining possible building arrangements.
+The \textbf{road network} is what ties the city together as a whole, and forms the grid that divides the lots, quarters and neighborhoods, giving the city grid its unique aesthetic character and defining possible building arrangements.
 Thus, for any procedural city reproduction model it will be necessary to have a robust procedure for extracting road geodata and generating the 3D representation of it without compromising the data integrity.
@@
 This creates a possible generative space of many different models that can be created from the same data and rule set, and with the induction of random fluctuations, many different models can be generated from the same dataset and rules.
-Using this procedure, a procedural model can structure the façade of a building, add texture and add minor detail models such as windows, doors and ornaments, and so on.
+Using this procedure, a procedural model can structure the facade of a building, add texture and add minor detail models such as windows, doors and ornaments, and so on.
 The maximum accuracy that can be obtained in these finer details is largely dependent on the availability of such data, and as most commercially available GIS data only serve crude building footprints and road placement, the model will often have to make guesses about the finer detail, which sacrifices accuracy to the real world.
@@
 
-The methods used for the geodata precision study were largely based on those used by \textbf{(Fan et al., 2014)} in their study of the quality of the OSM Munich dataslice in 2014.
+The methods used for the geodata precision study were largely based on those used by \textbf{(Fan et al., 2014)} in their study of the quality of the OSM Munich data slice in 2014.
 First a feature map was found, which mapped individual buildings between the OSM dataset and a reference dataset.
@@
 The motivation behind using the positional accuracy as an upper bound for correction is to tread the positional accuracy as a maximum resolution of the map.
-Making translation withing the map resolution will preserve the logical consistency to OSM source data, and preserve aestetic characteristics of the map, both between features and in the map as a whole
+Making translation withing the map resolution will preserve the logical consistency to OSM source data, and preserve aesthetic characteristics of the map, both between features and in the map as a whole
 
@@
 Once a feature map has been obtained, the next task will be to find a matching vertex set among the vertices of the corresponding building footprint polygons.
-Provided that the corresponding building footprints are true representations of the real life building, these pairs of matching verties must match the same part of the building.
+Provided that the corresponding building footprints are true representations of the real life building, these pairs of matching vertices must match the same part of the building.
 The most common case is that a vertex represents a corner in the footprint polygon, and in these cases it must hold that two matching vertices represent the same corner.
@@
 To avoid this effect, key points were extracted using the Douglas-Peucker algorithm, in order to create a simplified footprint with less points that still retain information about the rough features of the detailed footprint.
-This will eliminate dense vertex clusters, as well as any vertices that lie in roughly straight lines, in favour of creating polygons with reduced vertex counts.
+This will eliminate dense vertex clusters, as well as any vertices that lie in roughly straight lines, in favor of creating polygons with reduced vertex counts.
 A useful property of the Douglas-Peucker algorithm is that two highly similar polygons of different vertex counts will be reduced to resulting polygons with similar vertex counts.
@@
 
-The OSM project is a wiki-like geodatabase based on Volunteered Geographical Information (VGI).
+The OSM project is a wiki-like geographic database based on Volunteered Geographical Information (VGI).
 OSM road data and building footprints are commonly obtained by manually tracing features in commercially available satellite images.
@@
 
-It was found in the feature overlap study that the individual cases of overlap could be classified into types, where all collisions of the same type could in theory be corrected by the same algoritm.
+It was found in the feature overlap study that the individual cases of overlap could be classified into types, where all collisions of the same type could in theory be corrected by the same algorithm.
 Note however that since this study focused exclusively on the Stockholm area, and the feature collisions detected in Stockholm could be of types that are characteristic to Stockholm in particular.
-There could very well be additional types of collisions that are exclusive to cities with other unique city grid aestethics, which would not have been considered by this particular project.
+There could very well be additional types of collisions that are exclusive to cities with other unique city grid aesthetics, which would not have been considered by this particular project.
 
@@
 It turns out that, with the current means of data acquisition from OSM being limited to building footprints, using OSM data a model can only produce LOB-B models at best.
-If a modeler wishes to acquire building assets at a LOD-C level, they would have to complete the data with other sources, or rely on procedural methods to populate crude city models with additional details and building features that give the models more aestethic depth, but do not necessarily reflect reality.
+If a modeler wishes to acquire building assets at a LOD-C level, they would have to complete the data with other sources, or rely on procedural methods to populate crude city models with additional details and building features that give the models more aesthetic depth, but do not necessarily reflect reality.
 
@@
 It turns out however, that many studies in urban development have a highly case-by-case nature.
-Studies acroos many different areas have been done before that required very different knowledge about a city's geometry, e.g. for the purposes of modeling heat development
+Studies across many different areas have been done before that required very different knowledge about a city's geometry, e.g. for the purposes of modeling heat development
 \textbf{(Nakata-Osaki, Souza and Rodrigues, 2018)}
@@
 
-According to \textbf{(Stojanovski, 2019)}, software that strives to aid in the urban development process is commonly not sophisticated enough to fully represent the needs and workspaces of urban planners in practise.
+According to \textbf{(Stojanovski, 2019)}, software that strives to aid in the urban development process is commonly not sophisticated enough to fully represent the needs and workspaces of urban planners in practice.
 Stojanovski goes on to claim that a lot of research in this field requires detailed information that goes beyond simply knowing the geometry of the city.
@@
 It may also be necessary for a researcher to implement their own data structures on top of the existing city model, with will increase the complexity of the problem even further.
-The case-by-case nature of urban development projects also means that it is unlikely that a single piece of software can cater to the very differing needs of data and data structures for variuos projects.
+The case-by-case nature of urban development projects also means that it is unlikely that a single piece of software can cater to the very differing needs of data and data structures for various projects.
 
@@
 
-\section*{Acknowledgements}
+\section*{Acknowledgments}
 The author would like to thank KTH associate professor Christopher Peters for supervising this master's thesis project, as well as KTH postdoctoral researcher Todor Stojanovski for expert consultation in the field of urban development.

### 2020-08-25 report/report.tex
Remove SLU abbreviation from abstract

@@
 Any geographic dataset has a limit on its positional accuracy, which is determined by the theoretical precision of the measuring methods.
-In this project, the accuracy was obtained by comparing coordinate points of OSM building footprints to those in the Swedish University of Agricultural Sciences (SLU) building dataset, whose accuracy is known.
+In this project, the accuracy was obtained by comparing coordinate points of OSM building footprints to those in the Swedish University of Agricultural Sciences building dataset, whose accuracy is known.
 \end{abstract}
@@
 Varje geografisk datamängd har en gräns på hur väl den stämmer överrens med verkligheten, som bestäms av vilka mätmetoder som har använts i datans insamling.
-I detta projekt uppskattades felmarginalen genom att jämföra byggnaders koordinatpunkter i datamängen med de från en referenskarta från Svenska Lantbruksuniversitetet (SLU) med känd felmarginal.
+I detta projekt uppskattades felmarginalen genom att jämföra byggnaders koordinatpunkter i datamängen med de från en referenskarta från Svenska Lantbruksuniversitetet med känd felmarginal.
 \end{abstract}
